### 2014-06-15 paper_skeleton.tex
im about to change how ive written the hamiltonian

\documentclass{article}
%\documentclass[aps, twocolumn, tightenlines,amscd,amsmath,amssymb,verbatim]{revtex4}
\usepackage{latexsym}
\usepackage{amssymb,amsmath}
\usepackage{custom2}
\usepackage{graphicx} % for figures
\usepackage{epstopdf} % so can use EPS or PDF figures
%\usepackage{subfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{url}
\usepackage{amssymb,amsfonts}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage[pdftex]{hyperref}
\usepackage{lscape}
\captionsetup{justification=RaggedRight, singlelinecheck=false}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\argmax}{\text{argmax}}
\newcommand{\Tr}{\text{Tr}}
\newcommand{\z}{\mathscr{Z}}
%\newtheorem{claim}{Claim}

\addtolength{\evensidemargin}{-.5in}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\textwidth}{1.4in}
\addtolength{\textheight}{1.4in}
\addtolength{\topmargin}{-.5in}

\pagestyle{empty}

\begin{document}

\begin{center}
{\bf \LARGE{Information Gathering Strategies}}
\vspace{10pt}
\\ Eleanor Brush
\\ May 1, 2014
\end{center}

\tableofcontents

\section{Introduction}
Birds in flocks, like animals in many types of social groups, can use their peers to learn about the environment and update their opinions about where to move in that environment.  Previous work on starlings has measured the number of other birds an individuals pays attention to and shown that this number of ``neighbors" leads to a social network that is conducive to the whole flock reaching consensus.  It is unclear, however, why an individual would change its behavior, i.e. the number of neighbors it learns from, to achieve this outcome.  Individual birds should optimize how well they learn about the environment and should minimize the cognitive and time costs of learning.  I find the circumstances in which this individual level optimization can produce the observed group level optimization.

\section{Learning dynamics}
In this model, an individual's opinion changes with a probability that depends on its social interactions and the strength of the environmental stimulus. Specifically, an individual's opinion $\sigma_i$ can be either $1$ or $-1$ and $\sigma_i$ switches with probability $w_i(\sigma_i)=\frac{1}{2}(-\sigma_i\sum_jA_{ij}\sigma_j )(1-\beta_i\sigma_i)$ where $A_{ij}$ indicates the strength of the connection from $j$ to $i$ and $\beta_i$ indicates the strength of the environmental stimulus perceived by $i$.  The stochastic dynamics of the opinions can be written down with a master equation \cite{Glauber:1963fk}.  The vector $\vec{q}(t)$ of expected values, $q_i(t)=\langle \sigma_i(t)\rangle$, satisfies the following differential equations: \cite{Glauber:1963fk}
\begin{equation*}
\frac{d\vec{q}}{dt}=A\vec{q}+\vec{\beta}^T(\vec{1}-\vec{q}).
\end{equation*}
%Similarly, the matrix $R(t)$ of covariances, $R_{ij}(t)=\langle \sigma_i(t)\sigma_j\rangle$, satisfies the following differential equations:
%\begin{equation*}
%\frac{dR}{dt}=AR+RA^T
%\end{equation*}
These stochastic dynamics will lead to an equilibrium probability distribution over sets of opinions that agrees with the Boltzmann equilibrium distribution for a system with energy or Hamiltonian given by 
\begin{equation*}
E(\vec{q})=-\vec{q}^TA\vec{q}+\vec{\beta}^T(\vec{1}-\vec{q})
\end{equation*}

\section{Individual-level measures of performance}
The group is made up of $N$ individuals.  Each individual has a strategy that determines how many other individuals he pays attention to, $n_i=\sum_{j\neq i}{\bf I}(A_{ij}\neq 0)$.  For a given number of iterations, we distribute the individuals randomly in space.  Each individual then pays attention to its $n_i$ nearest neighbors.  This is meant to capture One individual is chosen to be a receiver; it and all those individuals within a radius $r$ of the receiver perceive an external signal of strength $\beta$.  We find the expected opinion of each individual $q_i$ after a period of time $T$.  We refer to each of these random iterations as a signaling event.

In the absence of an external signal, neither opinion is correct.  When an external signal is presented, some individuals will be able to perceive it ($\beta_i>0$) and some individuals will not ($\beta_i=0$).  In this case, having an opinion that matches the external signal ($\sigma_i=1$) is beneficial.  How beneficial it is depends on the kind of signal.  We consider two types of signals: predators and food.  If a predator is present, we assume that whichever individual is least aware of the signal will be predated.  We therefore find, over many signaling events, the probability that an individual will be eaten and an individual's fitness is given by the probability of surviving.  If the signal indicates the presence of food, on the other hand, fitness may be proportional to how well an individual learns the external signal.

\section{Optimization methods}
To understand what strategies we might expect to find, we are interested in identifying the optimal strategies.  We does this in two ways.  First, using the framework of adaptive dynamics in finite populations, we identify the evolutionarily stable strategy.  This framework assumes that there is a homogeneous population into which an mutant individual tries to invade.  If strategies change on a learning rather than an evolutionary timescale, or if there is just a lot of variation in the population, we might expect individuals to try to optimize their strategies in the context of a heterogeneous population.  This is our second method of optimization.

\section{ESS }
\subsection{Predation }

\subsection{Access to food }

\section{Learned optima }

\subsection{Predation }

\subsection{Access to food }
\section{Comparison to data}

\section{Discussion}

\section{Appendix}

\begin{align*}
H(\vec{v}))&=-\vec{v}^TA\vec{v}+\vec{\beta}^T(\vec{1}-\vec{v})
%\\&=\vec{v}A\vec{v}-\vec{\beta}^T\vec{v}+\vec{\beta}^T\vec{1}
\end{align*}
Define $V=\sum_kv_k/N$ and $\vec{y}=(\vec{v}-V\vec{1})/V$.  Then $\vec{v}=V(\vec{1}+\vec{y})$.  
\begin{align*}
H(\vec{y},V)&=-V^2(\vec{1}^T+\vec{y}^T)A(\vec{1}+\vec{y})+\vec{\beta}^T(\vec{1}-V(\vec{1}+\vec{y}))
\\&=-V^2(\vec{1}^T+\vec{y}^T)A\vec{y}-V\vec{\beta}^T\vec{y}+(1-V)\vec{\beta}^T\vec{1} \text{ since $A\vec{1}=\vec{0}$}
\\&=-V^2\vec{y}^TA\vec{y}+(-V^2\vec{1}^TA-V\vec{\beta}^T)\vec{y}+(1-V)\vec{\beta}^T\vec{1}
\end{align*}
Define $Q\in M_{N-1,N}$ that rotates an $N$-vector away from the consensus vector and define $\vec{z}=Q\vec{y}$ so that $\vec{y}=Q^T\vec{z}$ since $\sum_ky_k=0$.
\begin{align*}
H(\vec{z},V)&=-V^2\vec{z}^TQAQ^T\vec{z}+(-V^2\vec{1}^TAQ^T-V\vec{\beta}^TQ^T)\vec{z}+(1-V)\vec{\beta}^T\vec{1}
\\&=-V^2\vec{z}^TM\vec{z}+(-V^2\vec{\sigma}_1^T-V\vec{\sigma}_2^T)\vec{z}+(1-V)\vec{\beta}^T\vec{1}
\\ \Rightarrow P(\vec{z},V)&=\exp\left(V^2\vec{z}^TM\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+(V-1)\vec{\beta}^T\vec{1}\right)
\end{align*}
where $M=QAQ^T$, $\vec{\sigma}_1=QA^T\vec{1}$, and $\vec{\sigma}_2=Q\vec{\beta}$.

\begin{fact}
\begin{align*}
\int\exp\left(-\frac{1}{2}\vec{z}^TM\vec{z}+\vec{B}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{\det M}}\exp\left(-\frac{1}{2}\vec{B}^TM^{-1}\vec{B}\right)
\\\Rightarrow \int\exp\left(-\vec{z}^TM\vec{z}+\vec{B}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{2^{N-1}\det M}}\exp\left(-\frac{1}{4}\vec{B}^TM^{-1}\vec{B}\right)
\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det M}}\exp\left(-\frac{1}{4}\vec{B}^TM^{-1}\vec{B}\right)
\end{align*}
\end{fact}

\begin{align*}
\z&=\int_{-\infty}^\infty\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TM\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+(V-1)\vec{\beta}^T\vec{1}\right)d^NzdV
\\&=\int_{-\infty}^\infty \exp((V-1)\vec{\beta}^T\vec{1})\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TM\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}\right)d^NzdV
\\&=\sqrt{\pi^{N-1}}\int_{-\infty}^\infty\exp((V-1)\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}\det{M}}}\right)\exp\left(-\frac{1}{4}\vec{\sigma}_1^TM^{-1}\vec{\sigma}_1\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV
\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det M}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TM^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\exp((V-1)\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}}}\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV
\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det M}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TM^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left((V-1)\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV
\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\Pi_a\lambda_a}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TM^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left((V-1)\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV 
\\&\text{ where $\lambda_a$ are the eigenvalues of $M$}
\\ \Rightarrow -\log(\z)&=\text{constants}+\frac{1}{2}\sum_a\log(\lambda_a)+\frac{1}{4}\vec{\sigma}_1^TM^{-1}\vec{\sigma_1}-\log(f(M))
\end{align*}

where $f(M)=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left((V-1)\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV $.

\begin{align*}
\Rightarrow\frac{\partial -\log(\z)}{\partial M_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}M^{-1}_{ki}M^{-1}_{jl}\sigma_{1l}-\frac{\partial \log f}{\partial M_{ij}}
\\\Rightarrow\frac{\partial -\log(\z)}{\partial M_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}M^{-1}{ki}M^{-1}_{jl}\sigma_{1l}-\frac{1}{f}\frac{\partial f}{\partial M_{ij}}
\end{align*}

\begin{align*}
\frac{\partial f}{\partial M_{ij}}&=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left((V-1)\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)\left(-\frac{1}{4V}\sum_{kl}\sigma_{2k}M^{-1}_{ki}M^{-1}_{jl}\sigma_{2l}\right)dV
\\&=\left(-\frac{1}{4}\sum_{kl}\sigma_{2k}M^{-1}_{ki}M^{-1}_{jl}\sigma_{2l}\right)\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left((V-1)\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV
\end{align*}

\begin{align*}
\Rightarrow \langle z_iz_j\rangle &=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}M^{-1}{ki}M^{-1}_{jl}\sigma_{1l}+\left(\frac{1}{4}\sum_{kl}\sigma_{2k}M^{-1}_{ki}M^{-1}_{jl}\sigma_{2l}\right)\frac{\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left((V-1)\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV}{\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left((V-1)\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV}
\end{align*}

\nocite{*}
\bibliographystyle{plain}
\bibliography{info_evo}

\end{document}

### 2014-06-16 paper_skeleton.tex
about to write some appendix math

@@
 \begin{equation*}
-\frac{d\vec{q}}{dt}=A\vec{q}+\vec{\beta}^T(\vec{1}-\vec{q}).
+\frac{d\vec{q}}{dt}=M\vec{q}+\vec{\beta}^T(\vec{1}-\vec{q}).
 \end{equation*}
-%Similarly, the matrix $R(t)$ of covariances, $R_{ij}(t)=\langle \sigma_i(t)\sigma_j\rangle$, satisfies the following differential equations:
-%\begin{equation*}
-%\frac{dR}{dt}=AR+RA^T
-%\end{equation*}
+We simplify this by writing $\frac{d\vec{q}}{dt}=A\vec{q}+\vec{\beta}^T\vec{1}$, where $A=M-B$ and $B$ is a diagonal matrix with the values of $\vec{\beta}$ along the diagonal.
+
 These stochastic dynamics will lead to an equilibrium probability distribution over sets of opinions that agrees with the Boltzmann equilibrium distribution for a system with energy or Hamiltonian given by
@@
 \section{Appendix}
+\subsection{Equivalence of dynamics and Hamiltonian }
 
+\subsection{Numerically evaluating dynamics }
+
+\subsection{Calculating correlations }
 \begin{align*}
-H(\vec{v}))&=-\vec{v}^TA\vec{v}+\vec{\beta}^T(\vec{1}-\vec{v})
+H(\vec{v})&=-\vec{v}^TA\vec{v}-\vec{\beta}^T\vec{v}
 %\\&=\vec{v}A\vec{v}-\vec{\beta}^T\vec{v}+\vec{\beta}^T\vec{1}
@@
 \begin{align*}
-H(\vec{y},V)&=-V^2(\vec{1}^T+\vec{y}^T)A(\vec{1}+\vec{y})+\vec{\beta}^T(\vec{1}-V(\vec{1}+\vec{y}))
-\\&=-V^2(\vec{1}^T+\vec{y}^T)A\vec{y}-V\vec{\beta}^T\vec{y}+(1-V)\vec{\beta}^T\vec{1} \text{ since $A\vec{1}=\vec{0}$}
-\\&=-V^2\vec{y}^TA\vec{y}+(-V^2\vec{1}^TA-V\vec{\beta}^T)\vec{y}+(1-V)\vec{\beta}^T\vec{1}
+H(\vec{y},V)&=-V^2(\vec{1}^T+\vec{y}^T)A(\vec{1}+\vec{y})-\vec{\beta}^TV(\vec{1}+\vec{y})
+\\&=-V^2(\vec{1}^T+\vec{y}^T)A\vec{y}-V\vec{\beta}^T\vec{y}-V\vec{\beta}^T\vec{1} \text{ since $A\vec{1}=\vec{0}$}
+\\&=-V^2\vec{y}^TA\vec{y}+(-V^2\vec{1}^TA-V\vec{\beta}^T)\vec{y}-V\vec{\beta}^T\vec{1}
 \end{align*}
@@
 \begin{align*}
-H(\vec{z},V)&=-V^2\vec{z}^TQAQ^T\vec{z}+(-V^2\vec{1}^TAQ^T-V\vec{\beta}^TQ^T)\vec{z}+(1-V)\vec{\beta}^T\vec{1}
-\\&=-V^2\vec{z}^TM\vec{z}+(-V^2\vec{\sigma}_1^T-V\vec{\sigma}_2^T)\vec{z}+(1-V)\vec{\beta}^T\vec{1}
-\\ \Rightarrow P(\vec{z},V)&=\exp\left(V^2\vec{z}^TM\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+(V-1)\vec{\beta}^T\vec{1}\right)
+H(\vec{z},V)&=-V^2\vec{z}^TQAQ^T\vec{z}+(-V^2\vec{1}^TAQ^T-V\vec{\beta}^TQ^T)\vec{z}-V\vec{\beta}^T\vec{1}
+\\&=V^2\vec{z}^TP\vec{z}+(-V^2\vec{\sigma}_1^T-V\vec{\sigma}_2^T)\vec{z}-V\vec{\beta}^T\vec{1}
+\\ \Rightarrow P(\vec{z},V)&=\exp\left(-V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+V\vec{\beta}^T\vec{1}\right)
 \end{align*}
-where $M=QAQ^T$, $\vec{\sigma}_1=QA^T\vec{1}$, and $\vec{\sigma}_2=Q\vec{\beta}$.
+where $P=-QAQ^T$, $\vec{\sigma}_1=QA^T\vec{1}$, and $\vec{\sigma}_2=Q\vec{\beta}$.
 
@@
 \begin{align*}
-\int\exp\left(-\frac{1}{2}\vec{z}^TM\vec{z}+\vec{B}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{\det M}}\exp\left(-\frac{1}{2}\vec{B}^TM^{-1}\vec{B}\right)
-\\\Rightarrow \int\exp\left(-\vec{z}^TM\vec{z}+\vec{B}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{2^{N-1}\det M}}\exp\left(-\frac{1}{4}\vec{B}^TM^{-1}\vec{B}\right)
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det M}}\exp\left(-\frac{1}{4}\vec{B}^TM^{-1}\vec{B}\right)
+\int\exp\left(-\frac{1}{2}\vec{z}^TP\vec{z}+\vec{\sigma}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{\det P}}\exp\left(-\frac{1}{2}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
+\\\Rightarrow \int\exp\left(-\vec{z}^TP\vec{z}+\vec{\sigma}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{2^{N-1}\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
+\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
 \end{align*}
@@
 \begin{align*}
-\z&=\int_{-\infty}^\infty\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TM\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+(V-1)\vec{\beta}^T\vec{1}\right)d^NzdV
-\\&=\int_{-\infty}^\infty \exp((V-1)\vec{\beta}^T\vec{1})\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TM\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}\right)d^NzdV
-\\&=\sqrt{\pi^{N-1}}\int_{-\infty}^\infty\exp((V-1)\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}\det{M}}}\right)\exp\left(-\frac{1}{4}\vec{\sigma}_1^TM^{-1}\vec{\sigma}_1\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det M}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TM^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\exp((V-1)\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}}}\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det M}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TM^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left((V-1)\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\Pi_a\lambda_a}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TM^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left((V-1)\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV
-\\&\text{ where $\lambda_a$ are the eigenvalues of $M$}
-\\ \Rightarrow -\log(\z)&=\text{constants}+\frac{1}{2}\sum_a\log(\lambda_a)+\frac{1}{4}\vec{\sigma}_1^TM^{-1}\vec{\sigma_1}-\log(f(M))
+\z&=\int_{-\infty}^\infty\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+V\vec{\beta}^T\vec{1}\right)d^NzdV
+\\&=\int_{-\infty}^\infty \exp(V\vec{\beta}^T\vec{1})\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}\right)d^NzdV
+\\&=\sqrt{\pi^{N-1}}\int_{-\infty}^\infty\exp(V\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}\det{P}}}\right)\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
+\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det PP}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\exp(V\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}}}\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
+\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
+\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\Pi_a\lambda_a}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
+\\&\text{ where $\lambda_a$ are the eigenvalues of $P$}
+\\ \Rightarrow -\log(\z)&=\text{constants}+\frac{1}{2}\sum_a\log(\lambda_a)+\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma_1}-\log(f(P))
 \end{align*}
 
-where $f(M)=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left((V-1)\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV $.
+where $f(P)=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV $.
 
 \begin{align*}
-\Rightarrow\frac{\partial -\log(\z)}{\partial M_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}M^{-1}_{ki}M^{-1}_{jl}\sigma_{1l}-\frac{\partial \log f}{\partial M_{ij}}
-\\\Rightarrow\frac{\partial -\log(\z)}{\partial M_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}M^{-1}{ki}M^{-1}_{jl}\sigma_{1l}-\frac{1}{f}\frac{\partial f}{\partial M_{ij}}
+\Rightarrow\frac{\partial -\log(\z)}{\partial P_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{1l}-\frac{\partial \log f}{\partial P_{ij}}
+\\\Rightarrow\frac{\partial -\log(\z)}{\partial P_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}{ki}P^{-1}_{jl}\sigma_{1l}-\frac{1}{f}\frac{\partial f}{\partial P_{ij}}
 \end{align*}
@@
 \begin{align*}
-\frac{\partial f}{\partial M_{ij}}&=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left((V-1)\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)\left(-\frac{1}{4V}\sum_{kl}\sigma_{2k}M^{-1}_{ki}M^{-1}_{jl}\sigma_{2l}\right)dV
-\\&=\left(-\frac{1}{4}\sum_{kl}\sigma_{2k}M^{-1}_{ki}M^{-1}_{jl}\sigma_{2l}\right)\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left((V-1)\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV
+\frac{\partial f}{\partial P_{ij}}&=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)\left(-\frac{1}{4V}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)dV
+\\&=\left(-\frac{1}{4}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
 \end{align*}
@@
 \begin{align*}
-\Rightarrow \langle z_iz_j\rangle &=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}M^{-1}{ki}M^{-1}_{jl}\sigma_{1l}+\left(\frac{1}{4}\sum_{kl}\sigma_{2k}M^{-1}_{ki}M^{-1}_{jl}\sigma_{2l}\right)\frac{\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left((V-1)\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV}{\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left((V-1)\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TM^{-1}\vec{\sigma}_2\right)dV}
+\Rightarrow \langle z_iz_j\rangle &=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}{ki}P^{-1}_{jl}\sigma_{1l}+\left(\frac{1}{4}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)\frac{\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV}{\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV}
 \end{align*}

### 2014-09-25 paper_skeleton.tex
about to change signaling events to avoid problems with network connectedness

@@
 The stochastic dynamics of the opinions can be written down with a master equation \cite{Glauber:1963fk}.
-The vector $\vec{q}(t)$ of expected values, $q_i(t)=\langle \sigma_i(t)\rangle$, satisfies the following differential equations: \cite{Glauber:1963fk}
+The vector $\vec{v}(t)$ of expected values, $q_i(t)=\langle \sigma_i(t)\rangle$, satisfies the following differential equations: \cite{Glauber:1963fk}
 \begin{equation*}
-\frac{d\vec{q}}{dt}=M\vec{q}+\vec{\beta}^T(\vec{1}-\vec{q}).
+\frac{d\vec{v}}{dt}=M\vec{v}+\vec{\beta}^T(\vec{1}-\vec{v}).
 \end{equation*}
-We simplify this by writing $\frac{d\vec{q}}{dt}=A\vec{q}+\vec{\beta}^T\vec{1}$, where $A=M-B$ and $B$ is a diagonal matrix with the values of $\vec{\beta}$ along the diagonal.
+We simplify this by writing $\frac{d\vec{v}}{dt}=A\vec{v}+\vec{\beta}^T\vec{1}$, where $A=M-B$ and $B$ is a diagonal matrix with the values of $\vec{\beta}$ along the diagonal.
 
@@
 \begin{equation*}
-E(\vec{q})=-\vec{q}^TA\vec{q}+\vec{\beta}^T(\vec{1}-\vec{q})
+E(\vec{v})=-\vec{v}^TA\vec{v}-\vec{\beta}^T\vec{v}
 \end{equation*}
@@
 \subsection{Equivalence of dynamics and Hamiltonian }
+\begin{claim}
+If $H(\vec{v})=-\vec{v}^TA\vec{v}-\vec{\beta}^T\vec{v}$ and $\frac{d\vec{v}}{dt}=A\vec{v}+\vec{\beta}$ then $\nabla H\cdot \frac{d\vec{v}}{dt}\leq 0$.
+\end{claim}
+\begin{pf}
+\begin{align*}
+H(\vec{v})&=-\sum_{ij}A_{ij}v_iv_j-\sum_i\beta_iv_i
+\\\Rightarrow\frac{\partial H}{\partial v_i}&=-\sum_{j\neq i}A_{ij}v_j-\sum_{j\neq i}A_{ji}v_j-2A_{ii}v_i-\beta_i
+\\&=-\sum_j(A_{ij}+A^T_{ij})v_j-\beta_i
+\\ \Rightarrow \nabla H&=-(A+A^T)\vec{v}-\vec{\beta}
+\\ \Rightarrow \nabla H\cdot \frac{d\vec{v}}{dt}&=-\vec{v}^T(A+A^T)A\vec{v}-\vec{v}^T(A+A^T)\vec{\beta}-\vec{\beta}^TA\vec{v}-\vec{\beta}^T\vec{\beta}
+\\&=-\vec{v}^TA^2\vec{v}-\vec{v}^TA^TA\vec{v}-\vec{v}^TA\vec{\beta}-2\vec{v}^TA^T\vec{\beta}-\vec{\beta}^T\vec{\beta}
+\end{align*}
+If $A$ is negative semi-definite, then $A^2$ is positive semi-definite and $A^TA$ is positive semi-definite.
+(Consider that $\vec{v}^TA^TA\vec{v}=(A\vec{v})^T(A\vec{v})$ so that $\vec{v}A^TA\vec{v}=0$ if $A\vec{v}=0$ and $\vec{v}^TA^TA\vec{v}<0$ otherwise.) Therefore, $-\vec{v}^TA^2\vec{v}\leq 0$ and $-\vec{v}^TA^TA\vec{v}\leq 0$.
+At equilibrium, $\vec{v}$ will be ``close" to $\vec{\beta}$ so that $$-\vec{v}^TA\vec{\beta}-2\vec{v}^TA^T\vec{\beta}\sim-\vec{\beta}^TA\vec{\beta}$$
+\end{pf}
 
 \subsection{Numerically evaluating dynamics }
+\begin{claim}
+Let $\Phi(t)$ be the fundamental matrix solution to the homogeneous differential equations $\frac{d\vec{x}}{dt}=A\vec{x}$ so that $\Phi(t)=V\Lambda(t)$ where the columns of $V$ are the eigenvectors of $A$ and $$\Lambda(t)=\text{diag}(e^{\lambda_1 t},\dots,e^{\lambda_n t})$$ where $\{\lambda_i\}$ are the eigenvalues of $A$.
+(Note that if $\lambda_i=0$ then $e^{\lambda _i t}=1$ for all $t$.) Let $V^\dagger$ be the pseudoinverse of $V$.
+Let $\vec{c}=V^\dagger\vec{X_0}$.
+Let $$G(t)=\text{diag}(1/\lambda_1(1-e^{-\lambda_1t}),\dots,1/\lambda_n(1-e^{-\lambda_nt}))V^\dagger\vec{\beta}.$$
+(If $\lambda_i=0$ then the corresponding entry of the first part of $G(t)$ will be given by $t$ rather than the form above.)  THEN if we let $x(t)=\Phi(t)G(t)+\Phi(t)\vec{c}$, $X(t)$ solves the inhomogeneous equations $\frac{dx}{dt}=A\vec{x}+\vec{\beta}$ with initial conditions $x(0)=\vec{X_0}$.
+\end{claim}
+
+\begin{pf}
+It is clear that
+\begin{align*}
+\frac{d\Phi\vec{c}}{dt}&=A\Phi(t)\vec{c} \text{ and } \Phi(0)\vec{c}=\vec{X_0}.
+\end{align*}
+Now, 
+\begin{align*}
+\Phi(t)G(t)&=V
+\left(\begin{array}{ccccc}
+e^{\lambda_1 t}  &\dots &0
+\\ \vdots & \ddots &\vdots
+\\ 0 & \dots & e^{\lambda_nt}
+ \end{array}\right)
+ \left(\begin{array}{ccccc}
+\frac{1}{\lambda_1}(1-e^{\lambda_1 t})  &\dots &0
+\\ \vdots & \ddots &\vdots
+\\ 0 & \dots &\frac{1}{\lambda_n} (1-e^{\lambda_nt})
+ \end{array}\right)
+ V^\dagger\vec{\beta}
+ \\&=V
+ \left(\begin{array}{ccccc}
+\frac{1}{\lambda_1}(e^{\lambda_1 t}-1)  &\dots &0
+\\ \vdots & \ddots &\vdots
+\\ 0 & \dots &\frac{1}{\lambda_n} (e^{\lambda_nt}-1)
+ \end{array}\right)
+ V^\dagger\vec{\beta}
+ \\ \Rightarrow \frac{d\Phi(t)G(t)}{dt}&=V
+ \left(\begin{array}{ccccc}
+e^{\lambda_1 t}  &\dots &0
+\\ \vdots & \ddots &\vdots
+\\ 0 & \dots &e^{\lambda_nt}
+ \end{array}\right)
+ \\&=V
+ \left(\begin{array}{ccccc}
+e^{\lambda_1 t}  &\dots &0
+\\ \vdots & \ddots &\vdots
+\\ 0 & \dots &e^{\lambda_nt}
+ \end{array}\right)
+ V^\dagger\vec{\beta} -VV^\dagger\vec{\beta}+\vec{\beta}
+ \\&=V
+ \left(\begin{array}{ccccc}
+e^{\lambda_1 t}-1  &\dots &0
+\\ \vdots & \ddots &\vdots
+\\ 0 & \dots &e^{\lambda_nt}-1
+ \end{array}\right)
+ V^\dagger\vec{\beta} +\vec{\beta}
+  \\&=V
+ \left(\begin{array}{ccccc}
+\lambda_1e^{\lambda_1t}\cdot \frac{1}{\lambda_1}(1-e^{-\lambda_1 t})  &\dots &0
+\\ \vdots & \ddots &\vdots
+\\ 0 & \dots &\lambda_ne^{\lambda_nt}\cdot\frac{1}{\lambda_n}(1-e^{-\lambda_nt})
+ \end{array}\right)
+ V^\dagger\vec{\beta} +\vec{\beta}
+  \\&=AV\Lambda(t)
+ \left(\begin{array}{ccccc}
+ \frac{1}{\lambda_1}(1-e^{-\lambda_1 t})  &\dots &0
+\\ \vdots & \ddots &\vdots
+\\ 0 & \dots &\frac{1}{\lambda_n}(1-e^{-\lambda_nt})
+ \end{array}\right)
+ V^\dagger\vec{\beta} +\vec{\beta}
+ \\&=AV\Lambda(t)G(t)+\vec{\beta}
+ \\&=A\Phi(t)G(t)+\vec{\beta}
+\end{align*}
+Therefore $\frac{dx}{dt}=A\Phi(t)G(t)+A\Phi(t)\vec{c}+\vec{\beta}=Ax(t)+\vec{\beta}$ and since $G(0)=\vec{0}$, $x(0)=\vec{X_0}$.
+\end{pf}
 
 \subsection{Calculating correlations }
+The following is derived from the work in \cite{Bialek:2013fk}.
 \begin{align*}
@@
 \\&=V^2\vec{z}^TP\vec{z}+(-V^2\vec{\sigma}_1^T-V\vec{\sigma}_2^T)\vec{z}-V\vec{\beta}^T\vec{1}
-\\ \Rightarrow P(\vec{z},V)&=\exp\left(-V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+V\vec{\beta}^T\vec{1}\right)
+\\ \Rightarrow P(\vec{z},V)&=\frac{1}{\z}\exp\left(-V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+V\vec{\beta}^T\vec{1}\right)
 \end{align*}

### 2014-09-25 expected_spin.m

@@ -1,25 +1,31 @@
-function times = expected_spin(M,T,beta)
+function q = expected_spin(M,T,beta)
+%solves the differential equation dq/dt=M*q+beta*(1-q)  -->
+%dq/dt=A*q+beta where A=M-diag(beta)
+
 % N=length(strategy);
 % M=makenet_normed(strategy);
-
+A=M-diag(beta);
 N=size(M,1);
 
 if nargin==2
     beta=1*ones(N,1);
 end
 
-[vecs,vals]=eig(M);
+[vecs,vals]=eig(A);
 V=vecs;
 vals=diag(vals);
 nonzero=abs(vals)>0.00001;
 nonzerovals=vals(nonzero);
 
+%if there is a zero eigenvalue set the eigenvector to all ones
 %if there are two eigenvectors for eigenvalue 0 I specify what they should
 %be here
 indices=1:N;
 where=indices(~nonzero);
 vhat=1/sqrt(N)*ones(N,1);
-V(:,where(1))=ones(N,1);
+if sum(~nonzero)==1
+    V(:,where(1))=ones(N,1);
+end
 if sum(~nonzero)>1
     L=sum(~nonzero);
     for i=1:(L-1)
@@ -30,8 +36,11 @@ end
 
 % mu=pinv(V)*beta;s
 
-c=zeros(N,1);
-c(where(1))=1;
+% c=zeros(N,1);
+% c(where(1))=1;
+
+initialconditions=.5*ones(N,1);
+c=pinv(V)*initialconditions;
 
 l=exp(nonzerovals*T);
 Lhom=zeros(1,N);
@@ -59,8 +68,8 @@ instantaneous=Uinhom*beta+Uhom*c;
 % U=V*L/V;
 % integrated=U*beta;
 
-% times=[instantaneous,integrated];
+% q=[instantaneous,integrated];
 
-times=instantaneous;
+q=instantaneous;
 end
 

### 2014-09-25 finding_equilibria_v3.m

@@ -15,10 +15,10 @@ nummoves=1000;
 N=20;
 b=1;
 
-Tvals=[.01];
+Tvals=[1];
 NT=length(Tvals);
 % radvals=[.1 .5 1];
-radvals=.1;
+radvals=.25;
 Nr=length(radvals);
 
 ESSseaten=cell(NT,Nr);

### 2014-09-25 paircorrelations.m

@@ -1,4 +1,4 @@
-function [corrs]=paircorrelations(M,beta)
+function corrs=paircorrelations(M,beta)
 
 N=size(M,2);
 
@@ -14,47 +14,32 @@ for j=2:(N-1)
    q(j,:)=v;
 end
 
-% Mstar=q*M*transpose(q);
-% gamma=-transpose(M)*ones(N,1)-beta;
-% gammastar=q*gamma;
-% Mstarinv=inv(Mstar);
-% 
-% [vecs,vals]=eig(Mstar);
-% vals=diag(vals);
-% invvals=1./vals;
-% 
-% zcorrs=zeros(N-1,N-1);
-% for i=1:(N-1)
-%     for j=i:(N-1)
-%         soc=-sum(vecs(i,:).*invvals'.*vecs(j,:));
-%         sig=sum(sum((gammastar'*Mstarinv(:,i))*(gammastar'*Mstarinv(:,j))'))+sum(sum((gammastar'*Mstarinv(:,j))*(gammastar'*Mstarinv(:,i))'));
-% %         sig=0;
-%         zcorrs(i,j)=soc+sig;
-%         zcorrs(j,i)=soc+sig;
-%     end
-% end
+A=M-diag(beta);
+P=-q*A*transpose(q);
+Pinv=inv(P);
 
-negM=-M;
-Mstar=q*negM*transpose(q);
-Mstarinv=inv(Mstar);
-
-[vecs,vals]=eig(Mstar);
+[vecs,vals]=eig(P);
 vals=diag(vals);
 invvals=1./vals;
 
-sigma1=q*transpose(negM)*ones(N,1);
+sigma1=q*transpose(A)*ones(N,1);
 sigma2=q*beta;
 
 zcorrs=zeros(N-1,N-1);
 for i=1:(N-1)
     for j=i:(N-1)
         piece1=sum(vecs(i,:).*invvals'.*vecs(j,:));
-        piece2=sum(sum((sigma1'*Mstarinv(:,i))*(sigma1'*Mstarinv(:,j))'))+sum(sum((sigma1'*Mstarinv(:,j))*(sigma1'*Mstarinv(:,i))'));
-        piece3=sum(sum((sigma2'*Mstarinv(:,i))*(sigma2'*Mstarinv(:,j))'))+sum(sum((sigma2'*Mstarinv(:,j))*(sigma2'*Mstarinv(:,i))'));
[…]

### 2014-09-25 signalingevents.m

@@ -1,11 +1,11 @@
-function [meanscore scorevar probeaten probgettoeat]=signalingevents(strategy,numsigs_permove,nummoves,radius,b,T)
+function [meanH2, meancorrlength, probeaten, probgettoeat]=signalingevents(strategy,numsigs_permove,nummoves,radius,b,T)
 % strategy=randi([1 N/2-1],1,N)*2;
 numsigs_tot=numsigs_permove*nummoves;
 N=max(size(strategy));
 
 scores=zeros(N,numsigs_tot);
 H2norms=zeros(1,numsigs_tot);
-susceptvals=zeros(1,numsigs_tot);
+corrlengths=zeros(1,numsigs_tot);
 
 for i=1:nummoves
     positions=unifrnd(0,1,N,2);
@@ -26,24 +26,25 @@ for i=1:nummoves
         allreceivers=d(receiver,:)<=radius;
         beta(allreceivers)=b;
         v=real(expected_spin(M,T,beta));
-        
-        if sum(abs(v)>1000)>=1
-            j=j;
-        elseif sum(isnan(v))>=1
-            j=j;
-        else  
-            scores(:,(i-1)*numsigs_permove+j)=v;
-%             [h,s]=networkprops(M,'uniform',beta);
-%             H2norms((i-1)*numsigs_permove+j)=h;
-%             susceptvals((i-1)*numsigs_permove+j)=s;
+       
+        scores(:,(i-1)*numsigs_permove+j)=v;
+        h=H2norm(M,'additive');
+        if real(h)<10000 && imag(h)<1
+            H2norms((i-1)*numsigs_permove+j)=h;
+            [~,~,l,~]=correlationlength_mat_single_v3(M,d,b,radius,receiver);
+            corrlengths((i-1)*numsigs_permove+j)=l;
             j=j+1;
         end
     end
 end
-weird=max(abs(scores),[],1)>1000;
-scores(:,weird)=[];
-meanscore=mean(scores,2);
-scorevar=var(scores,[],2);
+
+meanH2=mean(H2norms);
+meancorrlength=mean(corrlengths);
+
+% weird=max(abs(scores),[],1)>1000;
+% scores(:,weird)=[];
+% meanscore=mean(scores,2);
+% scorevar=var(scores,[],2);
 
 [minvals,~]=min(scores);
 minmat=repmat(minvals,N,1);
[…]

### 2014-10-01 information_evolution_08_28_12.tex
trying to be more careful about disjointed networks when calculating correlation length

\documentclass{article}
\usepackage{latexsym}
\usepackage{amssymb,amsmath}
\usepackage{custom2}
\usepackage{graphicx} % for figures
\usepackage{epstopdf} % so can use EPS or PDF figures
\usepackage{subfig}
\usepackage{url}
\usepackage{amssymb,amsfonts}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage[pdftex]{hyperref}
\usepackage{lscape}
\captionsetup{justification=RaggedRight, singlelinecheck=false}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\argmax}{\text{argmax}}
\newcommand{\Tr}{\text{Tr}}
%\newtheorem{claim}{Claim}

\addtolength{\evensidemargin}{-.5in}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\textwidth}{1.4in}
\addtolength{\textheight}{1.4in}
\addtolength{\topmargin}{-.5in}

\pagestyle{empty}

\begin{document}
\begin{center}
\Large

\end{center}

\vspace{0pt}

\begin{center}
{\bf \LARGE{Evolution of Topological Strategies}}
\vspace{10pt}
\\ Eleanor Brush
\\ August 28, 2012
\end{center}

\vspace{0pt}
\normalsize

\section{Linear Algebra Preliminaries}

\begin{claim}
Let $v\in\R^n\setminus\{0\}$.  Let $A\in\R^{n-1\times n}$ be such that the rows of $A$ are an orthonormal basis to the orthogonal complement of $v$.  Then $A^TA=I_n-\frac{1}{||v||^2}vv^T$ (where $v$ is $v$ as a column vector so that $v\times v^T\in\R^{n\times n}$).
\end{claim}

\begin{pf}
Let $\tilde{v}$ and let $V=\left[\begin{array}{cc}A \\ \tilde{v}^T\end{array}\right].$  Then $V\in\R^{n\times n}$ and by construction the rows of $V$ are an orthonormal basis for $\R^n$.  First, we argue that $V^TV=I_n:$
\begin{align*}
V^TV&=M
\\ \Rightarrow VV^TV&=VM
\\ \Rightarrow V&=VM \text{ since the rows of $V$ are orthonormal} 
\\ \Rightarrow M&=I_n \text{ since $V$ is nonsingular}
\end{align*}
Then,
\begin{align*}
V^TV&=\left[A^T \tilde{v}\right]\times\left[ \begin{array}{cc}A\\ \tilde{v}^T\end{array}\right]
\\\Rightarrow (V_T^V)_{ij}&=\sum_kA_{ki}A_{kj}+v_iv_j \text{ for all $i,j$}
\\ \Rightarrow (V_TV)_{ij}&=(A^TA)_{ij}+(\tilde{v}\tilde{v}^T)_{ij} \text{ for all $i,j$}
\\ \Rightarrow V_TV&=A^TA+\tilde{v}\tilde{v^T}
\\ \Rightarrow A^TA&=V_TV-\tilde{v}\tilde{v^T}
\\ \Rightarrow A^TA&=I_n-\frac{1}{||v||^2}vv^T
\end{align*}
\end{pf}

\begin{claim}
Let $B\in\R^{n\times n}$ and $v$ be an eigenvector of $B$ with associated eigenvalue $\lambda$.  Let $A\in\R^{n-1\times n}$ be such that the rows of $A$ are an orthonormal basis to the orthogonal complement of $v$.  Then $\sigma(ABA^T)=\sigma(B)\setminus\{\lambda\}$ (where $\sigma(M)$ is the spectrum of $M$, i.e. the set of eigenvalues of $M$, with each eigenvalue appearing as many times as its algebraic multiplicity).
\end{claim}

\begin{pf}
Let $\tilde{v}=\frac{v}{||v||}$ and let $V=\left[\begin{array}{cc}A \\ \tilde{v}^T\end{array}\right].$  Since $V$ is orthonormal, $VBV^T$ has the same eigenvalues as $B$.
\begin{align*}
VBV^T&=\left[\begin{array}{cc}A \\ \tilde{v}^T\end{array}\right]\times B\times \left[A^T\tilde{v}\right]
\\&=\left[\begin{array}{cc} ABA^T & AB\tilde{v} \\ \tilde{v}^TBA^T & \tilde{v}^TB\tilde{v} \end{array}\right]
\\&=\left[\begin{array}{cc} ABA^T & \lambda A\tilde{v} \\ \tilde{v}^TBA^T & \lambda \tilde{v}^T\tilde{v} \end{array}\right]
\\&=\left[\begin{array}{cc} ABA^T & 0 \\ \tilde{v}^TBA^T & \lambda  \end{array}\right] \text{ by construction of $A$}
\end{align*}
Therefore, the eigenvalues of $VBV^T$ are the eigenvalues of $ABA^T$ and $\lambda$.  But, as noted above, the eigenvalues of $VBV^T$ are the same as the eigenvalues of $B$.  Therefore, $\sigma(ABA^T)=\sigma(B)\setminus\{\lambda\}$ (where $\sigma(M)$ includes each eigenvalue as many times as its algebraic multiplicity).
\end{pf}

\begin{claim}
Let $M\in\R^{n\times n}$ with eigenvalues $\lambda_1\geq\lambda_2\dots\geq\lambda_n$ and associated eigenvectors $u_1,\dots,u_n$.  Let $v\in\R^n$.  Let $u_i$ be the eigenvector such that 
$$i=\argmax_j\left\{\left\langle \frac{v}{||v||},\frac{u_j}{||u_j||}\right\rangle \right\}.$$
Let $A\in\R^{n-1\times n}$ be such that the rows of $A$ are an orthonormal basis to the orthogonal complement of $v$.  Let $\gamma_1,\dots,\gamma_{n-1}$ be the eigenvalues of $AMA^T$ and let $\tilde \lambda_j$, $j=1,\dots,n-1$ be $\{\lambda_j\}\setminus\{\lambda_i\}$ renumbered in decreasing order.  Then 
there is no $j$ such that 
\end{claim}

\begin{claim}\label{sum_bound} (As in \cite{Meyer:2001fk})
Let $A$ be an $n\times n$ nonnegative matrix, i.e. $A_{ij}\geq 0$ for all $i,j$.  Let $r=\rho(A)$ be the spectral radius of $A$.  Then
$$\min_i\sum_ja_{ij}\leq r\leq \max_i\sum_ja_{ij}.$$
\end{claim}

\begin{pf}
 Let $\rho=\max_i\sum_ja_{ij}$.  If $\rho=0$ then $A_{ij}=0$ for all $i,j$ and $\min_i\sum_ja_{ij}=r=\max_i\sum_ja_{ij}=0$.  Assume this is not the case so that $\rho\gneq0$. Then there exists an $E\geq 0$ such that every row sum of $B=A+E$ is $\rho$.

 Note that $A\leq B$ and $B\gneq 0$ (where the comparison is entry-wise).   This gives us that $\rho(A)\leq \rho(B)$.

 By definition of $B$, $B\times\vec{1}=\rho\vec{1}$ so that $(\rho,\vec{1})$ is an eigenpair of $B$.  Since $B$ is positive, there is a unique (up to scalar multiple) nonnegative eigenvector and its eigenvalue is $\rho(B)$.  Therefore, $\rho=\rho(B)$ and $\rho(A)\leq \rho \Rightarrow \rho(A)\leq\max_i\sum_ja_{ij}$.

 By the Collatz-Wielandt formula,
 \begin{align*}
 \rho(A)&=\max_{x\geq 0,x\neq 0}\min_{1\leq i\leq n,x_i\neq 0}\frac{[Ax]_i}{x_i}
 \\ &\geq \min_{1\leq i\leq n}[A\vec{1}]_i
 \\ &=\min_i\sum_ja_{ij}
 \end{align*}
 Therefore, $$\min_i\sum_ja_{ij}\leq r\leq \max_i\sum_ja_{ij}.$$
\end{pf}

\begin{lemma} (As in \cite{Meyer:2001fk})
For $A\in\C^{n\times n}$ let $|A|$ denote the matrix having entries $|a_{ij}|$ and for matrices $B,C\in\R^{n\times n}$ define $B\leq C$ to mean $b_{ij}\leq c_{ij}$ for each $i$ and $j$.  If $|A|\leq B$ than 
$$\rho(A)\leq \rho(|A|)\leq\rho(B).$$
\end{lemma}

\begin{pf}
The triangle inequality implies that $|A^k|\leq |A|^k$ for all positive integers $k$.  Also, $|A|\leq B$ implies that $|A|^k\leq B^k$.  Note that for any matrix norm $||\cdot||:\C^{n\times n}\to\R^+$, $\rho(A)=\lim_{k\to\infty}||A^k||^{1/k}.$  Consider the infinity norm, $||A||_\infty=\max_i\sum_j|a_{ij}|$:
\begin{align*}
||A^k||_\infty=|| (|A^k|) ||_\infty&\leq ||(|A|)^k||_\infty\leq ||B^k||_\infty \text{for all $k$, by the above observations}
\\ \Rightarrow ||A^k||_\infty^{1/k}&\leq ||(|A|)^k||_\infty^{1/k}\leq ||B^k||_\infty^{1/k}
\\ \Rightarrow\lim_{k\to\infty} ||A^k||_\infty^{1/k}&\lim_{k\to\infty}\leq ||(|A|)^k||_\infty^{1/k}\leq \lim_{k\to\infty}||B^k||_\infty^{1/k}
\\ \Rightarrow \rho(A)\leq \rho(|A|)&\leq\rho(B)
\end{align*}
\end{pf}

\begin{claim} \label{max_bounds}
Let $W$ be a stochastic matrix so that $\sum_jw_{ij}=1$ for all $i$ and let $W^1$ be $W$ without its first row and column.  Let the second largest eigenvalue of $W$ be $\mu$ and the largest eigenvalue of $W^1$ be $\lambda_1$.  Then $0\leq \lambda_1\leq 1$.  In particular,
\begin{enumerate}
\item $\lambda_1=1$ if and only if $w_{i1}=0$ for all $i$ and
\item $w_{i1}=1$ for all $i$ $\Rightarrow$ $\lambda_1=0$.
\end{enumerate}
\end{claim}

\begin{pf}
By Claim \ref{sum_bound}, 
\begin{align*}
0\leq \min_{i\neq 1}\sum_{j\neq 1} w_{ij}\leq \lambda_1&\leq \max_{i\neq 1}\sum_{j\neq 1}w_{ij}\leq 1
\end{align*}
To show 1.: $\lambda_1=1 \Rightarrow \max_{i\neq 1}\sum_{j\neq 1}w_{ij}=1 \Rightarrow \sum_{j\neq 1}w_{ij}=1$ for all $i$ $\Rightarrow$ $w_{i1}=0$ for all $i$

and $w_{i1}=0$ for all $i$ $\Rightarrow \sum_{j\neq i}w_{ij}=1$ for all $i$ $\Rightarrow$ $\min_{i\neq 1}\sum_{j\neq 1}w_{ij}=1$ $\Rightarrow$ $\lambda_1=1$.

To show 2.: $w_{i1}=1$ for all $i$ $\Rightarrow$ $\sum_{j\neq 1}w_{ij}=0$ for all $i$ $\Rightarrow $ $\max_i\sum_{j\neq 1}w_{ij}=0$ $\Rightarrow $ $\lambda_1=0$.
\end{pf}

%%%%%TO PROVE!!!!!
\begin{claim} \label{myconjecture}
Let $W,W^1$ be as in Claim \ref{max_bounds} and let $L=I_n-W$, $L^1=I_{n-1}-W^1$.  Let the second smallest eigenvalue of $L$ be $\mu$ and the smallest eigenvalue of $L^1$ be $\lambda_1$.  Then $1\geq \lambda_1\geq 0$.  In particular,
\begin{enumerate}
\item $\lambda_1=0$ if and only if $w_{i1}=0$ for all $i$ and
\item $w_{i1}=1$ for all $i$ $\Rightarrow$ $\lambda_1=1$.
\end{enumerate}
\end{claim}

\begin{pf}
The claim follows from Claim \ref{max_bounds} and the observation that the the eigenvalues of $W$ are are related to the eigenvalues of $L$ through $\lambda^W_i=1-\lambda^L_i$, and similarly for $W^1$ and $L^1$.
\end{pf}

\begin{theorem}{(Perron-Frobenius Theorem)} (As in \cite{Meyer:2001fk} and elsewhere) Let $A$ be an irreducible non-negative $n\times n$ matrix with spectral radius $r$.
\begin{enumerate}
\item $r$ is a positive real number and it is a simple eigenvalue of $A$
\item the only eigenvector whose components are all positive is the one associated with $r$
\item $r$ is the only eigenvalue on the spectral circle of $A$
\item (Collatz-Wielandt formula) $r=\max_{x\geq 0,x\neq 0}f(x)$ where $f(x)=\min_{1\leq i\leq n,x_i\neq 0}\frac{[Ax]_i}{x_i}.$
\item $\min_i\sum_ja_{ij}\leq r \leq \max_i\sum_ja_{ij}$
\end{enumerate}

\end{theorem}

\begin{theorem}{(Perron-Frobenius Theory for Nonnegative Matrices )} (As in \cite{Meyer:2001fk}) Let $A$ be a nonnegative $n\times n$ matrix with spectral radius $r$.  
\begin{enumerate}
\item $r$ is an eigenvalue of $A$, but $r=0$ is possible
\item the eigenvector associated with $r$ is nonnegative
\item $r=\max_{x\geq 0,x\neq 0}f(x)$ where $f(x)=\min_{1\leq i\leq n,x_i\neq 0}\frac{[Ax]_i}{x_i}.$
\end{enumerate}
\end{theorem}

\begin{theorem}{(Min-Max Theorem)} Let $A$ be an $n\times n$ Hermitian matrix with eigenvalues $\lambda_1\geq \lambda_2\geq \dots\geq\lambda_n$.  Let $R_A:\R^n\to\R$ be the Rayleigh-Ritz quotient,
$$R_A(x)=\frac{\langle Ax,x\rangle}{\langle x,x\rangle}.$$  
\begin{align*}
\lambda_k&=\max\{\min\{R_A(x):x\in U,x\neq 0\}:\dim(U)=k\}
\\ &=\min\{\max\{R_A(x):x\in U,x\neq 0\}:\dim(U)=n-k+1\}
\end{align*}
\end{theorem}

\begin{theorem}{(Cauchy Interlacing Theorem \cite{Meyer:2001fk})}
Let $A$ be a $n\times n$ hermitian matrix with eigenvalues $\lambda_1\geq \lambda_2\geq \dots\geq \lambda_n$ and for $c\in \C^{n}$ let $B$ be the bordered matrix
$$B=\left(\begin{array}{cccc} A & c \\ c^* & \alpha \end{array}\right) $$
with eigenvalues $\beta_1\geq \beta_2\geq \dots\geq \beta_n\geq \beta_{n+1}.$  Then the eigenvalues of $A$ interlace with those of $B$:
$$\beta_1\geq \lambda_1\geq \beta_2\geq \lambda_2\geq \dots\geq \beta_n\geq \lambda_n\geq\beta_{n+1}.$$
\end{theorem}

%%%%%%%
%%%%%%%%
%%%%%%%%%
%%%%%%%%%%
\section{The Model}

We consider a network where each agent has an interval variable, keeping track of, for example, direction or velocity of movement or a desired location, and uses the variables of the neighbors to whom it is connected to update its variable.  More concretely, there are $n$ nodes, which each have a variable $x_i$ and receive information from its neighbors weighted by a factor $w_{ij}$ such that $\sum_jw_{ij}=1$ for all $i=1,\dots,n$.  For instance, node $i$ may have $k$ connections, each of which it values equally so that $w_{ij}=1/k$ for those nodes from whom $i$ receives information and $w_{ij}=0$ otherwise.  In continuous time, the dynamics of the model are given by
$$\frac{d x_i}{dt}(t)=\sum_jw_{ij}(x_j(t)-x_i(t))+\xi_i$$ where $\xi_i$ is random noise.  Let $L$ be the Laplacian of the weighted matrix given by the $w_{ij}$ so that
$$L_{ij}=\left\{\begin{array}{cccc}
1& , & \text{ if } i=j \\
-w_{ij} & , & \text{ else}
\end{array}\right.$$
so that the model can be written as 
$$\dot{x}(t)=-Lx(t)+\xi$$

In discrete time, the dynamics of the model are given by
$$x_i(t+1)=x_i(t)+\sum_jw_{ij}(x_j(t)-x_i(t))+\xi_i$$
so that the whole system can be written as
$$x(t+1)=(I-L)x(t)+\xi=Wx(t)+\xi$$
where $W$ is the matrix given by the weights $w_{ij}$ and $I$ is the $n\times n$ identity matrix.

In either case, the model reaches equilibrium when every node has the same value, i.e. the system reaches consensus.  However, the consensus value can be any number.

Suppose there is external or environmental information that the agents can gather and that the agents perform better if they gather more accurate information.  In particular, suppose that a node receives an external signal $s$, which the rest of the agents want to learn as quickly and as accurately as possible.  We will suppose that the agent that receives the external signal is chosen at random.  For now, we assume that the first node is chosen.   Having received the external signal, that agent no longer listens to the other agents in the system, so that its dynamics are changed to $x_1(t)=s$ for all $t$.   The dynamics for the other nodes become
$$x_i(t+1)=x_i(t)+w_{i1}(s-x_i(t))+\sum_{j\neq 1}w_{ij}(x_j(t)-x_i(t))+\xi_i.$$
A consensus state is again the equilibrium, but now the consensus value is fixed at $s$, the value of the external signal.  We are particularly interested in deviations from that equilibrium, $\delta_i(t)=s-x_i(t)$:
\begin{align*}
\delta_i(t+1)&=s-x_i(t+1)
\\&=s-x_i(t)-w_{i1}(s-x_i(t))-\sum_{j\neq 1}w_{ij}(x_j(t)-x_i(t))-\xi_i
\\&=(1-w_{i1})\delta_j(t)-\sum_{j\neq 1}w_{ij}(\delta_i(t)-\delta_j(t))-\xi_i
\\&=(1-\sum_jw_{ji})\delta_i(t)+\sum_{i\neq 1}w_{ij}\delta_j(t)-\xi_i
\\&=\sum_{j\neq 1}w_{ij}\delta_j(t)-\xi_i
\end{align*}
so that the system of equations for the deviations from the external signal can be written
$$\delta(t+1)=W^1\delta(t)-\xi$$
where $W^1$ is the matrix $W$ with the first row and column removed.  The stability of the equilbrium $(s,s,\dots, s)$ is determined by the eigenvalue of $W^1$ that has the largest absolute value.  Let $\lambda_1$ be the dominant eigenvalue of $W^1$ and let $v_1$ be the corresponding eigenvector.  Then we can approximate
$$\delta(t)\sim\lambda_1^tv_1$$
where we assume the initial transient dynamics die out and the dynamics dictated by the principal eigenvector of $W^1$ take over rather quickly.

%%%%%%%%%%%%%%
\section{Measures of deviation\label{deviations}}
\cite{Young:2010fk}
\begin{align*}
||\delta(t)||&=\sqrt{\delta(t)^T\delta(t)}
\\&=\sqrt{\Tr(\delta(t)^T\delta(t))}
\\&=\sqrt{\Tr(\delta(t)\delta(t)^T)}
\\ \text{ so if we let }\Sigma(t)=\E[\delta(t)\delta(t)^T], \sqrt{\E[||\delta(t)||^2]}&=\sqrt{\Tr(\Sigma(t))}
\\ \Rightarrow \lim_{t\to\infty}\sqrt{\E[||\delta(t)||^2]}&=\sqrt{\Tr\left(\lim_{t\to\infty}\Sigma(t)\right)}
\end{align*}
As we found above,
\begin{align*}
\dot{\delta}(t)&=-L^1\delta(t)-\xi(t) \text{ , } \tag{*} \label{star}
\\ \text{ so that } \dot{\Sigma}(t)&=\E\left[\dot{\delta(t)}\delta(t)^T+\delta(t)\dot{\delta(t)^T}\right]
\\&=\E[-L^1\delta(t)\delta(t)^T-\xi(t)\delta(t)^T-\delta(t)\delta(t)^T(L^1)^T-\delta(t)\xi(t)^T]
\\&=-L^1\Sigma(t)-\E[\xi(t)\delta(t)^T]-\Sigma(t)(L^1)^T-\E[\delta(t)\xi(t)^T]
\end{align*}
Eq. (\ref{star}) allows us to write $\delta(t)=e^{-L^1t}\delta(0)-\int_0^te^{-L^1(t-\tau)}\xi(\tau)d\tau$, giving 
\begin{align*}
\E[\xi(t)\delta(t)^T]&=e^{-L^1t)}\E[\delta(0)\xi(t)^T]-\int_0^te^{-L^1(t-\tau)}\E[\xi(\tau)\xi(t)^T]d\tau
\\&=-\frac{1}{2}\int_0^te^{-L^1(t-\tau)}\delta(\tau-t)Id\tau
\\&=-\frac{1}{2}I
\\\dot{\Sigma}(t)&=-L^1\Sigma(t)-\Sigma(t)(L^1)^T+\frac{1}{2}I+\frac{1}{2}I
\\\Rightarrow 0&=-L^1\Sigma-\Sigma(L^1)^T+I \text{ at steady state}
\end{align*}
Again, learning dynamics are described as above by 
$$\dot{x}(t)=\sum_jw_{ij}(x_j(t)-x_i(t))+\xi_i(t).$$
Let $y$ represent the distance between population opinions and consensus, i.e.
$$y=QxQ^T,$$
where $Q\in\R^{n-1\times n}$ is such that $Q\vec{1}=0$ , $QQ^T=I_{n-1}$, and $Q^TQ=I_n-\frac{1}{n}\vec{1}\times\vec{1}^T$.   
If $\overline{L}=QLQ^T$ (so that $\dot{y}(t)=-\overline{L}y(t)+Q\xi(t)$) 
and if $\Sigma_y(t)=\E[y(t)y(t)^T]$, $\Sigma_y$ at steady state solves
$$0=-\overline{L}\Sigma_y-\Sigma_y\overline{L}+I.$$

%%%%%%%%%%%%%%
\section{Relationship between individual- and group-level performance}
If there is an external signal propagating across the network, it is in the best interest of the individuals of the network to position themselves so as to gather that information as quickly and reliably as possible.  While they may be interested in the network's ability to reach a consensus state, the benefits to the individual of that group-level measure of performance are less obvious.  However, in optimizing their own performance each individual may be contributing to the performance of the whole network.

First, in maximizing the speed with which an individual can learn the true external signal, the individual also affects the speed with which the network can come to consensus.  Let $L$ be the full matrix, $\overline{L}$ the transformed matrix describing the distance from consensus, and $L^i$ the reduced matrix in which individual $i$ remains convinced of the external signal; let $\overline{\lambda}$ and $\lambda^i$ be the smallest eigenvalues of the last two matrices.  The distance of the whole network from consensus shrinks as $e^{-\overline{\lambda}}$.  The size of the distance from the true signal shrinks as $e^{-\lambda^i}$.  As argued in Claim \ref{myconjecture}, $\lambda^i\leq\overline{\lambda}$.  Therefore, if an individual tries to maximize $\lambda_i$ he necessarily maximizes $\overline{\lambda}$ (Figure \ref{eigenvalue_comparison}).
\begin{figure}
\begin{center}
\includegraphics[width=.5\textwidth]{eigenvalue_comparison.pdf}
\end{center}
\caption{\label{eigenvalue_comparison} Here we plot the smallest eigenvalue $\lambda^i$ of the reduced Laplacian $L^i$ against the smallest eigenvalue $\overline{\lambda}$ of the transformed Laplacian $\overline{L}$, for several different matrices.  We find that $\lambda_i\leq \overline{\lambda}$.}
\end{figure}

In addition to the rate of convergence, the steady state distances from consensus can also be affected by individual-level optimizations.  If an individual is trying to minimize $\lim_{t\to\infty}\delta_j(t),$ he will also presumably minimize $\lim_{t\to\infty}\E[||\delta(t)||]$.  Section \nameref{deviations} shows how both $\lim_{t\to\infty}E[||y(t)||]$ and $\lim_{t\to\infty}E[||\delta(t)||]$ can be calculated from the Laplacian matrix $L$.  We find that $\lim_{t\to\infty}E[||\delta(t)||]\geq \lim_{t\to\infty}E[||y(t)||]$ (Figure \ref{lyap_comparison}).  Therefore, in trying to minimize $\lim_{t\to\infty}E[||\delta(t)||]$, an individual will necessarily minimize $\lim_{t\to\infty}E[||y(t)||]$.

Therefore, while a network's ability to come to consensus may not benefit the individual directly, in trying to gather information from the social network as quickly and reliably as possible, the individual is likely to affect the network in a way that would lead to more robust consensus dynamics.

\begin{figure}
\begin{center}
\includegraphics[width=.75\textwidth]{deviations.pdf}
\end{center}
\caption{\label{lyap_comparison} Here we plot the steady state distance from the external signal and the steady state distance from consensus, for many matrices, as calculated in Section \nameref{deviations}.}
\end{figure}

\section{Evolutionarily Stable Information Gathering Strategy?}
Is there an evolutionarily stable information gathering strategy?  For a given network configuration, we can score each node according to how quickly the node would converge to the external signal value.  This is determined by two properties of the reduced laplacian matrix $L^j$.

As derived above,
$$\dot{\delta}(t)=-L^j\delta(t)+\xi(t).$$
If we ignore noise for the moment, $\delta(t)=e^{-L^jt}\delta(0)$ and the largest eigenvalue of $-L^j$ (and hence the smallest eigenvalue of $L^1$) will determine the rate of convergence for the whole population.   Let $v^j$ be the eigenvector associated with the smallest eigenvalue $\lambda^j$, For sufficiently large $t$, we can approximate $\delta(t)\sim -v^j$ so that $|\delta_i(t)|\sim |v_i^j|$.  Therefore, the absolute value of the $i^{th}$ coordinate of $v^j$ indicates how close to the external signal node $i$ will be in the long run.  These two properties gives a convergence score for each node in the network:
$$e^{-\lambda^j}|v_i^j|.$$
Nodes with smaller convergence scores will gather the external signal more quickly.  Since each node in the network can receive the external signal and the nodes in the network would like to gather that information regardless of the receiver, we will give each node a score 
$$s_i=\langle e^{-\lambda^j}|v_i^j| \rangle_j,$$
where the average is across the different potential receivers $j$.

Before analyzing empirical networks, we'll look at a few idealized networks.  Imagine a cyclical graph of $N$ nodes places in a circle, each of which gathers information from its $k$ nearest neighbors.  We will introduce one invader with a different strategy $k'$ and see whether the natives or the invader get lower convergence scores.  To give a fitness score to the native and invader strategies, we look at the percent decrease in convergence scores between the average native score and the invader score.  Across network sizes there are three general patterns (see Figure \ref{ess_grid}):
\begin{enumerate}
\item if the native strategy is a low number of neighbors, invaders with more neighbors can invade
\item if the native strategy is low, having too many neighbors can be disadvantageous to the invader
\item if the native strategy is high, the dynamics are essentially neutral (except that an invader with a very low strategy can't invade)
\end{enumerate}

The fact that a very high invader strategy does poorly against a very low native strategy highlights the advantages and disadvantages of a high strategy.  If the signal comes to a node far away in the network, the invader can gather that signal more quickly than his poorly connected immediate neighbors.  However, if the signal comes to a node close to the invader, the invader is still collecting information from the nodes from across the circle and thus doesn't converge to the signal as quickly, a cost that is then relayed to his immediate neighbors who gather information from him (Figure \ref{convergence_in_space}).

For a given network size, we would like to know the ESS.  As the dynamics are essentially neutral for high native strategies, there are several high strategies that cannot be invaded by lower strategies but can drift between each other.  We take the minimum such strategy since presumably it's cognitively more parsimonious to attend to fewer rather than to more neighbors.  This ESS rises essentially linearly with the size of the network, so that the optimal number of neighbors is about $2/3$ total network size (Figure \ref{ess_vs_size}).

\begin{figure}
 \includegraphics[width=\textwidth]{ess_plot.pdf}
\caption{For various network sizes ($N$), we plot the percent difference in an invader's and a native's convergence score.  When the native strategyis to have few neighbors, an invader with more neighbors does generally (although not monotonically) better.  The landscape is essentially neutral when the native has many neighbors.}
\label{ess_grid}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=.5]{how_invader_succeeds.pdf}
\end{center}
\caption{For a network of size $20$ and a native population with a moderate number of neighbors, having too many neighbors cans actually be disadvantageous.  When the receiver is far away from the cheater, the cheater benefits by picking up that long-range signal more quickly than his close-minded neighbors.  However, when the receiver is close to the cheater, the cheater gets a noisy signal from the many sources of information he has.  The immediate neighbors of the cheater on average far worse because they do not benefit from the cheater's long range signal gathering and are harmed by gathering the spurious signal from the cheater when he gets it wrong.}
\label{convergence_in_space}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=.5]{ess_strategy.pdf}
\end{center}
\caption{For a given network size, we find the minimum neighbor strategy that cannot be invaded.}
\label{ess_vs_size}
\end{figure}

\nocite{*}
\bibliographystyle{plain}
\bibliography{info_evo}

\end{document}

### 2014-10-01 labtea(09_26_12).tex
trying to be more careful about disjointed networks when calculating correlation length

\documentclass{beamer}
%\usepackage{beamerthemeMarburg}
%\usepackage{beamerthemeboxes}
%\usepackage{beamerthemePittsburgh}
\usepackage{amsmath, amssymb,multicol}
\usepackage{concrete}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{url}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{relsize}
\newcommand{\R}{\mathbb{R}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

\setbeamertemplate{frametitle} {\begin{centering}\smallskip \insertframetitle\par \smallskip\end{centering}}
\setbeamertemplate{itemize item}{$\bullet$}
\setbeamertemplate{navigation symbols}{}
%\setbeamertemplate{footline}[text line]{%
%	\hfill\strut{%
%	\scriptsize\sf\color{black!60}%
%	\quad\insertframenumber
%	}%
%	\hfill}

\definecolor{DarkFern}{HTML}{407428}
\definecolor{DarkCharcoal}{HTML}{4D4944}
\colorlet{Fern}{DarkFern!85!white}
\colorlet{Charcoal}{DarkCharcoal!90!white}
\colorlet{LightCharcoal}{Charcoal!50!white}
\colorlet{AlertColor}{orange!80!black}
\colorlet{DarkRed}{red!70!black}
\colorlet{DarkBlue}{blue!70!black}
\colorlet{DarkGreen}{green!70!black}
\definecolor{MyColor}{HTML}{101BD2}
\colorlet{lightMyColor}{MyColor!70!white}

% Use the colors:

%\setbeamercolor{title}{fg=Fern}
\setbeamercolor{title}{fg=MyColor}
%\setbeamercolor{frametitle}{fg=Fern}
\setbeamercolor{frametitle}{fg=lightMyColor}
\setbeamercolor{normal text}{fg=Charcoal}
\setbeamercolor{block title}{fg=black,bg=MyColor!70!white}
\setbeamercolor{block body}{fg=black,bg=MyColor!35!white}
\setbeamercolor{alerted text}{fg=AlertColor}
\setbeamercolor{itemize item}{fg=lightMyColor}
\setbeamercolor{enumerate item}{fg=lightMyColor}

\title{Not `Won't you be my neighbor?' \\ But `Should you be my neighbor?'}
\subtitle{\textcolor{lightMyColor}{Optimal Information Gathering in Social Networks}}
\author{Eleanor Brush}

\begin{document}

\frame{\titlepage}

\begin{frame}

\begin{center}
\includegraphics[width=.8\textwidth]{social_network.pdf}
\end{center}

\end{frame}

\begin{frame}

\begin{center} {\Large What's the optimal way to gather information from a social network?}\end{center}

\begin{enumerate}
\pause \item a criterion of performance
\pause \item a strategy to maximize that criterion
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Overview}
\begin{enumerate}
\item a criterion for the whole network
\item a criterion for individuals
\item consequences of and strategy for optimizing individual performance
\item a measure of informational burden
\item future directions
\end{enumerate}
\end{frame}

\begin{frame}

\[ \scalebox{1.8}{$\dot{x}_i(t)=\sum_jw_{ij}\big(x_j(t)-x_i(t)\big)+\xi_i(t)$}\]
\\\[ \scalebox{1.25}{ where $\sum_jw_{ij} =1$ and $\xi_i(t)\sim N(0,1)$ i.i.d.} \]
\vspace{25pt}\[ \scalebox{1.5}{ $\Longrightarrow$ consensus with every $x_i=c$} \]

\end{frame}

\begin{frame}

\[ \scalebox{1.8}{$ \dot{x}(t)=-Lx(t)+\xi(t)$ }\]
\[\scalebox{1.25}{ where $L_{ij}=\left\{ \begin{array}{cccc} 1 & \text{ if } i=j \\ -w_{ij} & \text{ if } i\neq j \end{array}\right.$ }\]
%Without noise, an equilibrium is reached when $Lx=0$.  
\vspace{25pt}\[ \scalebox{1.5}{ $\Longrightarrow$ consensus with $x\propto \vec{1}$} \]
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[scale=.8]{vectors.pdf}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Deviation from Consensus}
Define the orthogonal transformation $Q\in\R^{n-1\times n}$ with 
\begin{enumerate}
\item $Q\vec{1}=0$
\item $QQ^T=I_{n-1}$ and 
\item $Q^TQ=I_n-\frac{1}{n}\vec{1}\times \vec{1}^T$
\end{enumerate}
Define $\textcolor{MyColor}{y=Qx}$ and $\textcolor{MyColor}{\overline{L}=QLQ^T}$.

$$\Longrightarrow \dot{y}(t)=-\overline{L}y(t)+Q\xi(t)$$

\vspace{50pt}\tiny{Starling flock networks manage uncertainty in consensus at low cost, \\George F. Young, Luca Scardovi, Andrea Cavagna, Irene Giardina, and Naomi E. Leonard. 2012.}

\end{frame}

\begin{frame}
\frametitle{Network Optimization}
One measure of performance:
$$H(L)=\lim_{t\to\infty}E[||y(t)||]$$

Objective: choose the number of neighbors $m$ that maximizes $$\frac{1}{\left(\frac{H}{\sqrt{N}}\right)m}$$

\end{frame}

\begin{frame}
\begin{center}
\includegraphics[width=\textwidth]{starling_robustness.pdf}
\end{center}
\vspace{30pt}\tiny{Starling flock networks manage uncertainty in consensus at low cost, \\ \vspace{-6pt} George F. Young, Luca Scardovi, Andrea Cavagna, Irene Giardina, and Naomi E. Leonard. 2012.}
\end{frame}

\begin{frame}
\frametitle{Learning about the world}
One node gets an external signal so that 
\begin{align*}
x_i(t)&=\text{signal } \text{ for all } t \text{ and }
\\ \dot{x}_j(t)&=\sum_kw_{jk}\big(x_k(t)-x_j(t)\big)+\xi_j(t) \text{ for } k\neq i
\\ \delta_j(t)&:=\text{ signal } -x_j(t)
\\ \Rightarrow \dot{\delta}(t)&=-L^i\delta(t)+L^i\xi(t) 
\\ \text{ where } L^i&=L \text{ without the $i^{th}$ row and column}
\end{align*}
\end{frame}

\begin{frame}

We can approximate $\delta_j(t)$ with $$\delta_j(t)\sim e^{-\lambda^it}v_j^i$$ where $\lambda^i$ is the smallest eigenvalue of $L^i$ and $v^i$ is its eigenvector.

\end{frame}

\begin{frame}
\begin{block}{{\small Deviations}}{
{\small from consensus: $\dot{y}(t)=-\overline{L}y(t)+Q\xi(t) $}
\\{\small from external signal: $\dot{\delta}(t)=-L^i\delta(t)-L^i\xi(t)$}
}\end{block}
Minimizing 
$$\delta_j(t)\sim e^{-\lambda^it}v_j^i$$
\textcolor{lightMyColor}{\emph{usually}} optimizes consensus in the whole group because 
\begin{enumerate}
\item the rate of convergence to consensus is given by $\overline{\lambda}$, the smallest eigenvalue of $\overline{L}$, and $\lambda^i\leq\overline{\lambda}$
\invisible<1>{\item $\lim_{t\to\infty}E[||\delta(t)||]\geq \lim_{t\to\infty}E[||y(t)||]$}
\end{enumerate}
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[width=.8\textwidth]{eigenvalue_comparison}
\end{center}
\end{frame}

\begin{frame}
\begin{block}{{\small Deviations}}{
{\small from consensus: $\dot{y}(t)=-\overline{L}y(t)+Q\xi(t) $}
\\{\small from external signal: $\dot{\delta}(t)=-L^i\delta(t)-L^i\xi(t)$}
}\end{block}
Minimizing 
$$\delta_j(t)\sim e^{-\lambda^it}v_j^i$$
\textcolor{lightMyColor}{\emph{usually}} optimizes consensus in the whole group because 
\begin{enumerate}
\item the rate of convergence to consensus is given by $\overline{\lambda}$, the smallest eigenvalue of $\overline{L}$, and $\lambda^i\leq\overline{\lambda}$
\item $\lim_{t\to\infty}E[||\delta(t)||]\geq \lim_{t\to\infty}E[||y(t)||]$
\end{enumerate}
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[width=.8\textwidth]{deviations.pdf}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Individual Optimization}
Objective: each node $j$ wants to minimize $s_j=\langle e^{-\lambda^i}v_j^i \rangle _i$
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[width=.6\textwidth]{circle.pdf}
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[width=.6\textwidth]{circle_w_invader.pdf}
\end{center}
\end{frame}

\begin{frame}
Optimizing $s_j$ affects everyone. 
%\begin{center}
\begin{columns}[c]
\column{.3\textwidth}
\column{.75\textwidth}
\includegraphics[width=\textwidth]{three_networks.pdf}
\end{columns}
%\end{center}
    \ra{1.3}
    \begin{tabular}{@{}clllllllllll@{}}
    $s_\text{invader}$  & & $0.2018$ & & & & $0.1993$ & & & & $0.1984$ 
    \\ $\langle s_j\rangle_{\text{native }j}$ & & $0.2018$ & & & & $0.2025$ & & & & $0.2029$
   \\ \small{$\lim_{t\to\infty} E[||y(t)||]$} & & $4.10$ & & & & $3.99$ & & & & $3.89$
    \\ \small{$\lim_{t\to\infty} E[||\delta(t)||]_{i=10}$} & & $5.98$ & & & & $5.48$ & & & & $5.09$
   \end{tabular}
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[width=.75\textwidth]{convergence_scores_one_network.pdf}
\end{center}
\end{frame}

\begin{frame}
\begin{center}$\small{\frac{\langle s_j\rangle _{\text{native }j}-s_\text{invader}}{\langle s_j\rangle _{\text{native }j}}}$ \end{center}
\begin{center}
\includegraphics[width=.75\textwidth]{invader_advantage.pdf}
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[width=.75\textwidth]{ESS_strategy.pdf}
\end{center}
\end{frame}

\begin{frame}
 \frametitle{Future Directions}
 \begin{itemize}
 \item allow for more variation among strategies across nodes and allow the strategies to evolve
 \item presence of multiple signals with different values 
 \item incorporate a principled measure of cognitive costs / informational burden of paying attention to more neighbors
 \item extend to other systems
 \item real network data
 \end{itemize}
 \end{frame}

\begin{frame}
\frametitle{Measure of Informational Burden}
\begin{align*}
&w_{ij}= \text{ weight of edge from } i\to j \ , \ \sum_jw_{ij}=1
\\& \Longrightarrow \text{ the vector $p$  such that } pW=p \text{ is the stationary distribution} 
 \\\\&\beta(W):= \text{ the entropy of the distribution }p \ , \ -\sum_ip_i\log(p_i)
\end{align*}

\end{frame}

\begin{frame}
\begin{center}
\includegraphics[width=.75\textwidth]{../../Documents/research/coarsegraining/compare_network_models.pdf}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Future Directions}
\begin{itemize}
\item real network data
\item how to minimize informational burden while maintaining network structure
\item how informational constraints affects network inference
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Acknowledgements}
Simon Levin
\\Naomi Leonard
\\George Young
\end{frame}

\begin{frame}
\frametitle{$H_2$ norm}
Consider a dynamical system $\dot{v}(t)=Av+Bw$ and $z=Cv$.  Let $\Sigma$ be the solution to the Lyapunov equation $$A^T\Sigma+\Sigma A+BB^T=0.$$  Then the 
$H_2$ norm\footnote{Robustness of Noisy Consensus Dynamics with Directed Communication, Proceedings of the American Control Conference, George Forrest Young, Luca Scardovi, and Naomi Ehrich Leonard, 2010. } of the system is 
$$\sqrt{\tr(C\Sigma C^T)}.$$ 
For $y$, this means that $\Sigma$ solves $$-\overline{L}^T\Sigma-\Sigma\overline{L}+I=0 \text{ and } H_2=\sqrt{\tr(\Sigma)}.$$
It can be shown\footnotemark[\value{footnote}] that $H(L)=\lim_{t\to\infty}E[||y(t)||]=H_2(L).$

\end{frame}

\end{document}

### 2014-10-01 notes_schooling_consensus_01_18_13.tex
trying to be more careful about disjointed networks when calculating correlation length

\documentclass{article}
%\documentclass[aps, twocolumn, tightenlines,amscd,amsmath,amssymb,verbatim]{revtex4}
\usepackage{latexsym}
\usepackage{amssymb,amsmath}
\usepackage{custom2}
\usepackage{graphicx} % for figures
\usepackage{epstopdf} % so can use EPS or PDF figures
%\usepackage{subfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{url}
\usepackage{amssymb,amsfonts}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage[pdftex]{hyperref}
\usepackage{lscape}
\captionsetup{justification=RaggedRight, singlelinecheck=false}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\argmax}{\text{argmax}}
\newcommand{\Tr}{\text{Tr}}
%\newtheorem{claim}{Claim}

\addtolength{\evensidemargin}{-.5in}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\textwidth}{1.4in}
\addtolength{\textheight}{1.4in}
\addtolength{\topmargin}{-.5in}

\pagestyle{empty}

\begin{document}

\begin{center}
{\bf \LARGE{Information Gathering Strategies in Social Networks}}
\vspace{10pt}
\\ Eleanor Brush
\\ January 18, 2013
\end{center}

\tableofcontents

\section{The Model}

We consider a network where each agent has an interval variable, keeping track of, for example, direction or velocity of movement or a desired location, and uses the variables of the neighbors to whom it is connected to update its variable.  More concretely, there are $n$ nodes, which each have a variable $x_i$ and receive information from its neighbors weighted by a factor $w_{ij}$ such that $\sum_jw_{ij}=1$ for all $i=1,\dots,n$.  For instance, node $i$ may have $k$ connections, each of which it values equally so that $w_{ij}=1/k$ for those nodes from whom $i$ receives information and $w_{ij}=0$ otherwise.  In continuous time, the dynamics of the model are given by
$$\frac{d x_i}{dt}(t)=\sum_jw_{ij}(x_j(t)-x_i(t))+\xi_i$$ 
where $\xi_i$ is random noise.  Let $L$ be the Laplacian of the weighted matrix given by the $w_{ij}$ so that
$$L_{ij}=\left\{\begin{array}{cccc}
1& , & \text{ if } i=j \\
-w_{ij} & , & \text{ else}
\end{array}\right.$$
so that the model can be written as 
$$\dot{x}(t)=-Lx(t)+\xi$$

Suppose there is external or environmental information that the agents can gather and that the agents perform better if they gather more accurate information.  In particular, suppose that a node receives an external signal $s$, which the rest of the agents want to learn as quickly and as accurately as possible.  We will suppose that the agent that receives the external signal is chosen at random.  For now, we assume that the first node is chosen.   Having received the external signal, that agent no longer listens to the other agents in the system, so that its dynamics are changed to $x_1(t)=s$ for all $t$.   The dynamics for the other nodes become
$$\frac{dx_i}{dt}=w_{i1}(s-x_i(t))+\sum_{j\neq 1}w_{ij}(x_j(t)-x_i(t))+\xi_i(t).$$
A consensus state is again the equilibrium, but now the consensus value is fixed at $s$, the value of the external signal.

%%%%%%%%%%%%%%
\section{Measures of deviation\label{deviations}}
Different nodes in the network might be subject to different sources and magnitudes of noise.  We will always assume that $\E[\xi_i]=0$.  Let $d_i=Var[\xi_i]$ and $D=diag(d_1,\dots,d_n)$.  In particular, we will set $d_i$ equal to the number of neighbors influencing node $i$, so that nodes with more neighbors make noisier estimates.

Let $y$ represent the distance between the vector of opinions and consensus, i.e.
$y=QxQ^T,$
where $Q\in\R^{n-1\times n}$ is such that $Q\vec{1}=0$ , $QQ^T=I_{n-1}$, and $Q^TQ=I_n-\frac{1}{n}\vec{1}\times\vec{1}^T$.   
Then 
\begin{equation}
\dot{y}(t)=-\overline{L}y(t)+Q\xi(t) \notag
\end{equation}
where $\overline{L}=QLQ^T$.  As shown in \cite{Young:2010fk}, if $\Sigma_y(t)=\E[y(t)y(t)^T]$,  $\Sigma_y$ at equilibrium solves
\begin{equation} 0=-\overline{L}\Sigma_y-\Sigma_y\overline{L}^T+D. \label{h2norm}
\end{equation}
If we instead consider deviation from the external signal, $\delta_i(t)=x_i(t)-s$,
\begin{align*}
\dot{\delta_i}(t)&=\dot{x_i}(t)
\\&=\sum_jw_{ij}(x_j(t)-x_i(t))+\xi_i(t)
\\&=\sum_jw_{ij}(\delta_j(t)-\delta_i(t))+\xi_i(t)
\\&=\sum_{j\neq 1}w_{ij}\delta_j(t)-\delta_i(t)\sum_jw_{ij}+\xi_i(t) \text{ since $\delta_1(t)\equiv 0$}
\\ \text{ so that }\dot{\delta}(t)&=-L^1\delta(t)+\xi(t)
\end{align*}
where $L^1$ is $L$ with the first row and column removed and $\delta(t)=(\delta_2(t)\dots\delta_n(t))$.
If $\Sigma(t)=\E[\delta(t)\delta(t)^T]$,  $\Sigma$ at equilibrium solves the analogous Lyapunov equation,
\begin{equation}
0=-L^1\Sigma-\Sigma(L^1)^T+D. \label{h2mod}
\end{equation}

\section{Relationship between consensus and external signal models}
\begin{claim} \label{myconjecture}
TO BE PROVED:  Let $L$ be the Laplacian of the weighted matrix and $L^1$ be $L$ without its first row and column.  Let the second smallest eigenvalue of $L$ be $\lambda$ and the smallest eigenvalue of $L^1$ be $\lambda_1$.  Then $1\geq \lambda \geq \lambda_1\geq 0$.  In particular,
\begin{enumerate}
\item $\lambda_1=0$ if and only if $w_{i1}=0$ for all $i$ and
\item if $w_{i1}=1$ for all $i$ then $\lambda_1=1$.
\end{enumerate}
\end{claim}

%Figure \ref{eigenvalue_comparison} shows that this is true for many randomly generated matrices.  
%\begin{figure}
%\begin{center}
%\includegraphics[width=.5\textwidth]{eigenvalue_comparison.pdf}
%\end{center}
%\caption{\label{eigenvalue_comparison} Here we plot the smallest eigenvalue $\lambda^i$ of the reduced Laplacian $L^i$ against the smallest eigenvalue $\overline{\lambda}$ of the transformed Laplacian $\overline{L}$, for several different matrices.  We find that $\lambda_i\leq \overline{\lambda}$.}
%\end{figure}

\begin{claim}
$||\delta(t)||\geq||y(t)||$ for all $t$.
\end{claim}

\begin{pf}
For the following, let $\langle x\rangle =\frac{\sum_ix_i}{n}$.
\begin{align*}
y&=x-\langle x\rangle \text{ and } \delta=x-s
\\ \Rightarrow ||y||^2&=\sum_i(x_i^2-2x_i\langle x\rangle+\langle x\rangle^2) \text{ and } ||\delta||^2=\sum_i(x_i^2-2x_is+s^2)
\\ \Rightarrow ||y||^2&=\sum_ix_i^2-2n\langle x\rangle^2+n\langle x\rangle^2 \text{ and } ||\delta||^2=\sum_ix_i^2-2ns\langle x\rangle+ns^2
\\ \Rightarrow ||\delta||^2-||y||^2&=ns^2-2ns\langle x\rangle+n\langle x\rangle^2
\\ &=n(\langle x\rangle-s)^2
\\ \Rightarrow ||\delta||^2-||y||^2&\geq 0 \text{ with equality iff } \langle x\rangle=s
\\ \Rightarrow ||\delta||^2&\geq||y||^2 \text{ with equality iff } \langle x\rangle=s
\\ \Rightarrow ||\delta||&\geq ||y|| \text{ with equality iff } \langle x\rangle=s
\end{align*}

\end{pf}

Therefore, $\lim_{t\to\infty}E[||\delta(t)||]\geq\lim_{t\to\infty}E[||y(t)||]$.

\section{Measures of Performance }

There are two general measures of performance: the speed with which the external signal can be learned and the steady state variance around the signal.  There is a tradeoff between these two measures.  As above let $L$ be the graph Laplacian of the network, $\overline{L}$ be the transformation of $L$, and $L^i$ be $L$ with the $i^{th}$ row and column removed, which describes how  a signal from individual $i$ propagates across the network.  Let $\lambda$ be the smallest eigenvalue of $\overline{L}$ and let $\lambda^i$ be the smallest eigenvalue of $L^i$ and $v^i$ its associated eigenvector.  Let $\Sigma_y$ and $\Sigma$ be the solutions to (\ref{h2norm}) and (\ref{h2mod}) respectively.  The eigenvalues $\lambda$ and $\lambda^i$ reflect how quickly consensus is reached or the external signal is learned and $\Sigma_y$ and $\Sigma$ are informative about the variance around consensus or the external signal.   We introduce a parameter $\alpha$ to represent the tradeoff between speed and variance.

Our measure of group performance will be 
\begin{equation}
p_G=\left(\frac{1}{e^{-\lambda}}\right)^{(1-\alpha)}\left(\frac{\sqrt{N}}{\sqrt{\Tr(\Sigma_y)}}\right)^\alpha  \label{groupperf}
\end{equation}
where we take reciprocals so that a higher value of $p_G$ is associated with better performance (or fitness).  Our measure of individual performance for node $j$ will be 
\begin{equation}
p_j=\left(\frac{1}{e^{-\lambda^i}v_j}\right)^{(1-\alpha)}\left(\frac{1}{\sqrt{\Sigma_{jj}}}\right)^\alpha  \label{indperf}
\end{equation}

\section{Measure of a Node's Position Predicts Individual Performance }

\section{Equilibrium Strategies}

\subsection{Group Optimization }

If we assume the whole group has the same strategy, we can ask what strategy gives the best performance for the group (and for every individual, since they are assumed to be identical and thus receive the same payoffs), for various levels of the tradeoff $\alpha$ between speed and variance (Figure \ref{group_performance}).  Unsurprisingly, when speed is more important (low $\alpha$), it is best for the group for every node to have many neighbors.  When variance is more important (high $\alpha$), it is best for every node to have few neighbors.  What is somewhat surprising is that at intermediate levels of tradeoff, there is not a optimum intermediate between low and high strategies.  Instead, the landscape is rather flat and every strategy performs about equally well.

\begin{figure}
\begin{subfigure}{\textwidth}
\begin{center}
\includegraphics[width=\textwidth]{group_performance.eps}
\end{center}
\caption{$N=20$}
\end{subfigure}
\begin{subfigure}{\textwidth}
\begin{center}
\includegraphics[width=\textwidth]{group_performance_v2.eps}
\end{center}
\caption{$N=40$}
\end{subfigure}
\caption{\label{group_performance} This figure shows how group performance depends on strategy and tradeoff, assuming the whole group adopts the same strategy.  When speed is more important than variance (low $\alpha$), it is best for the group to be highly connected.  When variance is more important (high $\alpha$), it is best for the group to be less well connected.}
\end{figure}

%\subsection{Individual Optimization }

\subsection{Adaptive Dynamics }
\subsubsection{Equilibria: ESS, NIS, CSS, dimorphic populations}

\begin{enumerate}
\item if only speed is valued, ESS is maximal
\item but if only accuracy is valued, there is no ESS
\item at intermediate levels of tradeoff, the population can become dimorphic at various pairs of strategies
\item scaling with group size
\end{enumerate}

As shown in \cite{Nowak:2006uq}, if $f_i$ is the fitness of an invader and $g_i$ is the fitness of a resident when there are $i$ invaders, the probability of the number of invaders increasing by one is given by 
\begin{equation}
p_{i\to i+1}=\frac{i(N-i) f_i}{N(if_i+(N-i)g_i)}\label{probup}
 \end{equation}
and the probability of the number of invaders decreasing by one is given by 
\begin{equation}
p_{i\to i-1}=\frac{i(N-i) g_i}{N(if_i+(N-i)g_i)}\label{probdown}
 \end{equation}
 Together these give the probability that the invader will come to fixation in the population:
 \begin{equation}
 p_f=\frac{1}{1+\sum_{k=1}^{N-1}\Pi_{j=1}^k\frac{p_{i\to i+1}}{p_{i\to i-1}}} \label{probfix}
 \end{equation}
Here, individual fitnesses are given by (\ref{indperf}).  Since individual fitness depends on where the node is in the network in addition to its strategy, we say that $f_i=\langle p_j\rangle_{j: \text{ $j$ is an invader} }$ and $g_i=\langle p_j\rangle_{j: \text{ $j$ is an resident} }$ when there are $i$ invaders.   For an intermediate level of tradeoff between speed and variance, invaders with intermediate strategies have the highest probability of reaching fixation (Figure \ref{fixation_prob}).

 \begin{figure}
\begin{center}
\includegraphics[width=.75\textwidth]{fixation_prob.eps}
\end{center}
\caption{ \label{fixation_prob} This figure shows, for each pair of resident and invader strategies, the difference between the probability of the invader strategy reaching fixation and the fixation probability under neutral dynamics, $p_f-\frac{1}{N}$.  }
\end{figure}

A reasonable definition of an steady state is a situation in which, for a given pair of resident and invader strategies, the probability of the number of invaders increasing by one is equal to the number of invaders decreasing by one so that the population is at a probabilistic steady state.  The probabilities may never be exactly equal.  However, there is a number of invaders such that the probability of increasing by one is greater than the probability of decreasing by one but if the number increases by one the probability of decreasing becomes greater than the probability of increasing so that the number of invaders will go back and forth between these two sizes.  This is our provisional definition of a steady state number of invaders.  For a given level of tradeoff between speed and variance ($\alpha=.5$) and network size ($N=20$) we find this steady state number (Figure \ref{prop_of_invaders}).

\begin{figure}
\begin{center}
\includegraphics[width=.75\textwidth]{proportion_of_invader.eps}
\end{center}
\caption{ \label{prop_of_invaders} This figure shows, for each pair of resident and invader strategies, the steady state number of invaders.  }
\end{figure}

If the population goes from a monomorphic one to one in which the invaders have reached their steady state size (as described above), does this improve or decrease the overall group performance?  In Figure (\ref{group_improvement}) we show this change in group performance.

\begin{figure}
\begin{center}
\includegraphics[width=.75\textwidth]{group_improvement.eps}
\end{center}
\caption{ \label{group_improvement} Here we show the effect on the group performance when a mutant invades and the invader population grows to the size that can be sustained by the group, as shown in Figure \ref{prop_of_invaders}.}
\end{figure}

\subsubsection{Greedy Optimization}

Rather than using adaptive dynamics to study how strategies evolve over time, we can allow the nodes themselves to optimize their strategies.  Given the set of strategies being used, each node can evaluate its performance and what its performance would have been had its strategy been one higher or lower and change its strategy accordingly.  We let the nodes perform this greedy optimization for several iterations until the strategies come to steady state.  (The steady state may oscillate between two adjacent strategies).  
Figures \ref{greedy_opt}-\ref{greedy_opt_v2} show the results of these optimizations, for different initial configurations of the network.

\begin{figure}

%\begin{subfigure}{.5 \textwidth}
%\begin{center}
%\includegraphics[width=\textwidth]{greedy_optimization_invader.eps}
%\end{center}
%\caption{The evolution of the invader's strategy over time: at first it increases and then oscillates between two high strategies.}
%\end{subfigure}
\begin{subfigure}{.5 \textwidth}
\begin{center}
\includegraphics[width=\textwidth]{end_of_greedy_optimization.eps}
\end{center}
\caption{The end result of a greedy optimization, where the $x$-axis represents the nodes and the $y$-axis the final strategy.  The node that started with a different strategy, $10$ as opposed to $2$ is first.  The greedy optimization results in spatial periodicity in strategy.}
\end{subfigure}
\begin{subfigure}{.5 \textwidth}
\begin{center}
\includegraphics[width=\textwidth]{greedy_optimization_group_perf.eps}
\end{center}
\caption{Group performance changes over time as the nodes perform the greedy optimization.  The greedy optimization tends to increase overall group performance.}
\end{subfigure}
\caption{\label{greedy_opt} The results of a greedy optimization, where at each time step each node increases its strategy by $2$, decreases its strategy by $2$, or stays the same, depending on which performs best given that the rest of the network has the strategies they do.  Initially, the population all has strategy $2$, except for the first node which has strategy $10$. ($\alpha=.5,$ $N=20$)}
\end{figure}

%\begin{figure}
%\begin{subfigure}{.5 \textwidth}
%\begin{center}
%\includegraphics[width=\textwidth]{greedy_optimization_invader_v2.eps}
%\end{center}
%\caption{The evolution of the invader's strategy over time: at first it increases and then oscillates between two high strategies.}
%\end{subfigure}
%\begin{subfigure}{.5 \textwidth}
%\begin{center}
%\includegraphics[width=\textwidth]{end_of_greedy_optimization_v2.eps}
%\end{center}
%\caption{The end result of a greedy optimization, where the $x$-axis represents the nodes and the $y$-axis the final strategy.  The node that started with a different strategy, $10$ as opposed to $2$ is first.  The greedy optimization results in spatial periodicity in strategy.}
%\end{subfigure}
%\begin{subfigure}{.5 \textwidth}
%\begin{center}
%\includegraphics[width=\textwidth]{greedy_optimization_group_perf_v2.eps}
%\end{center}
%\caption{Group performance changes over time as the nodes perform the greedy optimization.  There is a large drop immediately and little subsequent change.}
%\end{subfigure}
%\caption{\label{greedy_opt_v2} The results of a greedy optimization, where at each time step each node increases its strategy by $2$, decreases its strategy by $2$, or stays the same, depending on which performs best given that the rest of the network has the strategies they do.  Initially, the population all has strategy $10$, except for the first node which has strategy $2$. ($\alpha=.5,$ $N=20$)}
%\end{figure}

\begin{figure}
\begin{subfigure}{.5 \textwidth}
\begin{center}
\includegraphics[width=\textwidth]{end_of_greedy_optimization_v3.eps}
\end{center}
\caption{The end result of a greedy optimization, where the $x$-axis represents the nodes and the $y$-axis the final strategy.  The node that started with a different strategy, $10$ in a population of $2$, is first.  The greedy optimization results in spatial periodicity in strategy.}
\end{subfigure}
\begin{subfigure}{.5 \textwidth}
\begin{center}
\includegraphics[width=\textwidth]{greedy_optimization_group_perf_v3.eps}
\end{center}
\caption{Group performance changes over time as the nodes perform the greedy optimization.  In this case, greedy optimization tends to decrease group performance.}
\end{subfigure}
\caption{\label{greedy_opt_v2} The results of a greedy optimization, where at each time step each node increases its strategy by $2$, decreases its strategy by $2$, or stays the same, depending on which performs best given that the rest of the network has the strategies they do.  Initially, the population randomly drawn strategies. ($\alpha=.5,$ $N=20$)}
\end{figure}

\section{Future Work}

\begin{enumerate}
\item prove Claim \ref{myconjecture} about eigenvalues
\item find ESS and CIS 
\item dependence of equilibria on the size of the network
\item Information centrality predicts the accuracy of each individual's opinion and eigenvector centrality predicts the speed of each individual's opinion.  Is there a measure that predicts performance when both speed and accuracy are valued?
\item prove that spatial periodicity in strategy is optimal for group performance
\item perform these analysis on more (empirical) networks
\end{enumerate}

\section{Conclusions }

\begin{enumerate}
\item from the perspective of network design, measure X tells how to optimize individual performance
\item optimal topology for the group when we also consider speed to consensus
\item group performance does or doesn't or in certain conditions (i.e. size of group, tradeoff between speed and accuracy) emerge out of individual level considerations so group performance does or doesn't or in certain conditions have to be invoked
\end{enumerate}

\nocite{*}
\bibliographystyle{plain}
\bibliography{info_evo}

\end{document}

### 2014-10-01 notes_schooling_consensus_03_26_14.tex
trying to be more careful about disjointed networks when calculating correlation length

\documentclass{article}
%\documentclass[aps, twocolumn, tightenlines,amscd,amsmath,amssymb,verbatim]{revtex4}
\usepackage{latexsym}
\usepackage{amssymb,amsmath}
\usepackage{custom2}
\usepackage{graphicx} % for figures
\usepackage{epstopdf} % so can use EPS or PDF figures
%\usepackage{subfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{url}
\usepackage{amssymb,amsfonts}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage[pdftex]{hyperref}
\usepackage{lscape}
\captionsetup{justification=RaggedRight, singlelinecheck=false}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\argmax}{\text{argmax}}
\newcommand{\Tr}{\text{Tr}}
%\newtheorem{claim}{Claim}

\addtolength{\evensidemargin}{-.5in}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\textwidth}{1.4in}
\addtolength{\textheight}{1.4in}
\addtolength{\topmargin}{-.5in}

\pagestyle{empty}

\begin{document}

\begin{center}
{\bf \LARGE{Information Gathering Strategies}}
\vspace{10pt}
\\ Eleanor Brush
\\ March 18, 2013
\end{center}

\tableofcontents

\section{Introduction}
There are (at least) two reasons animals might update their opinions according to the opinions of the other animals around them.  They might want to be able to learn whatever information the other animals have learned about the environment or it might be advantageous for the animals to come to consensus, for example about where to go.

Previous work on how many neighbors birds have:
\begin{itemize} 
\item anisotropy in bird flocks suggest they pay attention to seven of their nearest neighbors \cite{Ballerini:2007uq}
\item statistical mechanical analysis of bird flocks suggest they pay attention to eleven of their nearest neighbors \cite{Bialek:2012fk}
\item if  birds pay attention to seven of their nearest neighbors they construct an interaction network that optimizes the group's ability to come to a robust consensus decision \cite{Young:2013kx}
\end{itemize}

One explanation is that the birds benefit if they are in a group that can come to a consensus decision.  Another explanation is that the birds are optimizing some other measure of performance and in so doing construct a network that enables the group's ability to come to consensus.  We are interested in seeing if there are other properties the individual's might be optimizing that leads to robust consensus.

{\bf The Question:} If individual birds are trying to optimize the speed and reliability with which they can learn an external signal that someone else has picked up, and change the number of birds they pay attention to optimize that performance, how does that affect the interaction network they develop?  In particular, does it improve or diminish the group's ability to come to consensus?

\section{The Model}

There are $n$ nodes in the system each have an opinion $x_i$ and receive information from their neighbors, which they weigh by factors $w_{ji}$ such that $\sum_jw_{ij}=1$.   A node updates its opinions by averaging the difference between its opinion and its neighbors'.  In continuous time, the dynamics of this learning process are given by
\begin{equation}
\frac{d x_i}{dt}(t)=\sum_jw_{ij}(x_j(t)-x_i(t))+\xi_i \label{model}
\end{equation}
where $\xi_i$ is random noise.  Let $L$ be the Laplacian of the weighted matrix given by the $w_{ij}$ so that
\begin{equation}
L_{ij}=\left\{\begin{array}{cccc}  \label{lap}
1& , & \text{ if } i=j \\ 
-w_{ij} & , & \text{ else}
\end{array}\right. \end{equation}
Then the vector of opinions ${\bf x}(t)$ is described by 
\begin{equation}
\dot{{\bf x}}(t)=-L{\bf x}(t)+\xi \label{vecmodel}
\end{equation}
If the graph described by $L$ is connected, then, in the absence of noise, the system will reach a consensus state in which all nodes have the same opinion, ${\bf x}(t)=\alpha{\bf 1}$.

The learning dynamics change if we introduce a stubborn node $r$ who stops listening to his neighbors, for instance if he learns about something in the environment that is more important than the information coming from his peers.  In this case, $x_r(t)=s$ for some constant $s$ and all the other nodes continue to update using their neighbors' opinions as before, so that $\dot{x_i}(t)$ is as in \ref{model} for $i\neq r$.  In this case, the consensus opinion will be $s$, the external signal.  This is equivalent to defining a new $L^r$ such that $L^r_{ri}=0$ for all $i$.
%, and we can choose $Q$ to be the $n\times n$ identity matrix with the $i^{th}$ diagonal element corresponding to the stubborn node set to $0$.  Then $\hat{L}$ is equal to $L$ with the $i^{th}$ row and column removed.  The distance from consensus now represents something with more easily justified fitness consequences: how closely a node's opinion matches the opinion of a node with true environmental information.
The Laplacian $L$ or $L^r$ affects how quickly consensus can reached and how close to the consensus opinion each node can be when there is noise in the system.

%%%%%%%%%%%%%%
\section{Measures of Performance \label{measures}}
There are two timescales on which an individual might care about its performance: a transient period during and a steady state period during which noise perturbs the opinions away from the external signal.

To measure the distance  from the external signal during a transient period (ignoring noise for the moment), we define, for each node $i$,
\begin{equation}
\tau_i^{ext}=\sqrt{\int_0^\infty |x_i(t)-s|^2dt}
\end{equation}
Let $\lambda^r$ be second smallest eigenvalue of $L^r$ and $v$ be the associated eigenvector.  Then we can approximate $x_i(t)-s\sim e^{-\lambda^rt}v_i$ and
\begin{equation}
\tau_i^{ext}=\sqrt{\int_0^\infty e^{-2\lambda^rt}v_i^2dt}=\frac{|v_i|}{\sqrt{2\lambda^r}}
\end{equation}

To measure perturbations away from the external signal due to noise in the steady state, we define, for each node $i$,
\begin{equation}
\sigma_i^{ext}=\sqrt{\lim_{t\to\infty}E[|x_i(t)-s|^2]}
\end{equation}
Different nodes in the network might be subject to different sources and magnitudes of noise.  We will always assume that $\E[\xi_i]=0$.  Let $d_i=Var[\xi_i]$ and $D=diag(d_1,\dots,d_n)$.  In particular, we will set $d_i$ equal to the number of neighbors influencing node $i$, so that nodes with more neighbors make noisier estimates.  This can be found from the $H_2$ norm of the matrix $L^r$ (see Appendix and \cite{Young:2010fk} for details).

These metrics, $\tau_i^{ext}$ and $\sigma_i^{ext}$, measure node $i$'s ability to learn the external signal either quickly during the transient period or reliably during the steady state period.  To combine them into one performance metric, we introduce a tradeoff parameter, $\alpha$ and measure performance according to $(1-\alpha)\tau_i^{ext}+\alpha\sigma_i^{ext}$.

We also need to quantify the group's ability to come to consensus, which again can be measured on a short or a long timescale.  Analogously to our metrics of individual performance, we define
\begin{equation}
\tau^{con}=\frac{1}{\sqrt{2\lambda}}
\end{equation}
where $\lambda$ is the second smallest eigenvalue of $L$.  In this case, there is no predefined consensus opinion so we need to find the distance between the vector of opinions and consensus in a way that doesn't depend on what consensus opinion is reached.  To do this, we define $y(t)$ to be $x(t)-\langle x(t)\rangle$ and 
\begin{equation}
\sigma^{con}=\sqrt{\frac{\lim_{t\to\infty}E[||y(t)||^2]}{n}}
\end{equation}
We have defined these measures so that $\tau^{con}$ is low when the group converges to consensus quickly and $\sigma^{con}$ is low when the steady state noise around the consensus state is low.  Again the metrics for each timescale can be combined into one metric by introducing a tradeoff parameter $\alpha$ and trying to minimize $(1-\alpha)\tau^{con}+\alpha\sigma^{con}$.

\section{Learning the external signal versus reaching consensus}
\begin{lemma} \label{myconjecture}
TO BE PROVED:  Let $L$ be the Laplacian of the weighted matrix and $L^r$ be $L$ with its $r^{th}$ row set equal to $0$.  Let the second smallest eigenvalue of $L$ be $\lambda$ and the smallest eigenvalue of $L^r$ be $\lambda^r$.  Then $1\geq \lambda \geq \lambda^r\geq 0$.  In particular,
\begin{enumerate}
\item $\lambda^r=0$ if and only if $w_{ir}=0$ for all $i$ and
\item if $w_{ir}=1$ for all $i$ then $\lambda^r=1$.
\end{enumerate}
\end{lemma}

In particular, this means that $\langle\tau_i^{ext}\rangle_i\geq \tau^{con}$.

\begin{lemma}
If we define $\delta(t)=x(t)-s\vec{1}$, then $||\delta(t)||^2\geq||y(t)||^2$ for all $t$.
\end{lemma}

\begin{pf}
For the following, let $\langle x\rangle =\frac{\sum_ix_i}{n}$.
\begin{align*}
y&=x-\langle x\rangle \text{ and } \delta=x-s
\\ \Rightarrow ||y||^2&=\sum_i(x_i^2-2x_i\langle x\rangle+\langle x\rangle^2) \text{ and } ||\delta||^2=\sum_i(x_i^2-2x_is+s^2)
\\ \Rightarrow ||y||^2&=\sum_ix_i^2-2n\langle x\rangle^2+n\langle x\rangle^2 \text{ and } ||\delta||^2=\sum_ix_i^2-2ns\langle x\rangle+ns^2
\\ \Rightarrow ||\delta||^2-||y||^2&=ns^2-2ns\langle x\rangle+n\langle x\rangle^2
\\ &=n(\langle x\rangle-s)^2
\\ \Rightarrow ||\delta||^2-||y||^2&\geq 0 \text{ with equality iff } \langle x\rangle=s
\\ \Rightarrow ||\delta||^2&\geq||y||^2 \text{ with equality iff } \langle x\rangle=s
%\\ \Rightarrow ||\delta||&\geq ||y|| \text{ with equality iff } \langle x\rangle=s
\end{align*}

\end{pf}

Therefore, $\lim_{t\to\infty}E[||\delta(t)||^2]\geq\lim_{t\to\infty}E[||y(t)||^2]$.  So that $\sum_{i=1}^n(\sigma_i^{ext})^2\geq n\sigma^{con}$ and finally this means that $\langle\sigma_i^{ext}\rangle_i\geq \sigma^{con}$.   It is therefore not unreasonable to hope that by minimizing $\tau_i^{ext}$ and $\sigma_i^{ext}$ the animals will tend to minimize $\tau^{con}$ and $\sigma^{con}$.

\section{Optimal Strategies}
If all $n$ nodes have the same number of neighbors, $2k$, then  
\begin{equation}
\tau^{con}=\frac{1}{\sqrt{2-\frac{2}{k}\sum_{\ell=1}^k\cos\left(\frac{2\pi \ell}{m}\right)}} 
\end{equation}
and 
\begin{equation}
\sigma^{con}=\sqrt{\frac{2k\sum_{j=1}^{n-1}\frac{1}{2-\frac{2}{k}\sum_{\ell=1}^k\cos\left(\frac{2\pi j\ell}{m}\right)}}{n}} 
\end{equation}

For a given tradeoff $\alpha$, we can find the number of neighbors that optimizes the animals' ability to learn the external signal or the group's ability to come to consensus (Figure \ref{opt}).
Not surprsingingly, as $\alpha$ increases (and long term noise matters more), it becomes better to pay attention to fewer neighbors (but more than $2$).  We find that the  optimal strategy scales sublinearly with network size.  Most importantly, the optimal strategy for gathering an external signal is to pay attention to fewer neighbors than the optimal strategy for reaching consensus.  This means that optimizing the ability to learn the external signal won't work!

\begin{figure}
\includegraphics[width=.5\textwidth]{../../Desktop/optstrats_alpha25.eps}
\includegraphics[width=.5\textwidth]{../../Desktop/optstrats_alpha1.eps}
\caption{Optimal strategies in a uniform network. \label{opt}}
\end{figure}

\section{Singular Strategies and Evolutionary Dynamics}

\section{Future Work}

\begin{enumerate}
\item prove Claim \ref{myconjecture} about eigenvalues
\item Information centrality predicts the accuracy of each individual's opinion and eigenvector centrality predicts the speed of each individual's opinion.  Is there a measure that predicts performance when both speed and accuracy are valued?
\item prove that spatial periodicity in strategy is optimal for group performance
\item allow nodes to actively optimize their strategies 
 \item allow for more variation among strategies across nodes and for more complicated network structures
  \item multiple signals with different values 
\item look at real networks
\end{enumerate}

%\section{Conclusions }
%
%\begin{enumerate}
%\item from the perspective of network design, measure X tells how to optimize individual performance
%\item optimal topology for the group when we also consider speed to consensus
%\item group performance does or doesn't or in certain conditions (i.e. size of group, tradeoff between speed and accuracy) emerge out of individual level considerations so group performance does or doesn't or in certain conditions have to be invoked
%\end{enumerate}

\section*{Appendix}
\subsection*{H2 Norm }

\ Let $y$ represent the distance between the vector of opinions and consensus, i.e.
$y=QxQ^T,$
where $Q\in\R^{n-1\times n}$ is such that $Q\vec{1}=0$ , $QQ^T=I_{n-1}$, and $Q^TQ=I_n-\frac{1}{n}\vec{1}\times\vec{1}^T$.   
Then 
\begin{equation}
\dot{y}(t)=-\overline{L}y(t)+Q\xi(t) \notag
\end{equation}
where $\overline{L}=QLQ^T$.  As shown in \cite{Young:2010fk}, if $\Sigma_y(t)=\E[y(t)y(t)^T]$,  $\Sigma_y$ at equilibrium solves
\begin{equation} 0=-\overline{L}\Sigma_y-\Sigma_y\overline{L}^T+D. \label{h2norm}
\end{equation}
%If we instead consider deviation from the external signal, $\delta_i(t)=x_i(t)-s$,
%\begin{align*}
%\dot{\delta_i}(t)&=\dot{x_i}(t)
%\\&=\sum_jw_{ij}(x_j(t)-x_i(t))+\xi_i(t)
%\\&=\sum_jw_{ij}(\delta_j(t)-\delta_i(t))+\xi_i(t)
%\\&=\sum_{j\neq 1}w_{ij}\delta_j(t)-\delta_i(t)\sum_jw_{ij}+\xi_i(t) \text{ since $\delta_1(t)\equiv 0$}
%\\ \text{ so that }\dot{\delta}(t)&=-L^1\delta(t)+\xi(t)
%\end{align*}
%where $L^1$ is $L$ with the first row and column removed and $\delta(t)=(\delta_2(t)\dots\delta_n(t))$.
%If $\Sigma(t)=\E[\delta(t)\delta(t)^T]$,  $\Sigma$ at equilibrium solves the analogous Lyapunov equation,
%\begin{equation}
%0=-L^1\Sigma-\Sigma(L^1)^T+D. \label{h2mod}
%\end{equation}

\nocite{*}
\bibliographystyle{plain}
\bibliography{info_evo}

\end{document}

### 2014-10-01 notes_schooling_consensus_10_01_12.tex
trying to be more careful about disjointed networks when calculating correlation length

--- 2014-10-01 information_evolution_08_28_12.tex
@@
 \begin{document}
-\begin{center}
-\Large
-
-\end{center}
-
-\vspace{0pt}
 
@@
 \\ Eleanor Brush
-\\ August 28, 2012
+\\ October 1, 2012
 \end{center}
 
-\vspace{0pt}
-\normalsize
+\tableofcontents
 
@@
 Let $W,W^1$ be as in Claim \ref{max_bounds} and let $L=I_n-W$, $L^1=I_{n-1}-W^1$.
-Let the second smallest eigenvalue of $L$ be $\mu$ and the smallest eigenvalue of $L^1$ be $\lambda_1$.
-Then $1\geq \lambda_1\geq 0$.
+Let the second smallest eigenvalue of $L$ be $\lambda$ and the smallest eigenvalue of $L^1$ be $\lambda_1$.
+Then $1\geq \lambda\geq \lambda_1\geq 0$.
 In particular,
@@
 
+\begin{claim}
+$||\delta(t)||\geq||y(t)||$ for all $t$.
+\end{claim}
+
+\begin{pf}
+For the following, let $\langle x\rangle =\frac{\sum_ix_i}{n}$.
+\begin{align*}
+y&=x-\langle x\rangle \text{ and } \delta=x-s
+\\ \Rightarrow ||y||^2&=\sum_i(x_i^2-2x_i\langle x\rangle+\langle x\rangle^2) \text{ and } ||\delta||^2=\sum_i(x_i^2-2x_is+s^2)
+\\ \Rightarrow ||y||^2&=\sum_ix_i^2-2n\langle x\rangle^2+n\langle x\rangle^2 \text{ and } ||\delta||^2=\sum_ix_i^2-2ns\langle x\rangle+ns^2
+\\ \Rightarrow ||\delta||^2-||y||^2&=ns^2-2ns\langle x\rangle+n\langle x\rangle^2
+\\ &=n(\langle x\rangle-s)^2
+\\ \Rightarrow ||\delta||^2-||y||^2&\geq 0 \text{ with equality iff } \langle x\rangle=s
+\\ \Rightarrow ||\delta||^2&\geq||y||^2 \text{ with equality iff } \langle x\rangle=s
+\\ \Rightarrow ||\delta||&\geq ||y|| \text{ with equality iff } \langle x\rangle=s
+\end{align*}
+
+\end{pf}
+
+Therefore, $\lim_{t\to\infty}E[||\delta(t)||]\geq\lim_{t\to\infty}E[||y(t)||]$.
+
 %%%%%%%%%%%%%%
@@
 Section \nameref{deviations} shows how both $\lim_{t\to\infty}E[||y(t)||]$ and $\lim_{t\to\infty}E[||\delta(t)||]$ can be calculated from the Laplacian matrix $L$.
-We find that $\lim_{t\to\infty}E[||\delta(t)||]\geq \lim_{t\to\infty}E[||y(t)||]$ (Figure \ref{lyap_comparison}).
+We find that $\lim_{t\to\infty}E[||\delta(t)||]\geq \lim_{t\to\infty}E[||y(t)||]$.
 Therefore, in trying to minimize $\lim_{t\to\infty}E[||\delta(t)||]$, an individual will necessarily minimize $\lim_{t\to\infty}E[||y(t)||]$.

### 2014-10-01 notes_schooling_consensus_12_04_12.tex
trying to be more careful about disjointed networks when calculating correlation length

--- 2014-10-01 notes_schooling_consensus_01_18_13.tex
@@
 \\ Eleanor Brush
-\\ January 18, 2013
+\\ December 27, 2012
 \end{center}
@@
 
-%\subsection{Individual Optimization }
+\subsection{Individual Optimization }
 
-\subsection{Adaptive Dynamics }
-\subsubsection{Equilibria: ESS, NIS, CSS, dimorphic populations}
-
-\begin{enumerate}
-\item if only speed is valued, ESS is maximal
-\item but if only accuracy is valued, there is no ESS
-\item at intermediate levels of tradeoff, the population can become dimorphic at various pairs of strategies
-\item scaling with group size
-\end{enumerate}
+\subsubsection{Adaptive Dynamics }
 
@@
 
-A reasonable definition of an steady state is a situation in which, for a given pair of resident and invader strategies, the probability of the number of invaders increasing by one is equal to the number of invaders decreasing by one so that the population is at a probabilistic steady state.
+A reasonable definition of an equilibrium is a situation in which, for a given pair of resident and invader strategies, the probability of the number of invaders increasing by one is equal to the number of invaders decreasing by one so that the population is at a probabilistic steady state.
 The probabilities may never be exactly equal.

### 2014-10-01 ESSN.m

@@ -1,0 +1,28 @@
+function Strat=ESSN(fitness,rho,N)
+s=size(fitness);
+S=s(2);
+d=diag(fitness);
+range=1:S;
+strat=[];
+for k=1:S
+    m=max(fitness(k,range~=k));
+    if fitness(k,k)>m
+       strat=[strat k]; 
+    end
+    if m==fitness(k,k)
+        v=range(fitness(k,:)==m);
+        if sum((fitness(v,k)>d(v))==0)==0
+            strat=[strat k];
+        end
+    end            
+end
+if nargin>1
+    Strat=[];
+   for k=strat
+       f=max(rho(k,range~=k));
+        if f<=1/N
+            Strat=[Strat k];
+        end
+   end
+else Strat=strat;
+end

### 2014-10-01 ESS_parallel.m

@@ -1,0 +1,110 @@
+numworkers=str2num(getenv('PROCS'));
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
+
+numsigs_permove=1;
+nummoves=5000;
+
+N=20;
+b=1;
+
+strats=1:1:(N-1);
+L=length(strats);
+
+Tvals=1;
+Nt=length(Tvals);
+
+radvals=[0:.25:2];
+% radvals=.1;
+Nr=length(radvals);
+
+ESSseaten=cell(Nt,Nr);
+ESSsgettoeat=cell(Nt,Nr);
+
+t0=tic;
+
+for q=1:Nt
+    for p=1:Nr
+        T=Tvals(q);
+        radius=radvals(p);
+
+        fitnesseaten=zeros(L,L);
+        rhoeaten=zeros(L,L);
+
+        fitnessgettoeat=zeros(L,L);
+        rhogettoeat=zeros(L,L);
+
+        featen=zeros(L,L,N-1);
+        geaten=zeros(L,L,N-1);
+
+        fgettoeat=zeros(L,L,N-1);
+        ggettoeat=zeros(L,L,N-1);
+        parfor ind=1:(L*L*(N-1))
+%         parfor ind=1:5
+            [u,v,k]=ind2sub([L,L,N-1],ind);
+            
+            resident=strats(u);
+            invader=strats(v);
+
+            strategy=resident*ones(1,N);
+            strategy(1:k)=invader;
+            [probeaten, probgettoeat]=signalingevents(strategy,numsigs_permove,nummoves,radius,b,T);
+
+            perfeaten=1-probeaten;
+            featen(ind)=mean(perfeaten(1:k));
+            geaten(ind)=mean(perfeaten((k+1):N));
+
+            perfgettoeat=probgettoeat;
+            fgettoeat(ind)=mean(perfgettoeat(1:k));
+            ggettoeat(ind)=mean(perfgettoeat((k+1):N));
+
+        end
+
+
+        for ind=1:L*L
+            [u,v]=ind2sub([L,L],ind);
+            where=zeros(N-1,1);
+            for k=1:N-1
+                where(k)=sub2ind([L,L,N-1],u,v,k);
+            end
+            featen_now=featen(where);
+            geaten_now=geaten(where);
+
[…]

### 2014-10-01 ESS_parallel_local.m

@@ -1,0 +1,97 @@
+t0=tic;
+numsigs_permove=1;
+nummoves=100;
+N=20;
+b=1;
+
+strats=2:2:(N-1);
+L=length(strats);
+
+Tvals=1;
+Nt=length(Tvals);
+
+% radvals=[.0 .1 .5 1.5];
+radvals=.1;
+Nr=length(radvals);
+
+ESSseaten=cell(Nt,Nr);
+ESSsgettoeat=cell(Nt,Nr);
+
+for q=1:Nt
+    for p=1:Nr
+        T=Tvals(q);
+        radius=radvals(p);
+
+        fitnesseaten=zeros(L,L);
+        rhoeaten=zeros(L,L);
+
+        fitnessgettoeat=zeros(L,L);
+        rhogettoeat=zeros(L,L);
+
+        featen=zeros(L,L,N-1);
+        geaten=zeros(L,L,N-1);
+
+        fgettoeat=zeros(L,L,N-1);
+        ggettoeat=zeros(L,L,N-1);
+        parfor ind=1:(L*L*(N-1))
+            [u,v,k]=ind2sub([L,L,N-1],ind);
+            squareind=sub2ind([L,L],u,v);
+            resident=strats(u);
+            invader=strats(v);
+
+            strategy=resident*ones(1,N);
+            strategy(1:k)=invader;
+            [meanscore, scorevar, probeaten, probgettoeat]=signalingevents(strategy,numsigs_permove,nummoves,radius,b,T);
+
+            perfeaten=1-probeaten;
+            featen(ind)=mean(perfeaten(1:k));
+            geaten(ind)=mean(perfeaten((k+1):N));
+
+            perfgettoeat=probgettoeat;
+            fgettoeat(ind)=mean(perfgettoeat(1:k));
+            ggettoeat(ind)=mean(perfgettoeat((k+1):N));
+
+        end
+
+
+        for ind=1:L*L
+            [u,v]=ind2sub([L,L],ind);
+            where=zeros(N-1,1);
+            for k=1:N-1
+                where(k)=sub2ind([L,L,N-1],u,v,k);
+            end
+            featen_now=featen(where);
+            geaten_now=geaten(where);
+
+            fgettoeat_now=fgettoeat(where);
+            ggettoeat_now=ggettoeat(where);
+
+            fitnesseaten(ind)=featen_now(1)/geaten_now(1);
[…]

### 2014-10-01 H2mod.m

@@ -1,0 +1,8 @@
+function h = H2mod(M,i)
+s=size(M);
+N=s(2);
+L=lap(M);
+range=1:N;
+Li=L(range~=i,range~=i);
+sigma = lyap(Li,-eye(N-1));
+h = sqrt(trace(sigma));

### 2014-10-01 NISN.m

@@ -1,0 +1,27 @@
+function Strat=NISN(fitness,rho,N)
+s=size(fitness);
+S=s(2);
+d=diag(fitness);
+range=1:S;
+strat=[];
+for k=1:S
+    if sum((d(range~=k)<fitness(range~=k,k))==1)==(S-1)
+       strat=[strat k]; 
+    end
+    if sum((d(range~=k)==fitness(range~=k,k))==1)>=1
+        v=range((d(range~=k)==fitness(range~=k,k)));
+        if max(fitness(k,v))<d(k)
+            strat=[strat k];
+        end
+    end            
+end
+if nargin>1
+    Strat=[];
+   for k=strat
+       f=max(rho(range~=k,k));
+        if f>=1/N
+            Strat=[Strat k];
+        end
+   end
+else Strat=strat;
+end

### 2014-10-01 analytical_eigenvectors.R

@@ -1,0 +1,196 @@
+library(igraph)
+
+f<-function(k,i,m,t){
+	x=0
+	if(is.element(i,c(1:(t+1),m-t-1,m-1))){
+		x=tan(pi/(m-2)*k*(i-1))+1
+	}
+	if(is.element(i,(t+2):(m-t-2))){
+		x=tan(pi/(m-2)*k*(i-m/2))+1
+	}
+	x
+}
+
+lambda<-function(k,m,t){
+	vec=matrix(1:t,nrow=1)
+	vec=apply(vec,2,f,k=k,m=m,t=t)
+	vec=cumprod(vec)
+	x=1-1/(2*t)*sum(vec)
+	return(x)
+}
+
+lap<-function(G){
+	n=length(V(G))
+	M=array(0,c(n,n))
+	for(i in 1:n){
+		M[i,neighbors(G,i,mode="in")]=-1/degree(G,i,mode="in")
+	}
+	diag(M)<-1+diag(M)
+	M
+}
+
+circle<-function(n,k){
+	m=array(0,c(n,n))
+	l=ceiling(k/2)
+	indices=rep(1:n,3)
+	if(k==1){
+		for(i in 1:(n-1)){
+			m[i,i+1]=1
+		}
+		m[n,1]=1
+	} else
+	for(i in 1:n){
+		m[i,indices[(i+1:l)+n]]=1
+		m[i,indices[(i-1:(k-l))+n]]=1
+	}
+	return(m)
+}
+
+
+#####
+####
+m=21
+t=2
+G=graph.adjacency(circle(n=m,k=2*t))
+L=lap(G)
+Lhat=L
+Lhat[1,]=0
+
+V=array(0,c(m,m))
+V[,1]=1
+for(i in 1:((m-1)/2)){
+	V[,i+1]=cos(2*pi*i/m*(0:(m-1)))
+	V[,i+(m-1)/2+1]=sin(2*pi*i/m*(0:(m-1)))
+}
+
+Lambda=(L%*%V/V)[m,]
+
+Vhat=eigen(Lhat)$vectors
+Vhat=cbind(Vhat[,m],Vhat)
+Vhat=Vhat[,-(m+1)]
+
+Lambdahat=(Lhat%*%Vhat/Vhat)[m,]
+
+M=V%*%diag(1-Lambda)
+
[…]

### 2014-10-01 consensusspeed.m

@@ -1,0 +1,8 @@
+function s=consensusspeed(M)
+        L=lap(M);
+        [~,d]=eig(L);
+        d=diag(d);
+        good=1-(abs(d)<0.0001);
+        d=d(logical(good));
+        lambda=min(real(d));
+        s=1/sqrt(2*lambda);

### 2014-10-01 correlationlength_mat.m

@@ -1,0 +1,88 @@
+function [distbins,avgcorr,corrlength] = correlationlength_mat(M,d,b,radius)
+% if nargin==5 || strcmp(path,'path')
+%     path=1;
+% else path=0;
+% end
+% 
+% if ~path
+N=size(M,2);
+corrvec=[];
+distvec=[];
+for i=1:N
+    receiver=i;
+    allreceivers=d(receiver,:)<=radius;
+    beta=zeros(N,1);
+    beta(allreceivers)=1;
+    bbeta=b*beta;
+    im=1;
+    while im>0
+        corrs=paircorrelations(M,bbeta);
+        im=max(max(abs(imag(corrs))));
+    end
+    uppercorr=triu(corrs,1);
+    upperdist=triu(d,1);
+    corrvectoadd=uppercorr(~~uppercorr);
+    distvectoadd=upperdist(~~upperdist);
+    corrvec=[corrvec; corrvectoadd; diag(corrs)];
+    distvec=[distvec; distvectoadd; diag(d)];
+end
+
+distbins=0:.02:1.5;
+[~,w]=histc(distvec,distbins);
+l=length(distbins);
+avgcorr=zeros(1,l);
+for i=1:l
+    now=(w==i);
+    avgcorr(i)=mean(corrvec(now));
+end
+i=sum(avgcorr>=0);
+if i<l
+%     corrlength=avgcorr(i)/(avgcorr(i)-avgcorr(i+1))*(distbins(i+1)-distbins(i))+distbins(i);
+    corrlength=find(avgcorr<0,1,'first');
+    corrlength=distbins(corrlength);
+else corrlength=max(distvec);
+end
+% end
+% 
+% if path
+% M(M==-1)=0;
+% M(~~M)=1;
+% g=sparse(M);
+% paths=graphallshortestpaths(g);
+% 
+% 
+% corrvec=[];
+% pathvec=[];
+% for i=1:N
+%     receiver=i;
+%     allreceivers=d(receiver,:)<=radius;
+%     beta=zeros(N,1);
+%     beta(allreceivers)=1;
+%     bbeta=b*beta;
+%     corrs=paircorrelations(M,bbeta);
+%     uppercorr=triu(corrs,1);
+%     upperpath=triu(paths,1);
+%     lowerpath=triu(transpose(paths),1);
+%     corrvectoadd=uppercorr(~~uppercorr);
[…]

### 2014-10-01 corrlength_dependence.m

@@ -1,0 +1,411 @@
+%%
+
+its=20;
+
+N=40;
+
+stratvals=[10 15 20];
+bvals=[0 0.01 .1:.1:1 1.5];
+radvals=[.01 .1 .5 1];
+
+Ns=length(stratvals);
+Nb=length(bvals);
+Nr=length(radvals);
+
+%%
+%find correlation length, correlation between correlation function at b and
+%at b=1000, and H2 norm for networks with various strategies and radii of
+%signal
+corrlengthmat=zeros(its,Ns,Nr);
+similaritymat=zeros(its,Ns,Nr,Nb,2);
+minbmat=zeros(its,Ns,Nr);
+H2mat=zeros(its,Ns);
+for itnow=1:its
+    positions=unifrnd(0,1,N,2);
+    d=squareform(pdist(positions));
+
+    for i=1:Ns
+        strat=stratvals(i);
+        strategy=strat*ones(N,1);
+
+        M=zeros(N);
+        for ind=1:N
+            [~, order]=sort(d(ind,:));
+            neighbors=order(2:strategy(ind)+1);
+            M(ind,neighbors)=1/strategy(ind);
+        end
+        M(1:N+1:end)=-1;
+        H2mat(itnow,i)=H2norm(M);
+        
+        for k=1:Nr
+            radius=radvals(k);
+            similarityvec=zeros(Nb,2);
+            [~,endavgcorr,corrlength] = correlationlength_mat(M,d,1000,radius);
+            stop1=find(isnan(endavgcorr),1,'first')-1;
+            corrlengthmat(itnow,i,k)=corrlength;
+            [~,beginavgcorr,~] = correlationlength_mat(M,d,00,radius);
+            stop2=find(isnan(beginavgcorr),1,'first')-1;
+            stop=min(stop1,stop2);
+            parfor j=1:Nb 
+                b=bvals(j);
+                [~,avgcorr,corrlength] = correlationlength_mat(M,d,b,radius);
+                similarityvec(j,:)=[corr(avgcorr(1:stop)',beginavgcorr(1:stop)'), corr(avgcorr(1:stop)',endavgcorr(1:stop)')]; %#ok<PFBNS>
+            end
[…]

### 2014-10-01 corrlength_parallel.m

@@ -1,0 +1,115 @@
+its=10;
+
+N=40;
+
+stratvals=[10 15 20];
+% stratvals=[10 20];
+bvals=[0 0.01 .1 .5 .75 1:5];
+radvals=[.01 .1 .5 1];
+% radvals=.5;
+
+Ns=length(stratvals);
+Nb=length(bvals);
+Nr=length(radvals);
+
+
+%find correlation length, correlation between correlation function at b and
+%at b=1000, and H2 norm for networks with various strategies and radii of
+%signal
+corrlengthmat=zeros(its,Ns,Nr);
+similaritymat=zeros(its,Ns,Nr,Nb,2);
+minbmat=zeros(its,Ns,Nr);
+H2mat=zeros(its,Ns);
+for itnow=1:its
+    positions=unifrnd(0,1,N,2);
+    d=squareform(pdist(positions));
+
+    for i=1:Ns
+        strat=stratvals(i);
+        strategy=strat*ones(N,1);
+
+        M=zeros(N);
+        for ind=1:N
+            [~, order]=sort(d(ind,:));
+            neighbors=order(2:strategy(ind)+1);
+            M(ind,neighbors)=1/strategy(ind);
+        end
+        M(1:N+1:end)=-1;
+        H2mat(itnow,i)=H2norm(M);
+        
+        for k=1:Nr
+            radius=radvals(k);
+            similarityvec=zeros(Nb,2);
+            [~,endavgcorr,corrlength] = correlationlength_mat(M,d,1000,radius);
+            corrlengthmat(itnow,i,k)=corrlength;
+            [~,beginavgcorr,~] = correlationlength_mat(M,d,00,radius);
+            valid=~isnan(beginavgcorr);
+            parfor j=1:Nb 
+                b=bvals(j);
+                [~,avgcorr,corrlength] = correlationlength_mat(M,d,b,radius);
+                similarityvec(j,:)=[corr(avgcorr(valid)',beginavgcorr(valid)'), corr(avgcorr(valid)',endavgcorr(valid)')]; %#ok<PFBNS>
+            end
+            similaritymat(itnow,i,k,:,:)=similarityvec;
+            v=similarityvec(:,1)-similarityvec(:,2);
+            boundary=sum(v>0);
+            if boundary==Nb
[…]

### 2014-10-01 covariance.m

@@ -1,0 +1,49 @@
+function x = covariance(strategy,beta,probs)
+x=cell(1,3);
+M=makenet(strategy);
+L=-lap(M);
+s=size(M);
+N=s(1);
+dist=zeros(N);
+% noise=diag(sum(transpose(M)));
+noise=eye(N);
+if nargin==2
+    probs=1/N*ones(1,N); %probability of each node being the receiver
+end
+if var(sum(transpose(M)))==0 %if everyone's the same don't bother looping over all receivers
+    L2=L;
+    L2(1,1)=L2(1,1)+beta;
+    sigma = lyap(L2,noise);
+    for i=1:N
+         dist(:,i) = circshift(reshape(diag(sigma),N,1),i-1);
+%          dist(:,i) = circshift(reshape(sigma(1,:),N,1),i-1);
+    end
+    dist=dist*diag(probs);
+    x{1}=sum(diag(sigma))/N;
+    x{2}=sum(sum(sigma))/N^2;
+    x{3}=sum(dist,2);
+end
+if var(sum(transpose(M)))~=0
+    consensus=zeros(1,N);
+    suscept=zeros(1,N);
+    for i=1:N
+        L2=L;
+        L2(i,i)=L2(i,i)+beta;
+        sigma = lyap(L2,noise);
+        dist(:,i) = diag(sigma);
+%         dist(:,i) = sigma(i,:);
+        consensus(i)=sum(diag(sigma))/N;
+        suscept(i)=sum(sum(sigma))/N^2;
+    end
+    dist=dist*diag(probs);
+    consensus=consensus.*probs;
+    suscept=suscept.*probs;
+    x{1}=sum(consensus);
+    x{2}=sum(suscept);
+    x{3}=sum(dist,2);
+end
+
+
+
+
+

### 2014-10-01 dynamics.m

@@ -1,0 +1,9 @@
+function v=dynamics(x,M)
+s=size(x);
+n=s(1);
+v=zeros(n,1);
+for i=1:n
+    neighbors=M(i,:)>0;
+    deg=sum(neighbors);
+   v(i,1)=x(i,1)+sum(x(neighbors,1)-x(i,1))/deg;
+end

### 2014-10-01 finding_equilibria_v2.m

@@ -1,0 +1,437 @@
+    %%
+    %%how group performance depends on the strategy of the group and the
+    %%tradeoff
+    N=20;
+    
+    strats=2:2:N;
+    probs=1/N*ones(1,N);
+    
+    alphavals=0:.25:1;
+    groupshort=zeros(1,length(strats));
+    grouplong=zeros(1,length(strats));
+    groupperf=zeros(length(alphavals),length(strats));
+    indshort=zeros(1,length(strats));
+    indlong=zeros(1,length(strats));
+    indperf=zeros(length(alphavals),length(strats));
+    
+    betavals=-.01:-.5:-2;
+    consensus.ability=zeros(length(betavals),length(strats));
+    
+    for s=1:length(strats)
+            
+        strat=strats(s);
+        strategy=strat*ones(1,N);
+        M=makenet(strategy);
+    
+%         groupshort(s)=consensusspeed(M);
+%         grouplong(s)=(H2norm(M,'additive')/sqrt(N));
+    
+%         v1=speeds(M,'both','mean');
+%         v2=vars(M,'additive','mean');
+%         indshort(s)=v1(1);
+%         indlong(s)=v2(1);
+        
+%         for a=1:length(alphavals)
+%             alpha=alphavals(a);
+%             groupperf(a,s)=1/((1-alpha)*groupshort(s)+alpha*grouplong(s));
+%             indperf(a,s)=1/((1-alpha)*v1(1)+alpha*v2(1));
+%         end
+        for b=1:length(betavals)
+            beta=betavals(b);
+            c=covariance(strategy,beta,probs);
+            consensus.ability(b,s)=1/c{1};
+        end
+    end
+        
+    
+    HM=HeatMap(groupperf,'RowLabels',alphavals,'ColumnLabels',strats);
+    P=HM.plot;
+    colormap hot
+    colorbar('Peer',P)
+    caxis(P,[0 max(max(groupperf))])
+    xlab=xlabel(P,'Strategy');
+    ylab=ylabel(P,'Tradeoff');
+    main=title(P,['Group Performance' ', N=' num2str(n)]);
+    set([xlab ylab],'FontName','Arial','FontSize',12);
+    set(main, 'FontName','Arial','FontSize',15);
+    xpos=get(xlab,'Position');
+    set(xlab,'Position',xpos+[0 .5 0]);
[…]

### 2014-10-01 fitness.m

@@ -1,0 +1,42 @@
+function m=fitness(N,alpha)
+strats=2:2:(N-1);
+    L=length(strats);
+    fitness=zeros(L,L);
+    rho=zeros(L,L);
+    
+
+        for i=1:L
+            for j=1:L
+            resident=strats(i);
+            invader=strats(j);
+
+            f=zeros(1,N-1);
+            g=zeros(1,N-1);
+
+            for k=1:(N-1)
+                strategy=resident*ones(1,N);
+                strategy(1:k)=invader;
+                M=makenet(strategy);
+                v1=speeds(M,'both','mean');
+                v2=vars(M,'additive','mean');
+                perf=1./((1-alpha)*v1+alpha*v2);
+                f(k)=mean(perf(1:k));
+                g(k)=mean(perf((k+1):N));
+                if k==1
+                    fitness(i,j)=f(k)/g(k);
+                end
+        
+            end
+            
+            ratio=g./f;
+            tosum=ones(1,N-1);
+            for k=1:(N-1)
+                tosum(k)=prod(ratio(1:k));
+            end
+            rho(i,j)=1/(1+sum(tosum));
+            
+            end
+        
+        end
+m{1}=fitness;
+m{2}=rho;

### 2014-10-01 greedy_opt.m

@@ -22,7 +22,6 @@ if t==1
 else strategy=evolution(:,t-1);
     strategy=reshape(strategy,1,N);
 
-
 M=makenet(strategy);
 
 v1=1./power(speeds(M,'both','max'),1-alpha);

### 2014-10-01 information_evolution_v2.R

@@ -1,0 +1,171 @@
+library(igraph)
+library(gplots)
+
+#####FUNCTIONS 
+dynamics<-function(x,G){
+	n<-length(x)
+	new<-x
+	for(i in 1:n){
+		new[i]=x[i]+sum((x[neighbors(G,i,mode="in")]-x[i])/degree(G,i,mode="in"))
+	}
+	new
+}
+
+lap<-function(G){
+	n=length(V(G))
+	M=array(0,c(n,n))
+	for(i in 1:n){
+		M[neighbors(G,i,mode="in"),i]=-1/degree(G,i,mode="in")
+	}
+	diag(M)<-1
+	M
+}
+
+Q<-function(n){
+	a=rep(1,n)
+	a=a/norm(a,type="2")
+	m=array(0,c(n-1,n))
+	m[1,1:2]=c(1/sqrt(2),-1/sqrt(2))
+	for(j in 2:(n-1)){
+		c=runif(n,-1,1)
+		c=c-sum(c)/n
+		for(k in 1:(j-1)){
+			c<-c-(c%*%m[k,])*m[k,]
+		}
+		c<-c/norm(c,type="2")
+		m[j,]=c
+	}
+	m
+}
+
+perp<-function(v){
+	n=length(v)
+	a=v/norm(v,type="2")
+	m=array(0,c(n-1,n))
+	c=runif(n,-1,1)
+	c=c-(c%*%a)*a
+	c=c/norm(c,type="2")
+	m[1,]=c
+	for(j in 2:(n-1)){
+		c=runif(n,-1,1)
+		c=c-(c%*%a)*a
+		for(k in 1:(j-1)){
+			c<-c-(c%*%m[k,])*m[k,]
+		}
+		c<-c/norm(c,type="2")
+		m[j,]=c
+	}
+	m
+}
+
+H2<-function(G){
+	vals<-eigen(lap(G))$values
+	vals<-Re(vals)
+	vals<-vals[-length(vals)]
+	h=sqrt(1/2*sum(1/vals))
+	h
+}
+
+nextgen<-function(strategy,fitness){
+	l<-floor(length(fitness)/2)
+	replace=order(fitness)[(length(fitness)-l+1):length(fitness)]
+	v=strategy
+	v[replace]=strategy[order(fitness)[1:l]]
+	v
+}
+
+id<-function(n){
+	m=array(0,c(n,n))
+	diag(m)=1
+	return(m)
+}
+
+N<-50
+
+rainb<-rainbow(N)
+
[…]

### 2014-10-01 paralleltestfun.m

@@ -1,0 +1,5 @@
+function A = paralleltestfun(n)
+
+A=randi(n,n);
+
+end

### 2014-10-01 pathlength_scratchwork.m

@@ -1,0 +1,22 @@
+N=10;
+strategy=4*ones(N,1);
+positions=unifrnd(0,1,N,2);
+    d=squareform(pdist(positions));
+
+    M=zeros(N);
+    for ind=1:N
+        [~, order]=sort(d(ind,:));
+        neighbors=order(2:strategy(ind)+1);
+        M(ind,neighbors)=1;
+    end
+    g=sparse(M);
+    
+    paths=graphallshortestpaths(g);
+    upperpaths=triu(paths,1);
+    lowerpaths=triu(transpose(paths),1);
+    pathsvec=[upperpaths(~~upperpaths), lowerpaths(~~lowerpaths)]
+    pathsvec=min(pathsvec')';
+    
+    [distbins,avgcorr,corrlength] = correlationlength(M,d,b,radius);
+    
+    

### 2014-10-01 second_eigenvalue.R

@@ -1,0 +1,99 @@
+library(igraph)
+library(R.matlab)
+
+lap<-function(G){
+	n=length(V(G))
+	M=array(0,c(n,n))
+	for(i in 1:n){
+		M[i,neighbors(G,i,mode="in")]=-1/degree(G,i,mode="in")
+	}
+	diag(M)<-1+diag(M)
+	M
+}
+
+id<-function(n){
+	m=array(0,c(n,n))
+	diag(m)=1
+	return(m)
+}
+
+N=50
+
+its=1000
+
+vals=array(,its)
+vals2=array(,its)
+sums=array(,its)
+
+j=1
+while(j<=its){
+
+positions<-array(,c(N,2))
+positions[,1]=runif(N,0,1)
+positions[,2]=runif(N,0,1)
+d=as.matrix(dist(positions))
+
+strategy=sample(1:(N-1),N,replace=TRUE)
+
+M<-array(0,c(N,N))
+for(i in 1:N){
+	neighbors=order(d[i,])[2:(strategy[i]+1)]
+	M[i,neighbors]=1/strategy[i]
+}
+L=id(N)-M
+
+# G=erdos.renyi.game(N,p=.4,type="gnp",directed=TRUE)
+# L=lap(G)
+i=which.max(abs(Re(eigen(t(L))$vectors[,N])))
+x=Re(eigen(L)$values)[N-1]
+x=round(x,5)
+y=array(0,N)
+for(k in 1:N){
+y[k]=Re(eigen(L[-k,-k])$values)[N-1]
+}
+sums[j]=sum(y)
+y=round(y,5)
+z=x-y[i]
+vals[j]=x
+vals2[j]=y[i]
+j=j+1
+print(c(x,z))
+if(z<0){
+	print("Negative")
+	j=its+1
+	}
+# if(z==0){
+	# print("interesting")
+	# j=its+1
+# }
+}
+
+# par(oma=c(0,4,0,0))
+plot(vals,vals2,bty="n",cex.lab=1.5,xlab=expression(bar(lambda)),ylab=expression(lambda^i),mgp=c(2.5,1,0))
+abline(0,1)
+
+claim<-function(L,v){
+	lambda=eigen(L)$values
+	us=eigen(L)$vectors
+	i=which.max(abs(v%*%us/norm(v,type='2')))
+	lambda=lambda[-i]
+	gap=round(min(diff(rev(lambda))),5)
+	p=perp(v)
[…]

### 2014-10-01 simulated_annealing.m

@@ -1,0 +1,59 @@
+%%
+
+N=20;
+its=10000;
+temp=2;
+
+randits=10;
+finalstrats=zeros(N,randits);
+H2vovertime=zeros(randits,its+1);
+corrlengthovertime=zeros(randits,its+1);
+
+numsigs_permove=1;
+nummoves=100;
+radius=.1;
+b=1;
+T=1;
+
+parfor r=1:randits
+    H2vec=zeros(its+1,1);
+    corrvec=zeros(its+1,1);
+    strategy=zeros(N,its+1);
+    perf=zeros(N,its+1);
+    mutant=zeros(1,its+1);
+
+    recordofchanges=zeros(1,its+1);
+
+    strategy(:,1)=2*randi([1 floor((N-1)/2)],N,1);
+    [meanH2, meancorrlength, probeaten, ~]=signalingevents(strategy(:,1),numsigs_permove,nummoves,radius,b,T);
+    perf(:,1)=probeaten;
+    H2vec(1)=meanH2;
+    corrvec(1)=meancorrlength;
+    for time=(1:its)+1
+        strategy(:,time)=strategy(:,time-1);
+
+        changer=randi([1 N],1);
+        mutant(time)=changer;
+        test=[reshape(strategy(:,time-1),N,1) reshape(strategy(:,time-1),N,1)];
+        v=exp(-abs((2:2:floor((N-1)/2))-test(changer,1))/temp);
+        probs=v/sum(v);
+        test(changer,2)=randchoice(2:2:floor((N-1)/2),probs);
+        [meanH2, meancorrlength, probeaten, ~]=signalingevents(test(:,2),numsigs_permove,nummoves,radius,b,T);
+        changed=probeaten;
+        performance=[perf(:,time-1) changed];
+        [~,w]=min(performance(changer,:));
+        strategy(changer,time)=test(changer,w);
+        H2vec(time)=meanH2;
+        corrvec(time)=meancorrlength;
+        if w==2 && strategy(changer,time-1)~=strategy(changer,time)
+            recordofchanges(time)=1;
+        end
+
+        [~, ~, probeaten, ~]=signalingevents(strategy(:,time),numsigs_permove,nummoves,radius,b,T);
+        perf(:,time)=probeaten;
+    end
+
+    finalstrats(:,r)=strategy(:,end);
+    H2overtime(r,:)=H2vec;
+    corrlengthovertime(r,:)=corrvec;
+end

### 2014-10-01 speeds.m

@@ -1,0 +1,68 @@
+function rates=speeds(M,lambdaorv,meanormax)
+if nargin==1
+    lambdaorv='both';
+    meanormax='mean';
+end
+s=size(M);
+N=s(2);
+L=lap(M);
+delta=zeros(N);
+if var(sum(transpose(M)))==0 %if everyone's the same don't bother looping over all receivers
+        [path_dists,~,~]=graphshortestpath(sparse(transpose(M)),1);
+        not_children=(path_dists==Inf);
+        children=(path_dists~=Inf);
+        children(1)=0;
+        delta(not_children,1)=0;
+        if sum(children>0)
+            W=L(children,children);
+            [v d]=eig(W);
+            [lambda,index] = min(real(diag(d)));
+            v=v(:,index);
+            v=v/mean(v);
+            v=abs(v);
+            if strcmp(lambdaorv,'both')==1
+            %delta(children,i)=exp(-minValue)*abs(v(:,index));
+            delta(children,1)=v/sqrt(2*lambda);
+            elseif strcmp(lambdaorv,'lambda')==1
+            %delta(children,i)=exp(-minValue);
+            delta(children,1)=1/sqrt(2*lambda);
+            else delta(children,1)=v;
+            end
+        end
+    if strcmp(meanormax,'mean')==1
+        rates=mean(delta(:,1))*ones(N,1);
+    else 
+        rates=max(delta(:,1))*ones(N,1);
+    end
+end 
+if var(sum(transpose(M)))~=0
+    for i=1:N
+        [path_dists,~,~]=graphshortestpath(sparse(transpose(M)),i);
+        not_children=(path_dists==Inf);
+        children=(path_dists~=Inf);
+        children(i)=0;
+        delta(not_children,i)=0;
+        if sum(children>0)
+            W=L(children,children);
+            [v d]=eig(W);
+            [lambda,index] = min(real(diag(d)));
+            v=v(:,index);
+            v=v/mean(v);
+            v=abs(v);
+            if strcmp(lambdaorv,'both')==1
+            %delta(children,i)=exp(-minValue)*abs(v(:,index));
+            delta(children,i)=v/sqrt(2*lambda);
+            elseif strcmp(lambdaorv,'lambda')==1
+            %delta(children,i)=exp(-minValue);
+            delta(children,i)=1/sqrt(2*lambda);
+            else delta(children,i)=v;
+            end
+        end
+    end
+    if strcmp(meanormax,'mean')==1
[…]

### 2014-10-01 triimage.m

@@ -1,0 +1,54 @@
+function triimage(scaledvals,vals,coords,cornerlabels,mainlabel)
+
+%%
+textfontsz=20;
+labfontsz=15;
+figure
+set(gcf,'Color','w')
+set(gcf,'PaperUnits','inches')
+v=get(gcf,'Position');
+ratio=v(4)/v(3);
+w=6.83;
+h=w*ratio;
+
+c=colormap;
+spaced=linspace(0,1,size(c,1));
+l=size(coords,1);
+
+xoffset=.25;
+yoffset=.2;
+
+hold on
+axis off
+axis equal
+
+for i=1:l
+    p=coords(i,:);
+    valnow=scaledvals(i);
+    ind=sum(spaced<=valnow);
+    if ind==size(c,1)
+        valcol=c(end,:);
+    else
+        valcol=c(ind,:)+(valnow-spaced(ind))/(spaced(ind+1)-spaced(ind))*(c(ind+1,:)-c(ind,:));
+    end
+    fill([p(1),p(1)+cstep,p(1)+cstep,p(1),p(1)-cstep,p(1)-cstep],[p(2)-2/sqrt(3)*cstep,p(2)-1/sqrt(3)*cstep,p(2)+1/sqrt(3)*cstep,p(2)+2/sqrt(3)*cstep,p(2)+1/sqrt(3)*cstep,p(2)-1/sqrt(3)*cstep],valcol,'EdgeColor','none');
+end
+
+text(0-xoffset,-yoffset,cellstr(cornerlabels(1)),'Color','k','HorizontalAlignment','left','FontSize',textfontsz)
+text(2+xoffset,-yoffset,cellstr(cornerlabels(2)),'Color','k','HorizontalAlignment','right','FontSize',textfontsz)
+text(1,sqrt(3)+yoffset,cellstr(cornerlabels(3)),'HorizontalAlignment','center','Color','k','Clipping','off','FontSize',textfontsz)
+% text(0,sqrt(3),'MUTUAL','HorizontalAlignment','center','Color','k','Clipping','off','FontSize',textfontsz)
+% text(0,sqrt(3)-yoffset,'INFORMATION','HorizontalAlignment','center','Color','k','Clipping','off','FontSize',textfontsz)
+
+    
+mybar=colorbar;
+tickvec=0:.2:1;
+ticklabvec=tickvec*(max(vals)-min(vals))+min(vals);
[…]

### 2014-10-01 trying_parallelization.m

@@ -1,0 +1,151 @@
+% matlabpool local 2
+
+numsigs_permove=1;
+nummoves=1000;
+N=20;
+b=1;
+
+strats=2:2:(N-1);
+L=length(strats);
+ 
+T=1;
+radius=.5;
+
+fitnesseaten=zeros(L,L);
+rhoeaten=zeros(L,L);
+
+fitnessgettoeat=zeros(L,L);
+rhogettoeat=zeros(L,L);
+
+featen=zeros(L,L,N-1);
+geaten=zeros(L,L,N-1);
+
+fgettoeat=zeros(L,L,N-1);
+ggettoeat=zeros(L,L,N-1);
+
+tic
+parfor ind=1:(L*L*(N-1))
+    [u,v,k]=ind2sub([L,L,N-1],ind);
+    squareind=sub2ind([L,L],u,v);
+    resident=strats(u);
+    invader=strats(v);
+
+    strategy=resident*ones(1,N);
+    strategy(1:k)=invader;
+    [meanscore, scorevar, probeaten, probgettoeat]=signalingevents(strategy,numsigs_permove,nummoves,radius,b,T);
+
+    perfeaten=1-probeaten;
+    featen(ind)=mean(perfeaten(1:k));
+    geaten(ind)=mean(perfeaten((k+1):N));
+
+    perfgettoeat=probgettoeat;
+    fgettoeat(ind)=mean(perfgettoeat(1:k));
+    ggettoeat(ind)=mean(perfgettoeat((k+1):N));
+         
+%     if k==N-1
+%         fitnesseaten(squareind)=featen(k)/geaten(k);
+%         fitnessgettoeat(squareind)=fgettoeat(k)/ggettoeat(k);
+%     end
+%     ratio=geaten./featen;
+%     tosum=ones(1,N-1);
+%     for k=1:(N-1)
+%         tosum(k)=prod(ratio(1:k));
+%     end
+%     rhoeaten(squareind)=1/(1+sum(tosum));
+% 
+%     ratio=ggettoeat./fgettoeat;
+%     tosum=ones(1,N-1);
+%     for k=1:(N-1)
+%         tosum(k)=prod(ratio(1:k));
+%     end
+%     rhogettoeat(squareind)=1/(1+sum(tosum));
+%     end
+end
+toc
+
+tic
+for ind=1:L*L
+    [u,v]=ind2sub([L,L],ind);
+    where=zeros(N-1,1);
+    for k=1:N-1
[…]

### 2014-11-03 paper_skeleton.tex
just updating the code for a regular check up

@@
 \\ Eleanor Brush
-\\ May 1, 2014
+\\ October 31, 2014
 \end{center}
@@
 
-\section{Learning dynamics}
+\section{Model }
+\subsection{Learning dynamics}
 In this model, an individual's opinion changes with a probability that depends on its social interactions and the strength of the environmental stimulus.
@@
 
-\section{Individual-level measures of performance}
+\subsection{Individual-level performance }
 The group is made up of $N$ individuals.
@@
 We therefore find, over many signaling events, the probability that an individual will be eaten and an individual's fitness is given by the probability of surviving.
-If the signal indicates the presence of food, on the other hand, fitness may be proportional to how well an individual learns the external signal.
+If there are resources available, we assume that fitness is proportional to the probability of being the first to reach the resources.
 
-\section{Optimization methods}
+\subsection{Group-level performance }
+We consider two group-level properties that indicate how cohesive the group is.
+The first property is $\mathscr{H}_2$ robustness.
+This is a measure of the robustness of the consensus state in which all birds have the same opinion to noise at equilibrium (see Appendix Sec.
+\ref{H2}).
+The second property is the correlation length of the flock.
+This is the distance at which the average correlations between birds changes from positive to negative, i.e.
+the distance over which birds' opinions tend to be positively correlated with each other (see Appendix Sec \ref{corr}).
+
+\subsection{Optimization methods }
 To understand what strategies we might expect to find, we are interested in identifying the optimal strategies.
@@
 
-\section{ESS }
-\subsection{Predation }
+\section{Results }
 
-\subsection{Access to food }
-
-\section{Learned optima }
-
-\subsection{Predation }
-
-\subsection{Access to food }
-\section{Comparison to data}
+\subsection{ESS }
+\subsection{Learned optima }
+\subsection{Relationship between group properties }
 
@@
 
+\section{Figures }
+\begin{figure}[ht]
+\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/ESSneighbors.pdf}
+\caption{\label{ESS} The ESS number of neighbors is always higher if selection is due to predation than due to resources.
+The ESS number of neighbors when selection is due to predation is a non-monotonic function of how public the signal is.
+The x-axis indicates the radius of the signal and the y-axis indicates the ESS number of neighbors.
+}
+\end{figure}
+
+\begin{figure}[ht]
+\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/greedyoptneighbors.pdf}
+\caption{\label{greedyopt} When the birds choose the optimal number of neighbors, given the strategies the rest of the flock are using, they settle on higher strategies when selection is due to predation than when selection is due to resources.
+Under selection due to predation, the $\mathscr{H}_2$ robustness and correlation length of the flock tend to decrease, whereas under selection due to resources, $\mathscr{H}_2$ robustness is relatively constant and correlation length tends to increase.
+The upper row shows results from implementing selection based to predation and the lower row shows results from implementing selection based on resources.
+In each panel, the x-axis represents the number of times the birds are allowed to choose optimal strategies.
+The first column shows one example of how the birds' strategies change over time.
+The second column shows, for many initial conditions, how the $\mathscr{H}_2$ robustness changes over time.
+The third column shows, for many initial conditions, how the correlation length changes over time.
+}
+\end{figure}
+
 \section{Appendix}
+\subsection{H2 Norm \label{H2}}
+
+\ Let $y$ represent the distance between the vector of opinions and consensus, i.e.
+$y=QxQ^T,$
+where $Q\in\R^{n-1\times n}$ is such that $Q\vec{1}=0$ , $QQ^T=I_{n-1}$, and $Q^TQ=I_n-\frac{1}{n}\vec{1}\times\vec{1}^T$.   
+Then 
+\begin{equation}
+\dot{y}(t)=-\overline{L}y(t)+Q\xi(t) \notag
+\end{equation}
+where $\overline{L}=QLQ^T$.  As shown in \cite{Young:2010fk}, if $\Sigma_y(t)=\E[y(t)y(t)^T]$,  $\Sigma_y$ at equilibrium solves
+\begin{equation} 0=-\overline{L}\Sigma_y-\Sigma_y\overline{L}^T+D. \label{h2norm}
+\end{equation}
+\begin{align*}
+\mathscr{H}_2&=\sqrt{\Tr(\Sigma_y)}
+\\\mathscr{H}_2 \text{ robustness} &=\frac{1}{\mathscr{H}_2}
+\end{align*}
+
+\subsection{Calculating correlations \label{corr}}
+The following is derived from the work in \cite{Bialek:2013fk}.
+\begin{align*}
+H(\vec{v})&=-\vec{v}^TA\vec{v}-\vec{\beta}^T\vec{v}
+%\\&=\vec{v}A\vec{v}-\vec{\beta}^T\vec{v}+\vec{\beta}^T\vec{1}
+\end{align*}
+Define $V=\sum_kv_k/N$ and $\vec{y}=(\vec{v}-V\vec{1})/V$.  Then $\vec{v}=V(\vec{1}+\vec{y})$.  
+\begin{align*}
+H(\vec{y},V)&=-V^2(\vec{1}^T+\vec{y}^T)A(\vec{1}+\vec{y})-\vec{\beta}^TV(\vec{1}+\vec{y})
+\\&=-V^2(\vec{1}^T+\vec{y}^T)A\vec{y}-V\vec{\beta}^T\vec{y}-V\vec{\beta}^T\vec{1} \text{ since $A\vec{1}=\vec{0}$}
+\\&=-V^2\vec{y}^TA\vec{y}+(-V^2\vec{1}^TA-V\vec{\beta}^T)\vec{y}-V\vec{\beta}^T\vec{1}
+\end{align*}
+Define $Q\in M_{N-1,N}$ that rotates an $N$-vector away from the consensus vector and define $\vec{z}=Q\vec{y}$ so that $\vec{y}=Q^T\vec{z}$ since $\sum_ky_k=0$.
+\begin{align*}
+H(\vec{z},V)&=-V^2\vec{z}^TQAQ^T\vec{z}+(-V^2\vec{1}^TAQ^T-V\vec{\beta}^TQ^T)\vec{z}-V\vec{\beta}^T\vec{1}
+\\&=V^2\vec{z}^TP\vec{z}+(-V^2\vec{\sigma}_1^T-V\vec{\sigma}_2^T)\vec{z}-V\vec{\beta}^T\vec{1}
+\\ \Rightarrow P(\vec{z},V)&=\frac{1}{\z}\exp\left(-V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+V\vec{\beta}^T\vec{1}\right)
+\end{align*}
+where $P=-QAQ^T$, $\vec{\sigma}_1=QA^T\vec{1}$, and $\vec{\sigma}_2=Q\vec{\beta}$.
+
+\begin{fact}
+\begin{align*}
+\int\exp\left(-\frac{1}{2}\vec{z}^TP\vec{z}+\vec{\sigma}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{\det P}}\exp\left(-\frac{1}{2}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
+\\\Rightarrow \int\exp\left(-\vec{z}^TP\vec{z}+\vec{\sigma}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{2^{N-1}\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
+\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
+\end{align*}
+\end{fact}
+
+\begin{align*}
+\z&=\int_{-\infty}^\infty\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+V\vec{\beta}^T\vec{1}\right)d^NzdV
+\\&=\int_{-\infty}^\infty \exp(V\vec{\beta}^T\vec{1})\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}\right)d^NzdV
+\\&=\sqrt{\pi^{N-1}}\int_{-\infty}^\infty\exp(V\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}\det{P}}}\right)\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
+\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det PP}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\exp(V\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}}}\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
+\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
+\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\Pi_a\lambda_a}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
+\\&\text{ where $\lambda_a$ are the eigenvalues of $P$}
+\\ \Rightarrow -\log(\z)&=\text{constants}+\frac{1}{2}\sum_a\log(\lambda_a)+\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma_1}-\log(f(P))
+\end{align*}
+
+where $f(P)=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV $.
+
+\begin{align*}
+\Rightarrow\frac{\partial -\log(\z)}{\partial P_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{1l}-\frac{\partial \log f}{\partial P_{ij}}
+\\\Rightarrow\frac{\partial -\log(\z)}{\partial P_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}{ki}P^{-1}_{jl}\sigma_{1l}-\frac{1}{f}\frac{\partial f}{\partial P_{ij}}
+\end{align*}
+
+\begin{align*}
+\frac{\partial f}{\partial P_{ij}}&=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)\left(-\frac{1}{4V}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)dV
+\\&=\left(-\frac{1}{4}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
+\end{align*}
+
+\small{
+\begin{align*}
+\Rightarrow\langle z_iz_j\rangle &=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}{ki}P^{-1}_{jl}\sigma_{1l}+\left(\frac{1}{4}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)\frac{\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV}{\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV}
+\end{align*}}
+
 \subsection{Equivalence of dynamics and Hamiltonian }
@@
 
-\subsection{Calculating correlations }
-The following is derived from the work in \cite{Bialek:2013fk}.
-\begin{align*}
-H(\vec{v})&=-\vec{v}^TA\vec{v}-\vec{\beta}^T\vec{v}
-%\\&=\vec{v}A\vec{v}-\vec{\beta}^T\vec{v}+\vec{\beta}^T\vec{1}
-\end{align*}
-Define $V=\sum_kv_k/N$ and $\vec{y}=(\vec{v}-V\vec{1})/V$.  Then $\vec{v}=V(\vec{1}+\vec{y})$.  
-\begin{align*}
-H(\vec{y},V)&=-V^2(\vec{1}^T+\vec{y}^T)A(\vec{1}+\vec{y})-\vec{\beta}^TV(\vec{1}+\vec{y})
-\\&=-V^2(\vec{1}^T+\vec{y}^T)A\vec{y}-V\vec{\beta}^T\vec{y}-V\vec{\beta}^T\vec{1} \text{ since $A\vec{1}=\vec{0}$}
-\\&=-V^2\vec{y}^TA\vec{y}+(-V^2\vec{1}^TA-V\vec{\beta}^T)\vec{y}-V\vec{\beta}^T\vec{1}
-\end{align*}
-Define $Q\in M_{N-1,N}$ that rotates an $N$-vector away from the consensus vector and define $\vec{z}=Q\vec{y}$ so that $\vec{y}=Q^T\vec{z}$ since $\sum_ky_k=0$.
-\begin{align*}
-H(\vec{z},V)&=-V^2\vec{z}^TQAQ^T\vec{z}+(-V^2\vec{1}^TAQ^T-V\vec{\beta}^TQ^T)\vec{z}-V\vec{\beta}^T\vec{1}
-\\&=V^2\vec{z}^TP\vec{z}+(-V^2\vec{\sigma}_1^T-V\vec{\sigma}_2^T)\vec{z}-V\vec{\beta}^T\vec{1}
-\\ \Rightarrow P(\vec{z},V)&=\frac{1}{\z}\exp\left(-V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+V\vec{\beta}^T\vec{1}\right)
-\end{align*}
-where $P=-QAQ^T$, $\vec{\sigma}_1=QA^T\vec{1}$, and $\vec{\sigma}_2=Q\vec{\beta}$.
-
-\begin{fact}
-\begin{align*}
-\int\exp\left(-\frac{1}{2}\vec{z}^TP\vec{z}+\vec{\sigma}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{\det P}}\exp\left(-\frac{1}{2}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
-\\\Rightarrow \int\exp\left(-\vec{z}^TP\vec{z}+\vec{\sigma}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{2^{N-1}\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
-\end{align*}
-\end{fact}
-
-\begin{align*}
-\z&=\int_{-\infty}^\infty\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+V\vec{\beta}^T\vec{1}\right)d^NzdV
-\\&=\int_{-\infty}^\infty \exp(V\vec{\beta}^T\vec{1})\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}\right)d^NzdV
-\\&=\sqrt{\pi^{N-1}}\int_{-\infty}^\infty\exp(V\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}\det{P}}}\right)\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det PP}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\exp(V\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}}}\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\Pi_a\lambda_a}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\\&\text{ where $\lambda_a$ are the eigenvalues of $P$}
-\\ \Rightarrow -\log(\z)&=\text{constants}+\frac{1}{2}\sum_a\log(\lambda_a)+\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma_1}-\log(f(P))
-\end{align*}
-
-where $f(P)=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV $.
-
-\begin{align*}
-\Rightarrow\frac{\partial -\log(\z)}{\partial P_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{1l}-\frac{\partial \log f}{\partial P_{ij}}
-\\\Rightarrow\frac{\partial -\log(\z)}{\partial P_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}{ki}P^{-1}_{jl}\sigma_{1l}-\frac{1}{f}\frac{\partial f}{\partial P_{ij}}
-\end{align*}
-
-\begin{align*}
-\frac{\partial f}{\partial P_{ij}}&=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)\left(-\frac{1}{4V}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)dV
-\\&=\left(-\frac{1}{4}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\end{align*}
-
-\begin{align*}
-\Rightarrow \langle z_iz_j\rangle &=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}{ki}P^{-1}_{jl}\sigma_{1l}+\left(\frac{1}{4}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)\frac{\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV}{\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV}
-\end{align*}
-
 \nocite{*}

### 2014-11-03 ESS_parallel_long.m

@@ -55,7 +55,7 @@ for q=1:Nt
 
             strategy=resident*ones(1,N);
             strategy(1:k)=invader;
-            [probeaten, probgettoeat]=signalingevents(strategy,numsigs_permove,nummoves,radius,b,T);
+            [probeaten, probgettoeat]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
 
             perfeaten=1-probeaten;
             featen(ind)=mean(perfeaten(1:k));

### 2014-11-03 ESS_parallel_test.m

@@ -1,0 +1,43 @@
+numworkers=str2num(getenv('PROCS'));
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
+
+numsigs_permove=str2num(getenv('numsigs'));
+nummoves=str2num(getenv('nummoves'));
+
+N=20;
+b=1;
+
+strats=1:1:(N-1);
+L=length(strats);
+
+Tvals=[.01 .1 1 10];
+Nt=length(Tvals);
+
+radvals=[0:.1:1.4];
+% radvals=.1;
+Nr=length(radvals);
+
+radius=radvals(2);
+T=Tvals(3);
+strategy=[1;19*ones(N-1,1)];
+
+t0=tic;
+[probeaten, probgettoeat]=signalingevents(strategy,numsigs_permove,nummoves,radius,b,T);
+t=toc(t0);
+
+t0=tic;
+[probeatenshort, probgettoeatshort]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+tshort=toc(t0);
+
+t0=tic;
+[probeatenlong, probgettoeatlong]=signalingevents_parallel(strategy,numsigs_permove*N,nummoves,radius,b,T);
+tlong=toc(t0);
+
+filename=strcat('/home/brush/schooling_consensus/ESStest','.mat');
+save(filename,'t','tshort','tlong','probeaten','probeatenshort','probeatenlong');
+
+delete(dellapool);
+
+exit ;

### 2014-11-03 correlationlength_mat_single_v3.m

@@ -19,10 +19,10 @@ l=length(distvec);
 
 pos=sum(corrvec>=0);
 mimag=max(abs(imag(corrvec)));
-if mimag~=0
+if mimag~=0 || sum(isnan(corrvec))>0
     corrlength=0;
 else
-    if pos<l
+    if pos<l && pos>1
         f=find(corrvec<0,1,'first');
         corrlength=interp1q(corrvec(f:-1:(f-1))',distvec(f:-1:(f-1))',0); 
     else corrlength=max(distvec);

### 2014-11-03 essparalleltest.m

@@ -1,0 +1,3 @@
+t0=tic;
+[probeaten, probgettoeat]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+t=toc(t0)

### 2014-11-03 expected_spin_test.m

@@ -1,0 +1,242 @@
+N=20;
+radius=1.4;
+T=.1;
+nummoves=1000;
+numsigs_permove=1;
+b=0;
+
+strategy=[17,16*ones(1,N-1)];
+numsigs_tot=numsigs_permove*nummoves;
+N=max(size(strategy));
+indvec=1:N;
+
+positions=unifrnd(0,1,N,2);
+d=squareform(pdist(positions));
+
+M=zeros(N);
+for ind=1:N
+    [~, order]=sort(d(ind,:));
+    neighbors=order(2:strategy(ind)+1);
+    M(ind,neighbors)=1/strategy(ind);
+end
+M(1:N+1:end)=-1; %sets diagonal equal to -1
+receivers=randsample(N,numsigs_permove,'true');
+j=1;
+beta=zeros(N,1);
+receiver=receivers(j);
+allreceivers=d(receiver,:)<=radius;
+beta(allreceivers)=b;
+v=real(expected_spin(M,T,beta));
+v2=real(expected_spin(M,T,beta));
+
+perm=randsample(20,20,'false');
+Mperm=M(perm,perm);
+betaperm=beta(perm);
+vperm=real(expected_spin(Mperm,T,betaperm));
+% [v(perm) vperm v2]
+
+%%
+A=M-diag(beta);
+N=size(M,1);
+
+[vecs,vals]=eig(A);
+V=vecs;
+vals=diag(vals);
+nonzero=abs(vals)>0.00001;
+nonzerovals=vals(nonzero);
+
+Aperm=Mperm-diag(betaperm);
+
+[vecsperm,valsperm]=eig(Aperm);
+Vperm=vecsperm;
+valsperm=diag(valsperm);
+nonzeroperm=abs(valsperm)>0.00001;
+nonzerovalsperm=valsperm(nonzeroperm);
+
+indices=1:N;
+vhat=1/sqrt(N)*ones(N,1);
+where=indices(~nonzero);
+if sum(~nonzero)==1
+    V(:,where(1))=ones(N,1);
+end
+if sum(~nonzero)>1
+    L=sum(~nonzero);
+    for i=1:(L-1)
+        vproj=V(:,where(i+1))-(V(:,where(i+1))'*vhat)*vhat;
+        V(:,where(i+1))=vproj;
+    end
+end
+
+whereperm=indices(~nonzeroperm);
+if sum(~nonzeroperm)==1
+    Vperm(:,whereperm(1))=ones(N,1);
+end
+if sum(~nonzeroperm)>1
[…]

### 2014-11-03 greedyopt_v2.m

@@ -1,0 +1,153 @@
+numworkers=str2num(getenv('PROCS'));
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
+
+numsigs_permove=1;
+nummoves=1000;
+
+N=20;
+b=1;
+radius=.1;
+T=1;
+
+timesteps=50;
+its=20;
+
+evolution_eaten=zeros(its,N,timesteps);
+groupspeed_eaten=zeros(its,timesteps);
+groupconsensus_eaten=zeros(its,timesteps);
+corrlengths_eaten=zeros(its,timesteps);
+disconnected_eaten=zeros(its,timesteps);
+
+evolution_gettoeat=zeros(its,N,timesteps);
+groupspeed_gettoeat=zeros(its,timesteps);
+groupconsensus_gettoeat=zeros(its,timesteps);
+corrlengths_gettoeat=zeros(its,timesteps);
+disconnected_gettoeat=zeros(its,timesteps);
+
+for num=1:its
+
+for t=1:timesteps
+    init=randi([1 N-1],1,N);
+if t==1
+    strategy=init;
+%     strategy=2*ones(1,N);
+%     strategy(1)=10; 
+    evolution_eaten(num,:,t)=strategy;
+else
+    strategy=evolution_eaten(num,:,t-1);
+    strategy=reshape(strategy,1,N);
+
+    [probeaten, ~]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+    perf=1-probeaten;
+    newstrategy=strategy;
+
+    for i=1:N
+        choosestrat=zeros(1,3);
+        choosestrat(2)=perf(i);
+
+        if strategy(i)==N-1
+            choosestrat(3)=perf(i);
+        else 
+            upone=strategy;
+            upone(i)=strategy(i)+1;
+
+            [probeaten, ~]=signalingevents_parallel(upone,numsigs_permove,nummoves,radius,b,T);
+            choosestrat(3)=1-probeaten(i);
+        end
+
+        if strategy(i)==1
+            choosestrat(1)=perf(i);
+        else
+            downone=strategy;
+            downone(i)=strategy(i)-1;
+
+            [probeaten, ~]=signalingevents_parallel(downone,numsigs_permove,nummoves,radius,b,T);
+            choosestrat(1)=1-probeaten(i);   
+        end 
+
[…]

### 2014-11-03 greedyopt_v2_2.m

@@ -3,26 +3,28 @@ dellacluster=parcluster('local');
 dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
 dellapool=parpool(dellacluster, numworkers) ;
 
-numsigs_permove=1;
-nummoves=1000;
+numsigs_permove=100;
+nummoves=10;
 
 N=20;
 b=1;
-radius=.5;
+radius=.1;
 T=1;
 
 timesteps=50;
-its=10;
+its=20;
 
 evolution_eaten=zeros(its,N,timesteps);
 groupspeed_eaten=zeros(its,timesteps);
 groupconsensus_eaten=zeros(its,timesteps);
 corrlengths_eaten=zeros(its,timesteps);
+disconnected_eaten=zeros(its,timesteps);
 
 evolution_gettoeat=zeros(its,N,timesteps);
 groupspeed_gettoeat=zeros(its,timesteps);
 groupconsensus_gettoeat=zeros(its,timesteps);
 corrlengths_gettoeat=zeros(its,timesteps);
+disconnected_gettoeat=zeros(its,timesteps);
 
 for num=1:its
 
@@ -32,16 +34,12 @@ if t==1
     strategy=init;
 %     strategy=2*ones(1,N);
 %     strategy(1)=10; 
-    [~, ~, meanlambda, meanH2, meancorrlength]=signalingevents_wgroupprops_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
-    groupspeed_eaten(num,1)=meanlambda;
-    groupconsensus_eaten(num,1)=meanH2;
-    corrlengths_eaten(num,1)=meancorrlength;
     evolution_eaten(num,:,t)=strategy;
 else
     strategy=evolution_eaten(num,:,t-1);
     strategy=reshape(strategy,1,N);
 
-    [probeaten, ~, ~, ~, ~]=signalingevents_wgroupprops_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+    [probeaten, ~]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
     perf=1-probeaten;
     newstrategy=strategy;
 
@@ -55,7 +53,7 @@ else
             upone=strategy;
             upone(i)=strategy(i)+1;
 
[…]

### 2014-11-03 greedyopt_v2_3.m

@@ -1,0 +1,153 @@
+numworkers=str2num(getenv('PROCS')); %#ok<ST2NM>
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
+
+numsigs_permove=10;
+nummoves=100;
+
+N=20;
+b=1;
+radius=.1;
+T=1;
+
+timesteps=50;
+its=20;
+
+evolution_eaten=zeros(its,N,timesteps);
+groupspeed_eaten=zeros(its,timesteps);
+groupconsensus_eaten=zeros(its,timesteps);
+corrlengths_eaten=zeros(its,timesteps);
+disconnected_eaten=zeros(its,timesteps);
+
+evolution_gettoeat=zeros(its,N,timesteps);
+groupspeed_gettoeat=zeros(its,timesteps);
+groupconsensus_gettoeat=zeros(its,timesteps);
+corrlengths_gettoeat=zeros(its,timesteps);
+disconnected_gettoeat=zeros(its,timesteps);
+
+for num=1:its
+
+for t=1:timesteps
+    init=randi([1 N-1],1,N);
+if t==1
+    strategy=init;
+%     strategy=2*ones(1,N);
+%     strategy(1)=10; 
+    evolution_eaten(num,:,t)=strategy;
+else
+    strategy=evolution_eaten(num,:,t-1);
+    strategy=reshape(strategy,1,N);
+
+    [probeaten, ~]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+    perf=1-probeaten;
+    newstrategy=strategy;
+
+    for i=1:N
+        choosestrat=zeros(1,3);
+        choosestrat(2)=perf(i);
+
+        if strategy(i)==N-1
+            choosestrat(3)=perf(i);
+        else 
+            upone=strategy;
+            upone(i)=strategy(i)+1;
+
+            [probeaten, ~]=signalingevents_parallel(upone,numsigs_permove,nummoves,radius,b,T);
+            choosestrat(3)=1-probeaten(i);
+        end
+
+        if strategy(i)==1
+            choosestrat(1)=perf(i);
+        else
+            downone=strategy;
+            downone(i)=strategy(i)-1;
+
[…]

### 2014-11-03 greedyopt_v2_4.m

@@ -1,0 +1,153 @@
+numworkers=str2num(getenv('PROCS')); %#ok<ST2NM>
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
+
+numsigs_permove=1000;
+nummoves=1;
+
+N=20;
+b=1;
+radius=.1;
+T=1;
+
+timesteps=50;
+its=20;
+
+evolution_eaten=zeros(its,N,timesteps);
+groupspeed_eaten=zeros(its,timesteps);
+groupconsensus_eaten=zeros(its,timesteps);
+corrlengths_eaten=zeros(its,timesteps);
+disconnected_eaten=zeros(its,timesteps);
+
+evolution_gettoeat=zeros(its,N,timesteps);
+groupspeed_gettoeat=zeros(its,timesteps);
+groupconsensus_gettoeat=zeros(its,timesteps);
+corrlengths_gettoeat=zeros(its,timesteps);
+disconnected_gettoeat=zeros(its,timesteps);
+
+for num=1:its
+
+for t=1:timesteps
+    init=randi([1 N-1],1,N);
+if t==1
+    strategy=init;
+%     strategy=2*ones(1,N);
+%     strategy(1)=10; 
+    evolution_eaten(num,:,t)=strategy;
+else
+    strategy=evolution_eaten(num,:,t-1);
+    strategy=reshape(strategy,1,N);
+
+    [probeaten, ~]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+    perf=1-probeaten;
+    newstrategy=strategy;
+
+    for i=1:N
+        choosestrat=zeros(1,3);
+        choosestrat(2)=perf(i);
+
+        if strategy(i)==N-1
+            choosestrat(3)=perf(i);
+        else 
+            upone=strategy;
+            upone(i)=strategy(i)+1;
+
+            [probeaten, ~]=signalingevents_parallel(upone,numsigs_permove,nummoves,radius,b,T);
+            choosestrat(3)=1-probeaten(i);
+        end
+
+        if strategy(i)==1
+            choosestrat(1)=perf(i);
+        else
+            downone=strategy;
+            downone(i)=strategy(i)-1;
+
[…]

### 2014-11-03 signalingevents_wgroupprops.m

@@ -1,13 +1,12 @@
-function [probeaten, probgettoeat, meanlambda, meanH2, meancorrlength, disconnectedcount]=signalingevents_wgroupprops(strategy,numsigs_permove,nummoves,radius,b,T)
+function [probeaten, probgettoeat, meanlambda, meanH2, meancorrlength, disconnectedcount, numreceivers]=signalingevents_wgroupprops(strategy,numsigs_permove,nummoves,radius,b,T)
 numsigs_tot=numsigs_permove*nummoves;
 N=max(size(strategy));
-index=1:N;
 
 scores=zeros(N,numsigs_permove,nummoves);
 lambdavals=zeros(1,nummoves);
 H2norms=zeros(1,nummoves);
 corrlengths=zeros(numsigs_permove,nummoves);
-
+keepreceivers=zeros(N,numsigs_permove,nummoves);
 disconnectedcount=0;
 
 % parfor pari=1:nummoves
@@ -29,26 +28,18 @@ for pari=1:nummoves
     Mbin(~~Mbin)=1;
     g=sparse(Mbin);
     [~,Ctotal]=graphconncomp(g,'Directed','false');
-%     paths=graphallshortestpaths(g);
 
-%     groups=netgroups(M);
-%     ng=size(groups,1);
-%     groupvec=zeros(N,1);
-%     for i=1:ng
-%         for j=1:length(groups{i})
-%             groupvec(Ctotal==groups{i}(j))=i;
-%         end
-%     end
+    L=lap(M);
+    [~,vals]=eig(L);
+    vals=diag(vals);
+    f=find(sigfig(vals,14)==0);
 
-    if max(Ctotal)==1
-        L=lap(M);
-        [~,vals]=eig(L);
-        vals=diag(vals);
-        vals=vals(2:length(vals));
+    if max(Ctotal)==1 && size(f,1)==1           
+        vals=vals(sigfig(vals,14)~=0);
         lambda=min(real(vals));
         lambdavals(pari)=lambda;
         h=H2norm(M,'additive');
-        H2norms(pari)=h;
+        H2norms(pari)=h; 
     else 
         disconnectedcount=disconnectedcount+1;
         lambdavals(pari)=NaN;
@@ -64,17 +55,7 @@ for pari=1:nummoves
         v=real(expected_spin(M,T,beta));
 
         scores(:,j,pari)=v;
-        
-%         Mgroup=M(groupvec==groupvec(receiver),groupvec==groupvec(receiver));
[…]

### 2014-11-19 paper_skeleton.tex
i just did a lot of work to make figures to show the ess strategies and why they look like they do, and now im trying to write that up

@@
 \begin{center}
-{\bf \LARGE{Information Gathering Strategies}}
+{\bf \LARGE{Evolution of Information Gathering Strategies}}
 \vspace{10pt}
@@
 
-\section{Introduction}
+\section{Abstract}
 Birds in flocks, like animals in many types of social groups, can use their peers to learn about the environment and update their opinions about where to move in that environment.
@@
 the number of neighbors it learns from, to achieve this outcome.
-Individual birds should optimize how well they learn about the environment and should minimize the cognitive and time costs of learning.
-I find the circumstances in which this individual level optimization can produce the observed group level optimization.
+Individual birds should optimize how well they learn about the environment.
+There are (at least) two features of the environment that are important to an individual bird's fitness---the location of a predator and the location of resources---and it can try to learn about these from its peers.
+In this work, we identify the optimal strategies for birds under selection pressure dictated by these types of information and explore the relationship between the robustness measure and the correlation length of the flock, a measure commonly used to describe how close a flock's structure is to criticality.
 
 \section{Model }
-\subsection{Learning dynamics}
-In this model, an individual's opinion changes with a probability that depends on its social interactions and the strength of the environmental stimulus.
+\subsection{Learning dynamics }
+An individual's opinion changes with a probability that depends on its social interactions and the strength of the environmental stimulus.
 Specifically, an individual's opinion $\sigma_i$ can be either $1$ or $-1$ and $\sigma_i$ switches with probability $w_i(\sigma_i)=\frac{1}{2}(-\sigma_i\sum_jA_{ij}\sigma_j )(1-\beta_i\sigma_i)$ where $A_{ij}$ indicates the strength of the connection from $j$ to $i$ and $\beta_i$ indicates the strength of the environmental stimulus perceived by $i$.
@@
 We simplify this by writing $\frac{d\vec{v}}{dt}=A\vec{v}+\vec{\beta}^T\vec{1}$, where $A=M-B$ and $B$ is a diagonal matrix with the values of $\vec{\beta}$ along the diagonal.
+In the absence of the environmental signal, there is no trend in the dynamics.
+In the presence of the environmental signal, however, those individuals who perceive the signal will be more likely to switch to having an opinion of $1$ so their expected value will increase, and consequently so will the values of the those paying attention to the informed individuals.
+Thus, having a higher expected value can be interpreted as having more information about the information and having a lower expected value can be interpreted as having less information.
 
-These stochastic dynamics will lead to an equilibrium probability distribution over sets of opinions that agrees with the Boltzmann equilibrium distribution for a system with energy or Hamiltonian given by
-\begin{equation*}
-E(\vec{v})=-\vec{v}^TA\vec{v}-\vec{\beta}^T\vec{v}
-\end{equation*}
+\subsection{Individual-level fitness }
+To find the optimal strategies from the individuals' perspectives, we need a measure of individual-level fitness, which we define as follows.
+Each individual has a strategy that determines how many other individuals he pays attention to, $n_i=\sum_{j\neq i}{\bf I}(A_{ij}\neq 0)$.
+We distribute the individuals randomly in space.
+Each individual then pays attention to its $n_i$ nearest neighbors.
+One individual is chosen to be a receiver; it and all those individuals within a radius $r$ of the receiver perceive an external signal of strength $\beta$.
+We find the expected opinion of each individual $q_i$ after a period of time $T$.
+We refer to each such set of random positions and receiver as a signaling event.
 
-\subsection{Individual-level performance }
-The group is made up of $N$ individuals.
-Each individual has a strategy that determines how many other individuals he pays attention to, $n_i=\sum_{j\neq i}{\bf I}(A_{ij}\neq 0)$.
-For a given number of iterations, we distribute the individuals randomly in space.
-Each individual then pays attention to its $n_i$ nearest neighbors.
-This is meant to capture One individual is chosen to be a receiver; it and all those individuals within a radius $r$ of the receiver perceive an external signal of strength $\beta$.
-We find the expected opinion of each individual $q_i$ after a period of time $T$.
-We refer to each of these random iterations as a signaling event.
-
-In the absence of an external signal, neither opinion is correct.
-When an external signal is presented, some individuals will be able to perceive it ($\beta_i>0$) and some individuals will not ($\beta_i=0$).
-In this case, having an opinion that matches the external signal ($\sigma_i=1$) is beneficial.
-How beneficial it is depends on the kind of signal.
 We consider two types of signals: predators and food.
-If a predator is present, we assume that whichever individual is least aware of the signal will be predated.
-We therefore find, over many signaling events, the probability that an individual will be eaten and an individual's fitness is given by the probability of surviving.
-If there are resources available, we assume that fitness is proportional to the probability of being the first to reach the resources.
+If the signal is information that a predator is present, we assume that whichever individual is least aware of the signal will be predated, i.e.
+the individual $i$ such that $q_i$ is lowest.
+We therefore find, over many signaling events, the probability that an individual will be eaten and an individual's fitness is given by the probability of surviving, i.e.
+$1$ minus the probability of being predated.
+If the signal is information that resources are available, we assume that whichever individual is most aware of the signal will get access to the resource, i.e.
+the individual $i$ such that $q_i$ is highest.
+In this case, fitness is proportional to the probability of being the first to reach the resources.
 
@@
 We consider two group-level properties that indicate how cohesive the group is.
-The first property is $\mathscr{H}_2$ robustness.
-This is a measure of the robustness of the consensus state in which all birds have the same opinion to noise at equilibrium (see Appendix Sec.
+The first property is $\mathscr{H}_2$ robustness, as used in \cite{Young:2010fk,Young:2013kx}.
+This is a measure of the robustness of the consensus state in which all birds have the same opinion to noise at equilibrium, as used in (see Appendix Sec.
 \ref{H2}).
@@
 First, using the framework of adaptive dynamics in finite populations, we identify the evolutionarily stable strategy.
-This framework assumes that there is a homogeneous population into which an mutant individual tries to invade.
+This framework assumes that there is a homogeneous population into which a mutant individual tries to invade.
 If strategies change on a learning rather than an evolutionary timescale, or if there is just a lot of variation in the population, we might expect individuals to try to optimize their strategies in the context of a heterogeneous population.
 This is our second method of optimization.
+Given a random set of initial strategies over the group, we allow each bird to choose the strategy that would be best given the rest of the strategies being used and repeat this process until the birds reach an equilibrium set of strategies.
+We repeat this over many initial sets of strategies to find average properties of this optimization process.
 
@@
 
-\subsection{ESS }
-\subsection{Learned optima }
+\subsection{ESS strategies }
+Surprisingly, even without imposing costs on paying attention to more neighbors, paying attention to the whole flock is not usually the best strategy.
+The ESS strategy of the number of neighbors to pay attention strongly depends on which selection pressure is being applied.
+If a focal bird pays attention to many others, its opinion will be similar to theirs.
+This will prevent it from being the most informed and hence the first to locate resources.
+Thus, the uninvadable strategy is to pay attention to few neighbors when selection is due to knowledge about resources.
+On the other hand, having many neighbors will prevent the focal bird from being the least informed and thus the most likely to be eaten.
+Thus, the uninvadable strategy is to pay attention to many neighbors when selection is due to knowledge about predators.
+Figure \ref{ESS} shows the ESS strategy when selection is due to knowledge about resources is always lower than the ESS strategy when selection is due to knowledge about predators (Figure \ref{ESS}).
+
+The ESS strategies when selection is due to knowledge about predators show two interesting features.
+First, there are many parameters for which there are multiple ESS strategies.
+It might seem that if a strategy outperforms all others when it is the resident strategy, it should also outperform other strategies as an invader.
+However, when there are multiple ESS strategies, if we take any two of those ESS strategies, each is uninvadable by the other.
+This indicates how important the social environment set by the resident strategy is.
+The resident strategy essentially sets a level of background noise.
+As an example, consider the strategies of having sixteen or seventeen neighbors when the radius of the signal is $r=0.1$.
+Figure \ref{ESS} shows how the probability of being predated for each of these invasion scenarios depends on the number of birds in the group and the number of informed neighbors the invader has.
+A bird with seventeen neighbors will have more uninformed neighbors than one with sixteen neighbors and thus will have less information itself.
+This poor information is  costly when the resident strategy is to have sixteen neighbors since the residents will have less noisy information, which prevents an mutant with seventeen neighbors to invade.
+Figure \ref{ESS} shows that by paying attention to more uninformed neighbors this mutant is likely to be eaten when there are many truly informed birds.
+However, having less noisy information does not allow a sixteen neighbor mutant to invade a seventeen neighbor group since, regardless of the number of uninformed neighbors the invader has, the background level of noise is much higher in a group all of whom have more uninformed neighbors.
+The sixteen neighbor mutant also pays a cost by being less likely to pay attention to the right neighbors when there are very few truly informed birds.
+Figure \ref{ESS} shows that by being less likely to pay attention to the right neighbors this mutant is likely to be eaten when there are few truly informed birds.
+
+These opposing forces caused by the costs of paying attention to more uninformed neighbors and the costs of not paying attention to informed neighbors also explains the second feature of the ESS strategies: they are non-monotonic with respect to the the radius of the signal, i.e.
+how public the signal is (Figure \ref{ESS}).
+
+%The ESS strategies when selection is due to knowledge about predators show two interesting features.
+First, there are many parameters for which there are multiple ESS strategies.
+For example, when the radius of the signal is $r=0.1$, the strategies of having fourteen, fifteen, sixteen, or seventeen neighbors are all ESS.
+This means that if the group is made up entirely of individuals having sixteen neighbors, then a mutant with any other strategy, including having seventeen neighbors, cannot invade.
+But if we instead consider a group made up entirely of individuals having seventeen neighbors, then a mutant with sixteen cannot invade either.
+This is because the success of any individual, resident or mutant, is strongly dependent on the environment set by the resident strategy.
+In general, if a focal bird has several informed neighbors (and does not have the signal itself), then its uninformed neighbors decrease the focal bird's performance and make it more likely to be eaten.
+Thus, for a given number of informed neighbors, having more neighbors rather than fewer is a worse strategy.
+On the other hand, it a focal bird has no informed neighbors, then averaging the naive opinions of more neighbors rather than fewer makes it more likely for it pick up some information about the signal and avoid being eaten.
+In Figure \ref{ESS} we show that a bird with seventeen neighbors is more likely to be eaten when it has informed neighbors than a bird with sixteen neighbors but less likely to be eaten when it has no informed neighbors.
+The magnitudes of these opposing effects---the cost of more naive neighbors when a bird has informed neighbors and the benefit of more naive neighbors when a bird has no informed neighbors---depend on how the rest of the group is gathering information.
+When the resident strategy is lower, e.g.
+sixteen, the noise picked up by a mutant with seventeen neighbors at an intermediate number of informed neighbors makes it very likely to be eaten in those situations.
+Conversely, when the resident strategy is higher, e.g.
+seventeen, the inability of a mutant with sixteen neighbors to pick up any information without informed neighbors makes it very likely to be eaten in those situations.
+For each radius of $r$ of the signal, these opposing forces equilibrate at different ESS strategies.
+%
+%The second feature of the ESS strategies is that they are non-monotonic with respect to the the radius of the signal, i.e.
+how public the signal is (Figure \ref{ESS}).
+This non-monotonic behavior is due to two opposing effects of paying attention to more neighbors.
+First, having more neighbors increases the probability that one of them will have the signal and thus the probability that the focal bird will have a good source of information.
+Second, having more neighbors increases the probability that a focal bird will be trying to learn from individuals who do not have true information about the environment.
+When few individuals have the true signal (i.e.
+low radius $r$), the second effect outweighs the first and the ESS strategies are to pay attention to fewer birds than are in the whole flock.
+However, as more individuals have the true signal (i.e.
+higher radius $r$), the first effect is more important.
+
+\subsection{Learned strategies }
+The equilibrium strategies when the individuals are allowed to learn are similar to the ESS strategies: learned strategies when selection is due to predation are always higher than learned strategies when selection is due to resources (Figure \ref{greedyopt}).
+Surprisingly, even when the group starts with heterogeneous strategies, they often reach a homogeneous equilibrium.
+Under selection due to predation, the $\mathscr{H}_2$ robustness and correlation length of the flock tend to decrease, whereas under selection due to resources, $\mathscr{H}_2$ robustness is relatively constant and correlation length tends to increase (Figure \ref{greedyopt}).
+
 \subsection{Relationship between group properties }
+Figure \ref{greedyopt} shows that there is a relationship between two group properties: the $\mathscr{H}_2$ norm and the correlation length of the flock.
 
 \section{Discussion}
+explore vs. exploit 
+risk aversion vs. risk tolerance
 
@@
 \begin{figure}[ht]
-\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/ESSneighbors.pdf}
+\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/ESSfigure.pdf}
 \caption{\label{ESS} The ESS number of neighbors is always higher if selection is due to predation than due to resources.
-The ESS number of neighbors when selection is due to predation is a non-monotonic function of how public the signal is.
-The x-axis indicates the radius of the signal and the y-axis indicates the ESS number of neighbors.
+When selection is due to predation, there can be multiple ESS strategies, which are a non-monotonic function of the radius of the signal.
+In (a), we show the (possibly multiple) ESS strategies as a function of the radius of the signal.
+In (b) and (c), each heat map shows the probability that an invader is eaten as a function of the number of informed birds in the flock on the horizontal axis and the number of informed neighbors the invader has on the vertical axis.
+In (b), the resident strategy is to have $16$ neighbors and the invader strategy is to have $17$ neighbors.
+In (c), the resident strategy is to have $17$ neighbors and the invader strategy is to have $16$ neighbors.
+Parameters:  in all panels $\beta=1$, $T=1$, in (b) and (c) $r=0.1$.
 }
@@
 The third column shows, for many initial conditions, how the correlation length changes over time.
+Parameters:  $\beta=1$, $r=.1$, $T=1$.
 }
@@
 
+\newpage
 \section{Appendix}

### 2014-12-15 paper_skeleton.tex
im about to try to write about the relationship between correlation length and ESS strategies

@@
 how public the signal is (Figure \ref{ESS}).
-
-%The ESS strategies when selection is due to knowledge about predators show two interesting features.
-First, there are many parameters for which there are multiple ESS strategies.
-For example, when the radius of the signal is $r=0.1$, the strategies of having fourteen, fifteen, sixteen, or seventeen neighbors are all ESS.
-This means that if the group is made up entirely of individuals having sixteen neighbors, then a mutant with any other strategy, including having seventeen neighbors, cannot invade.
-But if we instead consider a group made up entirely of individuals having seventeen neighbors, then a mutant with sixteen cannot invade either.
-This is because the success of any individual, resident or mutant, is strongly dependent on the environment set by the resident strategy.
-In general, if a focal bird has several informed neighbors (and does not have the signal itself), then its uninformed neighbors decrease the focal bird's performance and make it more likely to be eaten.
-Thus, for a given number of informed neighbors, having more neighbors rather than fewer is a worse strategy.
-On the other hand, it a focal bird has no informed neighbors, then averaging the naive opinions of more neighbors rather than fewer makes it more likely for it pick up some information about the signal and avoid being eaten.
-In Figure \ref{ESS} we show that a bird with seventeen neighbors is more likely to be eaten when it has informed neighbors than a bird with sixteen neighbors but less likely to be eaten when it has no informed neighbors.
-The magnitudes of these opposing effects---the cost of more naive neighbors when a bird has informed neighbors and the benefit of more naive neighbors when a bird has no informed neighbors---depend on how the rest of the group is gathering information.
-When the resident strategy is lower, e.g.
-sixteen, the noise picked up by a mutant with seventeen neighbors at an intermediate number of informed neighbors makes it very likely to be eaten in those situations.
-Conversely, when the resident strategy is higher, e.g.
-seventeen, the inability of a mutant with sixteen neighbors to pick up any information without informed neighbors makes it very likely to be eaten in those situations.
-For each radius of $r$ of the signal, these opposing forces equilibrate at different ESS strategies.
-%
-%The second feature of the ESS strategies is that they are non-monotonic with respect to the the radius of the signal, i.e.
-how public the signal is (Figure \ref{ESS}).
-This non-monotonic behavior is due to two opposing effects of paying attention to more neighbors.
-First, having more neighbors increases the probability that one of them will have the signal and thus the probability that the focal bird will have a good source of information.
-Second, having more neighbors increases the probability that a focal bird will be trying to learn from individuals who do not have true information about the environment.
-When few individuals have the true signal (i.e.
-low radius $r$), the second effect outweighs the first and the ESS strategies are to pay attention to fewer birds than are in the whole flock.
-However, as more individuals have the true signal (i.e.
-higher radius $r$), the first effect is more important.
+Figure \ref{ESS} shows that there are essentially three categories of radii: When the radius is low, the signal is private, with only a few birds having true information.
+When the radius is intermediate, the signal is public, with many birds having true information.
+When the radius is high, the signal is universal, with essentially all birds in the flock having true information.
+When the signal is private, paying attention to too few birds is costly because with more neighbors it is more likely to have a truly informed neighbor, but paying attention to too many uninformed birds can also be costly.
+This leads to a high but not maximal ESS strategy.
+When the signal is public, the costs of paying attention to uninformed birds are reduced and the ESS strategy increases to become as high as possible.
+When the signal is universal, not paying attention to informed birds does not impose a cost.
+Additionally, in a group in which the residents pay attention to all of their neighbors, any mutant strategy can get just as much information by paying attention to fewer neighbors so the maximal strategy is no longer an ESS (Figure \ref{ESS}).
 
@@
 
+expect single ESS strategies but social information depends on the social environment set by the resident strategy
+expect ESS strategies to decrease as the signal becomes more public, but we find the opposite
+
 \section{Figures }
@@
 In (a), we show the (possibly multiple) ESS strategies as a function of the radius of the signal.
-In (b) and (c), each heat map shows the probability that an invader is eaten as a function of the number of informed birds in the flock on the horizontal axis and the number of informed neighbors the invader has on the vertical axis.
-In (b), the resident strategy is to have $16$ neighbors and the invader strategy is to have $17$ neighbors.
-In (c), the resident strategy is to have $17$ neighbors and the invader strategy is to have $16$ neighbors.
-Parameters:  in all panels $\beta=1$, $T=1$, in (b) and (c) $r=0.1$.
+In (b) we show the probability of there being a given number of informed birds for signals with different radii.
+In (c) and (d), each heat map shows the probability that an invader is eaten as a function of the number of informed birds in the flock on the horizontal axis and the number of informed neighbors the invader has on the vertical axis.
+In (c), the resident strategy is to have $16$ neighbors and the invader strategy is to have $17$ neighbors.
+In (d), the resident strategy is to have $17$ neighbors and the invader strategy is to have $16$ neighbors.
+Parameters:  in all panels $\beta=1$, $T=1$, in (c) and (d) $r=0.1$.
 }

### 2014-12-15 groupprops.m

@@ -50,7 +50,8 @@ for pari=1:nummoves
         receiver=receivers(j);
         
         if max(Ctotal)==1
-            [~,~,l,~]=correlationlength_mat_single_v3(M,d,b,radius,receiver);
+%             [~,~,l,~]=correlationlength_mat_single_v3(M,d,b,radius,receiver);
+            [~,~,l] = correlationlength_mat_single(M,d,b,radius,receiver);
             corrlengths(j,pari)=l;
         else 
             corrlengths(j,pari)=NaN;

### 2014-12-15 info.py

@@ -1,121 +1,0 @@
-from __future__ import division
-import numpy as np
-import scipy
-import scipy.spatial
-import math, random
-
-def dynamics(x,M):
-	n = len(x)
-	new = list()
-	for i in range(n):
-		neighbors = 0
-		summed_diffs = 0
-		for j in range(n):
-			if M[j,i]>0:
-				neighbors += 1
-				summed_diffs += M[j,i]*(x[j]-x[i])
-		to_add = x[i] + summed_diffs / neighbors
-		new.append(to_add)
-	return new
-	
-def lap(M):
-	n = M.shape[1]
-	m = np.zeros_like(M)
-	m = m.astype(np.float)
-	for i in range(n):
-		neighbors = 0
-		for j in range(n):
-			if M[j,i]>0:
-				neighbors +=1
-		for j in range(n):
-			if M[j,i]>0:
-				m[j,i]=-1/neighbors
-		m[i,i]=1
-	return m
-	
-def H2(M):
-	vals,vec = np.linalg.eig(lap(M))
-	vals = vals.real
-	vals = sorted(vals)[1:]
-	vals = np.array(vals)
-	h = math.sqrt(.5*sum(1/vals))
-	return h
-	
-def nextgen(strategy,fitness):
-	l = len(fitness)//2	
-	to_return = list(strategy)
-	temp = np.array(fitness).argsort()
-	ranks = temp.argsort()
-	ranks = list(ranks)
-	w = 0
-	for j in range(len(fitness)):
-		if ranks[j]>=(len(fitness)-l):
-			to_return[j]=strategy[ranks.index(w)]
-			w += 1
-	return to_return
-
-def dist(N):
-	positions = np.empty(shape=(N,2))
-	for i in range(N):
-		positions[i,0]=random.uniform(0,1)
-		positions[i,1]=random.uniform(0,1)
-	D = scipy.spatial.distance.cdist(positions,positions)
-	return D
-
-N = 20
-its = 10
-D = dist(N)
-runs = 500
-
-dists = np.empty(shape=(its,N,N))
-strategy = np.empty(shape=(its+1,N))
-for j in range(N):
-	strategy[0,j] = random.randint(1,N-1)
-
-H2vals = np.empty(shape=its)
-
-for k in range(its):
-	strategy2 = strategy[k,:]
-	
-	#construct communication matrix
-	M = np.zeros(shape=(N,N))
[…]

### 2014-12-15 test3.m

@@ -1,6 +1,6 @@
 close all
 N=20;
-strategy=10*ones(N,1);
+strategy=2*ones(N,1);
 b=1;
 T=1;
 radius=.1;

### 2015-01-23 EBrush_notes_flocking_evolution.tex
im about to go through correlations and H2 more carefully

--- 2014-11-03 paper_skeleton.tex
@@
 
-\subsection{Individual-level performance }
-The group is made up of $N$ individuals.
+\subsection{Individual-level fitness }
+To find the optimal strategies from the individuals' perspectives, we need a measure of individual-level fitness, which we define as follows.
 Each individual has a strategy that determines how many other individuals he pays attention to, $n_i=\sum_{j\neq i}{\bf I}(A_{ij}\neq 0)$.
-For a given number of iterations, we distribute the individuals randomly in space.
+We distribute the individuals randomly in space.
 Each individual then pays attention to its $n_i$ nearest neighbors.
-This is meant to capture One individual is chosen to be a receiver; it and all those individuals within a radius $r$ of the receiver perceive an external signal of strength $\beta$.
+One individual is chosen to be a receiver; it and all those individuals within a radius $r$ of the receiver perceive an external signal of strength $\beta$.
 We find the expected opinion of each individual $q_i$ after a period of time $T$.
-We refer to each of these random iterations as a signaling event.
+We refer to each such set of random positions and receiver as a signaling event.
 
-In the absence of an external signal, neither opinion is correct.
-When an external signal is presented, some individuals will be able to perceive it ($\beta_i>0$) and some individuals will not ($\beta_i=0$).
-In this case, having an opinion that matches the external signal ($\sigma_i=1$) is beneficial.
-How beneficial it is depends on the kind of signal.
 We consider two types of signals: predators and food.
-If a predator is present, we assume that whichever individual is least aware of the signal will be predated.
-We therefore find, over many signaling events, the probability that an individual will be eaten and an individual's fitness is given by the probability of surviving.
-If there are resources available, we assume that fitness is proportional to the probability of being the first to reach the resources.
+If the signal is information that a predator is present, we assume that whichever individual is least aware of the signal will be predated, i.e.
+the individual $i$ such that $q_i$ is lowest.
+We therefore find, over many signaling events, the probability that an individual will be eaten and an individual's fitness is given by the probability of surviving, i.e.
+$1$ minus the probability of being predated.
+If the signal is information that resources are available, we assume that whichever individual is most aware of the signal will get access to the resource, i.e.
+the individual $i$ such that $q_i$ is highest.
+In this case, fitness is proportional to the probability of being the first to reach the resources.
 
@@
 
-\subsection{ESS }
-\subsection{Learned optima }
+\subsection{ESS strategies }
+Surprisingly, even without imposing costs on paying attention to more neighbors, having as many neighbors as possible is usually sub-optimal.
+The ESS strategies for the number of neighbors to pay attention strongly depends on which selection pressure is being applied.
+The ESS strategy when selection is due to knowledge of a predator is always higher than the ESS strategy when selection is due to knowledge about resources (Figure \ref{ESS}).
+When selection is due to predation, the ESS strategy is a non-monotonic function of the the radius of the signal, i.e.
+how public the signal is (Figure \ref{ESS}).
+This non-monotonic behavior is due to two opposing effects of paying attention to more neighbors.
+First, having more neighbors increases the probability that one of them will have the signal and thus the probability that the focal bird will have a good source of information.
+Second, having more neighbors increases the probability that a focal bird will be trying to learn from individuals who do not have true information about the environment.
+When few individuals have the true signal (i.e.
+low radius $r$), the second effect outweighs the first and the ESS strategies are to pay attention to fewer birds than are in the whole flock.
+However, as more individuals have the true signal (i.e.
+higher radius $r$), the first effect is more important
+
+\subsection{Learned strategies }
+The equilibrium strategies when the individuals are allowed to learn are similar to the ESS strategies: learned strategies when selection is due to predation are always higher than learned strategies when selection is due to resources (Figure \ref{greedyopt}).
+Surprisingly, even when the group starts with heterogeneous strategies, they often reach a homogeneous equilibrium.
+Under selection due to predation, the $\mathscr{H}_2$ robustness and correlation length of the flock tend to decrease, whereas under selection due to resources, $\mathscr{H}_2$ robustness is relatively constant and correlation length tends to increase (Figure \ref{greedyopt}).
+
 \subsection{Relationship between group properties }
+Figure \ref{greedyopt} shows that there is a relationship between two group properties: the $\mathscr{H}_2$ norm and the correlation length of the flock.
 
@@
 The x-axis indicates the radius of the signal and the y-axis indicates the ESS number of neighbors.
+Parameters:  $\beta=1$, $T=1$.
 }

### 2015-01-23 ESS_analysis.m

@@ -13,6 +13,8 @@ ESSeaten_total{2}=ESSseaten;
 ESSgettoeat_total{2}=ESSsgettoeat;
 
 load /Users/eleanorbrush/Desktop/greedyopt_rad=0.1_T=1_nummoves=10_numpermove=100.mat
+
+load /Users/eleanorbrush/Desktop/homogengroupprops.mat
 %% lots of fitness and rho heatmaps
 
 myseqmap=cbrewer('seq', 'YlOrRd',9);
@@ -226,3 +228,52 @@ set(gcf,'PaperPosition',[0 0 w h]);
 filename=strcat('/Users/eleanorbrush/Desktop/','greedyoptneighbors','.pdf');
 % print(filename,'-dpdf','-r300');
 
+
+%%
+figure
+w=6.83;
+h=.25*w;
+set(gcf,'Units','inches');
+set(gcf,'Position',[11.5 3 w h]);
+subplot(1,2,1)
+toplot=corrlengths;
+% toplot=corrlengths./repmat([1 radvals(2:end)],Lh,1);
+a=min(min(toplot))-1;
+A=max(max(toplot))-1;
+A=max(abs(a),abs(A))+1;
+mycolors=cbrewer('div','RdBu',20);
+imagesc(radvals,homogenstrats,toplot)
+% caxis manual
+% caxis([-A A])
+% colormap(mycolors)
+colorbar
+title('Correlation Length')
+xlabel('Radius of signal')
+ylabel('Strategy')
+
+subplot(1,2,2)
+hold on
+t=3;
+n=size(ESSseaten,2);
+for i=1:n
+    l=length(ESSseaten{t,i});
+    plot(corrlengths(5,i)*ones(l,1),ESSseaten{t,i},'o')
+end
+xlabel('Correlation length of highly connected group')
+ylabel('ESS strategy')
+
+set(gcf,'PaperSize',[w h]);
+set(gcf,'PaperPosition',[0 0 w h]);
+
+filename='/Users/eleanorbrush/Desktop/corrlength_exploration.pdf';
+% print(filename,'-dpdf','-r300')
+
+%%
+seqcols=colormap(jet(Nr));
+figure
+hold on
+
+for i=1:Nr
+    plot(ones(Lh,1)./(groupconsensus(:,i)./col(homogenstrats)),corrlengths(:,i),'-o','Color',seqcols(i,:),'LineWidth',2)
+end
+    

### 2015-01-23 ESS_figures.m

@@ -1,0 +1,314 @@
+N=20;
+radius=radvals(3);
+T=Tvals(3);
+nummoves=1000;
+numsigs_permove=1;
+b=1;
+perm=randsample(20,20,'false');
+
+strats=[16 19];
+invaderstrats=[17 16];
+
+maxrec=9;
+invader_scores=cell(maxrec+1,2);
+invader_counts=zeros(maxrec+1,2);
+invader_minscorer=cell(maxrec+1,2);
+resident_scores=cell(maxrec+1,2);
+resident_counts=zeros(maxrec+1,2);
+resident_minscorer=cell(maxrec+1,2);
+ninv=zeros(3,2);
+nres=zeros(3,2);
+
+for k=1:2
+    strat=strats(k);
+    invader=10;
+    invaderstrat=invaderstrats(k);
+    strategy=[strat*ones(1,N)];
+    strategy(invader)=invaderstrat;
+
+    numsigs_tot=numsigs_permove*nummoves;
+    N=max(size(strategy));
+    indvec=1:N;
+
+    scores=zeros(N,numsigs_permove,nummoves);
+    keepreceivers=zeros(N,numsigs_permove,nummoves);
+    keepM=zeros(N,N,numsigs_permove,nummoves);
+    eigcen=zeros(N,numsigs_permove,nummoves);
+    eigval=zeros(numsigs_permove,nummoves);
+
+    % parfor i=1:nummoves
+    for i=1:nummoves
+        positions=unifrnd(0,1,N,2);
+        d=squareform(pdist(positions));
+
+        M=zeros(N);
+        for ind=1:N
+            [~, order]=sort(d(ind,:));
+            neighbors=order(2:strategy(ind)+1);
+            M(ind,neighbors)=1/strategy(ind);
+        end
+        M(1:N+1:end)=-1; %sets diagonal equal to -1
+    %     M=M(perm,perm);
+        receivers=randsample(N,numsigs_permove,'true');
+        for j=1:numsigs_permove
+            beta=zeros(N,1);
+            receiver=receivers(j);
+            allreceivers=d(receiver,:)<=radius;
+            beta(allreceivers)=b;
+            v=real(expected_spin(M,T,beta));
+
+            scores(:,j,i)=v;
[…]

### 2015-01-23 ESS_figures_v4.m

@@ -1,0 +1,263 @@
+N = 20;
+radius = .1;
+T = 1;
+nummoves = 1000;
+numsigs_permove = 1;
+b = 1;
+perm = randsample(20,20,'false');
+
+strats = [16 17];
+% invaderstrats = [17 16];
+invaderstrats = fliplr(strats);
+
+maxrec = zeros(1,2);
+invader_counts = zeros(N,N,2);
+invader_minscores = cell(N,N,2);
+resident_counts = zeros(N,N,2);
+resident_minscores = cell(N,N,2);
+
+probeaten = zeros(N,2);
+
+
+for k = 1:2
+    strat = strats(k);
+    invader = 10;
+    invaderstrat = invaderstrats(k);
+    strategy = [strat*ones(1,N)];
+    strategy(invader) = invaderstrat;
+
+    numsigs_tot = numsigs_permove*nummoves;
+
+    scores = zeros(N,numsigs_permove,nummoves);
+    keepM = zeros(N,N,numsigs_permove,nummoves);
+    keepreceivers = zeros(N,numsigs_permove,nummoves);
+
+    % parfor i=1:nummoves
+    for i = 1:nummoves
+        positions = unifrnd(0,1,N,2);
+        d = squareform(pdist(positions));
+
+        M = zeros(N);
+        for ind = 1:N
+            [~, order] = sort(d(ind,:));
+            neighbors = order(2:strategy(ind)+1);
+            M(ind,neighbors) = 1/strategy(ind);
+        end
+        M(1:N+1:end) = -1; %sets diagonal equal to -1
+    %     M = M(perm,perm);
+        receivers = randsample(N,numsigs_permove,'true');
+        for j = 1:numsigs_permove
+            beta = zeros(N,1);
+            receiver = receivers(j);
+            allreceivers = d(receiver,:)<= radius;
+            beta(allreceivers) = b;
+            v = real(expected_spin(M,T,beta));
+
+            scores(:,j,i) = v;
+            keepM(:,:,j,i)= M;
+            keepreceivers(:,j,i) = (allreceivers);
+
+        end
+
+    end
+    
+    keepM=reshape(keepM,20,20,nummoves);
[…]

### 2015-01-23 ESS_figures_v5.m

@@ -1,0 +1,492 @@
+N = 20;
+nummoves = 1000;
+numsigs_permove = 1;
+b = 1;
+
+fontname='Times';
+textfontsz=12;
+labfontsz=9;
+%% load ESS data
+
+ESSeaten_total=cell(2,1);
+ESSgettoeat_total=cell(2,1);
+
+load /Users/eleanorbrush/Desktop/ESS_nummoves=1000_numpermove=1_fullstorage.mat
+
+ESSeaten_total{1}=ESSseaten;
+ESSgettoeat_total{1}=ESSsgettoeat;
+
+load /Users/eleanorbrush/Desktop/ESS_nummoves=10_numpermove=100.mat
+
+ESSeaten_total{2}=ESSseaten;
+ESSgettoeat_total{2}=ESSsgettoeat;
+
+load /Users/eleanorbrush/Desktop/greedyopt_rad=0.1_T=1_nummoves=10_numpermove=100.mat
+
+%% find prob eaten for number of informed birds and number of informed neighbors
+radius = .1;
+T = 1;
+strats = [16 17];
+invaderstrats = fliplr(strats);
+
+maxrec = zeros(1,2);
+invader_counts = zeros(N,N,2);
+invader_minscores = cell(N,N,2);
+resident_counts = zeros(N,N,2);
+resident_minscores = cell(N,N,2);
+
+probeaten = zeros(N,2);
+
+for k = 1:2
+    strat = strats(k);
+    invader = 10;
+    resident = [1:(invader-1) (invader+1):N];
+    invaderstrat = invaderstrats(k);
+    strategy = [strat*ones(1,N)];
+    strategy(invader) = invaderstrat;
+
+    numsigs_tot = numsigs_permove*nummoves;
+
+    scores = zeros(N,numsigs_permove,nummoves);
+    keepM = zeros(N,N,numsigs_permove,nummoves);
+    keepreceivers = zeros(N,numsigs_permove,nummoves);
+
+    % parfor i=1:nummoves
+    for i = 1:nummoves
+        positions = unifrnd(0,1,N,2);
+        d = squareform(pdist(positions));
+
+        M = zeros(N);
+        for ind = 1:N
+            [~, order] = sort(d(ind,:));
+            neighbors = order(2:strategy(ind)+1);
+            M(ind,neighbors) = 1/strategy(ind);
+        end
+        M(1:N+1:end) = -1; %sets diagonal equal to -1
[…]

### 2015-01-23 ESS_figures_v6.m

@@ -1,0 +1,565 @@
+%% setup
+N = 20;
+nummoves = 1000;
+numsigs_permove = 1;
+b = 1;
+
+fontname='Times';
+textfontsz=12;
+labfontsz=9;
+%% load ESS data
+
+ESSeaten_total=cell(2,1);
+ESSgettoeat_total=cell(2,1);
+
+load /Users/eleanorbrush/Desktop/ESS_nummoves=1000_numpermove=1_fullstorage.mat
+
+ESSeaten_total{1}=ESSseaten;
+ESSgettoeat_total{1}=ESSsgettoeat;
+
+load /Users/eleanorbrush/Desktop/ESS_nummoves=10_numpermove=100.mat
+
+ESSeaten_total{2}=ESSseaten;
+ESSgettoeat_total{2}=ESSsgettoeat;
+
+load /Users/eleanorbrush/Desktop/greedyopt_rad=0.1_T=1_nummoves=10_numpermove=100.mat
+
+load /Users/eleanorbrush/Desktop/homogengroupprops.mat
+
+%% find prob eaten for number of informed birds and number of informed neighbors
+radius = .1;
+T = 1;
+strats = [16 17];
+invaderstrats = fliplr(strats);
+
+maxrec = zeros(1,2);
+invader_counts = zeros(N,N,2);
+invader_minscores = cell(N,N,2);
+resident_counts = zeros(N,N,2);
+resident_minscores = cell(N,N,2);
+
+probeaten = zeros(N,2);
+
+for k = 1:2
+    strat = strats(k);
+    invader = 10;
+    resident = [1:(invader-1) (invader+1):N];
+    invaderstrat = invaderstrats(k);
+    strategy = [strat*ones(1,N)];
+    strategy(invader) = invaderstrat;
+
+    numsigs_tot = numsigs_permove*nummoves;
+
+    scores = zeros(N,numsigs_permove,nummoves);
+    keepM = zeros(N,N,numsigs_permove,nummoves);
+    keepreceivers = zeros(N,numsigs_permove,nummoves);
+
+    % parfor i=1:nummoves
+    for i = 1:nummoves
+        positions = unifrnd(0,1,N,2);
+        d = squareform(pdist(positions));
+
+        M = zeros(N);
+        for ind = 1:N
+            [~, order] = sort(d(ind,:));
+            neighbors = order(2:strategy(ind)+1);
+            M(ind,neighbors) = 1/strategy(ind);
[…]

### 2015-01-23 ESS_parallel_bigN_v2.m

@@ -1,0 +1,127 @@
+numworkers=str2num(getenv('PROCS')); %#ok<*ST2NM>
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
+
+numsigs_permove=str2num(getenv('numsigs'));
+nummoves=str2num(getenv('nummoves'));
+
+aid = getenv('SLURM_ARRAY_TASK_ID');
+aid = str2num(aid);
+
+N=50;
+b=1;
+
+strats=10:1:(N-1);
+L=length(strats);
+
+Tvals=[1]; %#ok<*NBRAK>
+Nt=length(Tvals);
+
+radvals=[0:.1:1.4];
+% radvals=.1;
+Nr=length(radvals); %#ok<NASGU>
+
+ESSseaten=cell(Nt);
+ESSsgettoeat=cell(Nt);
+
+storefitnesseaten=cell(Nt);
+storerhoeaten=cell(Nt);
+
+storefitnessgettoeat=cell(Nt);
+storerhogettoeat=cell(Nt);
+
+t0=tic;
+
+for q=1:Nt
+%     for p=1:Nr
+        T=Tvals(q);
+        radius=radvals(aid);
+%         radius=radvals(p);
+
+        fitnesseaten=zeros(L,L);
+        rhoeaten=zeros(L,L);
+
+        fitnessgettoeat=zeros(L,L);
+        rhogettoeat=zeros(L,L);
+
+        featen=zeros(L,L,N-1);
+        geaten=zeros(L,L,N-1);
+
+        fgettoeat=zeros(L,L,N-1);
+        ggettoeat=zeros(L,L,N-1);
+        parfor ind=1:(L*L*(N-1))
+%         parfor ind=1:5
+            [u,v,k]=ind2sub([L,L,N-1],ind);
+            
+            resident=strats(u); %#ok<PFBNS>
+            invader=strats(v);
+
+            strategy=resident*ones(1,N);
+            strategy(N+1-(1:k))=invader;
+            [probeaten, probgettoeat]=signalingevents(strategy,numsigs_permove,nummoves,radius,b,T);
+
+            perfeaten=1-probeaten;
+            featen(ind)=mean(perfeaten(N+1-(1:k)));
+            geaten(ind)=mean(perfeaten(N+1-((k+1):N)));
+
[…]

### 2015-01-23 ESS_vs_correlationlength.m

@@ -1,0 +1,425 @@
+%% load ESS data
+
+ESSeaten_total=cell(2,1);
+ESSgettoeat_total=cell(2,1);
+
+load /Users/eleanorbrush/Desktop/ESS_nummoves=1000_numpermove=1_fullstorage.mat
+
+ESSeaten_total{1}=ESSseaten;
+ESSgettoeat_total{1}=ESSsgettoeat;
+
+load /Users/eleanorbrush/Desktop/ESS_nummoves=10_numpermove=100.mat
+
+ESSeaten_total{2}=ESSseaten;
+ESSgettoeat_total{2}=ESSsgettoeat;
+
+load /Users/eleanorbrush/Desktop/greedyopt_rad=0.1_T=1_nummoves=10_numpermove=100.mat
+
+load /Users/eleanorbrush/Desktop/homogengroupprops.mat
+
+%% find prob eaten for number of informed birds and number of informed neighbors
+N=20;
+radius = 1.5;
+nummoves=1000;
+T = 1;
+strats = [16 18];
+invaderstrats = fliplr(strats);
+
+maxrec = zeros(1,2);
+invader_counts = zeros(N,N,2);
+invader_minscores = cell(N,N,2);
+resident_counts = zeros(N,N,2);
+resident_minscores = cell(N,N,2);
+keepscorestot = zeros(N,nummoves,2);
+keepMtot = zeros(N,N,nummoves,2);
+keepdtot = zeros(N,N,nummoves,2);
+keepreceivertot = zeros(nummoves,2);
+keepreceiverstot = zeros(N,nummoves,2);
+keepminscorestot = zeros(N,nummoves,2);
+
+probeaten = zeros(N,2);
+
+for k = 1:2
+    strat = strats(k);
+    invader = 10;
+    resident = [1:(invader-1) (invader+1):N];
+    invaderstrat = invaderstrats(k);
+    strategy = [strat*ones(1,N)];
+    strategy(invader) = invaderstrat;
+
+    scores = zeros(N,nummoves);
+    keepM = zeros(N,N,nummoves);
+    keepreceivers = zeros(N,nummoves);
+
+    % parfor i=1:nummoves
+    for i = 1:nummoves
+        positions = unifrnd(0,1,N,2);
+        d = squareform(pdist(positions));
+        
+        keepdtot(:,:,i,k) = d;
+
+        M = zeros(N);
+        for ind = 1:N
+            [~, order] = sort(d(ind,:));
[…]

### 2015-01-23 correlationlength_mat_single_fromrec.m

@@ -1,4 +1,4 @@
-function [distvec,corrvec,corrlength,crosscorrs] = correlationlength_mat_single_v3(M,d,b,radius,receiver)
+function [distvec,corrvec,corrlength,crosscorrs] = correlationlength_mat_single_fromrec(M,d,b,radius,receiver)
 
 N=size(M,2);
 

### 2015-01-23 correlationlength_mat_single_path.m

@@ -1,0 +1,49 @@
+function [pathbins,avgcorr,corrlength] = correlationlength_mat_single_path(M,d,b,radius,receiver)
+
+Mbin=M;
+Mbin(M==-1)=0;
+Mbin(~~M)=1;
+g=sparse(Mbin);
+paths=graphallshortestpaths(g);
+
+N=size(M,2);
+corrvec=[];
+pathvec=[];
+% for i=1:N
+%     receiver=i;
+    allreceivers=d(receiver,:)<=radius;
+    beta=zeros(N,1);
+    beta(allreceivers)=1;
+    bbeta=b*beta;
+    im=1;
+    while im>0
+        corrs=paircorrelations(M,bbeta);
+        im=max(max(abs(imag(corrs))));
+    end
+    uppercorr=triu(corrs,1);
+    upperpath=triu(paths,1);
+    lowerpath=triu(transpose(paths),1);
+    corrvectoadd=uppercorr(~~uppercorr);
+    pathvectoadd=[upperpath(~~upperpath) lowerpath(~~lowerpath)];
+    pathvectoadd=min(pathvectoadd,[],2);
+    corrvec=[corrvec; corrvectoadd; diag(corrs)]; 
+    pathvec=[pathvec; pathvectoadd; diag(paths)]; 
+% end
+pathbins=0:1:max(max(paths));
+
+[~,w]=histc(pathvec,pathbins);
+l=length(pathbins);
+avgcorr=zeros(1,l);
+for i=1:l
+    now=(w==i);
+    avgcorr(i)=mean(corrvec(now));
+end
+i=sum(avgcorr>=0);
+if i<l
+    f=find(avgcorr<0,1,'first');
+%     corrlength=avgcorr(i)/(avgcorr(i)-avgcorr(i+1))*(distbins(i+1)-distbins(i))+distbins(i);
+    corrlength=interp1q(avgcorr(f:-1:(f-1))',pathbins(f:-1:(f-1))',0); 
+else corrlength=max(pathvec);
+end
+
+end

### 2015-01-23 correlationlength_mat_single_v2.m

@@ -1,30 +1,0 @@
-function [distvec,corrvec,corrlength] = correlationlength_mat_single_v2(M,d,b,radius,i)
-% if nargin==5 || strcmp(path,'path')
-%     path=1;
-% else path=0;
-% end
-% 
-% if ~path
-N=size(M,2);
-
-receiver=i;
-allreceivers=d(receiver,:)<=radius;
-beta=zeros(N,1);
-beta(allreceivers)=1;
-bbeta=b*beta;
-corrs=paircorrelations(M,bbeta);
-
-corrvec=corrs(i,:);
-distvec=d(i,:);
-[~,o]=sort(distvec);
-distvec=distvec(o);
-corrvec=corrvec(o);
-
-l=length(distvec);
-
-i=sum(corrvec>=0);
-if i<l
-    corrlength=find(corrvec<0,1,'first');
-    corrlength=distvec(corrlength);
-else corrlength=max(distvec);
-end

### 2015-01-23 finding_criticality.m

@@ -1,4 +1,3 @@
-% [distvec,corrvec,corrlength,crosscorrs] = correlationlength_mat_single_v3(M,d,b,radius,i)
 
 T=1;
 teststrategy=2*randi([1 floor((N-1)/2)],N,1);
@@ -24,7 +23,6 @@ M(1:N+1:end)=-1;
     for j=1:Nb 
         b=bvals(j);
 %             [distvec,corrvec,corrlength]=correlationlength_mat(M,d,b,radius);
-%         [distvec,corrvec,corrlength,~] = correlationlength_mat_single_v3(M,d,b,radius,receiver);
         [distvec,corrvec,corrlength] = correlationlength_mat_single(M,d,b,radius,receiver);
         
         storeddistbins{j}=distvec;

### 2015-01-23 greedyopt_check.m

@@ -1,0 +1,37 @@
+num=10;
+f=find(col(sum(abs(diff(evolution_gettoeat(1,:,:),[],3)),2))==0,1,'first');
+strategy=evolution_gettoeat(num,:,f);
+[~, perf]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+newstrategy=strategy;
+for i=1:N
+    choosestrat=zeros(1,3);
+    choosestrat(2)=perf(i);
+
+    if strategy(i)==N-1
+        choosestrat(3)=perf(i);
+    else 
+        upone=strategy;
+        upone(i)=strategy(i)+1;
+
+        [~, probgettoeat]=signalingevents_parallel(upone,numsigs_permove,nummoves,radius,b,T);
+        choosestrat(3)=probgettoeat(i);
+    end
+
+    if strategy(i)==1
+        choosestrat(1)=perf(i);
+    else
+        downone=strategy;
+        downone(i)=strategy(i)-1;
+
+        [~, probgettoeat]=signalingevents_parallel(downone,numsigs_permove,nummoves,radius,b,T);
+        choosestrat(1)=1-probgettoeat(i);   
+    end 
+
+    if choosestrat(1)>choosestrat(2)
+        newstrategy(i)=strategy(i)-1;
+    end
+    if choosestrat(3)>choosestrat(2)
+        newstrategy(i)=strategy(i)+1;
+    end
+
+end

### 2015-01-23 greedyopt_v7.m

@@ -1,0 +1,206 @@
+numworkers=str2num(getenv('PROCS')); %#ok<ST2NM>
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
+
+aid = getenv('SLURM_ARRAY_TASK_ID');
+aid = str2num(aid); %#ok<ST2NM>
+
+numsigs_permove=1;
+nummoves=1000;
+
+N=20;
+b=1;
+radvals=[.1 .7];
+% Nr=length(radvals);
+T=1;
+
+timesteps=200;
+its=20;
+
+evolution_eaten=zeros(its,N,timesteps);
+groupspeed_eaten=zeros(its,timesteps);
+groupconsensus_eaten=zeros(its,timesteps);
+corrlengths_eaten=zeros(its,timesteps);
+disconnected_eaten=zeros(its,timesteps);
+
+evolution_gettoeat=zeros(its,N,timesteps);
+groupspeed_gettoeat=zeros(its,timesteps);
+groupconsensus_gettoeat=zeros(its,timesteps);
+corrlengths_gettoeat=zeros(its,timesteps);
+disconnected_gettoeat=zeros(its,timesteps);
+
+maxt_eaten=zeros(its,1);
+maxt_gettoeat=zeros(its,1);
+
+% for ir=1:Nr
+    radius=radvals(aid);
+for num=1:its
+    init=randi([1 N-1],1,N);
+    t=1;
+    while t<=timesteps
+
+    if t==1
+        strategy=init;
+    %     strategy=2*ones(1,N);
+    %     strategy(1)=10; 
+        evolution_eaten(num,:,t)=strategy;
+        t=t+1;
+    else
+        strategy=evolution_eaten(num,:,t-1);
+        strategy=reshape(strategy,1,N);
+
+        [probeaten, ~]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+        perf=1-probeaten;
+        newstrategy=strategy;
+
+        for i=1:N
+            choosestrat=zeros(1,3);
+            choosestrat(2)=perf(i);
+
+            if strategy(i)==N-1
+                choosestrat(3)=perf(i);
+            else 
+                upone=strategy;
+                upone(i)=strategy(i)+1;
+
[…]

### 2015-01-23 greedyopt_v7_groupprops.m

@@ -1,0 +1,50 @@
+numworkers=str2num(getenv('PROCS'));
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
+
+load /home/brush/schooling_consensus/greedyopt_T=1_nummoves=1000_numpermove=1_rad=0.1_timesteps=200.mato
+
+numsigs_permove=1;
+nummoves=1000;
+
+N=20;
+b=1;
+radius=.1;
+T=1;
+
+timesteps=200;
+its=20;
+
+groupspeed_eaten=zeros(its,timesteps);
+groupconsensus_eaten=zeros(its,timesteps);
+corrlengths_eaten=zeros(its,timesteps);
+disconnected_eaten=zeros(its,timesteps);
+
+groupspeed_gettoeat=zeros(its,timesteps);
+groupconsensus_gettoeat=zeros(its,timesteps);
+corrlengths_gettoeat=zeros(its,timesteps);
+disconnected_gettoeat=zeros(its,timesteps);
+
+for num=1:its
+    for t=1:timesteps
+        strategy_eaten=evolution_eaten(num,:,t);
+        [meanlambda, meanH2, meancorrlength, disconnectedcount]=groupprops_parallel(strategy_eaten,numsigs_permove,nummoves,radius,b,T);
+        groupspeed_eaten(num,t)=meanlambda;
+        groupconsensus_eaten(num,t)=meanH2;
+        corrlengths_eaten(num,t)=meancorrlength;
+        disconnected_eaten(num,t)=disconnectedcount;
+        strategy_gettoeat=evolution_gettoeat(num,:,t);
+        [meanlambda, meanH2, meancorrlength, disconnectedcount]=groupprops_parallel(strategy_gettoeat,numsigs_permove,nummoves,radius,b,T);
+        groupspeed_gettoeat(num,t)=meanlambda;
+        groupconsensus_gettoeat(num,t)=meanH2;
+        corrlengths_gettoeat(num,t)=meancorrlength;
+        disconnected_gettoeat(num,t)=disconnectedcount;
+    end
+end
+
[…]

### 2015-01-23 greedyopt_v7_test.m

@@ -1,0 +1,202 @@
+% numworkers=str2num(getenv('PROCS'));
+% dellacluster=parcluster('local');
+% dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+% dellapool=parpool(dellacluster, numworkers) ;
+
+numsigs_permove=1;
+nummoves=10;
+
+N=10;
+b=1;
+radius=.1;
+T=1;
+
+timesteps=5;
+its=2;
+
+evolution_eaten=zeros(its,N,timesteps);
+groupspeed_eaten=zeros(its,timesteps);
+groupconsensus_eaten=zeros(its,timesteps);
+corrlengths_eaten=zeros(its,timesteps);
+disconnected_eaten=zeros(its,timesteps);
+
+evolution_gettoeat=zeros(its,N,timesteps);
+groupspeed_gettoeat=zeros(its,timesteps);
+groupconsensus_gettoeat=zeros(its,timesteps);
+corrlengths_gettoeat=zeros(its,timesteps);
+disconnected_gettoeat=zeros(its,timesteps);
+
+max_eaten=zeros(1,its);
+max_gettoeat=zeros(1,its);
+
+%strategy evolution
+for num=1:its
+    init=randi([1 N-1],1,N);
+
+'predation regime'
+t=1;
+while t<=timesteps
+% for t=1:timesteps    
+    
+if t==1
+    strategy=init;
+%     strategy=2*ones(1,N);
+%     strategy(1)=10; 
+    evolution_eaten(num,:,t)=strategy;
+    t=t+1;
+else
+    strategy=evolution_eaten(num,:,t-1);
+    strategy=reshape(strategy,1,N);
+
+    [probeaten, ~]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+    perf=1-probeaten;
+    newstrategy=strategy;
+
+    for i=1:N
+        choosestrat=zeros(1,3);
+        choosestrat(2)=perf(i);
+
+        if strategy(i)==N-1
+            choosestrat(3)=perf(i);
+        else 
+            upone=strategy;
+            upone(i)=strategy(i)+1;
+
+            [probeaten, ~]=signalingevents_parallel(upone,numsigs_permove,nummoves,radius,b,T);
+            choosestrat(3)=1-probeaten(i);
+        end
+
+        if strategy(i)==1
+            choosestrat(1)=perf(i);
+        else
[…]

### 2015-01-23 greedyopt_v7_test2.m

@@ -1,0 +1,187 @@
+% numworkers=str2num(getenv('PROCS')); %#ok<ST2NM>
+% dellacluster=parcluster('local');
+% dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+% dellapool=parpool(dellacluster, numworkers) ;
+% 
+% aid = getenv('SLURM_ARRAY_TASK_ID');
+% aid = str2num(aid);
+
+numsigs_permove=1;
+nummoves=10;
+
+N=20;
+b=1;
+radvals=[.1 .7];
+Nr=length(radvals);
+T=1;
+
+timesteps=50;
+its=2;
+
+evolution_eaten=zeros(its,N,timesteps);
+groupspeed_eaten=zeros(its,timesteps);
+groupconsensus_eaten=zeros(its,timesteps);
+corrlengths_eaten=zeros(its,timesteps);
+disconnected_eaten=zeros(its,timesteps);
+
+evolution_gettoeat=zeros(its,N,timesteps);
+groupspeed_gettoeat=zeros(its,timesteps);
+groupconsensus_gettoeat=zeros(its,timesteps);
+corrlengths_gettoeat=zeros(its,timesteps);
+disconnected_gettoeat=zeros(its,timesteps);
+
+% for ir=1:Nr
+    radius=radvals(aid);
+for num=1:its
+    init=randi([1 N-1],1,N);
+    t=1;
+    while t<=timesteps
+
+    if t==1
+        strategy=init;
+    %     strategy=2*ones(1,N);
+    %     strategy(1)=10; 
+        evolution_eaten(num,:,t)=strategy;
+        t=t+1;
+    else
+        strategy=evolution_eaten(num,:,t-1);
+        strategy=reshape(strategy,1,N);
+
+        [probeaten, ~]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+        perf=1-probeaten;
+        newstrategy=strategy;
+
+        for i=1:N
+            choosestrat=zeros(1,3);
+            choosestrat(2)=perf(i);
+
+            if strategy(i)==N-1
+                choosestrat(3)=perf(i);
+            else 
+                upone=strategy;
+                upone(i)=strategy(i)+1;
+
+                [probeaten, ~]=signalingevents_parallel(upone,numsigs_permove,nummoves,radius,b,T);
+                choosestrat(3)=1-probeaten(i);
+            end
+
[…]

### 2015-01-23 paircorrelations_figuringitout.m

@@ -1,0 +1,65 @@
+strategy=(N-1)*ones(N,1);
+positions=unifrnd(0,1,N,2);
+d=squareform(pdist(positions));
+
+M=zeros(N);
+for ind=1:N
+    [~, order]=sort(d(ind,:));
+    neighbors=order(2:strategy(ind)+1);
+    M(ind,neighbors)=1/strategy(ind);
+end
+M(1:N+1:end)=-1;
+
+k=0;
+
+q=zeros(N-1,N);
+q(1,1:2)=[1/sqrt(2) -1/sqrt(2)];
+for j=2:(N-1)
+   v=rand(1,N);
+   v=v-sum(v)/N;
+   for k=1:(j-1)
+      v=v-(v*transpose(q(k,:)))*q(k,:); 
+   end
+   v=v/norm(v,2);
+   q(j,:)=v;
+end
+
+
+A=M-diag(beta);
+P=-q*A*transpose(q);
+Pinv=inv(P);
+
+[vecs,vals]=eig(P);
+vals=diag(vals);
+invvals=1./vals;
+U=vecs*diag(invvals)*transpose(vecs);
+
+sigma1=q*transpose(A)*ones(N,1);
+sigma2=q*beta;
+
+zcorrs=zeros(N-1,N-1);
+for i=1:(N-1)
+    for j=i:(N-1)
+%         piece1=sum(vecs(i,:).*invvals'.*vecs(j,:));
+%         piece1=sum(vecs(i,:).*reshape(invvals,1,[]).*vecs(j,:));
+%         piece2=sum(sum((sigma1'*Pinv(:,i))*(sigma1'*Pinv(:,j))'))+sum(sum((sigma1'*Pinv(:,j))*(sigma1'*Pinv(:,i))'));
+%         piece3=sum(sum((sigma2'*Pinv(:,i))*(sigma2'*Pinv(:,j))'))+sum(sum((sigma2'*Pinv(:,j))*(sigma2'*Pinv(:,i))'));
+%         piece2=(reshape(sigma1,1,[])*Pinv(:,i))*(reshape(sigma1,1,[])*col(Pinv(j,:)));
+%         piece3=(reshape(sigma2,1,[])*Pinv(:,i))*(reshape(sigma2,1,[])*col(Pinv(j,:)));
+        Pinv_deriv=kron(Pinv(:,i),Pinv(j,:));
+        piece1=U(i,j);
[…]

### 2015-01-23 signalingevents.m

@@ -1,11 +1,11 @@
-function [probeaten, probgettoeat]=signalingevents(strategy,numsigs_permove,nummoves,radius,b,T)
+% function [probeaten, probgettoeat]=signalingevents(strategy,numsigs_permove,nummoves,radius,b,T)
 numsigs_tot=numsigs_permove*nummoves;
 N=max(size(strategy));
 
 scores=zeros(N,numsigs_permove,nummoves);
 
 % parfor i=1:nummoves
-for i=1:nummoves
+for i=1:20000
     positions=unifrnd(0,1,N,2);
     d=squareform(pdist(positions));
 
@@ -53,4 +53,4 @@ for i=1:size(scores,2)
     maxscorer(rows(cols==i),i)=1/numsigs_tot;
 end
 probgettoeat=sum(maxscorer,2);
-end
+% end

### 2015-01-30 paper_skeleton.tex
im about to redo how i calculate h2 norm and correlations

@@
 Thus, the uninvadable strategy is to pay attention to many neighbors when selection is due to knowledge about predators.
-Figure \ref{ESS} shows the ESS strategy when selection is due to knowledge about resources is always lower than the ESS strategy when selection is due to knowledge about predators (Figure \ref{ESS}).
+Figure \ref{ESS}a shows the ESS strategy when selection is due to knowledge about resources is always lower than the ESS strategy when selection is due to knowledge about predators (Figure \ref{ESS}a).
 
@@
 This indicates how important the social environment set by the resident strategy is.
-The resident strategy essentially sets a level of background noise.
-As an example, consider the strategies of having sixteen or seventeen neighbors when the radius of the signal is $r=0.1$.
-Figure \ref{ESS} shows how the probability of being predated for each of these invasion scenarios depends on the number of birds in the group and the number of informed neighbors the invader has.
-A bird with seventeen neighbors will have more uninformed neighbors than one with sixteen neighbors and thus will have less information itself.
-This poor information is  costly when the resident strategy is to have sixteen neighbors since the residents will have less noisy information, which prevents an mutant with seventeen neighbors to invade.
-Figure \ref{ESS} shows that by paying attention to more uninformed neighbors this mutant is likely to be eaten when there are many truly informed birds.
-However, having less noisy information does not allow a sixteen neighbor mutant to invade a seventeen neighbor group since, regardless of the number of uninformed neighbors the invader has, the background level of noise is much higher in a group all of whom have more uninformed neighbors.
-The sixteen neighbor mutant also pays a cost by being less likely to pay attention to the right neighbors when there are very few truly informed birds.
-Figure \ref{ESS} shows that by being less likely to pay attention to the right neighbors this mutant is likely to be eaten when there are few truly informed birds.
+The pairwise invisibility plot in Figure \ref{ESS}b shows that low strategies are always costly and that high strategies are only costly when the resident strategy is high.
 
-These opposing forces caused by the costs of paying attention to more uninformed neighbors and the costs of not paying attention to informed neighbors also explains the second feature of the ESS strategies: they are non-monotonic with respect to the the radius of the signal, i.e.
-how public the signal is (Figure \ref{ESS}).
-Figure \ref{ESS} shows that there are essentially three categories of radii: When the radius is low, the signal is private, with only a few birds having true information.
+If a bird has the environmental signal, it will not be predated, regardless of the number of neighbors it has.
+The differences in performance only appear when comparing two birds, neither of whom have the signal.
+To understand the costs of having too few or too many neighbors, we compare the performance of two focal birds, a resident and an invader, neither of whom has the signal.
+There are two main cases: there can be a single informed bird or multiple informed birds.
+We further break down the first case depending on which of these two focal birds pays attention to the informed bird: neither, the resident, the invader, or both.
+For each of these cases, we find the probability that the invader is predated.
+An invader with one fewer neighbor than the resident always has probability of being predated greater than chance ($1/N$) because the resident often has an informed neighbor while the invader how no or fewer informed neighbors and thus a less informed opinion (Figure \ref{ESS}c).
+The total probability of being predated stays more or less constant regardless of the resident strategy, with a slight decrease for high resident strategies (Figure \ref{ESS}b and c).
+An invader with one more neighbor than the resident is only more likely than chance to be predated when the resident strategy is high enough  (Figure \ref{ESS}d).
+This invader performs most poorly when both it and the focal resident pay attention to the informed bird(s).
+In this case, an invader with one more neighbor has more naive neighbors and thus a less informed opinion.
+This is increasingly problematic as the resident strategy increases and both the focal birds are likely to pay attention to the informed bird(s).
+
+These opposing forces caused by the costs of not paying attention to informed neighbors  and the costs of paying attention to more uninformed neighbors  also explains the second feature of the ESS strategies: they are non-monotonic with respect to the the radius of the signal, i.e.
+how public the signal is (Figure \ref{ESS}a).
+Figure \ref{ESS}a shows that there are essentially three categories of radii: When the radius is low, the signal is private, with only a few birds having true information.
 When the radius is intermediate, the signal is public, with many birds having true information.
@@
 When the signal is universal, not paying attention to informed birds does not impose a cost.
-Additionally, in a group in which the residents pay attention to all of their neighbors, any mutant strategy can get just as much information by paying attention to fewer neighbors so the maximal strategy is no longer an ESS (Figure \ref{ESS}).
+Additionally, in a group in which the residents pay attention to all of their neighbors, any mutant strategy can get just as much information by paying attention to fewer neighbors so the maximal strategy is no longer an ESS (Figure \ref{ESS}a).
+
+Another way to explain the non-monotonic ESS number of neighbors is to consider the correlation length of a homogeneous flock as a function of the radius of the signal.
+If a flock is all using the same strategy, its average correlation length depends much more strongly on the radius of the signal than the strategy being used (Figure \ref{corrESS}a).
+In particular, the correlation length first rises and then falls as the radius is increased.
+The rise and fall of the correlation length correspond to the rise and fall of the ESS number of neighbors and if the we plot the ESS number of neighbors for various radii of the signal against the average correlation length of a homogeneous flock for that radii, we find a positive relationship (Figure \ref{corrESS}b).
+This is because when correlations in information die off very quickly (i.e.
+a low correlation length), having more neighbors is costly because birds even slightly away from the informed birds will have poor information.
+When there are strong correlations in information across the flock, on the other hand, these costs are reduced because every bird in the flock has on average higher quality information.
 
@@
 expect ESS strategies to decrease as the signal becomes more public, but we find the opposite
-
-\section{Figures }
-\begin{figure}[ht]
-\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/ESSfigure.pdf}
-\caption{\label{ESS} The ESS number of neighbors is always higher if selection is due to predation than due to resources.
-When selection is due to predation, there can be multiple ESS strategies, which are a non-monotonic function of the radius of the signal.
-In (a), we show the (possibly multiple) ESS strategies as a function of the radius of the signal.
-In (b) we show the probability of there being a given number of informed birds for signals with different radii.
-In (c) and (d), each heat map shows the probability that an invader is eaten as a function of the number of informed birds in the flock on the horizontal axis and the number of informed neighbors the invader has on the vertical axis.
-In (c), the resident strategy is to have $16$ neighbors and the invader strategy is to have $17$ neighbors.
-In (d), the resident strategy is to have $17$ neighbors and the invader strategy is to have $16$ neighbors.
-Parameters:  in all panels $\beta=1$, $T=1$, in (c) and (d) $r=0.1$.
-}
-\end{figure}
-
-\begin{figure}[ht]
-\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/greedyoptneighbors.pdf}
-\caption{\label{greedyopt} When the birds choose the optimal number of neighbors, given the strategies the rest of the flock are using, they settle on higher strategies when selection is due to predation than when selection is due to resources.
-Under selection due to predation, the $\mathscr{H}_2$ robustness and correlation length of the flock tend to decrease, whereas under selection due to resources, $\mathscr{H}_2$ robustness is relatively constant and correlation length tends to increase.
-The upper row shows results from implementing selection based to predation and the lower row shows results from implementing selection based on resources.
-In each panel, the x-axis represents the number of times the birds are allowed to choose optimal strategies.
-The first column shows one example of how the birds' strategies change over time.
-The second column shows, for many initial conditions, how the $\mathscr{H}_2$ robustness changes over time.
-The third column shows, for many initial conditions, how the correlation length changes over time.
-Parameters:  $\beta=1$, $r=.1$, $T=1$.
-}
-\end{figure}
 
@@
 \end{align*}
+
+\begin{claim}
+Given any $Q>0$ there exists a unique $\Sigma>0$ satisfyying $A\Sigma+\Sigma A^T=Z$ if and only if the system $\dot{x}=Ax$ is  global asymptotically stable.
+\end{claim}
+
+\begin{claim}
+Consider the dynamical system,
+\begin{equation}
+\dot{y}=Ay+B+C\xi  \label{tofindsigma}
+\end{equation}
+with $E[\xi(t)]=0$, $E[\xi(t)\xi(s)^T]=\frac{1}{2}\delta(t-s)I$, and $E[y(0)\xi^T(t)]=0$.
+If all eigenvalues of $A$ are negative and if $\Sigma(t)=E[y(t)y(t)^T]$ and $\Sigma_{ss}=\lim_{t\to\infty}\Sigma(t)$, then $\Sigma_{ss}$ satisfies the Lyapunov equation
+\begin{equation*}
+A\Sigma_{ss}+\Sigma_{ss}A^T=A^{-1}BB^T+BB^T(A^{-1})^T-CC^T
+\end{equation*}
+\end{claim}
+
+\begin{pf}
+\begin{align*}
+\Sigma(t)&=E[y(t)y(t)^T]
+\\ \Rightarrow \dot{\Sigma}&=E[\dot{y}(t)y(t)^T+y(t)\dot{y}(t)^T]
+\\&=E[(Ay+B+C\xi)y^T+y(y^TA^T+B^T+\xi^TC^T)]
+\\&=AE[yy^T]+E[yy^T]A^T+BE[y^T]+E[y]B^T+CE[\xi y^T]+E[y\xi^T]C^T
+\end{align*}
+Note that we can write a solution to (\ref{tofindsigma}) as 
+\begin{equation*}
+y(t)=\exp(At)y(0)+\int_0^t\exp(A(t-s))Bds+\int_0^t\exp(A(t-s))C\xi(s)ds.
+\end{equation*}
+This allows us to write $E[y]=\exp(At)y(0)+\int_0^t\exp(A(t-s))Bds$ and 
+\begin{align*}
+E[y \xi^T]&=\exp(At)E[y(0)\xi(t)^T]+\int_0^t\exp(A(t-s))BE[\xi(t)^T]ds+\int_0^t\exp(A(t-s))CE[\xi(s)\xi(t)]ds
+\\&=0+0+\frac{1}{2}C
+\\ \Rightarrow \dot{\Sigma}&=A\Sigma(t)+\Sigma(t)A^T+By(0)^T\exp(At)^T+\exp(At)y(0)B^T
+\\&+B\int_0^tB^T\exp(A(t-s))^Tds+\int_0^T\exp(A(t-s))BdsB^T+\frac{1}{2}CC^T+\frac{1}{2}CC^T
+\\&=A\Sigma(t)+\Sigma(t)A^T+By(0)^T\exp(At)^T+\exp(At)y(0)B^T
+\\&-BB^T(A^{-1})^T\exp(A(t-s))|_{s=0}^{s=t}-A^{-1}\exp(A(t-s))|_{s=0}^tBB^T+CC^T
+\\&=A\Sigma(t)+\Sigma(t)A^T+By(0)^T\exp(At)^T+\exp(At)y(0)B^T
+\\&-BB^T(A^{-1})^T(I-\exp(At))-A^{-1}(I-\exp(At))BB^T+CC^T
+\\ \lim_{t\to\infty}\dot{\Sigma}&=A\Sigma_{ss}+\Sigma_{ss}A^T-BB^T(A^{-1})^T-A^{-1}BB^T+CC^T \text{ since all eigenvalues of $A$ are negative}
+\\\Rightarrow A\Sigma_{ss}+\Sigma_{ss}A^T&=BB^T(A^{-1})^T+A^{-1}BB^T-CC^T \text{ since $\dot{\Sigma}=0$ at equilibrium}
+\end{align*}
+\end{pf}
 
@@
 (Consider that $\vec{v}^TA^TA\vec{v}=(A\vec{v})^T(A\vec{v})$ so that $\vec{v}A^TA\vec{v}=0$ if $A\vec{v}=0$ and $\vec{v}^TA^TA\vec{v}<0$ otherwise.) Therefore, $-\vec{v}^TA^2\vec{v}\leq 0$ and $-\vec{v}^TA^TA\vec{v}\leq 0$.
-At equilibrium, $\vec{v}$ will be ``close" to $\vec{\beta}$ so that $$-\vec{v}^TA\vec{\beta}-2\vec{v}^TA^T\vec{\beta}\sim-\vec{\beta}^TA\vec{\beta}$$
+At equilibrium, $\vec{v}$ will be ``close" to $\vec{\beta}$ so that $$-\vec{v}^TA\vec{\beta}-2\vec{v}^TA^T\vec{\beta}\sim-3\vec{\beta}^TA\vec{\beta}$$
 \end{pf}
@@
 \subsection{Numerically evaluating dynamics }
-\begin{claim}
+%\begin{claim}
 Let $\Phi(t)$ be the fundamental matrix solution to the homogeneous differential equations $\frac{d\vec{x}}{dt}=A\vec{x}$ so that $\Phi(t)=V\Lambda(t)$ where the columns of $V$ are the eigenvectors of $A$ and $$\Lambda(t)=\text{diag}(e^{\lambda_1 t},\dots,e^{\lambda_n t})$$ where $\{\lambda_i\}$ are the eigenvalues of $A$.
@@
 Let $$G(t)=\text{diag}(1/\lambda_1(1-e^{-\lambda_1t}),\dots,1/\lambda_n(1-e^{-\lambda_nt}))V^\dagger\vec{\beta}.$$
-(If $\lambda_i=0$ then the corresponding entry of the first part of $G(t)$ will be given by $t$ rather than the form above.)  THEN if we let $x(t)=\Phi(t)G(t)+\Phi(t)\vec{c}$, $X(t)$ solves the inhomogeneous equations $\frac{dx}{dt}=A\vec{x}+\vec{\beta}$ with initial conditions $x(0)=\vec{X_0}$.
+(If $\lambda_i=0$ then the corresponding entry of the first part of $G(t)$ will be given by $t$ rather than the form above.)  
+\begin{claim}
+If we let $x(t)=\Phi(t)G(t)+\Phi(t)\vec{c}$, $x(t)$ solves the inhomogeneous equations $\frac{dx}{dt}=A\vec{x}+\vec{\beta}$ with initial conditions $x(0)=\vec{X_0}$.
 \end{claim}
@@
 \\ 0 & \dots &e^{\lambda_nt}
- \end{array}\right)
+ \end{array}\right)V^\dagger\vec{\beta}
  \\&=V
@@
 
+\newpage
+\section{Figures }
+\begin{figure}
+\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/ESSfigure.pdf}
+\caption{\label{ESS} The ESS number of neighbors is always higher if selection is due to predation than due to resources.
+When selection is due to predation, there can be multiple ESS strategies, which are a non-monotonic function of the radius of the signal.
+In (a), we show the (possibly multiple) ESS strategies as a function of the radius of the signal.
+In (b) we show the relative fitness of an invader as a function of the invader and resident strategies when $r=0.1$.
+The dots show the multiple ESSs.
+For (c) and (d), we choose a focal resident bird and consider the situation where neither the focal bird nor the invader has the environmental signal.The total of the stacked bars is the total probability that the invader will be predated when neither the focal bird nor the invader have the signal.
+We further break down these cases into five cases: there is a single informed bird and neither focal bird, only the focal resident, only the invader, or both pay attention to it or there are multiple informed birds.
+The horizontal dashed line shows $1/N$, the probability of being predated if the strategies are neutral.
+In (c), the invader has one fewer neighbor than the resident.
+In (d), the invader has one more neighbor than the resident.
+Parameters:  in all panels $N=20$, $\beta=1$, $T=1$, in (b)-(d) $r=0.1$.
+}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/corrlength_vs_ESS.pdf}
+\caption{\label{corrESS}
+The correlation length of a homogeneous flock depends on both the strategy being used and the radius of the signal, but it depends much more strongly on the radius than the strategy.
+The ESS number of neighbors is an increasing function of correlation length of a homogeneous group.
+In (a), we show the average correlation length of a homogeneous flock as a function of the strategy being used and the radius of the signal.
+In (b), each point corresponds to a value of the radius of signal: the horizontal axis shows the correlation length of a flock paying attention to $N/2$ neighbors for that radius and the vertical axis shows the ESS number of neighbors for that radius.
+Parameters: $N=20$, $\beta=1$, $T=1$.
+}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/greedyoptneighbors.pdf}
+\caption{\label{greedyopt} When the birds choose the optimal number of neighbors, given the strategies the rest of the flock are using, they settle on higher strategies when selection is due to predation than when selection is due to resources.
+Under selection due to predation, the $\mathscr{H}_2$ robustness and correlation length of the flock tend to decrease, whereas under selection due to resources, $\mathscr{H}_2$ robustness is relatively constant and correlation length tends to increase.
+The upper row shows results from implementing selection based to predation and the lower row shows results from implementing selection based on resources.
+In each panel, the  horizontal axis represents the number of times the birds are allowed to choose optimal strategies.
+The first column shows one example of how the birds' strategies change over time.
+The second column shows, for many initial conditions, how the $\mathscr{H}_2$ robustness changes over time.
+The third column shows, for many initial conditions, how the correlation length changes over time.
+Parameters:  $N=20$, $\beta=1$, $r=.1$, $T=1$.
+}
+\end{figure}
+
 \nocite{*}

### 2015-01-30 H2norm.m

@@ -1,6 +1,6 @@
 function h = H2norm(M,scaled)
 % L=lap(M);
-L=-M;
+L=M;
 s=size(M);
 n=s(1);
 if nargin==1
@@ -25,5 +25,7 @@ for j=2:(n-1)
 end
 Lbar=q*L*transpose(q);
 noise=q*noise*transpose(q);
-sigma = lyap(Lbar,-noise);
+% X=lyap(A,Q) solves AX+XA^T+Q=0 and I want to solve
+% Lbar*Sigma+Sigma*Lbar^T=I
+sigma = lyap(Lbar,noise);
 h = sqrt(trace(sigma));

### 2015-01-30 expected_spin.m

@@ -11,65 +11,15 @@ if nargin==2
     beta=1*ones(N,1);
 end
 
-[vecs,vals]=eig(A);
-V=vecs;
-vals=diag(vals);
-nonzero=abs(vals)>0.00001;
-nonzerovals=vals(nonzero);
-
-%if there is a zero eigenvalue set the eigenvector to all ones
-%if there are two eigenvectors for eigenvalue 0 I specify what they should
-%be here
-indices=1:N;
-where=indices(~nonzero);
-vhat=1/sqrt(N)*ones(N,1);
-if sum(~nonzero)==1
-    V(:,where(1))=ones(N,1);
-%     V(:,where(1))=vhat;
-end
-if sum(~nonzero)>1
-    L=sum(~nonzero);
-    for i=1:(L-1)
-        vproj=V(:,where(i+1))-(V(:,where(i+1))'*vhat)*vhat;
-        V(:,where(i+1))=vproj;
-    end
-end
-
-% mu=pinv(V)*beta;s
-
-% c=zeros(N,1);
-% c(where(1))=1;
-
 initialconditions=.5*ones(N,1);
-c=pinv(V)*initialconditions;
 
-l=exp(nonzerovals*T);
-Lhom=zeros(1,N);
-Lhom(nonzero)=l;
-Lhom(~nonzero)=1;
-Lhom=diag(Lhom);
-Phi=V*Lhom;
-Uhom=Phi;
+fun = @(x)expm(-x*A)*beta;
+integrated = integral(fun,0,T,'ArrayValued',true);
 
-l=1./nonzerovals.*(exp(nonzerovals*T)-1);
-Linhom=zeros(1,N);
-Linhom(nonzero)=l;
-Linhom(~nonzero)=T;
-Linhom=diag(Linhom);
-% Uinhom=V*Linhom/V;
-Uinhom=V*Linhom*pinv(V);
-
-instantaneous=Uinhom*beta+Uhom*c;
-
-% l=1./power(nonzerovals,2).*(exp(nonzerovals*T)-1)-T./nonzerovals;
-% L=zeros(1,N);
-% L(nonzero)=l;
-% L(~nonzero)=T^2/2;
-% L=diag(L);
-% U=V*L/V;
-% integrated=U*beta;
-
-% q=[instantaneous,integrated];
+homogeneous=expm(T*A)*initialconditions;
+% particular=expm(T*A)*inv(A)*[eye(N)-expm(-T*A)]*beta;
[…]

### 2015-02-05 paper_skeleton.tex
i think things are clean now? though im worried about ess taking too long

@@
 
-\begin{center}
-{\bf \LARGE{Evolution of Information Gathering Strategies}}
-\vspace{10pt}
-\\ Eleanor Brush
-\\ October 31, 2014
-\end{center}
+%
+%\begin{center}
+%{\bf \LARGE{Evolution of Information Gathering Strategies}}
+%\vspace{10pt}
+%\\ Eleanor Brush
+%\\ October 31, 2014
+%\end{center}
 
@@
 \subsection{Learning dynamics }
-An individual's opinion changes with a probability that depends on its social interactions and the strength of the environmental stimulus.
-Specifically, an individual's opinion $\sigma_i$ can be either $1$ or $-1$ and $\sigma_i$ switches with probability $w_i(\sigma_i)=\frac{1}{2}(-\sigma_i\sum_jA_{ij}\sigma_j )(1-\beta_i\sigma_i)$ where $A_{ij}$ indicates the strength of the connection from $j$ to $i$ and $\beta_i$ indicates the strength of the environmental stimulus perceived by $i$.
-The stochastic dynamics of the opinions can be written down with a master equation \cite{Glauber:1963fk}.
-The vector $\vec{v}(t)$ of expected values, $q_i(t)=\langle \sigma_i(t)\rangle$, satisfies the following differential equations: \cite{Glauber:1963fk}
-\begin{equation*}
-\frac{d\vec{v}}{dt}=M\vec{v}+\vec{\beta}^T(\vec{1}-\vec{v}).
-\end{equation*}
-We simplify this by writing $\frac{d\vec{v}}{dt}=A\vec{v}+\vec{\beta}^T\vec{1}$, where $A=M-B$ and $B$ is a diagonal matrix with the values of $\vec{\beta}$ along the diagonal.
+An individual's opinion changes with a probability that depends on its social interactions and the strength of the environmental stimulus. Let $A$ be the adjacency matrix such that
+
+\begin{equation}
+A_{ij}=\left\{\begin{array}{l l l}
+1 & \text{ if $i$ pays attention to $j$ }
+\\0 & \text{ otherwise }
+\end{array}\right.
+\end{equation}
+
+We assume that having more neighbors means paying less attention to each so we normalize $A$ to find $\bar{A}$:
+\begin{equation}
+\bar{A}_{ij}=\frac{A_{ij}}{\sum_kA_{ik}}.
+\end{equation}
+
+If each bird averages the differences between its opinion and each of its neighbors' to update its opinion, the opinion dynamics will be described by 
+\begin{equation}
+\dot{x}=Lx
+\end{equation}
+where $L=\bar{A}-I$.
 In the absence of the environmental signal, there is no trend in the dynamics.
-In the presence of the environmental signal, however, those individuals who perceive the signal will be more likely to switch to having an opinion of $1$ so their expected value will increase, and consequently so will the values of the those paying attention to the informed individuals.
-Thus, having a higher expected value can be interpreted as having more information about the information and having a lower expected value can be interpreted as having less information.
+We introduce an environmental signal by changing the dynamics such that informed birds change their opinions to be closer to $1$.
+Having an opinion closer to $1$ can be interpreted as having more information about the environment and having a lower expected value can be interpreted as having less information.
+Specifically, we incorporate this environmental forcing with a diagonal matrix $B$ where $B_{ii}=b$ if $i$ perceives the environmental signal and $B_{ii}=0$ otherwise.
+The dynamics then become
+\begin{equation}
+\dot{x}=Lx+B(\vec{1}-x).
+\end{equation}
+If we let $L_\text{f}=L-B$, then we can simplify this by writing
+\begin{equation}
+\dot{x}=L_\text{f}x+\vec{B}
+\end{equation}
+where $\vec{B}=B\vec{1}$ indicates which birds perceive the environmental signal.
 
@@
 \subsection{Group-level performance }
+%%There are two ways to describe equilibrium properties
 We consider two group-level properties that indicate how cohesive the group is.
 The first property is $\mathscr{H}_2$ robustness, as used in \cite{Young:2010fk,Young:2013kx}.
-This is a measure of the robustness of the consensus state in which all birds have the same opinion to noise at equilibrium, as used in (see Appendix Sec.
-\ref{H2}).
+This is a measure of the robustness of the consensus state in which all birds have the same opinion to noise at equilibrium, as used in (see Appendix Section XXX).
 The second property is the correlation length of the flock.
 This is the distance at which the average correlations between birds changes from positive to negative, i.e.
-the distance over which birds' opinions tend to be positively correlated with each other (see Appendix Sec \ref{corr}).
+the distance over which birds' opinions tend to be positively correlated with each other (see Appendix Section XXX).
 
@@
 \section{Appendix}
-\subsection{H2 Norm \label{H2}}
+%%% APPENDIX NOW DISCUSSES H2 WITHOUT FORCING (AS IN GEORGE'S PAPER), H2 WITH FORCING, WHICH GIVES CORRELATIONS BUT NOT FLUCTUATIONS AROUND AVERAGE OPINION, STAT MECH CORRELATIONS WITHOUT FORCING AND STAT MECH CORRELATION WITH FORCING, BOTH OF WHICH GIVE FLUCTUATIONS AROUND AVERAGE OPINION.
+SO H2 WITH FORCING IS STILL INCOMPARABLE TO ANYTHING ELSE SINCE ITS CORRELATIONS NOT FLUCTUATIONS AROUND AVERAGE OPINIONS.
+IT'S OKAY FOR NOW.
+\subsection{Linear algebra background }
 
-\ Let $y$ represent the distance between the vector of opinions and consensus, i.e.
-$y=QxQ^T,$
-where $Q\in\R^{n-1\times n}$ is such that $Q\vec{1}=0$ , $QQ^T=I_{n-1}$, and $Q^TQ=I_n-\frac{1}{n}\vec{1}\times\vec{1}^T$.   
-Then 
-\begin{equation}
-\dot{y}(t)=-\overline{L}y(t)+Q\xi(t) \notag
-\end{equation}
-where $\overline{L}=QLQ^T$.  As shown in \cite{Young:2010fk}, if $\Sigma_y(t)=\E[y(t)y(t)^T]$,  $\Sigma_y$ at equilibrium solves
-\begin{equation} 0=-\overline{L}\Sigma_y-\Sigma_y\overline{L}^T+D. \label{h2norm}
-\end{equation}
-\begin{align*}
-\mathscr{H}_2&=\sqrt{\Tr(\Sigma_y)}
-\\\mathscr{H}_2 \text{ robustness} &=\frac{1}{\mathscr{H}_2}
-\end{align*}
+\begin{theorem} \label{lyapeq_theorem}
+Given any positive symmetric matrix $Q>0$ there exists a unique positive symmetric matrix $\Sigma>0$ satisfyying $A\Sigma+\Sigma A^T=Q$ if and only if the system $\dot{x}=Ax$ is globally asymptotically stable.
+\end{theorem}
 
-\begin{claim}
-Given any $Q>0$ there exists a unique $\Sigma>0$ satisfyying $A\Sigma+\Sigma A^T=Z$ if and only if the system $\dot{x}=Ax$ is  global asymptotically stable.
-\end{claim}
-
-\begin{claim}
+\begin{claim} \label{lyapeq_derivation}
 Consider the dynamical system,
 \begin{equation}
-\dot{y}=Ay+B+C\xi  \label{tofindsigma}
+\dot{y}=Ay+B+Q\xi  \label{tofindsigma}
 \end{equation}
@@
 \begin{equation*}
-A\Sigma_{ss}+\Sigma_{ss}A^T=A^{-1}BB^T+BB^T(A^{-1})^T-CC^T
+A\Sigma_{ss}+\Sigma_{ss}A^T=A^{-1}BB^T+BB^T(A^{-1})^T-QQ^T.
+\end{equation*}
+Further, if $C_{ij}=E[(y_i-E(y_i))(y_j-E(y_j))]$ and $C_{ss}=\lim_{t\to\infty}C$ then 
+\begin{equation*}
+AC_{ss}+C_{ss}A^T=-QQ^T
 \end{equation*}
@@
 \\ \Rightarrow \dot{\Sigma}&=E[\dot{y}(t)y(t)^T+y(t)\dot{y}(t)^T]
-\\&=E[(Ay+B+C\xi)y^T+y(y^TA^T+B^T+\xi^TC^T)]
-\\&=AE[yy^T]+E[yy^T]A^T+BE[y^T]+E[y]B^T+CE[\xi y^T]+E[y\xi^T]C^T
+\\&=E[(Ay+B+Q\xi)y^T+y(y^TA^T+B^T+\xi^TQ^T)]
+\\&=AE[yy^T]+E[yy^T]A^T+BE[y^T]+E[y]B^T+QE[\xi y^T]+E[y\xi^T]Q^T
 \end{align*}
@@
 \begin{equation*}
-y(t)=\exp(At)y(0)+\int_0^t\exp(A(t-s))Bds+\int_0^t\exp(A(t-s))C\xi(s)ds.
+y(t)=\exp(At)y(0)+\int_0^t\exp(A(t-s))Bds+\int_0^t\exp(A(t-s))Q\xi(s)ds.
 \end{equation*}
@@
 \begin{align*}
-E[y \xi^T]&=\exp(At)E[y(0)\xi(t)^T]+\int_0^t\exp(A(t-s))BE[\xi(t)^T]ds+\int_0^t\exp(A(t-s))CE[\xi(s)\xi(t)]ds
-\\&=0+0+\frac{1}{2}C
+E[y \xi^T]&=\exp(At)E[y(0)\xi(t)^T]+\int_0^t\exp(A(t-s))BE[\xi(t)^T]ds+\int_0^t\exp(A(t-s))QE[\xi(s)\xi(t)]ds
+\\&=0+0+\frac{1}{2}Q
 \\ \Rightarrow \dot{\Sigma}&=A\Sigma(t)+\Sigma(t)A^T+By(0)^T\exp(At)^T+\exp(At)y(0)B^T
-\\&+B\int_0^tB^T\exp(A(t-s))^Tds+\int_0^T\exp(A(t-s))BdsB^T+\frac{1}{2}CC^T+\frac{1}{2}CC^T
+\\&+B\int_0^tB^T\exp(A(t-s))^Tds+\int_0^T\exp(A(t-s))BdsB^T+\frac{1}{2}CC^T+\frac{1}{2}QQ^T
 \\&=A\Sigma(t)+\Sigma(t)A^T+By(0)^T\exp(At)^T+\exp(At)y(0)B^T
-\\&-BB^T(A^{-1})^T\exp(A(t-s))|_{s=0}^{s=t}-A^{-1}\exp(A(t-s))|_{s=0}^tBB^T+CC^T
+\\&-BB^T(A^{-1})^T\exp(A(t-s))|_{s=0}^{s=t}-A^{-1}\exp(A(t-s))|_{s=0}^tBB^T+QQ^T
 \\&=A\Sigma(t)+\Sigma(t)A^T+By(0)^T\exp(At)^T+\exp(At)y(0)B^T
-\\&-BB^T(A^{-1})^T(I-\exp(At))-A^{-1}(I-\exp(At))BB^T+CC^T
-\\ \lim_{t\to\infty}\dot{\Sigma}&=A\Sigma_{ss}+\Sigma_{ss}A^T-BB^T(A^{-1})^T-A^{-1}BB^T+CC^T \text{ since all eigenvalues of $A$ are negative}
-\\\Rightarrow A\Sigma_{ss}+\Sigma_{ss}A^T&=BB^T(A^{-1})^T+A^{-1}BB^T-CC^T \text{ since $\dot{\Sigma}=0$ at equilibrium}
+\\&-BB^T(A^{-1})^T(I-\exp(At))-A^{-1}(I-\exp(At))BB^T+QQ^T
+\\ \lim_{t\to\infty}\dot{\Sigma}&=A\Sigma_{ss}+\Sigma_{ss}A^T-BB^T(A^{-1})^T-A^{-1}BB^T+QQ^T \text{ since all eigenvalues of $A$ are negative}
+\\\Rightarrow A\Sigma_{ss}+\Sigma_{ss}A^T&=BB^T(A^{-1})^T+A^{-1}BB^T-QQ^T \text{ since $\dot{\Sigma}=0$ at equilibrium}.
+\end{align*}
+If we take the limit as $t$ goes to infinity, we find that 
+$$\lim_{t\to\infty}E[(y(t))]=\lim_{t\to\infty}(-A^{-1}(\exp(A(t-s))B|_{s=0}^{s=t}))=\lim_{t\to\infty}-A^{-1}(I-\exp(At))B=-A^{-1}B.$$
+Therefore, 
+\begin{align*}
+C&=\Sigma-(-A^{-1}B)(-A^{-1}B)^T
+\\&=\Sigma-A^{-1}BB^T(A^{-1})^T
+\\\Rightarrow AC+CA^T&=A\Sigma+\Sigma A^T-AA^{-1}BB^T(A^{-1})^T
+-A^{-1}BB^T(A^{-1})^TA^T
+\\&=A\Sigma+\Sigma A^T-BB^T(A^{-1})^T-A^{-1}BB^T
+\\&=-QQ^T.
 \end{align*}
@@
 
-\subsection{Calculating correlations \label{corr}}
-The following is derived from the work in \cite{Bialek:2013fk}.
+\subsection{Opinion dynamics with noise }
+Young et al.
+\cite{Young:2010fk} model opinion dynamics by assuming that each bird averages the differences between its opinion and each of the opinions of its neighbors and changes its opinion to minimize this difference and that each bird's opinion is subject to independent noise of unit intensity.
+The equations describing these dynamics can be written as
+\begin{equation}
+\dot{x_i}=\frac{\sum_{\text{neighbors $j$ of $i$ } }(x_j-x_i)}{K_i} +\xi_i=\sum_j\bar{A}_{ij}x_j-x_i+\xi_i=\sum_jL_{ij}x_j+\xi_i.
+\end{equation}
+In vector notation this becomes
+\begin{equation}
+\dot{x}=Lx+\xi.
+\end{equation}
+The equilibrium of these dynamics is a consensus state in which all birds have the some opinion, $x=\alpha\vec{1}$ for some $\alpha\in\R$.
+In the absence of any environmental signal, the consensus opinion can be any value $\alpha$.
+This means there is no single globally asymptotically stable equilibrium, so the variance of the opinions at steady state becomes difficult to analyze.
+However, we can rotate the opinion vector away from the consensus vector $\vec{1}$ to study deviations away from consensus.
+
+To rotate into the space orthogonal to the consensus vector, we can define a matrix $Q\in\R^{N-1}\times\R^N$ such that each row of $Q$ is orthogonal to $\vec{1}$ and the rows are pairwise orthogonal.
+Young et al.
+show in \cite{Young:2010fk} show that this equivalent to finding $Q$ such that $Q\vec{1}=\vec{0}$, $QQ^T=I_{N-1}$, and $Q^TQ+I_N-\frac{1}{N}\vec{1}_N\vec{1}_N^T$.
+If we then define $y=Qx$, $y$ is the component of $x$ orthogonal to the consensus vector.
+It follows that the dynamics of $y$ follow
 \begin{align*}
-H(\vec{v})&=-\vec{v}^TA\vec{v}-\vec{\beta}^T\vec{v}
-%\\&=\vec{v}A\vec{v}-\vec{\beta}^T\vec{v}+\vec{\beta}^T\vec{1}
+\dot{y}&=Q\dot{x}+Q\xi
+\\&=QLx+Q\xi
+\\&=QL\big(x-\frac{1}{N}\sum_kx_k\vec{1}\big)+Q\xi \text{ since } L\vec{1}=0
+\\&=QLQ^TQx+Q\xi
+\\&=QLQ^Ty+Q\xi
+\\&=\tilde{L}y+Q\xi.
 \end{align*}
-Define $V=\sum_kv_k/N$ and $\vec{y}=(\vec{v}-V\vec{1})/V$.  Then $\vec{v}=V(\vec{1}+\vec{y})$.  
+Now, $y=\vec{0}$ \emph{is} globally asymptotically stable, so analyzing its variance at steady state is a well-defined problem.
+Young et al.
+show in \cite{Young:2010fk} and we show below in Claim \ref{lyapeq_derivation} that if $\Sigma=\lim_{t\to\infty}E[y(t)y^T(t)]$, then $\Sigma$ satisfies $\tilde{L}\Sigma+\Sigma \tilde{L}^T=-I$.
+If we find such a $\Sigma$, then $\Tr(\Sigma)=\lim_{t\to\infty}\sum_iE[y_i(t)^2]\lim_{t\to\infty}=E[|y(t)|^2].$ If we define $\mathscr{H}=\Tr(\Sigma)$, then $H$ shows the expected distance between the opinion vector and the consensus vector at steady state.
+$\mathscr{H}$ is therefore a measure of the robustness of the consensus state to noise.
+
+In this paper, we consider the possibility of an environmental signal that drives the opinions to be equal to $1$. With this modification, the dynamics can be written
+\begin{equation}
+\dot{x}=Lx+B(\vec{1}-x)+\xi=L_\text{f}x+B\vec{1}+\xi
+\end{equation}
+In this case, $\vec{1}$ becomes a globally asymptotically stable equilibrium and we can find the steady state variance around this equilibrium without rotating into the subspace orthogonal to consensus.
+In particular, if $C_{ij}=\lim_{t\to\infty}E[(y_i(t)-\langle y_i(t)\rangle)(y_j(t)-\langle y_j(t)\rangle)]$, then the symmetric matrix $C$ satisfies $L_\text{f}C+CL_\text{f}^T=-I$ (Claim \ref{lyapeq_derivation}).
+
+\subsection{Probability distribution at equilibrium }
+A quite different way of modeling social interactions is to assume the opinions come to equilibrium and write down an expression for the probability of finding the opinion vector in every state, the method used by Bialek et al.
+in \cite{Bialek:2012fk}.
+In statistical physics, a Hamiltonian describes the energy of a given state, $H(v)$, and the Boltzmann distribution describes a probability distribution in which the most energetic states are least likely.
+Analogously, we can write a function $H(v)$ that gives a high energy when the opinions are not in consensus and a low energy when the opinions are in consensus and a probability distribution of opinion vectors that favors less energetic states.
+In order to do this, we must use a symmetric interaction matrix, rather than the possibly asymmetric $\bar{A}$ we used above.
+Bialek et al.
+study a model in which opinions are a three-dimensional velocity vector and the birds can weight the opinions of their nearest neighbors with any value $J$.
+In order to simplify this model and make it directly comparable to the opinion dynamics model we described above, we only consider a scalar-valued opinion $x_i$ and we impose the condition that the weights given to the neighbors' opinions are inversely proportional to the numb of neighbors.
+In particular, we define $S=\frac{1}{2}(A+A^T)$ and $P$ such that $P_{ij}=S_{ij}$ for $j\neq i$ and $P_{ii}=-\sum_kS_{ik}$.
+This allows us to write down a Hamiltonian
+\begin{equation}
+H(v)=-v^TPv
+\end{equation}
+where the negative sign makes opinions far from consensus give high energies and conversely.
+The probability distribution over opinion vectors $v$ then becomes $P(v)=1/Z\exp(-\frac{1}{2}H(v))$, where $Z=\int_\R^N\exp(-\frac{1}{2}H(v))dv$.
+We would like to use this probability distribution to find the covariance and correlation between the birds' opinion.
+
+As in the opinion dynamics framework, this requires some care when there is no environmental signal. If we write $P$ fully as 
+\begin{equation}
+P(v)=\frac{1}{Z}\exp\left(-\frac{1}{2}v^T(-P)v\right),
+\end{equation}
+we see that it is in the form of a multivariate Gaussian distribution.
+If $P$ is invertible, the covariance matrix would be given by $-P^{-1}$.
+However, since $P\vec{1}=\vec{0}$, one of the eigenvalues of $P$ is $0$ and it is therefore invertible.
+If we let $V=\sum_iv_i/N$ and $z=v-V\vec{1}$ so that $v=z+V\vec{1}$, then
 \begin{align*}
-H(\vec{y},V)&=-V^2(\vec{1}^T+\vec{y}^T)A(\vec{1}+\vec{y})-\vec{\beta}^TV(\vec{1}+\vec{y})
-\\&=-V^2(\vec{1}^T+\vec{y}^T)A\vec{y}-V\vec{\beta}^T\vec{y}-V\vec{\beta}^T\vec{1} \text{ since $A\vec{1}=\vec{0}$}
-\\&=-V^2\vec{y}^TA\vec{y}+(-V^2\vec{1}^TA-V\vec{\beta}^T)\vec{y}-V\vec{\beta}^T\vec{1}
+P(z,V)&=\frac{1}{Z}\exp\left(-\frac{1}{2}(z+V\vec{1})^T(-P)(z+V\vec{1})\right)
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}z^T(-P)z\right) \text{ since }P\vec{1}=P^T\vec{1}=\vec{0}.
 \end{align*}
-Define $Q\in M_{N-1,N}$ that rotates an $N$-vector away from the consensus vector and define $\vec{z}=Q\vec{y}$ so that $\vec{y}=Q^T\vec{z}$ since $\sum_ky_k=0$.
+To unpack this equation, we will write $z$ as a linear combination of eigenvectors of $P$.
+Since $P$ is symmetric and negative, there is an orthonormal set of eigenvectors $\{w_k\}$ such that $Pw_k=\lambda_kw_k$ with $w_1=\vec{1}$ and $\lambda_1=0$.
+If we let $W$ be a matrix whose columns are given by $\{w_k\}$ and $\Lambda$ be the diagonal matrix with $\{\lambda_k\}$ along the diagonal, then $W^T=W^{-1}$ and $P=W\Lambda W^T$ since $P$ is symmetric.
+If we write $z=\sum_kc_kw_k=W\vec{c}$, then $c_k=0$ since $\sum_iz_i=0$ implies that $z\cdot \vec{1}=0$.
+If we then write $\tilde{W}$ as the $N$ by $N-1$ matrix whose columns are given by $w_2,...w_N$,  $\tilde{\Lambda}$ be the diagonal matrix with $\lambda_2,\dots,\lambda_N$ along the diagonal, and $\tilde{P}=\tilde{W}\tilde{\Lambda}\tilde{W}^T$, then
 \begin{align*}
-H(\vec{z},V)&=-V^2\vec{z}^TQAQ^T\vec{z}+(-V^2\vec{1}^TAQ^T-V\vec{\beta}^TQ^T)\vec{z}-V\vec{\beta}^T\vec{1}
-\\&=V^2\vec{z}^TP\vec{z}+(-V^2\vec{\sigma}_1^T-V\vec{\sigma}_2^T)\vec{z}-V\vec{\beta}^T\vec{1}
-\\ \Rightarrow P(\vec{z},V)&=\frac{1}{\z}\exp\left(-V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+V\vec{\beta}^T\vec{1}\right)
+P(z)&=\frac{1}{Z}\exp\left(-\frac{1}{2}z^T(-\tilde{P})z\right).
 \end{align*}
-where $P=-QAQ^T$, $\vec{\sigma}_1=QA^T\vec{1}$, and $\vec{\sigma}_2=Q\vec{\beta}$.
+Note that $\tilde{W}^T(-\tilde{\Lambda}^{-1})\tilde{W}\tilde{P}=I$ so if we denote $\tilde{W}^T(-\tilde{\Lambda}^{-1})\tilde{W}$ as $\tilde{P}^{-1}$ then the covariance matrix of $z$ is given by $\tilde{P}^{-1}$.
+By dividing each entry $C_{ij}$ of the covariance matrix by $C_{ii}C_{jj}$, the covariance matrix can be used to find the correlation matrix.
 
-\begin{fact}
+The Hamiltonian framework, like the opinion dynamics framework, becomes easier to analyze if there is an environmental signal. In this case, the appropriate Hamiltonian is 
+\begin{equation}
+\begin{array}{lll}
+H(v)&=-(v-\vec{1})^TP_\text{f}(v-\vec{1})
+%\\&=-v^TP_\text{f}v+2v^TP_\text{f}\vec{1}-\vec{1}^TB\vec{1}
+%\\&=-v^TP_\text{f}v+2v^TB\vec{1}-\vec{1}^TB\vec{1}
+\end{array}
+\end{equation}
+since $P_\text{f}$ is symmetric and $P\vec{1}=\vec{0}$.
+It is clear that $H$ is minimized when $v=\vec{1}$.
+In this case, $P_\text{f}$ is invertible and $P_\text{f}^{-1}$ gives the covariance matrix of the opinions $v$, i.e.
+if $C=P_\text{f}^{-1}$, $C_{ij}=E[(v_i-\langle v_i\rangle)(v_j-\langle v_j\rangle)]$.
+
+We would like to compare the fluctuations around the average (consensus) opinion with and without the environmental signal, i.e.
+$E[(v_i-\frac{\sum_kv_k}{N})(v_j-\frac{\sum_kv_k}{N})]$.
+As above, let $V=\sum_iv_i/N$ and $z=v-V\vec{1}$ so that $v=z+V\vec{1}$.
 \begin{align*}
-\int\exp\left(-\frac{1}{2}\vec{z}^TP\vec{z}+\vec{\sigma}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{\det P}}\exp\left(-\frac{1}{2}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
-\\\Rightarrow \int\exp\left(-\vec{z}^TP\vec{z}+\vec{\sigma}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{2^{N-1}\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
+P(z,V)&=\frac{1}{Z}\exp\left(-\frac{1}{2}(v-\vec{1})^T(-P_\text{f})(v-\vec{1}\right)
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}(z-(1-V)\vec{1})^T(-P_\text{f})(z-(1-V)\vec{1}\right)
 \end{align*}
-\end{fact}
+Since the subspace of $z\in\R^N$ such that $\sum_iz_i=0$ is $N-1$ dimensional, it will be easier to rotate $z$ into $\R^{N-1}$ and integrate over all of $\R^{N-1}$.
+Above, we found a $Q\in\R^{N-1}\times\R^N$ such that each row of $Q$ was orthogonal to $\vec{1}$.
+Let $Q'\in\R^N\times\R^N$ be the matrix with $\frac{1}{\sqrt{N}}\vec{1}^T$ as its first row and the subsequent rows identical to the rows of $Q$.
+Let $\{q_1,\dots,q_N\}$ be the rows of $Q'$, so that $q_1=\vec{1}$.
+Since $Q'$ is orthonormal, it is invertible and we can write $y=(Q'^{-1})^Tz$.
+Then $Q'^Ty=z$ and $$0=\vec{1}^Tz=\vec{1}^TQ'^Ty=\sum_{k=1}^N\vec{1}^Tq_ky_k=y_1.$$
+Therefore, 
+$$\Omega=\{z\in\R^N:\sum_iz_i=0\}=\{z\in\R^N:z\cdot\vec{1}=0\}=\{z\in\R^N:\exists \ y\in\R^N \text{ such that } z=Q^Ty \text{ and } y_1=0\}.$$
+Further, if $e_1=(1,0,\dots,0)^T$, $Q'^Te_1=q_1=\frac{1}{\sqrt{N}}\vec{1}$ so that $\vec{1}=\sqrt{N}Q'^Te_1$.
+Now, this allows us to rewrite
+\begin{align*}
+P(z,V)&=\frac{1}{Z}\exp\left(-\frac{1}{2}(Q'^Ty-(1-V)\sqrt{N}Q'^Te_1)^T(-P_\text{f})(Q'^Ty-(1-V)\sqrt{N}Q'^Te_1)\right)
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}(y-(1-V)\sqrt{N}e_1)^T(-Q'P_\text{f}Q'^T)(y-(1-V)\sqrt{N}e_1)\right)
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}(y-(1-V)\sqrt{N}e_1)^TR(y-(1-V)\sqrt{N}e_1)\right)
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}(y+(V-1)\sqrt{N}e_1)^TR(y+(V-1)\sqrt{N}e_1)\right)
+\end{align*}
+where $R=-Q'P_\text{f}Q'^T$ and $y\in\R^N$ is such that $y_1=0$. To find $P(z)$ we can integrate out $V$:
+\begin{align*}
+P(z)&=\int_\R P(z,V)dV
+\\&=\int_\R \frac{1}{Z}\exp\left(-\frac{1}{2}((V-1)\sqrt{N}e_1+y)^TR((V-1)\sqrt{N}e_1+y)\right)dV
+\\&=\frac{1}{Z}\exp(-\frac{1}{2}y^TRy)\int_\R \exp\left(-\frac{1}{2}\left(R_{11}N\left(V-1\right)^2+(V-1)\sqrt{N}\sum_iR_{1i}y_i+(V-1)\sqrt{N}\sum_iy_iR_{i1}\right)\right)dV
+\\&=\frac{1}{Z}\exp(-\frac{1}{2}y^TRy)\int_\R \exp\left(-\frac{1}{2}\left(R_{11}N\left(V-1\right)^2+2(V-1)\sqrt{N}\sum_iR_{1i}y_i\right)\right)dV \text{ since $R$ is symmetric }
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^TRy-\left(\sum_iR_{1i}y_i/\sqrt{R_{11}}\right)^2\right)\right)\times
+\\&\int_\R \exp\left(-\frac{1}{2}\left(R_{11}N\left(V-1\right)^2+2(V-1)\sqrt{N}\sum_iR_{1i}y_i+\left(\sum_iR_{1i}y_i/\sqrt{R_{11}}\right)^2\right)\right)dV
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^TRy-\left(\sum_iR_{1i}y_i/\sqrt{R_{11}}\right)^2\right)\right)\int_\R \exp\left(-\frac{1}{2}\left(\sqrt{R_{11}}\sqrt{N}\left(V-1\right)+\sum_iR_{1i}y_i/\sqrt{R_{11}}\right)^2\right)dV
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^TRy-\left(\sum_iR_{1i}y_i/\sqrt{R_{11}}\right)^2\right)\right)\int_\R \exp\left(-\frac{1}{2}R_{11}N\left(\left(V-1\right)+\frac{\sum_iR_{1i}y_i}{R_{11}N}\right)^2\right)dV
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^TRy-\left(\sum_iR_{1i}y_i/\sqrt{R_{11}}\right)^2\right)\right)\times\sqrt{2\pi}\sqrt{R_{11}N}
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^TRy-y^TR_{\cdot 1}R_{1\cdot}y\right)\right)\times\sqrt{2\pi}\sqrt{R_{11}N} \text{ where $R_{1\cdot}$ is the first row of $R$}
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^T(R-\frac{1}{R_{11}}R_{\cdot 1}R_{1\cdot})y\right)\right)\times\sqrt{2\pi}\sqrt{R_{11}N}
+\\&=\frac{\sqrt{2\pi}\sqrt{R_{11}N}}{(2\pi)^{N/2}\det(-P_\text{f}^{-1})^{1/2}}\exp\left(-\frac{1}{2}\left(y^T(R-\frac{1}{R_{11}}R_{\cdot 1}R_{1\cdot})y\right)\right)
+\\&=\frac{\sqrt{2\pi}\sqrt{R_{11}N}}{(2\pi)^{N/2}\det(-P_\text{f}^{-1})^{1/2}}\exp\left(-\frac{1}{2}z^TQ'^{-1}(R-\frac{1}{R_{11}}R_{\cdot 1}R_{1\cdot})(Q'^{-1})^Tz\right)
+\\&=\frac{\sqrt{2\pi}\sqrt{R_{11}N}}{(2\pi)^{N/2}\det(-P_\text{f}^{-1})^{1/2}}\exp\left(-\frac{1}{2}z^T(-P_\text{f}-\frac{1}{R_{11}}Q'^{-1}R_{\cdot 1}R_{1\cdot}(Q'^{-1})^T)z\right)
+%\\&=\frac{\sqrt{2\pi}\sqrt{R_{11}N}}{(2\pi)^{N/2}\det(-P_\text{f}^{-1})^{1/2}}\exp\left(-\frac{1}{2}z^T(-P_\text{f}-\frac{1}{R_{11}}Q'^{-1}R_{\cdot 1}R_{1\cdot}(Q'^{-1})^T)z\right)
+\\&=\frac{\sqrt{2\pi}\sqrt{R_{11}N}}{(2\pi)^{N/2}\det(-P_\text{f}^{-1})^{1/2}}\exp\left(-\frac{1}{2}z^T(-P_\text{f}-\frac{1}{R_{11}}P_\text{f}Q'^T_{\cdot 1}Q'_{1\cdot}P_\text{f})z\right)
+\\&=\frac{\sqrt{2\pi}\sqrt{R_{11}N}}{(2\pi)^{N/2}\det(-P_\text{f}^{-1})^{1/2}}\exp\left(-\frac{1}{2}z^T(-P_\text{f}-\frac{1}{R_{11}}P_\text{f}(Q'_{1\cdot})^TQ'_{1\cdot}P_\text{f})z\right)
+%\\&=\frac{\sqrt{2\pi}\sqrt{R_{11}N}}{(2\pi)^{N/2}\det(-P_\text{f}^{-1})^{1/2}}\exp\left(-\frac{1}{2}z^T(-P_\text{f}-\frac{1}{R_{11}}P_\text{f}\Pi P_\text{f})z\right) \text{ where $\Pi_{ij}=1/N$ for all $i,j$}
+\\&=\frac{\sqrt{2\pi}\sqrt{R_{11}N}}{(2\pi)^{N/2}\det(-P_\text{f}^{-1})^{1/2}}\exp\left(-\frac{1}{2}z^T(-P_\text{f}-\frac{1}{R_{11}N}P_\text{f}\vec{1}\vec{1}^T P_\text{f})z\right) 
+\\&=\frac{\sqrt{2\pi}\sqrt{R_{11}N}}{(2\pi)^{N/2}\det(-P_\text{f}^{-1})^{1/2}}\exp\left(-\frac{1}{2}z^T(-P_\text{f}-\frac{1}{R_{11}N}B\vec{1}\vec{1}^T B)z\right)  \text{ since $P\vec{1}=\vec{0}$}
+\end{align*}
+By definition of $R$, $R_{11}=\sum_{kl}Q_{1k}P_{\text{f}kl}Q_{1l}$.
+Let $N_\text{I}$ be the number of birds that can perceive the signal, i.e.
+$N_\text{I}$ is the number of non-zero elements along the diagonal of $B$.
+Then, $R_{11}=\frac{b}{N}N_\text{I}$.
+Finally, we find that 
+\begin{equation}
+P(z)\propto \exp\left(-\frac{1}{2}z^T(-P_\text{f}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB)z\right)
+\end{equation}
+Now, let $P'=P_\text{f}+\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB$.
+Then $P'\vec{1}=-B\vec{1}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB\vec{1}=-B\vec{1}-\frac{bN_\text{I}}{bN_\text{I}}B\vec{1}=\vec{0}$, so we cannot invert $-P'$ to find the covariance matrix.
+However, as above, we can discard the $0$ eigenvector since we know that $z\cdot \vec{1}=0$.
+This gives us $E[(y_i-\frac{\sum_ky_k}{N})(y_j-\frac{\sum_ky_k}{N})]$.
 
+\subsection{Connecting two frameworks }
+To find $\mathscr{H}$, we found the matrix $\Sigma$ such that 
+\begin{equation}
+QLQ^T\Sigma+\Sigma QLQ^T=-I.
+\end{equation}
+If the vector $v$ is described by the dynamics $\dot{v}=Dv$, with $H(v)=-v^TPv$, then
 \begin{align*}
-\z&=\int_{-\infty}^\infty\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+V\vec{\beta}^T\vec{1}\right)d^NzdV
-\\&=\int_{-\infty}^\infty \exp(V\vec{\beta}^T\vec{1})\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}\right)d^NzdV
-\\&=\sqrt{\pi^{N-1}}\int_{-\infty}^\infty\exp(V\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}\det{P}}}\right)\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det PP}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\exp(V\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}}}\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\Pi_a\lambda_a}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\\&\text{ where $\lambda_a$ are the eigenvalues of $P$}
-\\ \Rightarrow -\log(\z)&=\text{constants}+\frac{1}{2}\sum_a\log(\lambda_a)+\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma_1}-\log(f(P))
+\dot{H}&=-\dot{v}^TPv-v^TP\dot{v}
+\\&=-v^TD^TPv-v^TPDv
+\\&=-v^T(D^TP+PD)v.
 \end{align*}
+Therefore, $H(v)$ decreases at all $v$ if and only if $D^TP+PD$ is a positive matrix.
+%MORE?!?
 
-where $f(P)=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV $.
-
-\begin{align*}
-\Rightarrow\frac{\partial -\log(\z)}{\partial P_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{1l}-\frac{\partial \log f}{\partial P_{ij}}
-\\\Rightarrow\frac{\partial -\log(\z)}{\partial P_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}{ki}P^{-1}_{jl}\sigma_{1l}-\frac{1}{f}\frac{\partial f}{\partial P_{ij}}
-\end{align*}
-
-\begin{align*}
-\frac{\partial f}{\partial P_{ij}}&=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)\left(-\frac{1}{4V}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)dV
-\\&=\left(-\frac{1}{4}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\end{align*}
-
-\small{
-\begin{align*}
-\Rightarrow\langle z_iz_j\rangle &=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}{ki}P^{-1}_{jl}\sigma_{1l}+\left(\frac{1}{4}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)\frac{\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV}{\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV}
-\end{align*}}
-
-\subsection{Equivalence of dynamics and Hamiltonian }
-\begin{claim}
-If $H(\vec{v})=-\vec{v}^TA\vec{v}-\vec{\beta}^T\vec{v}$ and $\frac{d\vec{v}}{dt}=A\vec{v}+\vec{\beta}$ then $\nabla H\cdot \frac{d\vec{v}}{dt}\leq 0$.
-\end{claim}
-\begin{pf}
-\begin{align*}
-H(\vec{v})&=-\sum_{ij}A_{ij}v_iv_j-\sum_i\beta_iv_i
-\\\Rightarrow\frac{\partial H}{\partial v_i}&=-\sum_{j\neq i}A_{ij}v_j-\sum_{j\neq i}A_{ji}v_j-2A_{ii}v_i-\beta_i
-\\&=-\sum_j(A_{ij}+A^T_{ij})v_j-\beta_i
-\\ \Rightarrow \nabla H&=-(A+A^T)\vec{v}-\vec{\beta}
-\\ \Rightarrow \nabla H\cdot \frac{d\vec{v}}{dt}&=-\vec{v}^T(A+A^T)A\vec{v}-\vec{v}^T(A+A^T)\vec{\beta}-\vec{\beta}^TA\vec{v}-\vec{\beta}^T\vec{\beta}
-\\&=-\vec{v}^TA^2\vec{v}-\vec{v}^TA^TA\vec{v}-\vec{v}^TA\vec{\beta}-2\vec{v}^TA^T\vec{\beta}-\vec{\beta}^T\vec{\beta}
-\end{align*}
-If $A$ is negative semi-definite, then $A^2$ is positive semi-definite and $A^TA$ is positive semi-definite.
+%Specifically, an individual's opinion $\sigma_i$ can be either $1$ or $-1$ and $\sigma_i$ switches with probability $w_i(\sigma_i)=\frac{1}{2}(-\sigma_i\sum_jA_{ij}\sigma_j )(1-\beta_i\sigma_i)$ where $A_{ij}$ indicates the strength of the connection from $j$ to $i$ and $\beta_i$ indicates the strength of the environmental stimulus perceived by $i$.
+The stochastic dynamics of the opinions can be written down with a master equation \cite{Glauber:1963fk}.
+The vector $\vec{v}(t)$ of expected values, $q_i(t)=\langle \sigma_i(t)\rangle$, satisfies the following differential equations: \cite{Glauber:1963fk}
+%\begin{equation*}
+%\frac{d\vec{v}}{dt}=M\vec{v}+\vec{\beta}^T(\vec{1}-\vec{v}).
+%\end{equation*}
+%We simplify this by writing $\frac{d\vec{v}}{dt}=A\vec{v}+\vec{\beta}^T\vec{1}$, where $A=M-B$ and $B$ is a diagonal matrix with the values of $\vec{\beta}$ along the diagonal. 
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%\subsection{H2 Norm \label{H2}} 
+%
+%\ Let $y$ represent the distance between the vector of opinions and consensus, i.e.
+%$y=QxQ^T,$
+%where $Q\in\R^{n-1\times n}$ is such that $Q\vec{1}=0$ , $QQ^T=I_{n-1}$, and $Q^TQ=I_n-\frac{1}{n}\vec{1}\times\vec{1}^T$.   
+%Then 
+%\begin{equation}
+%\dot{y}(t)=-\overline{L}y(t)+Q\xi(t) \notag
+%\end{equation}
+%where $\overline{L}=QLQ^T$.  As shown in \cite{Young:2010fk}, if $\Sigma_y(t)=\E[y(t)y(t)^T]$,  $\Sigma_y$ at equilibrium solves
+%\begin{equation} 0=-\overline{L}\Sigma_y-\Sigma_y\overline{L}^T+D. \label{h2norm}
+%\end{equation}
+%\begin{align*}
+%\mathscr{H}_2&=\sqrt{\Tr(\Sigma_y)}
+%\\\mathscr{H}_2 \text{ robustness} &=\frac{1}{\mathscr{H}_2}
+%\end{align*}
+%
+%\begin{claim}
+%Given any $Q>0$ there exists a unique $\Sigma>0$ satisfyying $A\Sigma+\Sigma A^T=Z$ if and only if the system $\dot{x}=Ax$ is  global asymptotically stable.
+%\end{claim}
+%
+%\begin{claim}
+%Consider the dynamical system,
+%\begin{equation}
+%\dot{y}=Ay+B+C\xi  \label{tofindsigma}
+%\end{equation}
+%with $E[\xi(t)]=0$, $E[\xi(t)\xi(s)^T]=\frac{1}{2}\delta(t-s)I$, and $E[y(0)\xi^T(t)]=0$.
+If all eigenvalues of $A$ are negative and if $\Sigma(t)=E[y(t)y(t)^T]$ and $\Sigma_{ss}=\lim_{t\to\infty}\Sigma(t)$, then $\Sigma_{ss}$ satisfies the Lyapunov equation
+%\begin{equation*}
+%A\Sigma_{ss}+\Sigma_{ss}A^T=A^{-1}BB^T+BB^T(A^{-1})^T-CC^T
+%\end{equation*}
+%\end{claim}
+%
+%\begin{pf}
+%\begin{align*}
+%\Sigma(t)&=E[y(t)y(t)^T]
+%\\ \Rightarrow \dot{\Sigma}&=E[\dot{y}(t)y(t)^T+y(t)\dot{y}(t)^T]
+%\\&=E[(Ay+B+C\xi)y^T+y(y^TA^T+B^T+\xi^TC^T)]
+%\\&=AE[yy^T]+E[yy^T]A^T+BE[y^T]+E[y]B^T+CE[\xi y^T]+E[y\xi^T]C^T
+%\end{align*}
+%Note that we can write a solution to (\ref{tofindsigma}) as 
+%\begin{equation*}
+%y(t)=\exp(At)y(0)+\int_0^t\exp(A(t-s))Bds+\int_0^t\exp(A(t-s))C\xi(s)ds.
+%\end{equation*}
+%This allows us to write $E[y]=\exp(At)y(0)+\int_0^t\exp(A(t-s))Bds$ and 
+%\begin{align*}
+%E[y \xi^T]&=\exp(At)E[y(0)\xi(t)^T]+\int_0^t\exp(A(t-s))BE[\xi(t)^T]ds+\int_0^t\exp(A(t-s))CE[\xi(s)\xi(t)]ds
+%\\&=0+0+\frac{1}{2}C
+%\\ \Rightarrow \dot{\Sigma}&=A\Sigma(t)+\Sigma(t)A^T+By(0)^T\exp(At)^T+\exp(At)y(0)B^T
+%\\&+B\int_0^tB^T\exp(A(t-s))^Tds+\int_0^T\exp(A(t-s))BdsB^T+\frac{1}{2}CC^T+\frac{1}{2}CC^T
+%\\&=A\Sigma(t)+\Sigma(t)A^T+By(0)^T\exp(At)^T+\exp(At)y(0)B^T
+%\\&-BB^T(A^{-1})^T\exp(A(t-s))|_{s=0}^{s=t}-A^{-1}\exp(A(t-s))|_{s=0}^tBB^T+CC^T
+%\\&=A\Sigma(t)+\Sigma(t)A^T+By(0)^T\exp(At)^T+\exp(At)y(0)B^T
+%\\&-BB^T(A^{-1})^T(I-\exp(At))-A^{-1}(I-\exp(At))BB^T+CC^T
+%\\ \lim_{t\to\infty}\dot{\Sigma}&=A\Sigma_{ss}+\Sigma_{ss}A^T-BB^T(A^{-1})^T-A^{-1}BB^T+CC^T \text{ since all eigenvalues of $A$ are negative}
+%\\\Rightarrow A\Sigma_{ss}+\Sigma_{ss}A^T&=BB^T(A^{-1})^T+A^{-1}BB^T-CC^T \text{ since $\dot{\Sigma}=0$ at equilibrium}
+%\end{align*}
+%\end{pf}
+%
+%
+%\subsection{Calculating correlations \label{corr}}
+%The following is derived from the work in \cite{Bialek:2013fk}.
+%\begin{align*}
+%H(\vec{v})&=-\vec{v}^TA\vec{v}-\vec{\beta}^T\vec{v}
+%%\\&=\vec{v}A\vec{v}-\vec{\beta}^T\vec{v}+\vec{\beta}^T\vec{1}
+%\end{align*}
+%Define $V=\sum_kv_k/N$ and $\vec{y}=(\vec{v}-V\vec{1})/V$.  Then $\vec{v}=V(\vec{1}+\vec{y})$.  
+%\begin{align*}
+%H(\vec{y},V)&=-V^2(\vec{1}^T+\vec{y}^T)A(\vec{1}+\vec{y})-\vec{\beta}^TV(\vec{1}+\vec{y})
+%\\&=-V^2(\vec{1}^T+\vec{y}^T)A\vec{y}-V\vec{\beta}^T\vec{y}-V\vec{\beta}^T\vec{1} \text{ since $A\vec{1}=\vec{0}$}
+%\\&=-V^2\vec{y}^TA\vec{y}+(-V^2\vec{1}^TA-V\vec{\beta}^T)\vec{y}-V\vec{\beta}^T\vec{1}
+%\end{align*}
+%Define $Q\in M_{N-1,N}$ that rotates an $N$-vector away from the consensus vector and define $\vec{z}=Q\vec{y}$ so that $\vec{y}=Q^T\vec{z}$ since $\sum_ky_k=0$.
+%\begin{align*}
+%H(\vec{z},V)&=-V^2\vec{z}^TQAQ^T\vec{z}+(-V^2\vec{1}^TAQ^T-V\vec{\beta}^TQ^T)\vec{z}-V\vec{\beta}^T\vec{1}
+%\\&=V^2\vec{z}^TP\vec{z}+(-V^2\vec{\sigma}_1^T-V\vec{\sigma}_2^T)\vec{z}-V\vec{\beta}^T\vec{1}
+%\\ \Rightarrow P(\vec{z},V)&=\frac{1}{\z}\exp\left(-V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+V\vec{\beta}^T\vec{1}\right)
+%\end{align*}
+%where $P=-QAQ^T$, $\vec{\sigma}_1=QA^T\vec{1}$, and $\vec{\sigma}_2=Q\vec{\beta}$.
+%
+%
+%\begin{fact}
+%\begin{align*}
+%\int\exp\left(-\frac{1}{2}\vec{z}^TP\vec{z}+\vec{\sigma}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{\det P}}\exp\left(-\frac{1}{2}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
+%\\\Rightarrow \int\exp\left(-\vec{z}^TP\vec{z}+\vec{\sigma}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{2^{N-1}\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
+%\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
+%\end{align*}
+%\end{fact}
+%
+%\begin{align*}
+%\z&=\int_{-\infty}^\infty\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+V\vec{\beta}^T\vec{1}\right)d^NzdV
+%\\&=\int_{-\infty}^\infty \exp(V\vec{\beta}^T\vec{1})\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}\right)d^NzdV
+%\\&=\sqrt{\pi^{N-1}}\int_{-\infty}^\infty\exp(V\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}\det{P}}}\right)\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
+%\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det PP}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\exp(V\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}}}\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
+%\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
+%\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\Pi_a\lambda_a}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
+%\\&\text{ where $\lambda_a$ are the eigenvalues of $P$}
+%\\ \Rightarrow -\log(\z)&=\text{constants}+\frac{1}{2}\sum_a\log(\lambda_a)+\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma_1}-\log(f(P))
+%\end{align*}
+%
+%where $f(P)=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV $.
+%
+%\begin{align*}
+%\Rightarrow\frac{\partial -\log(\z)}{\partial P_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{1l}-\frac{\partial \log f}{\partial P_{ij}}
+%\\\Rightarrow\frac{\partial -\log(\z)}{\partial P_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}{ki}P^{-1}_{jl}\sigma_{1l}-\frac{1}{f}\frac{\partial f}{\partial P_{ij}}
+%\end{align*}
+%
+%\begin{align*}
+%\frac{\partial f}{\partial P_{ij}}&=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)\left(-\frac{1}{4V}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)dV
+%\\&=\left(-\frac{1}{4}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
+%\end{align*}
+%
+%\small{
+%\begin{align*}
+%\Rightarrow\langle z_iz_j\rangle &=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}{ki}P^{-1}_{jl}\sigma_{1l}+\left(\frac{1}{4}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)\frac{\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV}{\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV}
+%\end{align*}}
+%
+%
+%\subsection{Equivalence of dynamics and Hamiltonian }
+%\begin{claim}
+%If $H(\vec{v})=-\vec{v}^TA\vec{v}-\vec{\beta}^T\vec{v}$ and $\frac{d\vec{v}}{dt}=A\vec{v}+\vec{\beta}$ then $\nabla H\cdot \frac{d\vec{v}}{dt}\leq 0$.
+%\end{claim}
+%\begin{pf}
+%\begin{align*}
+%H(\vec{v})&=-\sum_{ij}A_{ij}v_iv_j-\sum_i\beta_iv_i
+%\\\Rightarrow\frac{\partial H}{\partial v_i}&=-\sum_{j\neq i}A_{ij}v_j-\sum_{j\neq i}A_{ji}v_j-2A_{ii}v_i-\beta_i
+%\\&=-\sum_j(A_{ij}+A^T_{ij})v_j-\beta_i
+%\\ \Rightarrow \nabla H&=-(A+A^T)\vec{v}-\vec{\beta}
+%\\ \Rightarrow \nabla H\cdot \frac{d\vec{v}}{dt}&=-\vec{v}^T(A+A^T)A\vec{v}-\vec{v}^T(A+A^T)\vec{\beta}-\vec{\beta}^TA\vec{v}-\vec{\beta}^T\vec{\beta}
+%\\&=-\vec{v}^TA^2\vec{v}-\vec{v}^TA^TA\vec{v}-\vec{v}^TA\vec{\beta}-2\vec{v}^TA^T\vec{\beta}-\vec{\beta}^T\vec{\beta}
+%\end{align*}
+%If $A$ is negative semi-definite, then $A^2$ is positive semi-definite and $A^TA$ is positive semi-definite.
 (Consider that $\vec{v}^TA^TA\vec{v}=(A\vec{v})^T(A\vec{v})$ so that $\vec{v}A^TA\vec{v}=0$ if $A\vec{v}=0$ and $\vec{v}^TA^TA\vec{v}<0$ otherwise.) Therefore, $-\vec{v}^TA^2\vec{v}\leq 0$ and $-\vec{v}^TA^TA\vec{v}\leq 0$.
 At equilibrium, $\vec{v}$ will be ``close" to $\vec{\beta}$ so that $$-\vec{v}^TA\vec{\beta}-2\vec{v}^TA^T\vec{\beta}\sim-3\vec{\beta}^TA\vec{\beta}$$
-\end{pf}
-
-\subsection{Numerically evaluating dynamics }
-%\begin{claim}
-Let $\Phi(t)$ be the fundamental matrix solution to the homogeneous differential equations $\frac{d\vec{x}}{dt}=A\vec{x}$ so that $\Phi(t)=V\Lambda(t)$ where the columns of $V$ are the eigenvectors of $A$ and $$\Lambda(t)=\text{diag}(e^{\lambda_1 t},\dots,e^{\lambda_n t})$$ where $\{\lambda_i\}$ are the eigenvalues of $A$.
+%\end{pf}
+%
+%\subsection{Numerically evaluating dynamics }
+%%\begin{claim}
+%Let $\Phi(t)$ be the fundamental matrix solution to the homogeneous differential equations $\frac{d\vec{x}}{dt}=A\vec{x}$ so that $\Phi(t)=V\Lambda(t)$ where the columns of $V$ are the eigenvectors of $A$ and $$\Lambda(t)=\text{diag}(e^{\lambda_1 t},\dots,e^{\lambda_n t})$$ where $\{\lambda_i\}$ are the eigenvalues of $A$.
 (Note that if $\lambda_i=0$ then $e^{\lambda _i t}=1$ for all $t$.) Let $V^\dagger$ be the pseudoinverse of $V$.
@@
 Let $$G(t)=\text{diag}(1/\lambda_1(1-e^{-\lambda_1t}),\dots,1/\lambda_n(1-e^{-\lambda_nt}))V^\dagger\vec{\beta}.$$
-(If $\lambda_i=0$ then the corresponding entry of the first part of $G(t)$ will be given by $t$ rather than the form above.)  
-\begin{claim}
-If we let $x(t)=\Phi(t)G(t)+\Phi(t)\vec{c}$, $x(t)$ solves the inhomogeneous equations $\frac{dx}{dt}=A\vec{x}+\vec{\beta}$ with initial conditions $x(0)=\vec{X_0}$.
-\end{claim}
-
-\begin{pf}
-It is clear that
-\begin{align*}
-\frac{d\Phi\vec{c}}{dt}&=A\Phi(t)\vec{c} \text{ and } \Phi(0)\vec{c}=\vec{X_0}.
-\end{align*}
-Now, 
-\begin{align*}
-\Phi(t)G(t)&=V
-\left(\begin{array}{ccccc}
-e^{\lambda_1 t}  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots & e^{\lambda_nt}
- \end{array}\right)
- \left(\begin{array}{ccccc}
-\frac{1}{\lambda_1}(1-e^{\lambda_1 t})  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots &\frac{1}{\lambda_n} (1-e^{\lambda_nt})
- \end{array}\right)
- V^\dagger\vec{\beta}
- \\&=V
- \left(\begin{array}{ccccc}
-\frac{1}{\lambda_1}(e^{\lambda_1 t}-1)  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots &\frac{1}{\lambda_n} (e^{\lambda_nt}-1)
- \end{array}\right)
- V^\dagger\vec{\beta}
- \\ \Rightarrow \frac{d\Phi(t)G(t)}{dt}&=V
- \left(\begin{array}{ccccc}
-e^{\lambda_1 t}  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots &e^{\lambda_nt}
- \end{array}\right)V^\dagger\vec{\beta}
- \\&=V
- \left(\begin{array}{ccccc}
-e^{\lambda_1 t}  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots &e^{\lambda_nt}
- \end{array}\right)
- V^\dagger\vec{\beta} -VV^\dagger\vec{\beta}+\vec{\beta}
- \\&=V
- \left(\begin{array}{ccccc}
-e^{\lambda_1 t}-1  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots &e^{\lambda_nt}-1
- \end{array}\right)
- V^\dagger\vec{\beta} +\vec{\beta}
-  \\&=V
- \left(\begin{array}{ccccc}
-\lambda_1e^{\lambda_1t}\cdot \frac{1}{\lambda_1}(1-e^{-\lambda_1 t})  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots &\lambda_ne^{\lambda_nt}\cdot\frac{1}{\lambda_n}(1-e^{-\lambda_nt})
- \end{array}\right)
- V^\dagger\vec{\beta} +\vec{\beta}
-  \\&=AV\Lambda(t)
- \left(\begin{array}{ccccc}
- \frac{1}{\lambda_1}(1-e^{-\lambda_1 t})  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots &\frac{1}{\lambda_n}(1-e^{-\lambda_nt})
- \end{array}\right)
- V^\dagger\vec{\beta} +\vec{\beta}
- \\&=AV\Lambda(t)G(t)+\vec{\beta}
- \\&=A\Phi(t)G(t)+\vec{\beta}
-\end{align*}
-Therefore $\frac{dx}{dt}=A\Phi(t)G(t)+A\Phi(t)\vec{c}+\vec{\beta}=Ax(t)+\vec{\beta}$ and since $G(0)=\vec{0}$, $x(0)=\vec{X_0}$.
-\end{pf}
+%(If $\lambda_i=0$ then the corresponding entry of the first part of $G(t)$ will be given by $t$ rather than the form above.)  
+%\begin{claim}
+%If we let $x(t)=\Phi(t)G(t)+\Phi(t)\vec{c}$, $x(t)$ solves the inhomogeneous equations $\frac{dx}{dt}=A\vec{x}+\vec{\beta}$ with initial conditions $x(0)=\vec{X_0}$.
+%\end{claim}
+%
+%\begin{pf}
+%It is clear that
+%\begin{align*}
+%\frac{d\Phi\vec{c}}{dt}&=A\Phi(t)\vec{c} \text{ and } \Phi(0)\vec{c}=\vec{X_0}.
+%\end{align*}
+%Now, 
+%\begin{align*}
+%\Phi(t)G(t)&=V
+%\left(\begin{array}{ccccc}
+%e^{\lambda_1 t}  &\dots &0
+%\\ \vdots & \ddots &\vdots
+%\\ 0 & \dots & e^{\lambda_nt}
+% \end{array}\right)
+% \left(\begin{array}{ccccc}
+%\frac{1}{\lambda_1}(1-e^{\lambda_1 t})  &\dots &0
+%\\ \vdots & \ddots &\vdots
+%\\ 0 & \dots &\frac{1}{\lambda_n} (1-e^{\lambda_nt})
+% \end{array}\right)
+% V^\dagger\vec{\beta}
+% \\&=V
+% \left(\begin{array}{ccccc}
+%\frac{1}{\lambda_1}(e^{\lambda_1 t}-1)  &\dots &0
+%\\ \vdots & \ddots &\vdots
+%\\ 0 & \dots &\frac{1}{\lambda_n} (e^{\lambda_nt}-1)
+% \end{array}\right)
+% V^\dagger\vec{\beta}
+% \\ \Rightarrow \frac{d\Phi(t)G(t)}{dt}&=V
+% \left(\begin{array}{ccccc}
+%e^{\lambda_1 t}  &\dots &0
+%\\ \vdots & \ddots &\vdots
+%\\ 0 & \dots &e^{\lambda_nt}
+% \end{array}\right)V^\dagger\vec{\beta}
+% \\&=V
+% \left(\begin{array}{ccccc}
+%e^{\lambda_1 t}  &\dots &0
+%\\ \vdots & \ddots &\vdots
+%\\ 0 & \dots &e^{\lambda_nt}
+% \end{array}\right)
+% V^\dagger\vec{\beta} -VV^\dagger\vec{\beta}+\vec{\beta}
+% \\&=V
+% \left(\begin{array}{ccccc}
+%e^{\lambda_1 t}-1  &\dots &0
+%\\ \vdots & \ddots &\vdots
+%\\ 0 & \dots &e^{\lambda_nt}-1
+% \end{array}\right)
+% V^\dagger\vec{\beta} +\vec{\beta}
+%  \\&=V
+% \left(\begin{array}{ccccc}
+%\lambda_1e^{\lambda_1t}\cdot \frac{1}{\lambda_1}(1-e^{-\lambda_1 t})  &\dots &0
+%\\ \vdots & \ddots &\vdots
+%\\ 0 & \dots &\lambda_ne^{\lambda_nt}\cdot\frac{1}{\lambda_n}(1-e^{-\lambda_nt})
+% \end{array}\right)
+% V^\dagger\vec{\beta} +\vec{\beta}
+%  \\&=AV\Lambda(t)
+% \left(\begin{array}{ccccc}
+% \frac{1}{\lambda_1}(1-e^{-\lambda_1 t})  &\dots &0
+%\\ \vdots & \ddots &\vdots
+%\\ 0 & \dots &\frac{1}{\lambda_n}(1-e^{-\lambda_nt})
+% \end{array}\right)
+% V^\dagger\vec{\beta} +\vec{\beta}
+% \\&=AV\Lambda(t)G(t)+\vec{\beta}
+% \\&=A\Phi(t)G(t)+\vec{\beta}
+%\end{align*}
+%Therefore $\frac{dx}{dt}=A\Phi(t)G(t)+A\Phi(t)\vec{c}+\vec{\beta}=Ax(t)+\vec{\beta}$ and since $G(0)=\vec{0}$, $x(0)=\vec{X_0}$.
+%\end{pf}
+\begin{table}
+\caption{ \label{matrices} Table of matrices used in the text.}
+\ra{1.3}
+\begin{tabular}{@{}lllll@{}}
+$A$ & adjacency matrix & $A_{ij}=\left\{\begin{array}{l l l}
+1 & \text{ if $i$ pays attention to $j$ }
+\\0 & \text{ otherwise }
+\end{array}\right.$
+\\$\bar{A}$ & normalized adjacency matrix & $\bar{A}_{ij}=\left\{\begin{array}{l l l }
+\frac{1}{K_i} & \text{ if $i$ pays attention to $j$, where $K_i=$ \# of birds $i$ pays attention to}
+\\ 0 & \text{ otherwise }
+\end{array}\right. $
+\\$B$ & diagonal signal matrix & $B_{ij}=\left\{\begin{array}{l l l}
+b & \text{ if $i=j$ and $i$ perceives the environmental signal }
+\\0 & \text{ otherwise }
+\end{array}\right. $
+\\$L$ & opinion dynamics w/o environmental signal &$L=\bar{A}-I$
+\\$L_f$ & opinion dynamics with environmental signal & $L_\text{f}=L-B$
+\\$\tilde{L}$ & dynamics of deviations from consensus & $\tilde{L}=QLQ^T$
+\\$S$ & symmetrized interaction matrix & $S=\frac{1}{2}(\bar{A}+\bar{A}^T)$
+\\$P$ & energy interactions w/o environmental signal & $P_{ij}=\left\{\begin{array}{lll}
+S_{ij} & \text{ for } j\neq i
+\\-\sum_kS_{ik} & \text{ for } j=i
+\end{array}\right.
+$
+\\$P_\text{f}$ & energy interactions with environmental signal &  $P_\text{f}=P-B$
+\\$Q$ & rotates opinion vector away from consensus & $Q\in \R^{N-1}\times\R^N \text{ such that } Q\vec{1}=\vec{0}, \ QQ^T=I_{N-1}, \text{ and } Q^TQ=I_N-\frac{1}{N}\vec{1}_N\vec{1}_N^T $
+\end{tabular}
+\end{table}
 

### 2015-02-05 lyapunov_eq_with_forcing.m

@@ -155,10 +155,11 @@ strategy=[15,15*ones(1,N-1)];
 positions=unifrnd(0,1,N,2);
 d=squareform(pdist(positions));
 
-k=0;
+k=5;
 b=0.1;
 beta=b*[ones(k,1);zeros(N-k,1)];
 
+adj=zeros(N);
 L=zeros(N);
 M=zeros(N);
 for ind=1:N
@@ -166,19 +167,22 @@ for ind=1:N
     neighbors=order(2:strategy(ind)+1);
     M(ind,neighbors)=1/strategy(ind);
     L(ind,neighbors)=1/strategy(ind);
+    adj(ind,neighbors)=1;
 end
+L=L+transpose(L);
+Lsums=sum(L,2);
+L(1:N+1:end)=-Lsums;
 M(1:N+1:end)=-1;
 
-Nmat=.5*(L+transpose(L));
-Nbar=Nmat-diag(Nmat*ones(N,1));
+Nmat=L;
 B=diag(beta);
-P=Nbar-B;
+P=Nmat-B;
 A=M-B;
-% 
-% sigma=lyap(A,eye(N));
-% 
-% D=lyap(P,eye(N));
-% 
+
+sigma_forced=lyap(A,eye(N));
+
+D=lyap(P,eye(N));
+
 % figure
 % subplot(1,4,1)
 % imagesc(A)
@@ -187,25 +191,26 @@ A=M-B;
 % imagesc(D)
 % colorbar
 % subplot(1,4,3)
-% imagesc(sigma)
+% imagesc(sigma_forced)
 % colorbar
 % subplot(1,4,4)
 % imagesc(-inv(P))
 % colorbar
-% 
-% Q=A*inv(P)+inv(P)*transpose(A);
-% 
-% todivide=repmat(diag(-inv(P)),1,N).*repmat(reshape(diag(-inv(P)),1,[]),N,1);
-% todivide=power(todivide,.5);
-% corrs=-inv(P)./todivide;
-% 
-% figure
-% plot(d,corrs,'ok')
-% set(gca,'ylim',[0 1])
-% 
-% [vecs,vals]=eig(P);
+figure;plot(mean(d(1:5,:),1),-diag(inv(P)),'o')
 
-[W,Lambda]=eig(Nbar);
+Q=A*inv(P)+inv(P)*transpose(A);
+
[…]

### 2015-02-05 signalingevents.m

@@ -5,7 +5,7 @@ N=max(size(strategy));
 scores=zeros(N,numsigs_permove,nummoves);
 
 % parfor i=1:nummoves
-for i=1:20000
+for i=1:nummoves
     positions=unifrnd(0,1,N,2);
     d=squareform(pdist(positions));
 

### 2015-03-09 paper_skeleton.tex
ive changed how im calculating h2 and corr lengths and i think it all makes much more sense now... now i just have to write it up

@@
 Young et al.
-show in \cite{Young:2010fk} show that this equivalent to finding $Q$ such that $Q\vec{1}=\vec{0}$, $QQ^T=I_{N-1}$, and $Q^TQ+I_N-\frac{1}{N}\vec{1}_N\vec{1}_N^T$.
+show in \cite{Young:2010fk} show that this equivalent to finding $Q$ such that $Q\vec{1}=\vec{0}$, $QQ^T=I_{N-1}$, and $Q^TQ=I_N-\frac{1}{N}\vec{1}_N\vec{1}_N^T$.
 If we then define $y=Qx$, $y$ is the component of $x$ orthogonal to the consensus vector.
@@
 \begin{align*}
-\dot{y}&=Q\dot{x}+Q\xi
+\dot{y}&=Q\dot{x}
 \\&=QLx+Q\xi
@@
 Young et al.
-show in \cite{Young:2010fk} and we show below in Claim \ref{lyapeq_derivation} that if $\Sigma=\lim_{t\to\infty}E[y(t)y^T(t)]$, then $\Sigma$ satisfies $\tilde{L}\Sigma+\Sigma \tilde{L}^T=-I$.
+show in \cite{Young:2010fk} and we show below in Claim \ref{lyapeq_derivation} that if $\Sigma=\lim_{t\to\infty}E[y(t)y^T(t)]$, then $\Sigma$ satisfies $\tilde{L}\Sigma+\Sigma \tilde{L}^T=-QQ^T$.
 If we find such a $\Sigma$, then $\Tr(\Sigma)=\lim_{t\to\infty}\sum_iE[y_i(t)^2]\lim_{t\to\infty}=E[|y(t)|^2].$ If we define $\mathscr{H}=\Tr(\Sigma)$, then $H$ shows the expected distance between the opinion vector and the consensus vector at steady state.
@@
 In this case, $\vec{1}$ becomes a globally asymptotically stable equilibrium and we can find the steady state variance around this equilibrium without rotating into the subspace orthogonal to consensus.
-In particular, if $C_{ij}=\lim_{t\to\infty}E[(y_i(t)-\langle y_i(t)\rangle)(y_j(t)-\langle y_j(t)\rangle)]$, then the symmetric matrix $C$ satisfies $L_\text{f}C+CL_\text{f}^T=-I$ (Claim \ref{lyapeq_derivation}).
+In particular, if $C_{ij}=\lim_{t\to\infty}E[(x_i(t)-\langle x_i(t)\rangle)(x_j(t)-\langle x_j(t)\rangle)]$, then the symmetric matrix $C$ satisfies $L_\text{f}C+CL_\text{f}^T=-I$ (Claim \ref{lyapeq_derivation}).
 

### 2015-03-09 greedyopt_convergence_explanation.m

@@ -1,0 +1,85 @@
+timesteps=100;
+
+evolution_eaten=zeros(N,timesteps);
+groupconsensus_eaten=zeros(timesteps);
+groupconsensusforced_eaten=zeros(timesteps);
+corrlengths_eaten=zeros(timesteps);
+corrlengthsforced_eaten=zeros(timesteps);
+disconnected_eaten=zeros(timesteps);
+
+
+k=15;
+init=[14*ones(k,1);17*ones(N-k,1)];
+% [probeaten, ~]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+% [mean(1-probeaten(1:k)) mean(1-probeaten(k+1:end))]
+% 
+% strategy_upone=strategy;
+% strategy_upone(k+1)=strategy(end)+1;
+% 
+% strategy_downone=strategy;
+% strategy_downone(k+1)=strategy(end)-1;
+% 
+% [probeaten_upone, ~]=signalingevents_parallel(strategy_upone,numsigs_permove,nummoves,radius,b,T);
+% [probeaten_downone, ~]=signalingevents_parallel(strategy_downone,numsigs_permove,nummoves,radius,b,T);
+
+t=1;
+    while t<=timesteps
+
+    if t==1
+        strategy=init;
+    %     strategy=2*ones(1,N);
+    %     strategy(1)=10; 
+        evolution_eaten(:,t)=strategy;
+        t=t+1;
+    else
+        strategy=evolution_eaten(:,t-1);
+        strategy=reshape(strategy,1,N);
+
+        [probeaten, ~]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+        perf=1-probeaten;
+        newstrategy=strategy;
+        
+        i=randsample(1:N,1);
+%         for i=1:N
+            choosestrat=zeros(1,3);
+            choosestrat(2)=perf(i);
+
+            if strategy(i)==N-1
+                choosestrat(3)=perf(i);
+            else 
+                upone=strategy;
+                upone(i)=strategy(i)+1;
+
+                [probeaten, ~]=signalingevents_parallel(upone,numsigs_permove,nummoves,radius,b,T);
+                choosestrat(3)=1-probeaten(i);
+            end
+
+            if strategy(i)==1
+                choosestrat(1)=perf(i);
+            else
+                downone=strategy;
+                downone(i)=strategy(i)-1;
+
[…]

### 2015-03-12 paper_skeleton.tex
in the middle of straightening out correlations and H2...again

--- 2015-01-30 paper_skeleton.tex
@@
 \subsection{Learning dynamics }
-An individual's opinion changes with a probability that depends on its social interactions and the strength of the environmental stimulus.
-Specifically, an individual's opinion $\sigma_i$ can be either $1$ or $-1$ and $\sigma_i$ switches with probability $w_i(\sigma_i)=\frac{1}{2}(-\sigma_i\sum_jA_{ij}\sigma_j )(1-\beta_i\sigma_i)$ where $A_{ij}$ indicates the strength of the connection from $j$ to $i$ and $\beta_i$ indicates the strength of the environmental stimulus perceived by $i$.
-The stochastic dynamics of the opinions can be written down with a master equation \cite{Glauber:1963fk}.
-The vector $\vec{v}(t)$ of expected values, $q_i(t)=\langle \sigma_i(t)\rangle$, satisfies the following differential equations: \cite{Glauber:1963fk}
-\begin{equation*}
-\frac{d\vec{v}}{dt}=M\vec{v}+\vec{\beta}^T(\vec{1}-\vec{v}).
-\end{equation*}
-We simplify this by writing $\frac{d\vec{v}}{dt}=A\vec{v}+\vec{\beta}^T\vec{1}$, where $A=M-B$ and $B$ is a diagonal matrix with the values of $\vec{\beta}$ along the diagonal.
+An individual's opinion changes with a probability that depends on its social interactions and the strength of the environmental stimulus. Let $A$ be the adjacency matrix such that
+
+\begin{equation}
+A_{ij}=\left\{\begin{array}{l l l}
+1 & \text{ if $i$ pays attention to $j$ }
+\\0 & \text{ otherwise }
+\end{array}\right.
+\end{equation}
+
+We assume that having more neighbors means paying less attention to each so we normalize $A$ to find $\bar{A}$:
+\begin{equation}
+\bar{A}_{ij}=\frac{A_{ij}}{\sum_kA_{ik}}.
+\end{equation}
+
+If each bird averages the differences between its opinion and each of its neighbors' to update its opinion, the opinion dynamics will be described by 
+\begin{equation}
+\dot{x}=Lx
+\end{equation}
+where $L=\bar{A}-I$.
 In the absence of the environmental signal, there is no trend in the dynamics.
-In the presence of the environmental signal, however, those individuals who perceive the signal will be more likely to switch to having an opinion of $1$ so their expected value will increase, and consequently so will the values of the those paying attention to the informed individuals.
-Thus, having a higher expected value can be interpreted as having more information about the information and having a lower expected value can be interpreted as having less information.
+We introduce an environmental signal by changing the dynamics such that informed birds change their opinions to be closer to $1$.
+Having an opinion closer to $1$ can be interpreted as having more information about the environment and having a lower expected value can be interpreted as having less information.
+Specifically, we incorporate this environmental forcing with a diagonal matrix $B$ where $B_{ii}=b$ if $i$ perceives the environmental signal and $B_{ii}=0$ otherwise.
+The dynamics then become
+\begin{equation}
+\dot{x}=Lx+B(\vec{1}-x).
+\end{equation}
+If we let $L_\text{s}=L-B$, then we can simplify this by writing
+\begin{equation}
+\dot{x}=L_\text{s}x+\vec{B}
+\end{equation}
+where $\vec{B}=B\vec{1}$ indicates which birds perceive the environmental signal.
 
@@
 \section{Appendix}
-\subsection{H2 Norm \label{H2}}
+%%% APPENDIX NOW DISCUSSES H2 WITHOUT FORCING (AS IN GEORGE'S PAPER), H2 WITH FORCING, WHICH GIVES CORRELATIONS BUT NOT FLUCTUATIONS AROUND AVERAGE OPINION, STAT MECH CORRELATIONS WITHOUT FORCING AND STAT MECH CORRELATION WITH FORCING, BOTH OF WHICH GIVE FLUCTUATIONS AROUND AVERAGE OPINION.
+SO H2 WITH FORCING IS STILL INCOMPARABLE TO ANYTHING ELSE SINCE ITS CORRELATIONS NOT FLUCTUATIONS AROUND AVERAGE OPINIONS.
+IT'S OKAY FOR NOW.
+\subsection{Linear algebra background }
 
-\ Let $y$ represent the distance between the vector of opinions and consensus, i.e.
-$y=QxQ^T,$
-where $Q\in\R^{n-1\times n}$ is such that $Q\vec{1}=0$ , $QQ^T=I_{n-1}$, and $Q^TQ=I_n-\frac{1}{n}\vec{1}\times\vec{1}^T$.   
-Then 
-\begin{equation}
-\dot{y}(t)=-\overline{L}y(t)+Q\xi(t) \notag
-\end{equation}
-where $\overline{L}=QLQ^T$.  As shown in \cite{Young:2010fk}, if $\Sigma_y(t)=\E[y(t)y(t)^T]$,  $\Sigma_y$ at equilibrium solves
-\begin{equation} 0=-\overline{L}\Sigma_y-\Sigma_y\overline{L}^T+D. \label{h2norm}
-\end{equation}
-\begin{align*}
-\mathscr{H}_2&=\sqrt{\Tr(\Sigma_y)}
-\\\mathscr{H}_2 \text{ robustness} &=\frac{1}{\mathscr{H}_2}
-\end{align*}
+\begin{theorem} \label{lyapeq_theorem}
+Given any positive symmetric matrix $Q>0$ there exists a unique positive symmetric matrix $\Sigma>0$ satisfyying $A\Sigma+\Sigma A^T=Q$ if and only if the system $\dot{x}=Ax$ is globally asymptotically stable.
+\end{theorem}
 
-\begin{claim}
-Given any $Q>0$ there exists a unique $\Sigma>0$ satisfyying $A\Sigma+\Sigma A^T=Z$ if and only if the system $\dot{x}=Ax$ is  global asymptotically stable.
-\end{claim}
-
-\begin{claim}
+\begin{claim} \label{lyapeq_derivation}
 Consider the dynamical system,
 \begin{equation}
-\dot{y}=Ay+B+C\xi  \label{tofindsigma}
+\dot{y}=Ay+v+Q\xi  \label{tofindsigma}
 \end{equation}
 with $E[\xi(t)]=0$, $E[\xi(t)\xi(s)^T]=\frac{1}{2}\delta(t-s)I$, and $E[y(0)\xi^T(t)]=0$.
-If all eigenvalues of $A$ are negative and if $\Sigma(t)=E[y(t)y(t)^T]$ and $\Sigma_{ss}=\lim_{t\to\infty}\Sigma(t)$, then $\Sigma_{ss}$ satisfies the Lyapunov equation
+If all eigenvalues of $A$ are negative and if $\Sigma_t=E[y(t)y(t)^T]$ and $\Sigma=\lim_{t\to\infty}\Sigma_t$, then $\Sigma$ satisfies the Lyapunov equation
 \begin{equation*}
-A\Sigma_{ss}+\Sigma_{ss}A^T=A^{-1}BB^T+BB^T(A^{-1})^T-CC^T
+A\Sigma+\Sigma A^T=A^{-1}vv^T+vv^T(A^{-1})^T-QQ^T.
+\end{equation*}
+Further, if $C_{ij}=\lim_{t\to\infty} E[(y_i(t)-E(y_i(t)))(y_j(t)-E(y_j(t)))]$ then 
+\begin{equation*}
+AC+CA^T=-QQ^T
 \end{equation*}
@@
 \begin{align*}
-\Sigma(t)&=E[y(t)y(t)^T]
+\Sigma_t&=E[y(t)y(t)^T]
 \\ \Rightarrow \dot{\Sigma}&=E[\dot{y}(t)y(t)^T+y(t)\dot{y}(t)^T]
-\\&=E[(Ay+B+C\xi)y^T+y(y^TA^T+B^T+\xi^TC^T)]
-\\&=AE[yy^T]+E[yy^T]A^T+BE[y^T]+E[y]B^T+CE[\xi y^T]+E[y\xi^T]C^T
+\\&=E[(Ay+v+Q\xi)y^T+y(y^TA^T+v^T+\xi^TQ^T)]
+\\&=AE[yy^T]+E[yy^T]A^T+vE[y^T]+E[y]v^T+QE[\xi y^T]+E[y\xi^T]Q^T
+\\&=A\Sigma_t+\Sigma_tA^T+vE[y^T]+E[y]v^T+QE[\xi y^T]+E[y\xi^T]Q^T.
 \end{align*}
 Note that we can write a solution to (\ref{tofindsigma}) as 
-\begin{equation*}
-y(t)=\exp(At)y(0)+\int_0^t\exp(A(t-s))Bds+\int_0^t\exp(A(t-s))C\xi(s)ds.
-\end{equation*}
-This allows us to write $E[y]=\exp(At)y(0)+\int_0^t\exp(A(t-s))Bds$ and 
+\begin{equation}
+y(t)=\exp(At)y(0)+\int_0^t\exp(A(t-s))vds+\int_0^t\exp(A(t-s))Q\xi(s)ds. \label{sde_solution}
+\end{equation}
+Eq. (\ref{sde_solution}) allows us to write 
 \begin{align*}
-E[y \xi^T]&=\exp(At)E[y(0)\xi(t)^T]+\int_0^t\exp(A(t-s))BE[\xi(t)^T]ds+\int_0^t\exp(A(t-s))CE[\xi(s)\xi(t)]ds
-\\&=0+0+\frac{1}{2}C
-\\ \Rightarrow \dot{\Sigma}&=A\Sigma(t)+\Sigma(t)A^T+By(0)^T\exp(At)^T+\exp(At)y(0)B^T
-\\&+B\int_0^tB^T\exp(A(t-s))^Tds+\int_0^T\exp(A(t-s))BdsB^T+\frac{1}{2}CC^T+\frac{1}{2}CC^T
-\\&=A\Sigma(t)+\Sigma(t)A^T+By(0)^T\exp(At)^T+\exp(At)y(0)B^T
-\\&-BB^T(A^{-1})^T\exp(A(t-s))|_{s=0}^{s=t}-A^{-1}\exp(A(t-s))|_{s=0}^tBB^T+CC^T
-\\&=A\Sigma(t)+\Sigma(t)A^T+By(0)^T\exp(At)^T+\exp(At)y(0)B^T
-\\&-BB^T(A^{-1})^T(I-\exp(At))-A^{-1}(I-\exp(At))BB^T+CC^T
-\\ \lim_{t\to\infty}\dot{\Sigma}&=A\Sigma_{ss}+\Sigma_{ss}A^T-BB^T(A^{-1})^T-A^{-1}BB^T+CC^T \text{ since all eigenvalues of $A$ are negative}
-\\\Rightarrow A\Sigma_{ss}+\Sigma_{ss}A^T&=BB^T(A^{-1})^T+A^{-1}BB^T-CC^T \text{ since $\dot{\Sigma}=0$ at equilibrium}
+E[y(t)]&=\exp(At)y(0)+\int_0^t\exp(A(t-s))vds
+\\&=\exp(At)y(0)-A^{-1}\exp(A(t-s))v|_{s=0}^{s=t} \text{ which exists because $A$ has no $0$ eigenvalues }
+\\&=\exp(At)y(0)-A^{-1}(I-\exp(At))v.
+\end{align*}
+Note that if we take the limit as $t$ goes to infinity, 
+$$\lim_{t\to\infty}E[(y(t))]=\lim_{t\to\infty}\exp(At)y(0)-A^{-1}(I-\exp(At))v=-A^{-1}v $$ 
+since all eigenvalues of $A$ are negative.  Eq. (\ref{sde_solution}) also allows us to write
+\begin{align*}
+E[y \xi^T]&=\exp(At)y(0)E[\xi(t)^T]+\int_0^t\exp(A(t-s))BE[\xi(t)^T]ds+\int_0^t\exp(A(t-s))QE[\xi(s)\xi(t)]ds
+\\&=0+0+\frac{1}{2}Q
+\\ \Rightarrow \dot{\Sigma} &= A\Sigma_t+\Sigma_tA^T+vE[y(t)^T]+E[y(t)]v^T+\frac{1}{2}QQ^T+\frac{1}{2}QQ^T
+\\ &= A\Sigma_t+\Sigma_tA^T+vE[y(t)^T]+E[y(t)]v^T+QQ^T
+\\ \Rightarrow \lim_{t\to\infty} \dot{\Sigma}&=A\Sigma+\Sigma A-vv^T(A^{-1})^T-A^{-1}vv^T+QQ^T
+\\ \Rightarrow A\Sigma+\Sigma A&=vv^T(A^{-1})^T+A^{-1}vv^T-QQ^T \text{ since $\dot{\Sigma}=0$ at equilibrium}.
+\end{align*}
+
+Further, by definition of $C$, 
+\begin{align*}
+C&=\Sigma-\lim_{t\to\infty}E[y(t)]E[y(t)^T]
+\\&=\Sigma-(-A^{-1}v)(-A^{-1}v)^T
+\\&=\Sigma-A^{-1}vv^T(A^{-1})^T
+\\\Rightarrow AC+CA^T&=A\Sigma+\Sigma A^T-AA^{-1}vv^T(A^{-1})^T
+-A^{-1}vv^T(A^{-1})^TA^T
+\\&=A\Sigma+\Sigma A^T-vv^T(A^{-1})^T-A^{-1}vv^T
+\\&=-QQ^T.
 \end{align*}
@@
 
-\subsection{Calculating correlations \label{corr}}
-The following is derived from the work in \cite{Bialek:2013fk}.
+\subsection{Opinion dynamics with noise }
+In our model, each bird in the flock as an opinion $x_i$, giving an opinion vector $x$.
+The $\mathscr{H}_2$ robustness of the opinion dynamics is defined by the inverse of the expected distance at equilibrium from the opinion vector to a consensus state in which all birds have the same opinion, i.e.
+$x_i=\alpha$ for all $i$ and for some $\alpha\in\R$.
+That is, the dynamics are robust if, at equilibrium, the opinions are close to being in consensus.
+We can find this distance by rotating the opinion vector into the space orthogonal to the consensus vector, $\vec{1}$.
+We define a matrix $Q\in\R^{N-1}\times\R^N$ such that each row of $Q$ is orthogonal to $\vec{1}$ and the rows are pairwise orthogonal.
+Young et al.
+show in \cite{Young:2010fk} show that this equivalent to finding $Q$ such that $Q\vec{1}=\vec{0}$, $QQ^T=I_{N-1}$, and $Q^TQ=I_N-\frac{1}{N}\vec{1}_N\vec{1}_N^T$.
+If $y=Qx$, then $||y||=y^Ty$ describes how far away from consensus $x$ is.
+Specifically, robustness as in \cite{Young:2010fk} is given by
+\begin{equation}
+r=\lim_{t\to\infty}\frac{1}{E[||y||]}=\lim_{t\to\infty}\frac{1}{\Tr E[yy^T]}. \label{H2_robustness}
+\end{equation}
+We also define a full rotation matrix $R\in\R^N\times\R^N$ such that $R=\left(\begin{array}{cc} Q \\ \frac{1}{\sqrt{N}}\vec{1}^T \end{array}\right)$.
+
+Young et al.
+\cite{Young:2010fk} model opinion dynamics by assuming that each bird averages the differences between its opinion and each of the opinions of its neighbors and changes its opinion to minimize this difference and that each bird's opinion is subject to independent noise of unit intensity.
+The equations describing these dynamics can be written as
+\begin{equation}
+\dot{x_i}=\frac{\sum_{\text{neighbors $j$ of $i$ } }(x_j-x_i)}{K_i} +\xi_i=\sum_j\bar{A}_{ij}x_j-x_i+\xi_i=\sum_jL_{ij}x_j+\xi_i.
+\end{equation}
+In vector notation this becomes
+\begin{equation}
+\dot{x}=Lx+\xi.
+\end{equation}
+The equilibrium of these dynamics is a consensus state in which all birds have the some opinion, $x=\alpha\vec{1}$ for some $\alpha\in\R$.
+In the absence of any environmental signal, the consensus opinion can be any value $\alpha$.
+In the presence of an environmental signal, some birds in the flock perceive the signal and change their opinions to be closer to $1$.
+The environmental forcing is described by a diagonal matrix $B$ where $B_{ii}=b$ if $i$ perceives the environmental signal and $B_{ii}=0$ otherwise and the dynamics then become
+\begin{equation}
+\dot{x}=Lx+B(\vec{1}-x)=L_\text{s}x+v.
+\end{equation}
+where $L_\text{s}=L-B$ and $v=B\vec{1}$. If $b=0$, then $L_\text{s}=L$ and $v=\vec{0}$, so the case with the signal reduces to the case without the signal.
+
+If $z=Rx$, then the first $N-1$ coordinates of $z$ are $y=Qx$ and the last is $\sum_ix_i/\sqrt{N}$. The dynamics of $z$ follow
 \begin{align*}
-H(\vec{v})&=-\vec{v}^TA\vec{v}-\vec{\beta}^T\vec{v}
-%\\&=\vec{v}A\vec{v}-\vec{\beta}^T\vec{v}+\vec{\beta}^T\vec{1}
+\dot{z}&=R\dot{x}
+\\&=RL_\text{s}x+Rv+R\xi.
 \end{align*}
-Define $V=\sum_kv_k/N$ and $\vec{y}=(\vec{v}-V\vec{1})/V$.  Then $\vec{v}=V(\vec{1}+\vec{y})$.  
+Since the rows of $R$ are orthonormal, $RR^T=I$. Further, since the rows of $R$ are orthonormal, $R$ is invertible, which implies that $R^T=R^{-1}$.  Then we can write
 \begin{align*}
-H(\vec{y},V)&=-V^2(\vec{1}^T+\vec{y}^T)A(\vec{1}+\vec{y})-\vec{\beta}^TV(\vec{1}+\vec{y})
-\\&=-V^2(\vec{1}^T+\vec{y}^T)A\vec{y}-V\vec{\beta}^T\vec{y}-V\vec{\beta}^T\vec{1} \text{ since $A\vec{1}=\vec{0}$}
-\\&=-V^2\vec{y}^TA\vec{y}+(-V^2\vec{1}^TA-V\vec{\beta}^T)\vec{y}-V\vec{\beta}^T\vec{1}
+\dot{z}&=RL_\text{s}R^Tz+Rv+R\xi
+\\&=\tilde{L}_\text{s}z+\tilde{v}+R\xi 
 \end{align*}
-Define $Q\in M_{N-1,N}$ that rotates an $N$-vector away from the consensus vector and define $\vec{z}=Q\vec{y}$ so that $\vec{y}=Q^T\vec{z}$ since $\sum_ky_k=0$.
+where $\tilde{L}_\text{s}=RL_\text{s}R^T$ and $\tilde{v}=Rv=RB\vec{1}$.
+If $b\neq 0$, $L_\text{s}$ and therefore $\tilde{L}_\text{s}$ have all negative eigenvalues and $z=(0, 0, \dots 0, \sqrt{N})^T$ is a globally asymptotically stable equilibrium.
+Therefore, by Claim \ref{lyapeq_derivation} and Theorem \ref{lyapeq_theorem}, $\lim_{t\to\infty}E[z(t)z(t)^T]=\Sigma_z$ where $\tilde{L}_\text{s}\Sigma+\Sigma\tilde{L}_\text{s}^T=-RR^T$.
+We can then find $\Sigma_y=\lim_{t\to\infty}E[y(t)y(t)^T]$
+by removing the last row and column of $\Sigma_z$.
+
+If $b=0$, $L$ has a $0$ eigenvalue corresponding to the eigenvector $\vec{1}$.
+Therefore, $RLR^T$ also has a $0$ eigenvalue, so we can therefore no longer find the desired $\Sigma_z$.
+Since the last row of $R=\frac{1}{\sqrt{N}}\vec{1}$, all entries of the the last column of $LR^T$ are $0$.
+Therefore, all entries of the last column of $RLR^T$ are $0$.
+In other words, the dynamics of the first $N-1$ elements of $z$, i.e.
+$y$, do not depend on the last.
+This lets us write
 \begin{align*}
-H(\vec{z},V)&=-V^2\vec{z}^TQAQ^T\vec{z}+(-V^2\vec{1}^TAQ^T-V\vec{\beta}^TQ^T)\vec{z}-V\vec{\beta}^T\vec{1}
-\\&=V^2\vec{z}^TP\vec{z}+(-V^2\vec{\sigma}_1^T-V\vec{\sigma}_2^T)\vec{z}-V\vec{\beta}^T\vec{1}
-\\ \Rightarrow P(\vec{z},V)&=\frac{1}{\z}\exp\left(-V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+V\vec{\beta}^T\vec{1}\right)
+\dot{y}&=QLQ^Ty+Q\xi=\tilde{L}+Q\xi
 \end{align*}
-where $P=-QAQ^T$, $\vec{\sigma}_1=QA^T\vec{1}$, and $\vec{\sigma}_2=Q\vec{\beta}$.
+where $\tilde{L}=QLQ^T$. Another way to see this is to write 
+\begin{align*}
+ \dot{y}&=Q\dot{x}
+ \\ &=QLx+Q\xi
+ \\&=QL(x-\langle x\rangle \vec{1})+Q\xi \text{ since $L\vec{1}=\vec{0}$}
+ \\&=QLQ^TQx+Q\xi
+ \\&=\tilde{L}y+Q\xi.
+\end{align*}
+Again, by Claim \ref{lyapeq_derivation} and Theorem \ref{lyapeq_theorem}, we can now find $\lim_{t\to\infty}E[y(t)y(t)^T]=\Sigma_y$ by finding a solution to $\tilde{L}\Sigma_y+\Sigma_y\tilde{L}=-QQ^T$.
 
-\begin{fact}
+To summarize, in the absence of an environmental signal, $\Sigma_y$ satisfies $\tilde{L}\Sigma_y+\Sigma_y\tilde{L}=-QQ^T$ where $\tilde{L}=QLQ^T$ and the presence of an environmental signal, $\Sigma_y$ can be found in the first $N-1$ rows and columns of $\Sigma_z$ where $\Sigma_z$ satisfies $\tilde{L}_\text{s}\Sigma_z+\Sigma_z\tilde{L}_\text{s}=-RR^T$ where $\tilde{L}_\text{s}=RL_\text{s}R^T$.
+
+\subsection{Probability distribution at equilibrium }
+A quite different way of modeling social interactions the method used by Bialek et al.
+in \cite{Bialek:2012fk}, which is to assume the opinions come to equilibrium and write down an expression for the probability of finding the opinion vector in every state.
+In statistical physics, a Hamiltonian describes the energy of a given state, $H(v)$, and the Boltzmann distribution describes a probability distribution in which the most energetic states are least likely.
+Analogously, we can write a function $H(v)$ that gives a high energy when the opinions are not in consensus and a low energy when the opinions are in consensus and a probability distribution of opinion vectors that favors less energetic states.
+In order to do this, we must use a symmetric interaction matrix, rather than the possibly asymmetric $\bar{A}$ we used above.
+Bialek et al.
+study a model in which opinions are a three-dimensional velocity vector and the birds can weight the opinions of their nearest neighbors with any value $J$.
+In order to simplify this model and make it directly comparable to the opinion dynamics model we described above, we only consider a scalar-valued opinion $x_i$ and we impose the condition that the weights given to the neighbors' opinions are inversely proportional to the number of neighbors.
+In particular, we define $S=\frac{1}{2}(A+A^T)$ and $P$ such that $P_{ij}=S_{ij}$ for $j\neq i$ and $P_{ii}=-\sum_kS_{ik}$.
+This allows us to write down a Hamiltonian
+\begin{equation}
+H(v)=-v^TPv
+\end{equation}
+where the negative sign makes opinions far from consensus give high energies and conversely. The probability distribution over opinion vectors $v$ then becomes 
+\begin{equation}
+P(v)=\frac{1}{Z}\exp\left(-\frac{1}{2}v^T(-P)v\right),
+\end{equation}
+where $Z=\int_\R^N\exp\left(-\frac{1}{2}H(v)\right)dv$.
+We can now use this probability distribution to find the  covariances and correlations between the birds' fluctuations around consensus.
+If we let $V=\sum_iv_i/N=\langle v\rangle$ and $z=v-V\vec{1}$ so that $z$ represents the deviations in each bird's opinion from the consensus opinion.
+Then
 \begin{align*}
-\int\exp\left(-\frac{1}{2}\vec{z}^TP\vec{z}+\vec{\sigma}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{\det P}}\exp\left(-\frac{1}{2}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
-\\\Rightarrow \int\exp\left(-\vec{z}^TP\vec{z}+\vec{\sigma}^T\vec{z}\right)d^{N-1}z&=\sqrt{\frac{(2\pi)^{N-1}}{2^{N-1}\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}^TP^{-1}\vec{\sigma}\right)
+P(z,V)&=\frac{1}{Z}\exp\left(-\frac{1}{2}(z+V\vec{1})^T(-P)(z+V\vec{1})\right)
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}z^T(-P)z\right) \text{ since }P\vec{1}=P^T\vec{1}=\vec{0}.
 \end{align*}
-\end{fact}
+If $P$ were invertible, the covariance matrix would be given by $-P^{-1}$. However, since $P\vec{1}=\vec{0}$, one of the eigenvalues of $P$ is $0$ and it is therefore not invertible.
+To proceed, we will write $z$ as a linear combination of eigenvectors of $P$.
+Since $P$ is symmetric and negative, there is an orthonormal set of eigenvectors $\{w_k\}$ such that $Pw_k=\lambda_kw_k$ with $w_N=\vec{1}$ and $\lambda_N=0$.
+If we let $W$ be a matrix whose columns are given by $\{w_k\}$ and $\Lambda$ be the diagonal matrix with $\{\lambda_k\}$ along the diagonal, then $W^T=W^{-1}$ and $P=W\Lambda W^T$ since $P$ is symmetric.
+If we write $z=\sum_kc_kw_k=W{c}$, then $c_N=0$ since $\sum_iz_i=0$.
+If we then denote the $N$ by $N-1$ matrix whose columns are given by $w_1,...w_{N-1}$ as $\tilde{W}$ and the diagonal matrix with $\lambda_1,\dots,\lambda_{N-1}$ along the diagonal as $\tilde{\Lambda}$, then $z=\tilde{W}\tilde{c}$ for a $\tilde{c}\in\R^{N-1}$ and
+%and $\tilde{P}=\tilde{W}\tilde{\Lambda}\tilde{W}^T$, then 
+\begin{align*}
+P(z)&=\frac{1}{Z}\exp\left(-\frac{1}{2}z^T(-{P})z\right)
+\\\Rightarrow P(\tilde{c})&=\frac{1}{Z}\exp\left(-\frac{1}{2}\tilde{c}^T\tilde{W}^T(-W\Lambda W^T)\tilde{W}\tilde{c}\right)
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}c^T\tilde{W}^T(-\tilde{W}\tilde{\Lambda}\tilde{W}^T)\tilde{W}c\right) 
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\tilde{c}^T(-\tilde{\Lambda})\tilde{c}\right)  
+\end{align*}
+Therefore, $E[\langle c \rangle ]=\vec{0}$ and $E[\tilde{c}\tilde{c}^T]=(-\tilde{\Lambda})^{-1}$.
+Therefore, $C=E[zz^T]=E[\tilde{W}\tilde{c}\tilde{c}^T\tilde{W}^T]=-\tilde{W}\tilde{\Lambda}\tilde{W}^T$.
+By dividing each entry $C_{ij}$ of the covariance matrix by $C_{ii}C_{jj}$, the covariance matrix can be used to find the correlation matrix.
+
+The Hamiltonian framework, like the opinion dynamics framework, can also accommodate an environmental signal. In this case, the appropriate Hamiltonian is 
+\begin{equation}
+\begin{array}{lll}
+H(v)&=-(v-\vec{1})^TP_\text{s}(v-\vec{1})
+\end{array}
+\end{equation}
+where $P_\text{s}=P-B$.
+It is clear that $H$ is minimized when $v=\vec{1}$. 
+Again, we are interested in the deviations from the consensus opinion. As above, let $V=\sum_iv_i/N$ and $z=v-V\vec{1}$ so that $v=z+V\vec{1}$. 
+\begin{align*}
+P(v)&=\frac{1}{Z}\exp\left(-\frac{1}{2}(v-\vec{1})^T(-P_\text{s})(v-\vec{1}\right)
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}(z-(1-V)\vec{1})^T(-P_\text{s})(z-(1-V)\vec{1}\right)
+\end{align*}
+Since the subspace of $z\in\R^N$ such that $\sum_iz_i=0$ is $N-1$ dimensional, it will be easier to rotate $z$ into $\R^{N-1}$ and integrate over all of $\R^{N-1}$.
+Again, we will use a rotation matrix $R\in\R^N\times\R^N$ such that the last row of $R$ is $\frac{1}{\sqrt{N}}\vec{1}$ and the other rows are normal and orthogonal to $\vec{1}$.
+Above, we found a $Q\in\R^{N-1}\times\R^N$ such that each row of $Q$ was orthogonal to $\vec{1}$ and $R\in\R^N\times\R^N$ be the matrix with $\frac{1}{\sqrt{N}}\vec{1}^T$ as its last row and the first $N-1$ rows identical to the rows of $Q$.
+Then $R$ is a rotation matrix, so $RR^T=R^TR=I$ and $\det(R)=1$.
+This allows us to write $z-(1-V)\vec{1}=Q^TQv$ and $V=\frac{1}{N}\vec{1}^Tv$.
+If $y=Rz$, then $y_N=0$ since $\vec{1}^Tz=0$.
+Therefore,
+$$\Omega=\{z\in\R^N:\sum_iz_i=0\}=\{z\in\R^N:z\cdot\vec{1}=0\}=\{z\in\R^N:\exists \ y\in\R^N \text{ such that } z=R^Ty \text{ and } y_N=0\}.$$
+Further, if $e_N=(0,0,\dots,0,1)^T$, $R^Te_N=q_N=\frac{1}{\sqrt{N}}\vec{1}$ so that $\vec{1}=\sqrt{N}R^Te_N$.This allows us to write
+\begin{align*}
+P(v)&=\frac{1}{Z}\exp\left(-\frac{1}{2}(R^Ty-(1-V)\sqrt{N}R^Te_N)^T(-P_\text{s})(R^Ty-(1-V)\sqrt{N}R^Te_N)\right)
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}(y-(1-V)\sqrt{N}e_N)^T(-RP_\text{s}R^T)(y-(1-V)\sqrt{N}e_N)\right)
+\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}(y+\sqrt{N}Ve_N-\sqrt{N}e_N)^T\Tilde{P}_\text{s}(y+\sqrt{N}Ve_N-\sqrt{N}e_N)\right)
+\end{align*}
+where $\Tilde{P}_\text{s}=-RP_\text{s}R^T$.
+To find $P(z)$ we can integrate out $V$.
+However, to consider this as a probability density function over $y$ and $V$, we have to consider the determinant of the transformation from $v$ to $y$ and $V$.
+Combining the transformations from $v$ to $(z,V)$ and from $z$ to $(yV)$,
+\begin{align*}
+y=RQ^TQv=Qv \text{ and } V=\frac{1}{N}\vec{1}^Tv=\frac{1}{\sqrt{N}}R_{N\cdot }v
+\end{align*}
+so that the entire transformation can be represented by $\bar{R}=\left(\begin{array}{cc}Q \\1/\sqrt{N}R_{N\cdot }\end{array}\right)$, so that $\det(\bar{R})=\frac{1}{\sqrt{N}}\det(R)=\frac{1}{\sqrt{N}}$.
+Therefore, $P(y,V)=\sqrt{N}P(v)$ is the appropriate density function for $y$ and $V$.
+Finally, note that $Z=\sqrt{(2\pi)^N\det(-P_\text{s}^{-1})}$.
+This gives
+
+%\begin{align*}
+%P(z)&=\int_\R P(z,V)dV
+%\\&=\int_\R \frac{1}{Z}\exp\left(-\frac{1}{2}((V-1)\sqrt{N}e_N+y)^T\Tilde{P}_\text{s}((V-1)\sqrt{N}e_N+y)\right)dV
+%\\&=\frac{1}{Z}\exp(-\frac{1}{2}y^T\Tilde{P}_\text{s}y)\int_\R \exp\left(-\frac{1}{2}\left(\Tilde{P}_{\text{s}NN}N\left(V-1\right)^2+(V-1)\sqrt{N}\sum_{i=1}^{N-1}\Tilde{P}_{\text{s}Ni}y_i+(V-1)\sqrt{N}\sum_{i=1}^{N-1}y_i\Tilde{P}_{\text{s}iN}\right)\right)dV
+%\\&=\frac{1}{Z}\exp(-\frac{1}{2}y^T\Tilde{P}_\text{s}y)\int_\R \exp\left(-\frac{1}{2}\left(\Tilde{P}_{\text{s}NN}N\left(V-1\right)^2+2(V-1)\sqrt{N}\sum_{i=1}^{N-1}\Tilde{P}_{\text{s}Ni}y_i\right)\right)dV \text{ since $\Tilde{P}_\text{s}$ is symmetric }
+%\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^T\Tilde{P}_\text{s}y-\left(\sum_{i=1}^{N-1}\Tilde{P}_{\text{s}Ni}y_i/\sqrt{\Tilde{P}_{\text{s}NN}}\right)^2\right)\right)\times
+%\\&\int_\R \exp\left(-\frac{1}{2}\left(\Tilde{P}_{\text{s}NN}N\left(V-1\right)^2+2(V-1)\sqrt{N}\sum_{i=1}^{N-1}\Tilde{P}_{\text{s}Ni}y_i+\left(\sum_{i=1}^{N-1}\Tilde{P}_{\text{s}Ni}y_i/\sqrt{\Tilde{P}_{\text{s}NN}}\right)^2\right)\right)dV
+%\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^T\Tilde{P}_\text{s}y-\left(\sum_i\Tilde{P}_{\text{s}Ni}y_i/\sqrt{\Tilde{P}_{\text{s}NN}}\right)^2\right)\right)\int_\R \exp\left(-\frac{1}{2}\left(\sqrt{\Tilde{P}_{\text{s}NN}}\sqrt{N}\left(V-1\right)+\sum_i\Tilde{P}_{\text{s}Ni}y_i/\sqrt{\Tilde{P}_{\text{s}NN}}\right)^2\right)dV
+%\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^T\Tilde{P}_\text{s}y-\left(\sum_i\Tilde{P}_{\text{s}Ni}y_i/\sqrt{\Tilde{P}_{\text{s}NN}}\right)^2\right)\right)\int_\R \exp\left(-\frac{1}{2}\tilde{P}_{\text{s}NN}N\left(V-1+\frac{\sum_i\tilde{P}_{\text{s}Ni}y_i}{\tilde{P}_{\text{s}NN}\sqrt{N}}\right)^2\right)dV
+%\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^T\tilde{P}_{\text{s}}y-\left(\sum_i\tilde{P}_{\text{s}Ni}y_i/\sqrt{\tilde{P}_{\text{s}NN}}\right)^2\right)\right)\times\sqrt{\frac{2\pi}{\tilde{P}_{\text{s}NN}N}}
+%\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^T\tilde{P}_{\text{s}}y-\frac{1}{\tilde{P}_{\text{s}NN}}y^T\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot}y\right)\right)\times\sqrt{\frac{2\pi}{\tilde{P}_{\text{s}NN}N}} \text{ where $\tilde{P}_{\text{s}N\cdot}$ is the last row of $\tilde{P}_{\text{s}}$}
+%\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^T(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot})y\right)\right)\times\sqrt{\frac{2\pi}{\tilde{P}_{\text{s}NN}N}}
+%\\&=\frac{\sqrt{2\pi}}{\sqrt{(2\pi)^N\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}N}}\exp\left(-\frac{1}{2}\left(y^T(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot })y\right)\right)
+%\\&=\frac{1}{\sqrt{(2\pi)^{N-1}\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}N}}\exp\left(-\frac{1}{2}z^TR^T(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot})Rz\right)
+%\\&=\frac{1}{\sqrt{(2\pi)^{N-1}\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}N}}\exp\left(-\frac{1}{2}z^T(-P_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}}R^T\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot}R)z\right)
+%%\\&=\frac{\sqrt{2\pi}\sqrt{R_{11}N}}{(2\pi)^{N/2}\det(-P_\text{s}^{-1})^{1/2}}\exp\left(-\frac{1}{2}z^T(-P_\text{s}-\frac{1}{R_{11}}Q'^{-1}R_{\cdot 1}R_{1\cdot}(Q'^{-1})^T)z\right)
+%\\&=\frac{1}{\sqrt{(2\pi)^{N-1}\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}N}}\exp\left(-\frac{1}{2}z^T(-P_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}}P_\text{s}R^T_{\cdot N}R_{N\cdot}P_\text{s})z\right)
+%\\&=\frac{1}{\sqrt{(2\pi)^{N-1}\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}N}}\exp\left(-\frac{1}{2}z^T(-P_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}}P_\text{s}(R_{N\cdot})^TR_{N\cdot}P_\text{s})z\right)
+%%\\&=\frac{\sqrt{2\pi}\sqrt{R_{11}N}}{(2\pi)^{N/2}\det(-P_\text{s}^{-1})^{1/2}}\exp\left(-\frac{1}{2}z^T(-P_\text{s}-\frac{1}{R_{11}}P_\text{s}\Pi P_\text{s})z\right) \text{ where $\Pi_{ij}=1/N$ for all $i,j$}
+%\\&=\frac{1}{\sqrt{(2\pi)^{N-1}\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}N}}\exp\left(-\frac{1}{2}z^T(-P_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}N}P_\text{s}\vec{1}\vec{1}^T P_\text{s})z\right) 
+%\\&=\frac{1}{\sqrt{(2\pi)^{N-1}\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}N}}\exp\left(-\frac{1}{2}z^T(-P_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}N}B\vec{1}\vec{1}^T B)z\right)  \text{ since $P\vec{1}=\vec{0}$}
+%\end{align*}
 
 \begin{align*}
-\z&=\int_{-\infty}^\infty\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}+V\vec{\beta}^T\vec{1}\right)d^NzdV
-\\&=\int_{-\infty}^\infty \exp(V\vec{\beta}^T\vec{1})\int_{\R^{N-1}}\exp\left(V^2\vec{z}^TP\vec{z}+(V^2\vec{\sigma}_1^T+V\vec{\sigma}_2^T)\vec{z}\right)d^NzdV
-\\&=\sqrt{\pi^{N-1}}\int_{-\infty}^\infty\exp(V\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}\det{P}}}\right)\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det PP}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\exp(V\vec{\beta}^T\vec{1})\times\left(\sqrt{\frac{1}{V^{2(N-1)}}}\right)\exp\left(-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\det P}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\\&=\sqrt{\pi^{N-1}}\sqrt{\frac{1}{\Pi_a\lambda_a}}\exp\left(-\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma}_1\right)\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\\&\text{ where $\lambda_a$ are the eigenvalues of $P$}
-\\ \Rightarrow -\log(\z)&=\text{constants}+\frac{1}{2}\sum_a\log(\lambda_a)+\frac{1}{4}\vec{\sigma}_1^TP^{-1}\vec{\sigma_1}-\log(f(P))
+P(y,V)&=\sqrt{N}P(v)
+\\&=\frac{\sqrt{N}}{Z}\exp\left(-\frac{1}{2}((V-1)\sqrt{N}e_N+y)^T\Tilde{P}_\text{s}((V-1)\sqrt{N}e_N+y)\right)
+\\ \Rightarrow P(y)&=\int_\R P(y,V)dV
+\\&=\int_\R \frac{\sqrt{N}}{Z}\exp\left(-\frac{1}{2}((V-1)\sqrt{N}e_N+y)^T\Tilde{P}_\text{s}((V-1)\sqrt{N}e_N+y)\right)dV
+\\&=\frac{\sqrt{N}}{Z}\exp\left(-\frac{1}{2}y^T\Tilde{P}_\text{s}y\right)\int_\R \exp\left(-\frac{1}{2}\left(\Tilde{P}_{\text{s}NN}N\left(V-1\right)^2+(V-1)\sqrt{N}\tilde{P}_{\text{s}N\cdot}y+(V-1)\sqrt{N}y^T\tilde{P}_{\text{s}\cdot N}\right)\right)dV
+\\&=\frac{\sqrt{N}}{Z}\exp\left(-\frac{1}{2}y^T\Tilde{P}_\text{s}y\right)\int_\R \exp\left(-\frac{1}{2}\left(\Tilde{P}_{\text{s}NN}N\left(V-1\right)^2+2(V-1)\sqrt{N}\tilde{P}_{\text{s}N\cdot}y\right)\right)dV \text{ since $\Tilde{P}_\text{s}$ is symmetric }
+\\&=\frac{\sqrt{N}}{Z}\exp\left(-\frac{1}{2}\left(y^T\Tilde{P}_\text{s}y-\left(\frac{\tilde{P}_{\text{s}N\cdot}y}{\sqrt{\tilde{P}_\text{s}NN}}\right)^2\right)\right)\times
+\\&\int_\R \exp\left(-\frac{1}{2}\left(\Tilde{P}_{\text{s}NN}N\left(V-1\right)^2+2(V-1)\sqrt{N}\tilde{P}_{\text{s}N\cdot}y+\left(\frac{\tilde{P}_{\text{s}N\cdot}y}{\sqrt{\Tilde{P}_{\text{s}NN}}}\right)^2\right)\right)dV
+\\&=\frac{\sqrt{N}}{Z}\exp\left(-\frac{1}{2}\left(y^T\left(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot}\right)y\right)\right)\int_\R \exp\left(-\frac{1}{2}\left(\sqrt{\Tilde{P}_{\text{s}NN}N}\left(V-1\right)+\frac{\tilde{P}_{\text{s}N\cdot}y}{\sqrt{\Tilde{P}_{\text{s}NN}}}\right)^2\right)dV
+\\&=\frac{\sqrt{N}}{Z}\exp\left(-\frac{1}{2}\left(y^T(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot})y\right)\right)\int_\R \exp\left(-\frac{1}{2}\tilde{P}_{\text{s}NN}N\left(V-1+\frac{\sum_i\tilde{P}_{\text{s}Ni}y_i}{\tilde{P}_{\text{s}NN}\sqrt{N}}\right)^2\right)dV
+\\&=\frac{\sqrt{N}}{Z}\exp\left(-\frac{1}{2}\left(y^T(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot})y\right)\right)\times\sqrt{\frac{2\pi}{\tilde{P}_{\text{s}NN}N}}
+%\\&=\frac{\sqrt{2\pi}}{\sqrt{(2\pi)^N\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}}}\exp\left(-\frac{1}{2}\left(y^T(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot })y\right)\right)
+\\&=\frac{1}{\sqrt{(2\pi)^{N-1}\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}}}\exp\left(-\frac{1}{2}\left(y^T(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot })y\right)\right)
 \end{align*}
 
-where $f(P)=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV $.
+Therefore, $E[y_i]=0$ for all $i$ and $E[yy^T]=(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot })^{-1}$. Then, 
+\begin{equation*}
+E[zz^T]=R^T\left(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot }\right)^{-1}R=\left(P_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}N}B\vec{1}\vec{1}^TB\right)^{-1}
+\end{equation*}
+Let $N_\text{I}$ be the number of birds that can perceive the signal, i.e.
+$N_\text{I}$ is the number of non-zero elements along the diagonal of $B$.
+Then, $\tilde{P}_{\text{s}NN}=\frac{b}{N}N_\text{I}$.
+Finally, we find that 
+\begin{equation}
+E[zz^T]=\left(P_\text{s}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB\right)^{-1}
+\end{equation}
+Now, let $P'=P_\text{s}+\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB$.
+Then $P'\vec{1}=-B\vec{1}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB\vec{1}=-B\vec{1}-\frac{bN_\text{I}}{bN_\text{I}}B\vec{1}=\vec{0}$, so we cannot invert $-P'$ to find the covariance matrix.
+However, as above, we can discard the $0$ eigenvector since we know that $z\cdot \vec{1}=0$.
+This gives us $E[(v_i-\frac{\sum_kv_k}{N})(v_j-\frac{\sum_kv_k}{N})]$.
 
+\subsection{Connecting two frameworks }
+To find $\mathscr{H}$, we found the matrix $\Sigma$ such that 
+\begin{equation}
+QLQ^T\Sigma+\Sigma QLQ^T=-I.
+\end{equation}
+If the vector $v$ is described by the dynamics $\dot{v}=Dv$, with $H(v)=-v^TPv$, then
 \begin{align*}
-\Rightarrow\frac{\partial -\log(\z)}{\partial P_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{1l}-\frac{\partial \log f}{\partial P_{ij}}
-\\\Rightarrow\frac{\partial -\log(\z)}{\partial P_{ij}}&=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}{ki}P^{-1}_{jl}\sigma_{1l}-\frac{1}{f}\frac{\partial f}{\partial P_{ij}}
+\dot{H}&=-\dot{v}^TPv-v^TP\dot{v}
+\\&=-v^TD^TPv-v^TPDv
+\\&=-v^T(D^TP+PD)v.
 \end{align*}
+Therefore, $H(v)$ decreases at all $v$ if and only if $D^TP+PD$ is a positive matrix.
+%MORE?!?
 
-\begin{align*}
-\frac{\partial f}{\partial P_{ij}}&=\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)\left(-\frac{1}{4V}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)dV
-\\&=\left(-\frac{1}{4}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV
-\end{align*}
+%Specifically, an individual's opinion $\sigma_i$ can be either $1$ or $-1$ and $\sigma_i$ switches with probability $w_i(\sigma_i)=\frac{1}{2}(-\sigma_i\sum_jA_{ij}\sigma_j )(1-\beta_i\sigma_i)$ where $A_{ij}$ indicates the strength of the connection from $j$ to $i$ and $\beta_i$ indicates the strength of the environmental stimulus perceived by $i$.
+The stochastic dynamics of the opinions can be written down with a master equation \cite{Glauber:1963fk}.
+The vector $\vec{v}(t)$ of expected values, $q_i(t)=\langle \sigma_i(t)\rangle$, satisfies the following differential equations: \cite{Glauber:1963fk}
+%\begin{equation*}
+%\frac{d\vec{v}}{dt}=M\vec{v}+\vec{\beta}^T(\vec{1}-\vec{v}).
+%\end{equation*}
+%We simplify this by writing $\frac{d\vec{v}}{dt}=A\vec{v}+\vec{\beta}^T\vec{1}$, where $A=M-B$ and $B$ is a diagonal matrix with the values of $\vec{\beta}$ along the diagonal. 
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
+%
 
-\small{
-\begin{align*}
-\Rightarrow\langle z_iz_j\rangle &=\frac{1}{2}\sum_a\frac{w_iw_j}{\lambda_a}+\frac{1}{4}\sum_{kl}\sigma_{1k}P^{-1}{ki}P^{-1}_{jl}\sigma_{1l}+\left(\frac{1}{4}\sum_{kl}\sigma_{2k}P^{-1}_{ki}P^{-1}_{jl}\sigma_{2l}\right)\frac{\int_{-\infty}^\infty\frac{1}{V^{(N)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV}{\int_{-\infty}^\infty\frac{1}{V^{(N-1)}}\exp\left(V\vec{\beta}^T\vec{1}-\frac{1}{4V}\vec{\sigma}_2^TP^{-1}\vec{\sigma}_2\right)dV}
-\end{align*}}
-
-\subsection{Equivalence of dynamics and Hamiltonian }
-\begin{claim}
-If $H(\vec{v})=-\vec{v}^TA\vec{v}-\vec{\beta}^T\vec{v}$ and $\frac{d\vec{v}}{dt}=A\vec{v}+\vec{\beta}$ then $\nabla H\cdot \frac{d\vec{v}}{dt}\leq 0$.
-\end{claim}
-\begin{pf}
-\begin{align*}
-H(\vec{v})&=-\sum_{ij}A_{ij}v_iv_j-\sum_i\beta_iv_i
-\\\Rightarrow\frac{\partial H}{\partial v_i}&=-\sum_{j\neq i}A_{ij}v_j-\sum_{j\neq i}A_{ji}v_j-2A_{ii}v_i-\beta_i
-\\&=-\sum_j(A_{ij}+A^T_{ij})v_j-\beta_i
-\\ \Rightarrow \nabla H&=-(A+A^T)\vec{v}-\vec{\beta}
-\\ \Rightarrow \nabla H\cdot \frac{d\vec{v}}{dt}&=-\vec{v}^T(A+A^T)A\vec{v}-\vec{v}^T(A+A^T)\vec{\beta}-\vec{\beta}^TA\vec{v}-\vec{\beta}^T\vec{\beta}
-\\&=-\vec{v}^TA^2\vec{v}-\vec{v}^TA^TA\vec{v}-\vec{v}^TA\vec{\beta}-2\vec{v}^TA^T\vec{\beta}-\vec{\beta}^T\vec{\beta}
-\end{align*}
-If $A$ is negative semi-definite, then $A^2$ is positive semi-definite and $A^TA$ is positive semi-definite.
-(Consider that $\vec{v}^TA^TA\vec{v}=(A\vec{v})^T(A\vec{v})$ so that $\vec{v}A^TA\vec{v}=0$ if $A\vec{v}=0$ and $\vec{v}^TA^TA\vec{v}<0$ otherwise.) Therefore, $-\vec{v}^TA^2\vec{v}\leq 0$ and $-\vec{v}^TA^TA\vec{v}\leq 0$.
-At equilibrium, $\vec{v}$ will be ``close" to $\vec{\beta}$ so that $$-\vec{v}^TA\vec{\beta}-2\vec{v}^TA^T\vec{\beta}\sim-3\vec{\beta}^TA\vec{\beta}$$
-\end{pf}
-
-\subsection{Numerically evaluating dynamics }
-%\begin{claim}
-Let $\Phi(t)$ be the fundamental matrix solution to the homogeneous differential equations $\frac{d\vec{x}}{dt}=A\vec{x}$ so that $\Phi(t)=V\Lambda(t)$ where the columns of $V$ are the eigenvectors of $A$ and $$\Lambda(t)=\text{diag}(e^{\lambda_1 t},\dots,e^{\lambda_n t})$$ where $\{\lambda_i\}$ are the eigenvalues of $A$.
-(Note that if $\lambda_i=0$ then $e^{\lambda _i t}=1$ for all $t$.) Let $V^\dagger$ be the pseudoinverse of $V$.
-Let $\vec{c}=V^\dagger\vec{X_0}$.
-Let $$G(t)=\text{diag}(1/\lambda_1(1-e^{-\lambda_1t}),\dots,1/\lambda_n(1-e^{-\lambda_nt}))V^\dagger\vec{\beta}.$$
-(If $\lambda_i=0$ then the corresponding entry of the first part of $G(t)$ will be given by $t$ rather than the form above.)  
-\begin{claim}
-If we let $x(t)=\Phi(t)G(t)+\Phi(t)\vec{c}$, $x(t)$ solves the inhomogeneous equations $\frac{dx}{dt}=A\vec{x}+\vec{\beta}$ with initial conditions $x(0)=\vec{X_0}$.
-\end{claim}
-
-\begin{pf}
-It is clear that
-\begin{align*}
-\frac{d\Phi\vec{c}}{dt}&=A\Phi(t)\vec{c} \text{ and } \Phi(0)\vec{c}=\vec{X_0}.
-\end{align*}
-Now, 
-\begin{align*}
-\Phi(t)G(t)&=V
-\left(\begin{array}{ccccc}
-e^{\lambda_1 t}  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots & e^{\lambda_nt}
- \end{array}\right)
- \left(\begin{array}{ccccc}
-\frac{1}{\lambda_1}(1-e^{\lambda_1 t})  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots &\frac{1}{\lambda_n} (1-e^{\lambda_nt})
- \end{array}\right)
- V^\dagger\vec{\beta}
- \\&=V
- \left(\begin{array}{ccccc}
-\frac{1}{\lambda_1}(e^{\lambda_1 t}-1)  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots &\frac{1}{\lambda_n} (e^{\lambda_nt}-1)
- \end{array}\right)
- V^\dagger\vec{\beta}
- \\ \Rightarrow \frac{d\Phi(t)G(t)}{dt}&=V
- \left(\begin{array}{ccccc}
-e^{\lambda_1 t}  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots &e^{\lambda_nt}
- \end{array}\right)V^\dagger\vec{\beta}
- \\&=V
- \left(\begin{array}{ccccc}
-e^{\lambda_1 t}  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots &e^{\lambda_nt}
- \end{array}\right)
- V^\dagger\vec{\beta} -VV^\dagger\vec{\beta}+\vec{\beta}
- \\&=V
- \left(\begin{array}{ccccc}
-e^{\lambda_1 t}-1  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots &e^{\lambda_nt}-1
- \end{array}\right)
- V^\dagger\vec{\beta} +\vec{\beta}
-  \\&=V
- \left(\begin{array}{ccccc}
-\lambda_1e^{\lambda_1t}\cdot \frac{1}{\lambda_1}(1-e^{-\lambda_1 t})  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots &\lambda_ne^{\lambda_nt}\cdot\frac{1}{\lambda_n}(1-e^{-\lambda_nt})
- \end{array}\right)
- V^\dagger\vec{\beta} +\vec{\beta}
-  \\&=AV\Lambda(t)
- \left(\begin{array}{ccccc}
- \frac{1}{\lambda_1}(1-e^{-\lambda_1 t})  &\dots &0
-\\ \vdots & \ddots &\vdots
-\\ 0 & \dots &\frac{1}{\lambda_n}(1-e^{-\lambda_nt})
- \end{array}\right)
- V^\dagger\vec{\beta} +\vec{\beta}
- \\&=AV\Lambda(t)G(t)+\vec{\beta}
- \\&=A\Phi(t)G(t)+\vec{\beta}
-\end{align*}
-Therefore $\frac{dx}{dt}=A\Phi(t)G(t)+A\Phi(t)\vec{c}+\vec{\beta}=Ax(t)+\vec{\beta}$ and since $G(0)=\vec{0}$, $x(0)=\vec{X_0}$.
-\end{pf}
+\begin{table}
+\caption{ \label{matrices} Table of matrices and variables used in the text.}
+\ra{1.3}
+\begin{tabular}{@{}lllll@{}}
+$A$ & adjacency matrix & $A_{ij}=\left\{\begin{array}{l l l}
+1 & \text{ if $i$ pays attention to $j$ }
+\\0 & \text{ otherwise }
+\end{array}\right.$
+\\$\bar{A}$ & normalized adjacency matrix & $\bar{A}_{ij}=\left\{\begin{array}{l l l }
+\frac{1}{K_i} & \text{ if $i$ pays attention to $j$, where $K_i=$ \# of birds $i$ pays attention to}
+\\ 0 & \text{ otherwise }
+\end{array}\right. $
+\\$B$ & diagonal signal matrix & $B_{ij}=\left\{\begin{array}{l l l}
+b & \text{ if $i=j$ and $i$ perceives the environmental signal }
+\\0 & \text{ otherwise }
+\end{array}\right. $
+\\$L$ & opinion dynamics w/o environmental signal &$L=\bar{A}-I$
+\\$L_\text{s}$ & opinion dynamics with environmental signal & $L_\text{s}=L-B$
+\\$\tilde{L}$ & dynamics of deviations from consensus w/o signal & $\tilde{L}=QLQ^T$
+\\$\tilde{L}_\text{s}$ & dynamics of deviations from consensus w/ signal & $\tilde{L}=RL_\text{s}R^T$
+\\$S$ & symmetrized interaction matrix & $S=\frac{1}{2}(\bar{A}+\bar{A}^T)$
+\\$P$ & energy interactions w/o environmental signal & $P_{ij}=\left\{\begin{array}{lll}
+S_{ij} & \text{ for } j\neq i
+\\-\sum_kS_{ik} & \text{ for } j=i
+\end{array}\right.
+$
+\\$P_\text{s}$ & energy interactions with environmental signal &  $P_\text{s}=P-B$
+\\$\tilde{P}$ 
+\\$\tilde{P}_\text{s}$
+\\$Q$ & rotates opinion vector away from consensus & $Q\in \R^{N-1}\times\R^N \text{ such that } Q\vec{1}=\vec{0}, \ QQ^T=I_{N-1}, \text{ and } Q^TQ=I_N-\frac{1}{N}\vec{1}_N\vec{1}_N^T $
+\\$R$ & full rotation matrix whose last row is $\frac{1}{\sqrt{N}}\vec{1}^T$ & $\left(\begin{array}{cc} Q \\ \frac{1}{\sqrt{N}}\vec{1}^T \end{array}\right)$
+\\$r$ & $\mathscr{H}_2$ robustness & $\lim_{t\to\infty}\frac{1}{E[||y||]}$ where $y=Qx$
+\\$W$ & 
+\end{tabular}
+\end{table}
 
@@
 \begin{figure}
-\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/corrlength_vs_ESS.pdf}
+\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/robustness_vs_ESS.pdf}
 \caption{\label{corrESS}
-The correlation length of a homogeneous flock depends on both the strategy being used and the radius of the signal, but it depends much more strongly on the radius than the strategy.
+%The correlation length of a homogeneous flock depends on both the strategy being used and the radius of the signal, but it depends much more strongly on the radius than the strategy.
 The ESS number of neighbors is an increasing function of correlation length of a homogeneous group.
@@
 \begin{figure}
-\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/greedyoptneighbors.pdf}
+%\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/greedyoptneighbors.pdf}
 \caption{\label{greedyopt} When the birds choose the optimal number of neighbors, given the strategies the rest of the flock are using, they settle on higher strategies when selection is due to predation than when selection is due to resources.

### 2015-03-12 paper_skeleton.tex
im about to change an old validation code file

@@
 \section{Appendix}
-%%% APPENDIX NOW DISCUSSES H2 WITHOUT FORCING (AS IN GEORGE'S PAPER), H2 WITH FORCING, WHICH GIVES CORRELATIONS BUT NOT FLUCTUATIONS AROUND AVERAGE OPINION, STAT MECH CORRELATIONS WITHOUT FORCING AND STAT MECH CORRELATION WITH FORCING, BOTH OF WHICH GIVE FLUCTUATIONS AROUND AVERAGE OPINION.
-SO H2 WITH FORCING IS STILL INCOMPARABLE TO ANYTHING ELSE SINCE ITS CORRELATIONS NOT FLUCTUATIONS AROUND AVERAGE OPINIONS.
-IT'S OKAY FOR NOW.
+
 \subsection{Linear algebra background }
@@
 
-If $z=Rx$, then the first $N-1$ coordinates of $z$ are $y=Qx$ and the last is $\sum_ix_i/\sqrt{N}$. The dynamics of $z$ follow
+If $\bar{y}=Rx$, then the first $N-1$ coordinates of $\bar{y}$ are $y=Qx$ and the last is $\sum_ix_i/\sqrt{N}$. The dynamics of $\bar{y}$ follow
 \begin{align*}
-\dot{z}&=R\dot{x}
+\dot{\bar{y}}&=R\dot{x}
 \\&=RL_\text{s}x+Rv+R\xi.
@@
 \begin{align*}
-\dot{z}&=RL_\text{s}R^Tz+Rv+R\xi
-\\&=\tilde{L}_\text{s}z+\tilde{v}+R\xi 
+\dot{\bar{y}}&=RL_\text{s}R^T\bar{y}+Rv+R\xi
+\\&=\tilde{L}_\text{s}\bar{y}+\tilde{v}+R\xi 
 \end{align*}
 where $\tilde{L}_\text{s}=RL_\text{s}R^T$ and $\tilde{v}=Rv=RB\vec{1}$.
-If $b\neq 0$, $L_\text{s}$ and therefore $\tilde{L}_\text{s}$ have all negative eigenvalues and $z=(0, 0, \dots 0, \sqrt{N})^T$ is a globally asymptotically stable equilibrium.
-Therefore, by Claim \ref{lyapeq_derivation} and Theorem \ref{lyapeq_theorem}, $\lim_{t\to\infty}E[z(t)z(t)^T]=\Sigma_z$ where $\tilde{L}_\text{s}\Sigma+\Sigma\tilde{L}_\text{s}^T=-RR^T$.
-We can then find $\Sigma_y=\lim_{t\to\infty}E[y(t)y(t)^T]$
-by removing the last row and column of $\Sigma_z$.
+If $b\neq 0$, $L_\text{s}$ and therefore $\tilde{L}_\text{s}$ have all negative eigenvalues and $\bar{y}=(0, 0, \dots 0, \sqrt{N})^T$ is a globally asymptotically stable equilibrium.
+Therefore, by Claim \ref{lyapeq_derivation} and Theorem \ref{lyapeq_theorem}, $\lim_{t\to\infty}E[\bar{y}(t)\bar{y}(t)^T]=\Sigma_{\bar{y}}$ where $\tilde{L}_\text{s}\Sigma_{\bar{y}}+\Sigma_{\bar{y}}\tilde{L}_\text{s}^T=-RR^T=-I_N$.
+We can  find $\Sigma_y=\lim_{t\to\infty}E[y(t)y(t)^T]$
+by removing the last row and column of $\Sigma_{\bar{y}}$.
 
 If $b=0$, $L$ has a $0$ eigenvalue corresponding to the eigenvector $\vec{1}$.
-Therefore, $RLR^T$ also has a $0$ eigenvalue, so we can therefore no longer find the desired $\Sigma_z$.
+Therefore, $RLR^T$ also has a $0$ eigenvalue, so we can therefore no longer find the desired $\Sigma_{\bar{y}}$.
 Since the last row of $R=\frac{1}{\sqrt{N}}\vec{1}$, all entries of the the last column of $LR^T$ are $0$.
 Therefore, all entries of the last column of $RLR^T$ are $0$.
-In other words, the dynamics of the first $N-1$ elements of $z$, i.e.
+In other words, the dynamics of the first $N-1$ elements of $\bar{y}$, i.e.
 $y$, do not depend on the last.
@@
 \begin{align*}
-\dot{y}&=QLQ^Ty+Q\xi=\tilde{L}+Q\xi
+\dot{y}&=QLQ^Ty+Q\xi=\bar{L}+Q\xi
 \end{align*}
-where $\tilde{L}=QLQ^T$. Another way to see this is to write 
+where $\bar{L}=QLQ^T$. Another way to see this is to write 
 \begin{align*}
@@
  \\&=QLQ^TQx+Q\xi
- \\&=\tilde{L}y+Q\xi.
+ \\&=\bar{L}y+Q\xi.
 \end{align*}
-Again, by Claim \ref{lyapeq_derivation} and Theorem \ref{lyapeq_theorem}, we can now find $\lim_{t\to\infty}E[y(t)y(t)^T]=\Sigma_y$ by finding a solution to $\tilde{L}\Sigma_y+\Sigma_y\tilde{L}=-QQ^T$.
+Again, by Claim \ref{lyapeq_derivation} and Theorem \ref{lyapeq_theorem}, we can now find $\lim_{t\to\infty}E[y(t)y(t)^T]=\Sigma_y$ by finding a solution to $\bar{L}\Sigma_y+\Sigma_y\bar{L}=-QQ^T=-I_{N-1}$.
 
-To summarize, in the absence of an environmental signal, $\Sigma_y$ satisfies $\tilde{L}\Sigma_y+\Sigma_y\tilde{L}=-QQ^T$ where $\tilde{L}=QLQ^T$ and the presence of an environmental signal, $\Sigma_y$ can be found in the first $N-1$ rows and columns of $\Sigma_z$ where $\Sigma_z$ satisfies $\tilde{L}_\text{s}\Sigma_z+\Sigma_z\tilde{L}_\text{s}=-RR^T$ where $\tilde{L}_\text{s}=RL_\text{s}R^T$.
+To summarize, in the absence of an environmental signal, $\Sigma_y$ satisfies $\bar{L}\Sigma_y+\Sigma_y\bar{L}^T=-I_{N-1}$ where $\bar{L}=QLQ^T$ and in the presence of an environmental signal, $\Sigma_y$ can be found in the first $N-1$ rows and columns of $\Sigma_{\bar{y}}$ where $\Sigma_{\bar{y}}$ satisfies $\tilde{L}_\text{s}\Sigma_{\bar{y}}+\Sigma_{\bar{y}}\tilde{L}_\text{s}^T=-I_N$ where $\tilde{L}_\text{s}=RL_\text{s}R^T$.
 
 \subsection{Probability distribution at equilibrium }
+%Specifically, an individual's opinion $\sigma_i$ can be either $1$ or $-1$ and $\sigma_i$ switches with probability $w_i(\sigma_i)=\frac{1}{2}(-\sigma_i\sum_jA_{ij}\sigma_j )(1-\beta_i\sigma_i)$ where $A_{ij}$ indicates the strength of the connection from $j$ to $i$ and $\beta_i$ indicates the strength of the environmental stimulus perceived by $i$.
+The stochastic dynamics of the opinions can be written down with a master equation \cite{Glauber:1963fk}.
+The vector $\vec{v}(t)$ of expected values, $q_i(t)=\langle \sigma_i(t)\rangle$, satisfies the following differential equations: \cite{Glauber:1963fk}
 A quite different way of modeling social interactions the method used by Bialek et al.
@@
 Therefore, $E[\langle c \rangle ]=\vec{0}$ and $E[\tilde{c}\tilde{c}^T]=(-\tilde{\Lambda})^{-1}$.
-Therefore, $C=E[zz^T]=E[\tilde{W}\tilde{c}\tilde{c}^T\tilde{W}^T]=-\tilde{W}\tilde{\Lambda}\tilde{W}^T$.
-By dividing each entry $C_{ij}$ of the covariance matrix by $C_{ii}C_{jj}$, the covariance matrix can be used to find the correlation matrix.
+Therefore, $C_z=E[zz^T]=E[\tilde{W}\tilde{c}\tilde{c}^T\tilde{W}^T]=-\tilde{W}\tilde{\Lambda}^{-1}\tilde{W}^T$.
+By dividing each entry $C_{zij}$ of the covariance matrix by $C_{zii}C_{zjj}$, the covariance matrix can be used to find the correlation matrix.
 
@@
 
-%\begin{align*}
-%P(z)&=\int_\R P(z,V)dV
-%\\&=\int_\R \frac{1}{Z}\exp\left(-\frac{1}{2}((V-1)\sqrt{N}e_N+y)^T\Tilde{P}_\text{s}((V-1)\sqrt{N}e_N+y)\right)dV
-%\\&=\frac{1}{Z}\exp(-\frac{1}{2}y^T\Tilde{P}_\text{s}y)\int_\R \exp\left(-\frac{1}{2}\left(\Tilde{P}_{\text{s}NN}N\left(V-1\right)^2+(V-1)\sqrt{N}\sum_{i=1}^{N-1}\Tilde{P}_{\text{s}Ni}y_i+(V-1)\sqrt{N}\sum_{i=1}^{N-1}y_i\Tilde{P}_{\text{s}iN}\right)\right)dV
-%\\&=\frac{1}{Z}\exp(-\frac{1}{2}y^T\Tilde{P}_\text{s}y)\int_\R \exp\left(-\frac{1}{2}\left(\Tilde{P}_{\text{s}NN}N\left(V-1\right)^2+2(V-1)\sqrt{N}\sum_{i=1}^{N-1}\Tilde{P}_{\text{s}Ni}y_i\right)\right)dV \text{ since $\Tilde{P}_\text{s}$ is symmetric }
-%\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^T\Tilde{P}_\text{s}y-\left(\sum_{i=1}^{N-1}\Tilde{P}_{\text{s}Ni}y_i/\sqrt{\Tilde{P}_{\text{s}NN}}\right)^2\right)\right)\times
-%\\&\int_\R \exp\left(-\frac{1}{2}\left(\Tilde{P}_{\text{s}NN}N\left(V-1\right)^2+2(V-1)\sqrt{N}\sum_{i=1}^{N-1}\Tilde{P}_{\text{s}Ni}y_i+\left(\sum_{i=1}^{N-1}\Tilde{P}_{\text{s}Ni}y_i/\sqrt{\Tilde{P}_{\text{s}NN}}\right)^2\right)\right)dV
-%\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^T\Tilde{P}_\text{s}y-\left(\sum_i\Tilde{P}_{\text{s}Ni}y_i/\sqrt{\Tilde{P}_{\text{s}NN}}\right)^2\right)\right)\int_\R \exp\left(-\frac{1}{2}\left(\sqrt{\Tilde{P}_{\text{s}NN}}\sqrt{N}\left(V-1\right)+\sum_i\Tilde{P}_{\text{s}Ni}y_i/\sqrt{\Tilde{P}_{\text{s}NN}}\right)^2\right)dV
-%\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^T\Tilde{P}_\text{s}y-\left(\sum_i\Tilde{P}_{\text{s}Ni}y_i/\sqrt{\Tilde{P}_{\text{s}NN}}\right)^2\right)\right)\int_\R \exp\left(-\frac{1}{2}\tilde{P}_{\text{s}NN}N\left(V-1+\frac{\sum_i\tilde{P}_{\text{s}Ni}y_i}{\tilde{P}_{\text{s}NN}\sqrt{N}}\right)^2\right)dV
-%\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^T\tilde{P}_{\text{s}}y-\left(\sum_i\tilde{P}_{\text{s}Ni}y_i/\sqrt{\tilde{P}_{\text{s}NN}}\right)^2\right)\right)\times\sqrt{\frac{2\pi}{\tilde{P}_{\text{s}NN}N}}
-%\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^T\tilde{P}_{\text{s}}y-\frac{1}{\tilde{P}_{\text{s}NN}}y^T\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot}y\right)\right)\times\sqrt{\frac{2\pi}{\tilde{P}_{\text{s}NN}N}} \text{ where $\tilde{P}_{\text{s}N\cdot}$ is the last row of $\tilde{P}_{\text{s}}$}
-%\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\left(y^T(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot})y\right)\right)\times\sqrt{\frac{2\pi}{\tilde{P}_{\text{s}NN}N}}
-%\\&=\frac{\sqrt{2\pi}}{\sqrt{(2\pi)^N\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}N}}\exp\left(-\frac{1}{2}\left(y^T(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot })y\right)\right)
-%\\&=\frac{1}{\sqrt{(2\pi)^{N-1}\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}N}}\exp\left(-\frac{1}{2}z^TR^T(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot})Rz\right)
-%\\&=\frac{1}{\sqrt{(2\pi)^{N-1}\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}N}}\exp\left(-\frac{1}{2}z^T(-P_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}}R^T\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot}R)z\right)
-%%\\&=\frac{\sqrt{2\pi}\sqrt{R_{11}N}}{(2\pi)^{N/2}\det(-P_\text{s}^{-1})^{1/2}}\exp\left(-\frac{1}{2}z^T(-P_\text{s}-\frac{1}{R_{11}}Q'^{-1}R_{\cdot 1}R_{1\cdot}(Q'^{-1})^T)z\right)
-%\\&=\frac{1}{\sqrt{(2\pi)^{N-1}\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}N}}\exp\left(-\frac{1}{2}z^T(-P_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}}P_\text{s}R^T_{\cdot N}R_{N\cdot}P_\text{s})z\right)
-%\\&=\frac{1}{\sqrt{(2\pi)^{N-1}\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}N}}\exp\left(-\frac{1}{2}z^T(-P_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}}P_\text{s}(R_{N\cdot})^TR_{N\cdot}P_\text{s})z\right)
-%%\\&=\frac{\sqrt{2\pi}\sqrt{R_{11}N}}{(2\pi)^{N/2}\det(-P_\text{s}^{-1})^{1/2}}\exp\left(-\frac{1}{2}z^T(-P_\text{s}-\frac{1}{R_{11}}P_\text{s}\Pi P_\text{s})z\right) \text{ where $\Pi_{ij}=1/N$ for all $i,j$}
-%\\&=\frac{1}{\sqrt{(2\pi)^{N-1}\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}N}}\exp\left(-\frac{1}{2}z^T(-P_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}N}P_\text{s}\vec{1}\vec{1}^T P_\text{s})z\right) 
-%\\&=\frac{1}{\sqrt{(2\pi)^{N-1}\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}N}}\exp\left(-\frac{1}{2}z^T(-P_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}N}B\vec{1}\vec{1}^T B)z\right)  \text{ since $P\vec{1}=\vec{0}$}
-%\end{align*}
-
 \begin{align*}
@@
 \\&=\frac{\sqrt{N}}{Z}\exp\left(-\frac{1}{2}\left(y^T(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot})y\right)\right)\times\sqrt{\frac{2\pi}{\tilde{P}_{\text{s}NN}N}}
-%\\&=\frac{\sqrt{2\pi}}{\sqrt{(2\pi)^N\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}}}\exp\left(-\frac{1}{2}\left(y^T(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot })y\right)\right)
 \\&=\frac{1}{\sqrt{(2\pi)^{N-1}\det(-P_\text{s}^{-1})\tilde{P}_{\text{s}NN}}}\exp\left(-\frac{1}{2}\left(y^T(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot })y\right)\right)
@@
 \begin{equation*}
-E[zz^T]=R^T\left(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot }\right)^{-1}R=\left(P_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}N}B\vec{1}\vec{1}^TB\right)^{-1}
+E[zz^T]=R^T\left(\tilde{P}_{\text{s}}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot }\right)^{-1}R=\left(-P_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}N}B\vec{1}\vec{1}^TB\right)^{-1}
 \end{equation*}
@@
 \begin{equation}
-E[zz^T]=\left(P_\text{s}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB\right)^{-1}
+C_z=E[zz^T]=\left(-P_\text{s}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB\right)^{-1}
 \end{equation}
@@
 \subsection{Connecting two frameworks }
-To find $\mathscr{H}$, we found the matrix $\Sigma$ such that 
+To find $\mathscr{H}$, we found the matrix $\Sigma_y$ such that 
 \begin{equation}
-QLQ^T\Sigma+\Sigma QLQ^T=-I.
+QLQ^T\Sigma_y+\Sigma_y QLQ^T=-I.
 \end{equation}
@@
 Therefore, $H(v)$ decreases at all $v$ if and only if $D^TP+PD$ is a positive matrix.
-%MORE?!?
 
-%Specifically, an individual's opinion $\sigma_i$ can be either $1$ or $-1$ and $\sigma_i$ switches with probability $w_i(\sigma_i)=\frac{1}{2}(-\sigma_i\sum_jA_{ij}\sigma_j )(1-\beta_i\sigma_i)$ where $A_{ij}$ indicates the strength of the connection from $j$ to $i$ and $\beta_i$ indicates the strength of the environmental stimulus perceived by $i$.
-The stochastic dynamics of the opinions can be written down with a master equation \cite{Glauber:1963fk}.
-The vector $\vec{v}(t)$ of expected values, $q_i(t)=\langle \sigma_i(t)\rangle$, satisfies the following differential equations: \cite{Glauber:1963fk}
-%\begin{equation*}
-%\frac{d\vec{v}}{dt}=M\vec{v}+\vec{\beta}^T(\vec{1}-\vec{v}).
-%\end{equation*}
-%We simplify this by writing $\frac{d\vec{v}}{dt}=A\vec{v}+\vec{\beta}^T\vec{1}$, where $A=M-B$ and $B$ is a diagonal matrix with the values of $\vec{\beta}$ along the diagonal. 
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
+Table \ref{correspondence} shows how to find the appropriate covariance matrices depending on the framework being used and whether or not a signal is present.
+If we use $P$ instead of $L$ with the noisy dynamics framework, $C_y=2\Sigma_y$.
+%MORE HERE!!!!
+
+ \[
+ \boxed{\bar{y}=Rx$, $y=Qx=\bar{y}_{i=1}^{N-1}$ (or $y=RQ^TQx$), $z=Q^TQx$, $V=\frac{1}{N}\sum_ix_i}
+ \]
+
+\begin{table}
+\caption{\label{correspondence} $P^\dagger$ is the inverse of $P$ with the $0$ eigenvector / value removed. }
+\ra{1.3}
+\begin{tabular}{@{}l|llll@{}}
+& Without signal & With signal 
+\\\hline From noisy dynamics &  $\bar{L}\Sigma_y+\Sigma_y\bar{L}^T=-I_{N-1}$ & $\tilde{L}_\text{s}\Sigma_{\bar{y}}+\Sigma_{\bar{y}}\tilde{L}_\text{s}^T=-I_N$, $\Sigma_y=(\Sigma_{\bar{y}})_{i,j=1}^{N-1}$
+\\\hline From Boltzmann equilibrium & $C_y=QP^{\dagger}Q^T$ & $C_y=(\tilde{P}_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot })^\dagger$
+\\ & & or $C_y=Q(-P_\text{s}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB)^{\dagger}Q^T$
+\end{tabular}
+\end{table}
 
@@
 \end{array}\right. $
+\\ $\Lambda$ & diagonal matrix of eigenvalues of $P$
 \\$L$ & opinion dynamics w/o environmental signal &$L=\bar{A}-I$
 \\$L_\text{s}$ & opinion dynamics with environmental signal & $L_\text{s}=L-B$
-\\$\tilde{L}$ & dynamics of deviations from consensus w/o signal & $\tilde{L}=QLQ^T$
+\\$\bar{L}$ & dynamics of deviations from consensus w/o signal & $\bar{L}=QLQ^T$
 \\$\tilde{L}_\text{s}$ & dynamics of deviations from consensus w/ signal & $\tilde{L}=RL_\text{s}R^T$
@@
 \\$P_\text{s}$ & energy interactions with environmental signal &  $P_\text{s}=P-B$
-\\$\tilde{P}$ 
 \\$\tilde{P}_\text{s}$
@@
 \\$r$ & $\mathscr{H}_2$ robustness & $\lim_{t\to\infty}\frac{1}{E[||y||]}$ where $y=Qx$
-\\$W$ & 
+\\$W$ & matrix of eigenvectors of $P$
 \end{tabular}

### 2015-03-19 paper_skeleton.tex
reworking on description of math in the paper

@@
 
+\section{Introduction }
+Questions:
+\begin{enumerate}
+\item How are the two group properties---$\mathscr{H}_2$ robustness and correlation length---related?
+\item What are the equilibrium strategies of individuals seeking to optimize their own fitness?
+\item What group properties emerge from individual-level optimization?
+\end{enumerate}
+
 \section{Model }
-\subsection{Learning dynamics }
-An individual's opinion changes with a probability that depends on its social interactions and the strength of the environmental stimulus. Let $A$ be the adjacency matrix such that
+
+\subsection{Modeling interactions and group-level properties }
+There are two ways to describe equilibrium properties of a flock sharing information over a network.
+The first is to write down a set of differential equations with noise to describe how each bird's opinion changes over time and then to analyze the distribution of these opinions after the dynamics have reached equilibrium \cite{Young:2010fk,Young:2013kx}.
+The second is to write down an explicitly probability distribution of the opinions at equilibrium and to analyze this distribution.
+WIth either approach, we consider a vector of opinions $v$ where $v_i$ a a scalar that represents the $i^\text{th}$'s bird opinion about an unspecified behavior, for example where or how fast to fly.
+Both approaches rely on a network of connections between the birds, which we will describe first.
+Let $A$ be the adjacency matrix such that
 
@@
 \end{equation}
+If $n_i$ is the number of neighbors node $i$ has, then $n_i=\sum_kA_{ik}$.
+We assume that having more neighbors means paying less attention to each so we normalize $A$ to find $\bar{A}$: $\bar{A}_{ij}=\frac{A_{ij}}{n_i}$.
 
-We assume that having more neighbors means paying less attention to each so we normalize $A$ to find $\bar{A}$:
+The Laplacian of this matrix is given by $L=\bar{A}-I$, which is symmetric only if $A$ is.
+We can symmetrize the interactions by defining $S=\frac{1}{2}(A+A^T)$ and $P$ such that $P_{ij}=S_{ij}$ for $j\neq i$ and $P_{ii}=-\sum_kS_{ik}$.
+Then $L$ and $P$ share the properties that $L\vec{1}=P\vec{1}=\vec{0}$ and that each is negative semi-definite, with the difference being that $P$ is necessarily symmetric.
+It is more common to use $L$ in a dynamics system approach, so that
 \begin{equation}
-\bar{A}_{ij}=\frac{A_{ij}}{\sum_kA_{ik}}.
+\dot{v}=Lv+D\xi \label{opinion_dynamics}
 \end{equation} 
+where $D$ is a diagonal matrix with entires $\{\sqrt{n_i}\}$ and $E[\xi(t)]=\vec{0}$, $E[\xi(t)\xi(s)^T]=\frac{1}{2}\delta(t-s)I$, and $E[v(0)\xi(t)^T]=\vec{0}$.
+This describes a system in which each bird changes averages the differences between its own opinions and those of its neighbors and changes its opinion to minimize these differences.
+Our choice of $D$ assumes that birds with more neighbors make noisier estimates of how to change their opinions.
 
-If each bird averages the differences between its opinion and each of its neighbors' to update its opinion, the opinion dynamics will be described by 
+The Boltzmann distribution in statistical physics is a standard way in which to write an equilibrium distribution.
+In physics, those states with the highest energy are least likely.
+In our system, those sets of opinions that are least in consensus will be less likely.
+Specifically, we will use the following expression to represent the ``energy'' of an opinion vector
 \begin{equation}
-\dot{x}=Lx
+H(v)=-v^TPv \label{Hamiltonian}
 \end{equation}
-where $L=\bar{A}-I$.
-In the absence of the environmental signal, there is no trend in the dynamics.
-We introduce an environmental signal by changing the dynamics such that informed birds change their opinions to be closer to $1$.
-Having an opinion closer to $1$ can be interpreted as having more information about the environment and having a lower expected value can be interpreted as having less information.
-Specifically, we incorporate this environmental forcing with a diagonal matrix $B$ where $B_{ii}=b$ if $i$ perceives the environmental signal and $B_{ii}=0$ otherwise.
-The dynamics then become
+and the Boltzmann distribution
 \begin{equation}
-\dot{x}=Lx+B(\vec{1}-x).
+P(v)=\frac{1}{Z}\exp\left(-\frac{1}{2}H(v)\right) \label{Boltzmann}
 \end{equation}
-If we let $L_\text{s}=L-B$, then we can simplify this by writing
+where $Z=\int_\R^N\exp\left(-\frac{1}{2}H(v)\right)dv$.
+
+With the dynamical approach, in the absence of noise, the system would reach consensus with each $v_i=\alpha$ for some $\alpha\in\R$.
+Similarly, with the statistical physics approach, consensus states with each $v_i=\alpha$ are most likely.
+However, all consensus values $\alpha$ are equally likely.
+It is therefore difficult to describe correlations between the bird's opinions at equilibrium.
+It becomes easier, however, to consider deviations from consensus.
+If all $v_i$ are equal, then $v=\alpha\vec{1}$.
+To consider deviations from this consensus state, it is useful to define a matrix $Q\in\R^{N-1}\times\R^N$ such that each row of $Q$ is orthogonal to $\vec{1}$ and the rows are pairwise orthogonal.
+Young et al.
+show in \cite{Young:2010fk,Young:2013kx} that this is equivalent to a $Q$ such that $QQ^T=I_{N-1}$ and $Q^TQ=I_N-\frac{1}{N}\vec{1}\vec{1}^T$.
+To simply consider absolute deviations from the average opinion of the flock, we can define $z=v-\langle v\rangle \vec{1}=Q^TQv$, where $\langle v \rangle =\frac{v_i}{N}$.
+To consider the component of the opinion vector that is orthogonal to consensus, we can define $y=Qv=Qz$.
+Let $\Sigma_z=\lim_{t\to\infty}E[zz^T]$ and $\Sigma_y=\lim_{t\to\infty}E[yy^T]$ using the dynamical approach and $C_z=E[zz^T]$ and $C_y=E[yy^T]$ using the statistical physics approach.
+Since $y=Qz$ and $z=Q^Ty$, in either case, we can translate easily between $\Sigma_z$ and $\Sigma_y$ or between $C_z$ and $C_y$.
+In the Appendix, we derive equations for each of these covariance matrices and summarize the expression for the $y$ covariance matrices in Table \ref{correspondence}.
+The first thing to note is that, if instead of using $L$ we use $P$ in the dynamical approach, then $C_y=2\Sigma_y$, as shown in Claim XXX.
+There is therefore a close correspondence between the two approaches.
+
+These covariance matrices allow us to define two group properties.
+The first is $\mathscr{H}_2$ robustness, as used in \cite{Young:2010fk,Young:2013kx}.
+This is a measure of the robustness of the consensus state in which all birds have the same opinion to noise at equilibrium.
+Specifically,
 \begin{equation}
-\dot{x}=L_\text{s}x+\vec{B}
+\rho=\frac{1}{\lim_{t\to\infty}E[||y||]}=\frac{1}{\lim_{t\to\infty}\Tr(\Sigma_y)}. \label{H2def}
 \end{equation}
-where $\vec{B}=B\vec{1}$ indicates which birds perceive the environmental signal.
+The second property is the correlation length of the flock.
+This is the distance over which birds' opinions tend to be positively correlated with each other.
+If we know the distances between the birds and the covariance matrix $C_z$, we can plot the correlations $\{E[z_iz_j]\}$ between each pair of birds' deviations from consensus against the distance between the birds and then identify the correlation length $\lambda$ by finding the distance at which these correlations become $0$.
+
+So far, we have considered birds whose opinions only depend on the opinions of the rest of the flock.
+We can additionally incorporate environmental information by assuming that there is an environmental signal such that more informed birds have opinions closer to $1$.
+If we let $B$ be a matrix with $B_ii=1$ if $i$ can perceive the signal and $0$ otherwise, then the dynamical system becomes
+\begin{equation}
+\dot{v}=Lv+B(\vec{1}-v)+\xi =L_\text{s}v+B\vec{1}+n_i\xi \label{opinion_dynamics2}
+\end{equation}
+where $L_\text{s}=L-B$. Similarly, the statistical physics model becomes
+\begin{equation}
+P(v)=\frac{1}{Z}\exp\left(-\frac{1}{2}(v-\vec{1})^T(-P+B)(v-\vec{1})\right)=\frac{1}{Z}\exp\left(\frac{1}{2}(v-\vec{1})^T(-P_\text{s})(v-\vec{1})\right)
+\end{equation}
+where $P_\text{s}=P-B$.
+Again, we can find the covariance matrices $\Sigma_z$, $\Sigma_y$, $C_z$, and $C_y$.
+As was the case without a signal, if we use $P$ in the dynamical approach, then we find that $C_y=2\Sigma_y$.
+We define $\rho$ and $\lambda$ as above using the matrices $\Sigma_y$ and $C_z$ in the presence of a signal.
 
 \subsection{Individual-level fitness }
-To find the optimal strategies from the individuals' perspectives, we need a measure of individual-level fitness, which we define as follows.
-Each individual has a strategy that determines how many other individuals he pays attention to, $n_i=\sum_{j\neq i}{\bf I}(A_{ij}\neq 0)$.
-We distribute the individuals randomly in space.
+To find the optimal strategies from the individuals' perspectives, we need a measure of individual-level fitness.
+We use the dynamical system to describe how well each bird learns about the environmental signal.
+For a given matrix $L$, the expected opinion $q_i$ of each individual satisfies the equations
+\begin{equation}
+\dot{q}=L_\text{s}q+B\vec{1}. \label{expected_opinion}
+\end{equation}
+To generate a matrix $L$, we distribute the individuals randomly in space (specifically, in a unit square).
+has a strategy $n_i$ that dictates how many neighbors it has, so that $n_i=\sum_{j\neq i}{\bf I}(A_{ij}\neq 0)$.
 Each individual then pays attention to its $n_i$ nearest neighbors.
-One individual is chosen to be a receiver; it and all those individuals within a radius $r$ of the receiver perceive an external signal of strength $\beta$.
-We find the expected opinion of each individual $q_i$ after a period of time $T$.
-We refer to each such set of random positions and receiver as a signaling event.
+One individual is chosen to be a receiver; it and all those individuals within a radius $r$ of the receiver perceive, so that $B_{ii}=1$ for all those birds within this radius.
+We do this $1000$ times and find $q(1)$ for each matrix, which we refer to as a signaling event.
+This is meant to simulate a situation in which such that information is present only intermittently and otherwise the birds are moving randomly through space and restructuring their network.
 
-We consider two types of signals: predators and food.
+We consider signals that provide information about two types of content: predators and food.
 If the signal is information that a predator is present, we assume that whichever individual is least aware of the signal will be predated, i.e.
 the individual $i$ such that $q_i$ is lowest.
-We therefore find, over many signaling events, the probability that an individual will be eaten and an individual's fitness is given by the probability of surviving, i.e.
-$1$ minus the probability of being predated.
+Then, fitness is proportional to the probability of surviving, i.e.
+$1$ minus the probability of being the least informed over many signaling events.
 If the signal is information that resources are available, we assume that whichever individual is most aware of the signal will get access to the resource, i.e.
 the individual $i$ such that $q_i$ is highest.
-In this case, fitness is proportional to the probability of being the first to reach the resources.
-
-\subsection{Group-level performance }
-%%There are two ways to describe equilibrium properties
-We consider two group-level properties that indicate how cohesive the group is.
-The first property is $\mathscr{H}_2$ robustness, as used in \cite{Young:2010fk,Young:2013kx}.
-This is a measure of the robustness of the consensus state in which all birds have the same opinion to noise at equilibrium, as used in (see Appendix Section XXX).
-The second property is the correlation length of the flock.
-This is the distance at which the average correlations between birds changes from positive to negative, i.e.
-the distance over which birds' opinions tend to be positively correlated with each other (see Appendix Section XXX).
+In this case, fitness is proportional to the probability of getting the resource, i.e.
+the probability of being the most informed over many signaling events.
+The behavior that the birds can change to improve their fitness is the number of other birds they pay attention to, $n_i$.
+In particular, each individual has a strategy $n_i$ that dictates how many neighbors it has, so that $n_i=\sum_{j\neq i}{\bf I}(A_{ij}\neq 0)$.
 
@@
 
-\subsection{ESS strategies }
+\subsection{Robustness and correlation length are positively related. }
+In a homogeneous flock, where every bird is using the same strategy, as both the strategy and the radius of the signal change, both $\mathscr{H}_2$ and the correlation length are affected.
+As the Figure \ref{compare_props} shows that there is a relationship between two group properties: the $\mathscr{H}_2$ norm and the correlation length of the flock.
+
+\subsection{There are multiple ESS, which depend on the content of the signal. }
 Surprisingly, even without imposing costs on paying attention to more neighbors, paying attention to the whole flock is not usually the best strategy.
@@
 
+\subsection{The predation ESS is to pay attention to more neighbors when the signal is more public.}
 These opposing forces caused by the costs of not paying attention to informed neighbors  and the costs of paying attention to more uninformed neighbors  also explains the second feature of the ESS strategies: they are non-monotonic with respect to the the radius of the signal, i.e.
@@
 
-Another way to explain the non-monotonic ESS number of neighbors is to consider the correlation length of a homogeneous flock as a function of the radius of the signal.
-If a flock is all using the same strategy, its average correlation length depends much more strongly on the radius of the signal than the strategy being used (Figure \ref{corrESS}a).
-In particular, the correlation length first rises and then falls as the radius is increased.
-The rise and fall of the correlation length correspond to the rise and fall of the ESS number of neighbors and if the we plot the ESS number of neighbors for various radii of the signal against the average correlation length of a homogeneous flock for that radii, we find a positive relationship (Figure \ref{corrESS}b).
-This is because when correlations in information die off very quickly (i.e.
-a low correlation length), having more neighbors is costly because birds even slightly away from the informed birds will have poor information.
-When there are strong correlations in information across the flock, on the other hand, these costs are reduced because every bird in the flock has on average higher quality information.
+Figure \ref{ESS_v_H2} shows how the ESS strategy changes as a function of the $\mathscr{H}_2$ robustness.
 
@@
 Under selection due to predation, the $\mathscr{H}_2$ robustness and correlation length of the flock tend to decrease, whereas under selection due to resources, $\mathscr{H}_2$ robustness is relatively constant and correlation length tends to increase (Figure \ref{greedyopt}).
-
-\subsection{Relationship between group properties }
-Figure \ref{greedyopt} shows that there is a relationship between two group properties: the $\mathscr{H}_2$ norm and the correlation length of the flock.
 
@@
 
+We define a matrix $Q\in\R^{N-1}\times\R^N$ such that each row of $Q$ is orthogonal to $\vec{1}$ and the rows are pairwise orthogonal.
+Young et al.
+show in \cite{Young:2010fk} show that this equivalent to finding $Q$ such that $Q\vec{1}=\vec{0}$, $QQ^T=I_{N-1}$, and $Q^TQ=I_N-\frac{1}{N}\vec{1}_N\vec{1}_N^T$.
+Since the rows of $R$ are orthonormal, $RR^T=I$.
+Further, since the rows of $R$ are orthonormal, $R$ is invertible, which implies that $R^T=R^{-1}$.
+Then we can write
+
+Again, we will use a rotation matrix $R\in\R^N\times\R^N$ such that the last row of $R$ is $\frac{1}{\sqrt{N}}\vec{1}$ and the other rows are normal and orthogonal to $\vec{1}$.
+Above, we found a $Q\in\R^{N-1}\times\R^N$ such that each row of $Q$ was orthogonal to $\vec{1}$ and $R\in\R^N\times\R^N$ be the matrix with $\frac{1}{\sqrt{N}}\vec{1}^T$ as its last row and the first $N-1$ rows identical to the rows of $Q$.
+Then $R$ is a rotation matrix, so $RR^T=R^TR=I$ and $\det(R)=1$.
+This allows us to
+\[
+ \boxed{\bar{y}=Rx$, $y=Qx=\bar{y}_{i=1}^{N-1}$ (or $y=RQ^TQx$), $z=Q^TQx$, $V=\frac{1}{N}\sum_ix_i}
+ \]
+
 \subsection{Opinion dynamics with noise }
@@
 We can find this distance by rotating the opinion vector into the space orthogonal to the consensus vector, $\vec{1}$.
-We define a matrix $Q\in\R^{N-1}\times\R^N$ such that each row of $Q$ is orthogonal to $\vec{1}$ and the rows are pairwise orthogonal.
-Young et al.
-show in \cite{Young:2010fk} show that this equivalent to finding $Q$ such that $Q\vec{1}=\vec{0}$, $QQ^T=I_{N-1}$, and $Q^TQ=I_N-\frac{1}{N}\vec{1}_N\vec{1}_N^T$.
 If $y=Qx$, then $||y||=y^Ty$ describes how far away from consensus $x$ is.
@@
 Young et al.
-\cite{Young:2010fk} model opinion dynamics by assuming that each bird averages the differences between its opinion and each of the opinions of its neighbors and changes its opinion to minimize this difference and that each bird's opinion is subject to independent noise of unit intensity.
+\cite{Young:2010fk} model opinion dynamics by assuming that each bird averages the differences between its opinion and each of the opinions of its neighbors and changes its opinion to minimize this difference and that each bird's opinion is subject to independent noise with variance proportional to the number of neighbors each bird has.
 The equations describing these dynamics can be written as
 \begin{equation}
-\dot{x_i}=\frac{\sum_{\text{neighbors $j$ of $i$ } }(x_j-x_i)}{K_i} +\xi_i=\sum_j\bar{A}_{ij}x_j-x_i+\xi_i=\sum_jL_{ij}x_j+\xi_i.
+\dot{x_i}=\frac{\sum_{\text{neighbors $j$ of $i$ } }(x_j-x_i)}{K_i} +\sqrt{n_i}\xi_i=\sum_j\bar{A}_{ij}x_j-x_i+\sqrt{n_i}\xi_i=\sum_jL_{ij}x_j+\sqrt{n_i}\xi_i.
 \end{equation}
@@
 \begin{equation}
-\dot{x}=Lx+\xi.
+\dot{x}=Lx+D\xi.
 \end{equation}
@@
 \begin{equation}
-\dot{x}=Lx+B(\vec{1}-x)=L_\text{s}x+v.
+\dot{x}=Lx+B(\vec{1}-x)+D\xi=L_\text{s}x+v+D\xi.
 \end{equation}
@@
 
-If $\bar{y}=Rx$, then the first $N-1$ coordinates of $\bar{y}$ are $y=Qx$ and the last is $\sum_ix_i/\sqrt{N}$. The dynamics of $\bar{y}$ follow
+If $\bar{y}=Rx$, then the first $N-1$ coordinates of $\bar{y}$ are $y=Qx$ and the last is $\sum_ix_i/\sqrt{N}$. The dynamics of $\bar{y}$ are described by 
 \begin{align*}
 \dot{\bar{y}}&=R\dot{x}
-\\&=RL_\text{s}x+Rv+R\xi.
+\\&=RL_\text{s}x+Rv+RD\xi.
+\\&=RL_\text{s}R^T\bar{y}+Rv+RD\xi
+\\&=\tilde{L}_\text{s}\bar{y}+\tilde{v}+\tilde{D}\xi 
 \end{align*}
-Since the rows of $R$ are orthonormal, $RR^T=I$. Further, since the rows of $R$ are orthonormal, $R$ is invertible, which implies that $R^T=R^{-1}$.  Then we can write
-\begin{align*}
-\dot{\bar{y}}&=RL_\text{s}R^T\bar{y}+Rv+R\xi
-\\&=\tilde{L}_\text{s}\bar{y}+\tilde{v}+R\xi 
-\end{align*}
-where $\tilde{L}_\text{s}=RL_\text{s}R^T$ and $\tilde{v}=Rv=RB\vec{1}$.
+where $\tilde{L}_\text{s}=RL_\text{s}R^T$, $\tilde{v}=Rv=RB\vec{1}$, and $\tilde{D}=RD$.
 If $b\neq 0$, $L_\text{s}$ and therefore $\tilde{L}_\text{s}$ have all negative eigenvalues and $\bar{y}=(0, 0, \dots 0, \sqrt{N})^T$ is a globally asymptotically stable equilibrium.
-Therefore, by Claim \ref{lyapeq_derivation} and Theorem \ref{lyapeq_theorem}, $\lim_{t\to\infty}E[\bar{y}(t)\bar{y}(t)^T]=\Sigma_{\bar{y}}$ where $\tilde{L}_\text{s}\Sigma_{\bar{y}}+\Sigma_{\bar{y}}\tilde{L}_\text{s}^T=-RR^T=-I_N$.
+Therefore, by Claim \ref{lyapeq_derivation} and Theorem \ref{lyapeq_theorem}, $\lim_{t\to\infty}E[\bar{y}(t)\bar{y}(t)^T]=\Sigma_{\bar{y}}$ where $\tilde{L}_\text{s}\Sigma_{\bar{y}}+\Sigma_{\bar{y}}\tilde{L}_\text{s}^T=-RD^2R^T$.
 We can  find $\Sigma_y=\lim_{t\to\infty}E[y(t)y(t)^T]$
@@
 \end{align*}
-Again, by Claim \ref{lyapeq_derivation} and Theorem \ref{lyapeq_theorem}, we can now find $\lim_{t\to\infty}E[y(t)y(t)^T]=\Sigma_y$ by finding a solution to $\bar{L}\Sigma_y+\Sigma_y\bar{L}=-QQ^T=-I_{N-1}$.
+Again, by Claim \ref{lyapeq_derivation} and Theorem \ref{lyapeq_theorem}, we can now find $\lim_{t\to\infty}E[y(t)y(t)^T]=\Sigma_y$ by finding a solution to $\bar{L}\Sigma_y+\Sigma_y\bar{L}=-QD^2Q^T$.
 
-To summarize, in the absence of an environmental signal, $\Sigma_y$ satisfies $\bar{L}\Sigma_y+\Sigma_y\bar{L}^T=-I_{N-1}$ where $\bar{L}=QLQ^T$ and in the presence of an environmental signal, $\Sigma_y$ can be found in the first $N-1$ rows and columns of $\Sigma_{\bar{y}}$ where $\Sigma_{\bar{y}}$ satisfies $\tilde{L}_\text{s}\Sigma_{\bar{y}}+\Sigma_{\bar{y}}\tilde{L}_\text{s}^T=-I_N$ where $\tilde{L}_\text{s}=RL_\text{s}R^T$.
+To summarize, in the absence of an environmental signal, $\Sigma_y$ satisfies $\bar{L}\Sigma_y+\Sigma_y\bar{L}^T=-QD^2Q^T$ where $\bar{L}=QLQ^T$ and in the presence of an environmental signal, $\Sigma_y$ can be found in the first $N-1$ rows and columns of $\Sigma_{\bar{y}}$ where $\Sigma_{\bar{y}}$ satisfies $\tilde{L}_\text{s}\Sigma_{\bar{y}}+\Sigma_{\bar{y}}\tilde{L}_\text{s}^T=-RD^2R^T$ where $\tilde{L}_\text{s}=RL_\text{s}R^T$.
 
@@
 \end{align*}
-If $P$ were invertible, the covariance matrix would be given by $-P^{-1}$. However, since $P\vec{1}=\vec{0}$, one of the eigenvalues of $P$ is $0$ and it is therefore not invertible.
-To proceed, we will write $z$ as a linear combination of eigenvectors of $P$.
+Since $P$ has a $0$ eigenvalue, it is not invertible. However, $z$ is orthogonal to the eigenvector of $P$ with eigenvalue $0$. Therefore, the covariance of $z$ is $C_z=-P^\dagger$.
+%If $P$ were invertible, the covariance matrix would be given by $-P^{-1}$. However, since $P\vec{1}=\vec{0}$, one of the eigenvalues of $P$ is $0$ and it is therefore not invertible.
+%To proceed, we will write $z$ as a linear combination of eigenvectors of $P$.
 Since $P$ is symmetric and negative, there is an orthonormal set of eigenvectors $\{w_k\}$ such that $Pw_k=\lambda_kw_k$ with $w_N=\vec{1}$ and $\lambda_N=0$.
@@
 If we then denote the $N$ by $N-1$ matrix whose columns are given by $w_1,...w_{N-1}$ as $\tilde{W}$ and the diagonal matrix with $\lambda_1,\dots,\lambda_{N-1}$ along the diagonal as $\tilde{\Lambda}$, then $z=\tilde{W}\tilde{c}$ for a $\tilde{c}\in\R^{N-1}$ and
-%and $\tilde{P}=\tilde{W}\tilde{\Lambda}\tilde{W}^T$, then 
-\begin{align*}
-P(z)&=\frac{1}{Z}\exp\left(-\frac{1}{2}z^T(-{P})z\right)
-\\\Rightarrow P(\tilde{c})&=\frac{1}{Z}\exp\left(-\frac{1}{2}\tilde{c}^T\tilde{W}^T(-W\Lambda W^T)\tilde{W}\tilde{c}\right)
-\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}c^T\tilde{W}^T(-\tilde{W}\tilde{\Lambda}\tilde{W}^T)\tilde{W}c\right) 
-\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\tilde{c}^T(-\tilde{\Lambda})\tilde{c}\right)  
-\end{align*}
-Therefore, $E[\langle c \rangle ]=\vec{0}$ and $E[\tilde{c}\tilde{c}^T]=(-\tilde{\Lambda})^{-1}$.
-Therefore, $C_z=E[zz^T]=E[\tilde{W}\tilde{c}\tilde{c}^T\tilde{W}^T]=-\tilde{W}\tilde{\Lambda}^{-1}\tilde{W}^T$.
+%%and $\tilde{P}=\tilde{W}\tilde{\Lambda}\tilde{W}^T$, then 
+%\begin{align*}
+%P(z)&=\frac{1}{Z}\exp\left(-\frac{1}{2}z^T(-{P})z\right)
+%\\\Rightarrow P(\tilde{c})&=\frac{1}{Z}\exp\left(-\frac{1}{2}\tilde{c}^T\tilde{W}^T(-W\Lambda W^T)\tilde{W}\tilde{c}\right)
+%\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}c^T\tilde{W}^T(-\tilde{W}\tilde{\Lambda}\tilde{W}^T)\tilde{W}c\right) 
+%\\&=\frac{1}{Z}\exp\left(-\frac{1}{2}\tilde{c}^T(-\tilde{\Lambda})\tilde{c}\right)  
+%\end{align*}
+%Therefore, $E[\langle c \rangle ]=\vec{0}$ and $E[\tilde{c}\tilde{c}^T]=(-\tilde{\Lambda})^{-1}$.
+Therefore, $C_z=E[zz^T]=E[\tilde{W}\tilde{c}\tilde{c}^T\tilde{W}^T]=-\tilde{W}\tilde{\Lambda}^{-1}\tilde{W}^T=-P^\dagger$.
 By dividing each entry $C_{zij}$ of the covariance matrix by $C_{zii}C_{zjj}$, the covariance matrix can be used to find the correlation matrix.
@@
 Since the subspace of $z\in\R^N$ such that $\sum_iz_i=0$ is $N-1$ dimensional, it will be easier to rotate $z$ into $\R^{N-1}$ and integrate over all of $\R^{N-1}$.
-Again, we will use a rotation matrix $R\in\R^N\times\R^N$ such that the last row of $R$ is $\frac{1}{\sqrt{N}}\vec{1}$ and the other rows are normal and orthogonal to $\vec{1}$.
-Above, we found a $Q\in\R^{N-1}\times\R^N$ such that each row of $Q$ was orthogonal to $\vec{1}$ and $R\in\R^N\times\R^N$ be the matrix with $\frac{1}{\sqrt{N}}\vec{1}^T$ as its last row and the first $N-1$ rows identical to the rows of $Q$.
-Then $R$ is a rotation matrix, so $RR^T=R^TR=I$ and $\det(R)=1$.
-This allows us to write $z-(1-V)\vec{1}=Q^TQv$ and $V=\frac{1}{N}\vec{1}^Tv$.
-If $y=Rz$, then $y_N=0$ since $\vec{1}^Tz=0$.
-Therefore,
-$$\Omega=\{z\in\R^N:\sum_iz_i=0\}=\{z\in\R^N:z\cdot\vec{1}=0\}=\{z\in\R^N:\exists \ y\in\R^N \text{ such that } z=R^Ty \text{ and } y_N=0\}.$$
-Further, if $e_N=(0,0,\dots,0,1)^T$, $R^Te_N=q_N=\frac{1}{\sqrt{N}}\vec{1}$ so that $\vec{1}=\sqrt{N}R^Te_N$.This allows us to write
+Therefore, we define $y=Rz=RQ^TQv=Qv$.
+Since $\sum_iz_i=0$, $y_N=0$ so $V$ describes the mean of $v$ and $y$ describes the $N-1$ dimensional space of deviations from the consensus opinion.
+%We first write $z-(1-V)\vec{1}=Q^TQv$ and $V=\frac{1}{N}\vec{1}^Tv$. If $y=Rz$, then $y_N=0$ since $\vec{1}^Tz=0$. Therefore, 
+%$$\Omega=\{z\in\R^N:\sum_iz_i=0\}=\{z\in\R^N:z\cdot\vec{1}=0\}=\{z\in\R^N:\exists \ y\in\R^N \text{ such that } z=R^Ty \text{ and } y_N=0\}.$$
+%Further, if $e_N=(0,0,\dots,0,1)^T$, $R^Te_N=q_N=\frac{1}{\sqrt{N}}\vec{1}$ so that $\vec{1}=\sqrt{N}R^Te_N$.This allows us to write
 \begin{align*}
@@
 However, to consider this as a probability density function over $y$ and $V$, we have to consider the determinant of the transformation from $v$ to $y$ and $V$.
-Combining the transformations from $v$ to $(z,V)$ and from $z$ to $(yV)$,
+Combining the transformations from $v$ to $(z,V)$ and from $z$ to $(y,V)$,
 \begin{align*}
@@
 \begin{equation}
-QLQ^T\Sigma_y+\Sigma_y QLQ^T=-I.
+QLQ^T\Sigma_y+\Sigma_y QLQ^T=-QD^2Q^T.
 \end{equation}
@@
 Table \ref{correspondence} shows how to find the appropriate covariance matrices depending on the framework being used and whether or not a signal is present.
-If we use $P$ instead of $L$ with the noisy dynamics framework, $C_y=2\Sigma_y$.
-%MORE HERE!!!!
-
- \[
- \boxed{\bar{y}=Rx$, $y=Qx=\bar{y}_{i=1}^{N-1}$ (or $y=RQ^TQx$), $z=Q^TQx$, $V=\frac{1}{N}\sum_ix_i}
- \]
+If we use $P$ instead of $L$ with the noisy dynamics framework and all birds have the same number of neighbors, then $C_y=2\Sigma_y$.
 
@@
 & Without signal & With signal 
-\\\hline From noisy dynamics &  $\bar{L}\Sigma_y+\Sigma_y\bar{L}^T=-I_{N-1}$ & $\tilde{L}_\text{s}\Sigma_{\bar{y}}+\Sigma_{\bar{y}}\tilde{L}_\text{s}^T=-I_N$, $\Sigma_y=(\Sigma_{\bar{y}})_{i,j=1}^{N-1}$
-\\\hline From Boltzmann equilibrium & $C_y=QP^{\dagger}Q^T$ & $C_y=(\tilde{P}_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot })^\dagger$
+\\\hline From noisy dynamics &  $\bar{L}\Sigma_y+\Sigma_y\bar{L}^T=-QD^2Q^T$ & $\tilde{L}_\text{s}\Sigma_{\bar{y}}+\Sigma_{\bar{y}}\tilde{L}_\text{s}^T=-RD^2R^T$, $\Sigma_y=(\Sigma_{\bar{y}})_{i,j=1}^{N-1}$
+\\\hline From Boltzmann equilibrium & $C_y=-QP^{\dagger}Q^T$ & $C_y=(\tilde{P}_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot })^\dagger$
 \\ & & or $C_y=Q(-P_\text{s}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB)^{\dagger}Q^T$
@@
 \end{array}\right. $
+\\ $\lambda$ & correlation length
 \\ $\Lambda$ & diagonal matrix of eigenvalues of $P$
@@
 \\$R$ & full rotation matrix whose last row is $\frac{1}{\sqrt{N}}\vec{1}^T$ & $\left(\begin{array}{cc} Q \\ \frac{1}{\sqrt{N}}\vec{1}^T \end{array}\right)$
-\\$r$ & $\mathscr{H}_2$ robustness & $\lim_{t\to\infty}\frac{1}{E[||y||]}$ where $y=Qx$
+\\$\rho$ & $\mathscr{H}_2$ robustness & $\lim_{t\to\infty}\frac{1}{E[||y||]}$ where $y=Qx$
 \\$W$ & matrix of eigenvectors of $P$
@@
 \includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/robustness_vs_ESS.pdf}
-\caption{\label{corrESS}
+\caption{\label{ESS_v_H2}
 %The correlation length of a homogeneous flock depends on both the strategy being used and the radius of the signal, but it depends much more strongly on the radius than the strategy.
@@
 
+\begin{figure}
+\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/H2_v_corrlength.pdf}
+\caption{\label{compare_props} }
+\end{figure}
+
 \nocite{*}

### 2015-03-19 ESS_figures_v9.m

@@ -23,26 +23,9 @@ yoffset=.06;
 
 %% load ESS data
 
-ESSeaten_total=cell(2,1);
-ESSgettoeat_total=cell(2,1);
-storefitnesseaten_total=cell(2,1);
-storefitnessgettoeat_total=cell(2,1);
+load /Users/eleanorbrush/Desktop/ESS_nummoves=1000_numpermove=1.mat
 
-load /Users/eleanorbrush/Desktop/ESS_nummoves=1000_numpermove=1_fullstorage.mat
-
-ESSeaten_total{1}=ESSseaten;
-ESSgettoeat_total{1}=ESSsgettoeat;
-storefitnesseaten_total{1}=storefitnesseaten;
-storefitnessgettoeat_total{1}=storefitnessgettoeat;
-
-load /Users/eleanorbrush/Desktop/ESS_nummoves=10_numpermove=100.mat
-
-ESSeaten_total{2}=ESSseaten;
-ESSgettoeat_total{2}=ESSsgettoeat;
-storefitnesseaten_total{2}=storefitnesseaten;
-storefitnessgettoeat_total{2}=storefitnessgettoeat;
-
-load /Users/eleanorbrush/Desktop/greedyopt_rad=0.1_T=1_nummoves=10_numpermove=100.mat
+load /Users/eleanorbrush/Desktop/greedyopt_simultaneous_T=1_nummoves=1000_numpermove=1_rad=0.2_timesteps=100.mat
 
 load /Users/eleanorbrush/Desktop/homogengroupprops.mat
 %% find number of receivers for different radius
@@ -105,12 +88,7 @@ set(gcf,'Position',[11.5 3 w h]);
 
 n=size(ESSseaten,2);
 k=1;
-t=3;
-
-ESSseaten=ESSeaten_total{k};
-ESSsgettoeat=ESSgettoeat_total{k};
-storefitnesseaten=storefitnesseaten_total{k};
-storefitnessgettoeat=storefitnessgettoeat_total{k};
+t=1;
 
 subplot(2,2,1)
 hold on
@@ -249,11 +227,12 @@ set(gcf,'Position',[11.5 3 w h]);
 subplot(1,2,1)
 % toplot=corrlengths_forced;
 % toplot=corrlengths./repmat([1 radvals(2:end)],Lh,1);
-toplot=groupconsensus_forced;
-toplot_vec=1./(groupconsensus_forced(1,:)*6);
[…]

### 2015-03-19 ESS_parallel.m

@@ -12,7 +12,7 @@ b=1;
 strats=1:1:(N-1);
 L=length(strats);
 
-Tvals=[.01 .1 1 10];
+Tvals=[1 .01 .1 10];
 Nt=length(Tvals);
 
 radvals=[0:.1:1.4]; %#ok<NBRAK>
@@ -49,7 +49,8 @@ for q=1:Nt
         parfor ind=1:(L*L*(N-1))
 %         parfor ind=1:5
             [u,v,k]=ind2sub([L,L,N-1],ind);
-            
+            % rows are residents, columns are invaders, transpose to plot
+            % normally
             resident=strats(u);
             invader=strats(v);
 

### 2015-03-19 H2_corrlength_comparison_v2.m

@@ -1,11 +1,13 @@
-function [meanH2, meanH2_forced, corrlength, corrlength_forced, disconnectedcount]=groupprops(strategy,numsigs_permove,nummoves,radius,b,T) %#ok<INUSD>
 
-N=max(size(strategy));
+N=20;
+strategy=6*ones(N,1);
 
 dvec=[];
 corrvec=[];
+corrlengthvec=[];
 dvec_forced=[];
 corrvec_forced=[];
+corrlengthvec_forced=[];
 disconnectedcount=0;
 H2vec=zeros(nummoves,1);
 H2vec_forced=zeros(nummoves,numsigs_permove);
@@ -21,7 +23,7 @@ for j=2:(N-1)
    v=v/norm(v,2);
    Q(j,:)=v;
 end
-Qfull=[Q; 1/sqrt(N)*ones(1,N)];
+% Qfull=[1/sqrt(N)*ones(1,N); Q];
 
 
 for i=1:nummoves   
@@ -55,19 +57,37 @@ for i=1:nummoves
         disconnectedcount=disconnectedcount+numsigs_permove;
     else
         
-        sigmay=lyap(Ltilde,Q*noise*transpose(Q));
-        H=sqrt(trace(sigmay));
+        sigma=lyap(Ltilde,Q*noise*transpose(Q));
+        H=sqrt(trace(sigma));
         H2vec(i)=H;
-  
+        
+%         sigma_forced=lyap(Lf,eye(N));
+%         H_forced=trace(sigma_forced);
+%         H2vec_forced(i)=H_forced;
+%         
         Wtilde=W(:,[1:(w-1) (w+1):end]);
         Lambdatilde=Lambda([1:(w-1) (w+1):end],[1:(w-1) (w+1):end]);
-        Cov=Wtilde*inv(Lambdatilde)*transpose(Wtilde); %#ok<*MINV>
+        Ptildeinv=Wtilde*inv(Lambdatilde)*transpose(Wtilde); %#ok<*MINV>
         
-        todivide=repmat(diag(-Cov),1,N).*repmat(reshape(diag(-Cov),1,[]),N,1);
+        todivide=repmat(diag(-Ptildeinv),1,N).*repmat(reshape(diag(-Ptildeinv),1,[]),N,1);
         todivide=power(todivide,.5);
-        Corr=-Cov./todivide;
+        C=-Ptildeinv./todivide;
[…]

### 2015-03-19 expected_spin_lessold.m

@@ -1,0 +1,26 @@
+function q = expected_spin(M,T,beta)
+%solves the differential equation dq/dt=M*q+beta*(1-q)  -->
+%dq/dt=A*q+beta where A=M-diag(beta)
+
+% N=length(strategy);
+% M=makenet_normed(strategy);
+A=M-diag(beta);
+N=size(M,1);
+
+if nargin==2
+    beta=1*ones(N,1);
+end
+
+initialconditions=.5*ones(N,1);
+
+fun = @(x)expm(-x*A)*beta;
+integrated = integral(fun,0,T,'ArrayValued',true);
+
+homogeneous=expm(T*A)*initialconditions;
+% particular=expm(T*A)*inv(A)*[eye(N)-expm(-T*A)]*beta;
+particular=expm(T*A)*integrated;
+instantaneous=homogeneous+particular;
+
+q=instantaneous;
+end
+

### 2015-03-19 figuringout_groupprops.m

@@ -1,0 +1,92 @@
+strategy=3*ones(N,1);
+
+N=max(size(strategy));
+
+dvec=[];
+corrvec=[];
+corrvec_old=[];
+% dvec_forced=[];
+% corrvec_forced=[];
+disconnectedcount=0;
+H2vec=zeros(nummoves,1);
+H2vec_old=zeros(nummoves,1);
+% H2vec_forced=zeros(nummoves,numsigs_permove);
+
+Q=zeros(N-1,N);
+Q(1,1:2)=[1/sqrt(2) -1/sqrt(2)];
+for j=2:(N-1)
+   v=rand(1,N);
+   v=v-sum(v)/N;
+   for k=1:(j-1)
+      v=v-(v*transpose(Q(k,:)))*Q(k,:); 
+   end
+   v=v/norm(v,2);
+   Q(j,:)=v;
+end
+Qfull=[1/sqrt(N)*ones(1,N); Q];
+
+
+for i=1:nummoves   
+    positions=unifrnd(0,1,N,2);
+    d=squareform(pdist(positions));
+
+    A=zeros(N);
+    Abar=zeros(N);
+    for ind=1:N
+        [~, order]=sort(d(ind,:));
+        neighbors=order(2:strategy(ind)+1);
+        A(ind,neighbors)=1;
+        Abar(ind,neighbors)=1/strategy(ind);
+    end
+  
+    L=Abar-eye(N);
+    Ltilde=Q*L*transpose(Q);
+    [~,vals]=eig(Ltilde);
+    problem=find(sigfig(diag(vals),13)==0,1);
+    
+    S=.5*(Abar+transpose(Abar));
+    P=S;
+    P(1:N+1:end)=-sum(S,2);
+    [W,Lambda]=eig(P);
+    w=find(sigfig(diag(Lambda),13)==0);
+    
+    if length(w)>1 || ~isempty(problem)
+        disconnectedcount=disconnectedcount+numsigs_permove;
+    else
+        
+        sigma=lyap(Ltilde,eye(N-1));
+        H=sqrt(trace(sigma));
+        H2vec(i)=H;
+        
+        h=H2norm(L,'non');
+        H2vec_old(i)=h; 
+        
+        Wtilde=W(:,[1:(w-1) (w+1):end]);
+        Lambdatilde=Lambda([1:(w-1) (w+1):end],[1:(w-1) (w+1):end]);
[…]

### 2015-03-19 greedyopt_v7_test2.m

@@ -15,8 +15,8 @@ radvals=[.1 .7];
 Nr=length(radvals);
 T=1;
 
-timesteps=50;
-its=2;
+timesteps=10;
+its=1;
 
 evolution_eaten=zeros(its,N,timesteps);
 groupspeed_eaten=zeros(its,timesteps);
@@ -33,7 +33,7 @@ disconnected_gettoeat=zeros(its,timesteps);
 % for ir=1:Nr
     radius=radvals(aid);
 for num=1:its
-    init=randi([1 N-1],1,N);
+%     init=randi([1 N-1],1,N);
     t=1;
     while t<=timesteps
 
@@ -50,8 +50,10 @@ for num=1:its
         [probeaten, ~]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
         perf=1-probeaten;
         newstrategy=strategy;
-
-        for i=1:N
+        
+        changeorder=randsample(1:N,N,'false');
+%         for i=1:N
+        for i=changeorder
             choosestrat=zeros(1,3);
             choosestrat(2)=perf(i);
 
@@ -76,9 +78,11 @@ for num=1:its
             end 
 
             if choosestrat(1)>choosestrat(2)
+                strategy(i)=strategy(i)-1;
                 newstrategy(i)=strategy(i)-1;
             end
             if choosestrat(3)>choosestrat(2)
+                strategy(i)=strategy(i)-1;
                 newstrategy(i)=strategy(i)+1;
             end
 

### 2015-03-19 greedyopt_v8.m

@@ -15,7 +15,7 @@ radvals=[.1 .2 .7];
 % Nr=length(radvals);
 T=1;
 
-timesteps=100;
+timesteps=200;
 its=20;
 
 evolution_eaten=zeros(its,N,timesteps);
@@ -24,6 +24,7 @@ groupconsensusforced_eaten=zeros(its,timesteps);
 corrlengths_eaten=zeros(its,timesteps);
 corrlengthsforced_eaten=zeros(its,timesteps);
 disconnected_eaten=zeros(its,timesteps);
+changeorder_eaten=zeros(its,N,timesteps);
 
 evolution_gettoeat=zeros(its,N,timesteps);
 groupconsensus_gettoeat=zeros(its,timesteps);
@@ -31,6 +32,7 @@ groupconsensusforced_gettoeat=zeros(its,timesteps);
 corrlengths_gettoeat=zeros(its,timesteps);
 corrlengthsforced_gettoeat=zeros(its,timesteps);
 disconnected_gettoeat=zeros(its,timesteps);
+changeorder_gettoeat=zeros(its,N,timesteps);
 
 maxt_eaten=zeros(its,1);
 maxt_gettoeat=zeros(its,1);
@@ -50,14 +52,20 @@ for num=1:its
         t=t+1;
     else
         strategy=evolution_eaten(num,:,t-1);
-        strategy=reshape(strategy,1,N);
+        strategy=reshape(strategy,N,1);
 
-        [probeaten, ~]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
-        perf=1-probeaten;
-        newstrategy=strategy;
+%         [probeaten, ~]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+%         perf=1-probeaten;
+%         newstrategy=strategy;
 
-        i=randsample(1:N,1);
+%         i=randsample(1:N,1);
 %         for i=1:N
+        changeorder=randsample(1:N,N,'false');
+        changeorder_eaten(num,:,t)=changeorder;     
+        for i=changeorder
+            [probeaten, ~]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+            perf=1-probeaten;
+            
             choosestrat=zeros(1,3);
             choosestrat(2)=perf(i);
 
@@ -70,7 +78,7 @@ for num=1:its
                 [probeaten, ~]=signalingevents_parallel(upone,numsigs_permove,nummoves,radius,b,T);
[…]

### 2015-03-19 groupprops_test.m

@@ -1,6 +1,16 @@
-function [meanH2, meanH2_forced, corrlength, corrlength_forced, disconnectedcount]=groupprops(strategy,numsigs_permove,nummoves,radius,b,T) %#ok<INUSD>
+numworkers=str2num(getenv('PROCS')); %#ok<ST2NM>
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
 
-N=max(size(strategy));
+numsigs_permove=1;
+nummoves=1000;
+
+N=20;
+b=1;
+T=1;
+radius=.7;
+strategy=6*ones(N,1);
 
 dvec=[];
 corrvec=[];
@@ -21,7 +31,7 @@ for j=2:(N-1)
    v=v/norm(v,2);
    Q(j,:)=v;
 end
-Qfull=[Q; 1/sqrt(N)*ones(1,N)];
+Qfull=[1/sqrt(N)*ones(1,N); Q];
 
 
 for i=1:nummoves   
@@ -42,9 +52,6 @@ for i=1:nummoves
     [~,vals]=eig(Ltilde);
     problem=find(sigfig(diag(vals),13)==0,1);
     
-%     noise=eye(N);
-    noise=diag(strategy);
-    
     S=.5*(Abar+transpose(Abar));
     P=S;
     P(1:N+1:end)=-sum(S,2);
@@ -55,19 +62,24 @@ for i=1:nummoves
         disconnectedcount=disconnectedcount+numsigs_permove;
     else
         
-        sigmay=lyap(Ltilde,Q*noise*transpose(Q));
-        H=sqrt(trace(sigmay));
+        sigma=lyap(Ltilde,eye(N-1));
+        H=trace(sigma);
         H2vec(i)=H;
-  
+        
+%         sigma_forced=lyap(Lf,eye(N));
+%         H_forced=trace(sigma_forced);
+%         H2vec_forced(i)=H_forced;
+%         
         Wtilde=W(:,[1:(w-1) (w+1):end]);
         Lambdatilde=Lambda([1:(w-1) (w+1):end],[1:(w-1) (w+1):end]);
-        Cov=Wtilde*inv(Lambdatilde)*transpose(Wtilde); %#ok<*MINV>
[…]

### 2015-03-19 homogengroupprops_test.m

@@ -1,17 +1,17 @@
-numworkers=str2num(getenv('PROCS')); %#ok<ST2NM>
-dellacluster=parcluster('local');
-dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
-dellapool=parpool(dellacluster, numworkers) ;
+% numworkers=str2num(getenv('PROCS')); %#ok<ST2NM>
+% dellacluster=parcluster('local');
+% dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+% dellapool=parpool(dellacluster, numworkers) ;
 
 numsigs_permove=1;
-nummoves=1000;
+nummoves=100;
 
 N=20;
 b=1;
 T=1;
 
-homogenstrats=6:1:(N-1);
-% homogenstrats=6;
+% homogenstrats=6:1:(N-1);
+homogenstrats=6;
 Lh=length(homogenstrats);
 
 radvals=0:.1:1.4;
@@ -39,9 +39,8 @@ for i=1:Nr
     end
 end
 
-save('/home/brush/schooling_consensus/homogengroupprops.mat','groupconsensus','groupconsensus_forced','corrlengths','corrlengths_forced','disconnected','homogenstrats','Lh','radvals','Nr','N')
-% save('/home/brush/schooling_consensus/homogengroupprops.mat','groupconsensus','corrlengths','disconnected','homogenstrats','Lh','radvals','Nr','N')
-
-delete(dellapool);
-
-exit ;
+% save('/home/brush/schooling_consensus/homogengroupprops.mat','groupconsensus','groupconsensus_forced','corrlengths','corrlengths_forced','disconnected','homogenstrats','Lh','radvals','Nr','N')
+% 
+% delete(dellapool);
+% 
+% exit ;

### 2015-03-19 new_correlations.m

@@ -1,0 +1,171 @@
+N=20;
+b=2;
+informed=15;
+radius=.5;
+strategy=[10,10*ones(1,N-1)];
+
+dvec=[];
+corrvec=[];
+corrvec_forced=[];
+disconnectedcount=0;
+nummoves=100;
+H2vec1=zeros(1,nummoves);
+H2vec_forced1=zeros(1,nummoves);
+
+Q=zeros(N-1,N);
+Q(1,1:2)=[1/sqrt(2) -1/sqrt(2)];
+for j=2:(N-1)
+   v=rand(1,N);
+   v=v-sum(v)/N;
+   for k=1:(j-1)
+      v=v-(v*transpose(Q(k,:)))*Q(k,:); 
+   end
+   v=v/norm(v,2);
+   Q(j,:)=v;
+end
+Qfull=[1/sqrt(N)*ones(1,N); Q];
+
+for i=1:nummoves    
+    positions=unifrnd(0,1,N,2);
+    d=squareform(pdist(positions));
+   
+    A=zeros(N);
+    Abar=zeros(N);
+    for ind=1:N
+        [~, order]=sort(d(ind,:));
+        neighbors=order(2:strategy(ind)+1);
+        A(ind,neighbors)=1;
+        Abar(ind,neighbors)=1/strategy(ind);
+    end
+  
+    L=Abar-eye(N);
+    Ltilde=Q*L*transpose(Q);
+    [~,vals]=eig(Ltilde);
+    prob=find(sigfig(diag(vals),13)==0);
+    
+    S=.5*(Abar+transpose(Abar));
+    P=S;
+    P(1:N+1:end)=-sum(S,2);
+    [W,Lambda]=eig(P);
+    w=find(sigfig(diag(Lambda),13)==0);
+    
+    if length(w)>1 || ~isempty(prob)
+        disconnectedcount=disconnectedcount+1;
+    else
+        
+        sigma=lyap(Ltilde,eye(N-1));
+        H=trace(sigma);
+        H2vec1(i)=H;
+         
+        Wtilde=W(:,[1:(w-1) (w+1):end]);
+        Lambdatilde=Lambda([1:(w-1) (w+1):end],[1:(w-1) (w+1):end]);
[…]

### 2015-03-20 paper_skeleton.tex
just added the possibility of 0 neighbors in signaling_events

@@
 In the Appendix, we derive equations for each of these covariance matrices and summarize the expression for the $y$ covariance matrices in Table \ref{correspondence}.
-The first thing to note is that, if instead of using $L$ we use $P$ in the dynamical approach, then $C_y=2\Sigma_y$, as shown in Claim XXX.
+The first thing to note is that, if instead of using $L$ we use $P$ in the dynamical approach and all birds have the same number of neighbors, then $C_y=2\Sigma_y$, as shown in Claim \ref{cov_equivalence}.
 There is therefore a close correspondence between the two approaches.
@@
 Again, we can find the covariance matrices $\Sigma_z$, $\Sigma_y$, $C_z$, and $C_y$.
-As was the case without a signal, if we use $P$ in the dynamical approach, then we find that $C_y=2\Sigma_y$.
+As was the case without a signal, if we use $P$ in the dynamical approach and all birds have the same number of neighbors, then we find that $C_y=2\Sigma_y$ (Claim \ref{cov_equivalence}).
 We define $\rho$ and $\lambda$ as above using the matrices $\Sigma_y$ and $C_z$ in the presence of a signal.
@@
 In a homogeneous flock, where every bird is using the same strategy, as both the strategy and the radius of the signal change, both $\mathscr{H}_2$ and the correlation length are affected.
-As the Figure \ref{compare_props} shows that there is a relationship between two group properties: the $\mathscr{H}_2$ norm and the correlation length of the flock.
+Figure \ref{compare_props} shows that there is a relationship between two group properties: the $\mathscr{H}_2$ norm and the correlation length of the flock.
+%%WHY? intuition for why each changes as they do with strategy and radius?!?!?
 
@@
 The equilibrium strategies when the individuals are allowed to learn are similar to the ESS strategies: learned strategies when selection is due to predation are always higher than learned strategies when selection is due to resources (Figure \ref{greedyopt}).
-Surprisingly, even when the group starts with heterogeneous strategies, they often reach a homogeneous equilibrium.
+%Surprisingly, even when the group starts with heterogeneous strategies, they often reach a homogeneous equilibrium.
 Under selection due to predation, the $\mathscr{H}_2$ robustness and correlation length of the flock tend to decrease, whereas under selection due to resources, $\mathscr{H}_2$ robustness is relatively constant and correlation length tends to increase (Figure \ref{greedyopt}).
@@
 
+%%ORGANIZE INTRODUCTION TO THE ROTATION MATRICES. 
 We define a matrix $Q\in\R^{N-1}\times\R^N$ such that each row of $Q$ is orthogonal to $\vec{1}$ and the rows are pairwise orthogonal.
@@
  \]
+
+ Depending on the model framework, we will find either $E[yy^T]$ or $E[zz^T]$. This are interchangeable since $E[yy^t]=E[Rzz^TR^T]=RE[zz^T]R^T$ and $E[zz^T]=E[R^Tyy^TR]=R^TE[yy^T]R$.
 
@@
 We can now use this probability distribution to find the  covariances and correlations between the birds' fluctuations around consensus.
-If we let $V=\sum_iv_i/N=\langle v\rangle$ and $z=v-V\vec{1}$ so that $z$ represents the deviations in each bird's opinion from the consensus opinion.
-Then
+If we let $V=\sum_iv_i/N=\langle v\rangle$ and $z=v-V\vec{1}$, $z$ represents the deviations in each bird's opinion from the consensus opinion.
+We are interested in finding $C_z=\left(E[z_iz_j]\right)_{ij}$.
+To do this, we first find the probability distribution over $(z,V)$:
 \begin{align*}
@@
 Since $\sum_iz_i=0$, $y_N=0$ so $V$ describes the mean of $v$ and $y$ describes the $N-1$ dimensional space of deviations from the consensus opinion.
-%We first write $z-(1-V)\vec{1}=Q^TQv$ and $V=\frac{1}{N}\vec{1}^Tv$. If $y=Rz$, then $y_N=0$ since $\vec{1}^Tz=0$. Therefore, 
-%$$\Omega=\{z\in\R^N:\sum_iz_i=0\}=\{z\in\R^N:z\cdot\vec{1}=0\}=\{z\in\R^N:\exists \ y\in\R^N \text{ such that } z=R^Ty \text{ and } y_N=0\}.$$
-%Further, if $e_N=(0,0,\dots,0,1)^T$, $R^Te_N=q_N=\frac{1}{\sqrt{N}}\vec{1}$ so that $\vec{1}=\sqrt{N}R^Te_N$.This allows us to write
 \begin{align*}
@@
 However, to consider this as a probability density function over $y$ and $V$, we have to consider the determinant of the transformation from $v$ to $y$ and $V$.
-Combining the transformations from $v$ to $(z,V)$ and from $z$ to $(y,V)$,
-\begin{align*}
-y=RQ^TQv=Qv \text{ and } V=\frac{1}{N}\vec{1}^Tv=\frac{1}{\sqrt{N}}R_{N\cdot }v
-\end{align*}
-so that the entire transformation can be represented by $\bar{R}=\left(\begin{array}{cc}Q \\1/\sqrt{N}R_{N\cdot }\end{array}\right)$, so that $\det(\bar{R})=\frac{1}{\sqrt{N}}\det(R)=\frac{1}{\sqrt{N}}$.
+The  transformation can be represented by $\bar{R}=\left(\begin{array}{cc}Q \\1/\sqrt{N}R_{N\cdot }\end{array}\right)$, so that $\det(\bar{R})=\frac{1}{\sqrt{N}}\det(R)=\frac{1}{\sqrt{N}}$.
 Therefore, $P(y,V)=\sqrt{N}P(v)$ is the appropriate density function for $y$ and $V$.
@@
 \begin{equation}
-C_z=E[zz^T]=\left(-P_\text{s}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB\right)^{-1}
+C_z=\left(-P_\text{s}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB\right)^{-1}
 \end{equation}
-Now, let $P'=P_\text{s}+\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB$.
-Then $P'\vec{1}=-B\vec{1}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB\vec{1}=-B\vec{1}-\frac{bN_\text{I}}{bN_\text{I}}B\vec{1}=\vec{0}$, so we cannot invert $-P'$ to find the covariance matrix.
+If $P'=P_\text{s}+\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB$, then $P'\vec{1}=-B\vec{1}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB\vec{1}=-B\vec{1}-\frac{bN_\text{I}}{bN_\text{I}}B\vec{1}=\vec{0}$, so we cannot invert $-P'$ to find the covariance matrix.
 However, as above, we can discard the $0$ eigenvector since we know that $z\cdot \vec{1}=0$.
-This gives us $E[(v_i-\frac{\sum_kv_k}{N})(v_j-\frac{\sum_kv_k}{N})]$.
+This gives us $C_z=(-P_\text{s}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB)^{\dagger}$.
 
 \subsection{Connecting two frameworks }
-To find $\mathscr{H}$, we found the matrix $\Sigma_y$ such that 
+Table \ref{correspondence} shows how to find the covariance of $y$ depending on the framework being used and whether or not a signal is present. 
+\begin{claim}\label{cov_equivalence}If we use $P$ instead of $L$ with the noisy dynamics framework and all birds have the same number of neighbors $n$, then $C_y=\frac{2}{n}\Sigma_y$.
+\end{claim}
+\begin{pf}
+If we use $P$ in the noisy dynamics framework, $\Sigma_y$ satisfies
 \begin{equation}
-QLQ^T\Sigma_y+\Sigma_y QLQ^T=-QD^2Q^T.
+QPQ^T\Sigma_y+\Sigma_y QP^TQ^T=-QD^2Q^T.
 \end{equation}
-If the vector $v$ is described by the dynamics $\dot{v}=Dv$, with $H(v)=-v^TPv$, then
+If all birds have $n$ neighbors, then $D^2$ has $n$ along the diagonal and $0$ elsewhere and $-QD^2Q^T=-nI_{N-1}$.  Now consider
 \begin{align*}
-\dot{H}&=-\dot{v}^TPv-v^TP\dot{v}
-\\&=-v^TD^TPv-v^TPDv
-\\&=-v^T(D^TP+PD)v.
+QPQ^TC_y+C_yQP^TQ^T&=QPQ^T(-QP^\dagger Q^T)+(-QP^\dagger Q^T)QP^T
+\\&=-QP(I_N-\frac{1}{N}\vec{1}\vec{1}^T)P^\dagger Q^T-QP^\dagger (I_N-\frac{1}{N}\vec{1}\vec{1}^T)P^T
+\\&=-QPP^\dagger Q^T-QP^\dagger P^TQ^T \text{ since $P\vec{1}=\vec{0}$ }
+\\&=-Q(I_N-\frac{1}{N}\vec{1}\vec{1}^T)Q^T-Q(I_N-\frac{1}{N}\vec{1}\vec{1}^T)Q^T
+\\&=-QQ^T-QQ^T \text{ since $Q\vec{1}=\vec{0}$}
+\\&=-2I_{N-1}
+\\ \Rightarrow QPQ^T\left(\frac{n}{2}C_y\right)+\left(\frac{n}{2}C_y\right)QP^TQ^T&=-nI_{N-1}
+\\ \Rightarrow\frac{n}{2}C_y&=\Sigma_y
 \end{align*}
-Therefore, $H(v)$ decreases at all $v$ if and only if $D^TP+PD$ is a positive matrix.
-
-Table \ref{correspondence} shows how to find the appropriate covariance matrices depending on the framework being used and whether or not a signal is present.
-If we use $P$ instead of $L$ with the noisy dynamics framework and all birds have the same number of neighbors, then $C_y=2\Sigma_y$.
+since there is a unique solution to this Lyapunov equation.
+\end{pf}
+A similar proof shows the equivalence of the two covariance matrices in the presence of a signal, as long as $P_\text{s}$ is used in the noisy dynamical system and all birds have the same number of neighbors.
+%
+%We can also make a connection between the two frameworks by considering the dynamics that minimize the Hamiltonian that gives rise to the Boltzmann distribution. 
+%If the vector $v$ is described by the dynamics $\dot{v}=Wv$, with $H(v)=-v^TPv$, then
+%\begin{align*}
+%\dot{H}&=-\dot{v}^TPv-v^TP\dot{v}
+%\\&=-v^TW^TPv-v^TPWv
+%\\&=-v^T(W^TP+PW)v.
+%\end{align*}
+%Therefore, $H(v)$ decreases at all $v$ if and only if $D^TP+PD$ is a positive matrix.
+%
 
@@
 \begin{figure}
+\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/H2_v_corrlength.pdf}
+\caption{\label{compare_props} }
+\end{figure}
+
+\begin{figure}
 \includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/ESSfigure.pdf}
@@
 
-\begin{figure}
-\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/H2_v_corrlength.pdf}
-\caption{\label{compare_props} }
-\end{figure}
-
 \nocite{*}

### 2015-03-20 greedyopt_v8.m

@@ -100,14 +100,14 @@ for num=1:its
 
         evolution_eaten(num,:,t)=strategy;
         
-        if sum(evolution_eaten(num,:,t)==evolution_eaten(num,:,t-1))==N
-            evolution_eaten(num,:,(t+1):end)=repmat(col(strategy),1,timesteps-t);
-            t_eaten=t;
-            t=timesteps+1;
-        else
+%         if sum(evolution_eaten(num,:,t)==evolution_eaten(num,:,t-1))==N
+%             evolution_eaten(num,:,(t+1):end)=repmat(col(strategy),1,timesteps-t);
+%             t_eaten=t;
+%             t=timesteps+1;
+%         else
             t_eaten=t;
             t=t+1;
-        end
+%         end
     end
     end
     maxt_eaten(num)=t_eaten;
@@ -169,14 +169,14 @@ for num=1:its
 
         end
         evolution_gettoeat(num,:,t)=strategy;
-        if sum(evolution_gettoeat(num,:,t)==evolution_gettoeat(num,:,t-1))==N
-            evolution_gettoeat(num,:,(t+1):end)=repmat(col(strategy),1,timesteps-t);
-            t_gettoeat=t;
-            t=timesteps+1;
-        else 
+%         if sum(evolution_gettoeat(num,:,t)==evolution_gettoeat(num,:,t-1))==N
+%             evolution_gettoeat(num,:,(t+1):end)=repmat(col(strategy),1,timesteps-t);
+%             t_gettoeat=t;
+%             t=timesteps+1;
+%         else 
             t_gettoeat=t;
             t=t+1;
-        end
+%         end
     end
     end
     maxt_gettoeat(num)=t_gettoeat;

### 2015-03-20 groupprops.m

@@ -123,44 +123,51 @@ for i=1:nummoves
     end
 end
 
-meanH2=mean(H2vec(~isnan(H2vec)));
-H2vec_forced=col(H2vec_forced);
-meanH2_forced=mean(H2vec_forced(~isnan(H2vec_forced)));
-% meanH2_forced=meanH2;
+if disconnectedcount<nummoves
+    meanH2=mean(H2vec(~isnan(H2vec)));
+    H2vec_forced=col(H2vec_forced);
+    meanH2_forced=mean(H2vec_forced(~isnan(H2vec_forced)));
+    % meanH2_forced=meanH2;
 
-deltad=.05;
-dbins=(0:deltad:1.4)+deltad/2;
-Nbins=length(dbins);
-avgcorrs=zeros(1,Nbins);
-avgcorrs_forced=zeros(1,Nbins);
+    deltad=.05;
+    dbins=(0:deltad:1.4)+deltad/2;
+    Nbins=length(dbins);
+    avgcorrs=zeros(1,Nbins);
+    avgcorrs_forced=zeros(1,Nbins);
 
-for i=1:Nbins
-    w1=find(dvec>=dbins(i)-deltad/2);
-    w2=find(dvec<dbins(i)+deltad/2);
-    w=intersect(w1,w2);
-    avgcorrs(i)=mean(corrvec(w));
-end
+    for i=1:Nbins
+        w1=find(dvec>=dbins(i)-deltad/2);
+        w2=find(dvec<dbins(i)+deltad/2);
+        w=intersect(w1,w2);
+        avgcorrs(i)=mean(corrvec(w));
+    end
 
-for i=1:Nbins
-    w1=find(dvec_forced>=dbins(i)-deltad/2);
-    w2=find(dvec_forced<dbins(i)+deltad/2);
-    w=intersect(w1,w2);
-    avgcorrs_forced(i)=mean(corrvec_forced(w));
-end
+    for i=1:Nbins
+        w1=find(dvec_forced>=dbins(i)-deltad/2);
+        w2=find(dvec_forced<dbins(i)+deltad/2);
+        w=intersect(w1,w2);
+        avgcorrs_forced(i)=mean(corrvec_forced(w));
+    end
 
-f=find(avgcorrs<0,1,'first');
-if isempty(f)
-    corrlength=max(dvec);
+    f=find(avgcorrs<0,1,'first');
+    if isempty(f)
+        corrlength=max(dvec);
+    else
[…]

### 2015-03-20 corrlength_v_radius.m
rewrote groupprops so that robustness is 0 if network is unconnected and i define corrlength no matterwhat by removing 0 eigenvectors... now can run homogengroupprops for any strategy

@@ -1,5 +1,5 @@
 N=20;
-strategy=12*ones(N,1);
+strategy=2*ones(N,1);
 
 deltad=.05;
 
@@ -29,8 +29,16 @@ end
 S=.5*(Abar+transpose(Abar));
 P=S;
 P(1:N+1:end)=-sum(S,2);
-
-
+[W,Lambda]=eig(P);
+w=find(sigfig(diag(Lambda),13)==0);
+if length(w)>1
+    'prob'
+end
+Wtilde=W(:,setdiff(1:N,w));
+Lambdatilde=Lambda(setdiff(1:N,w),setdiff(1:N,w));
+Cov=Wtilde*inv(Lambdatilde)*transpose(Wtilde); %#ok<*MINV>
+plot(d,-Cov,'o');set(gca,'ylim',[-.5 5])      
+%%
 %     noise=eye(N);
 noise=diag(strategy);
     

### 2015-04-14 paper_skeleton.tex
rewrote groupprops so that robustness is 0 if network is unconnected and i define corrlength no matterwhat by removing 0 eigenvectors... now can run homogengroupprops for any strategy
about to change paper_skeleton significantly to add an explanation due to the hypergeometric distribution and new greedy opt results

@@
 \section{Introduction }
+%Introduction to schools and flocks
+%
+%Reasons for flocking / paying attention to your peers.
+Predation \cite{Ballerini:2007uq} Empirical evidence for reduced predator efficiency when the prey are a in a bigger group (captive fish) \cite{Landeau:1986fk}.
+Foraging (find citations).
+Foraging (cite bat / information center papers).
+I have to make some statement to the effect that I assume predator efficiency is constant irrespective of the type of network the flock forms or if it's even connected.
+Which is conservative.
+%
+%How to construct a social network: Zone model a la Couzin uses metric interaction rules.
+Metric zones seem to reproduce flocking dynamics in surf scoters \cite{Lukeman:2010vn}.
+Evidence for topological network in starlings \cite{Ballerini:2007uq}, which they suggest must rely on vision \cite{Ballerini:2007uq}, although this can produced by a model in which an interaction range is governed by absolute distance instead of number of neighbors \cite{Bode:2010ys}.
+Evidence that the ability to see one's neighbors should be included in constructing social networks (golden shiners) \cite{Strandburg-Peshkin:2013fk}
+%
+%evidence for six to seven neighbors from anisotropy of positions of neighbors \cite{Ballerini:2007uq}  Ballerini argues that six / seven represents a cognitive limitation \cite{Ballerini:2007uq}.
+maximum likelihood interaction range is around ten \cite{Bialek:2012fk}
+%
+%assertion that individuals are trying to maintain a cohesive group "The main goal of the interaction among individuals is to maintain cohesion of the group.
+This cohesion is a very strong biological requirement, shaped by the evolutionary pressure for survival: Stragglers and small groups are significantly more prone to predation than animals belonging to large and highly cohesive aggregations." \cite{Ballerini:2007uq}, the support for which comes from a paper about teleost fishes.
+But why should the group maintain cohesion due to a predator response if predation is the thing causing cohesion?
+There's something circular here.
+And "An alternative interpretation of the interaction range we find is that the specific value of $n_\text{c}$ may be functional to optimize anti predatory response: If each individual interacts with too few neighbors, information is non noisy, but it is too short-ranged; conversely, if the interaction involves too many neighbors, information is averaged over several ill-informed individuals, and it is too noisy" \cite{Ballerini:2007uq} --- BUT WE FIND THE OPPOSITE.
+"The interaction mechanism shaped by evolution must keep cohesion in the face of strong perturbations, of which predation is the most relevant.
+We believe that topological interaction is the only mechanisms granting such robust cohesion and, therefore, higher biological fitness." \cite{Ballerini:2007uq}
+%
+%Evidence for how collective create interesting structures \cite{Cavagna:2009kx} \cite{Bialek:2013fk} \cite{Young:2013kx}
+%
+%Individual evolution, consequences for group structure\cite{Torney:2015fk}
+%
+%NOTE: A similar model of random flock dynamics is in \cite{Shang:2014uq}.
+They randomly place nodes in a unit square repeatedly, but they let them keep their opinions across spatial shuffles.
+They prove that the system exponentially approaches consensus in the absence of noise and show that the system almost always reaches consensus in the presence of noise \cite{Shang:2014uq}.
+With or without noise, the time to consensus decreases as the number of neighbors increases, but with diminishing returns, as in \cite{Young:2013kx}.
+%
+%Can individual optimization reproduce interesting group structure?
+%
+%CRITICALITY AS AN INFORMATIONAL TRAGEDY OF THE COMMONS.
+%
+%Two points to make about optimization in a group: 1) \cite{Torney:2015fk} found that when one individual changes its strategy it does not change the overall social environment and thus can optimize without regard for the group.
+However, to find the group-level optimum, one must change the group's strategy and thus the social environment, leading to an group optimum different from the individual's.
+Same here.
+2) The existence of multiple ESS is evidence for how important social environment is because what is an un-invadable strategy when it is at high frequency does poorly when another strategy is at high frequency.
+%
 Questions:
@@
 \end{equation}
-If $n_i$ is the number of neighbors node $i$ has, then $n_i=\sum_kA_{ik}$.
-We assume that having more neighbors means paying less attention to each so we normalize $A$ to find $\bar{A}$: $\bar{A}_{ij}=\frac{A_{ij}}{n_i}$.
+If $k_i$ is the number of neighbors node $i$ has, then $k_i=\sum_kA_{ik}$.
+We assume that having more neighbors means paying less attention to each so we normalize $A$ to find $\bar{A}$: $\bar{A}_{ij}=\frac{A_{ij}}{k_i}$.
 
@@
 \begin{equation}
+\dot{v}_i=\frac{1}{k_i}\sum_i\left(v_j-v_i\right)+\sqrt{k_i}\xi_i
+\end{equation}
+\begin{equation}
 \dot{v}=Lv+D\xi \label{opinion_dynamics}
 \end{equation} 
-where $D$ is a diagonal matrix with entires $\{\sqrt{n_i}\}$ and $E[\xi(t)]=\vec{0}$, $E[\xi(t)\xi(s)^T]=\frac{1}{2}\delta(t-s)I$, and $E[v(0)\xi(t)^T]=\vec{0}$.
+where $D$ is a diagonal matrix with entires $\{\sqrt{k_i}\}$ and $E[\xi(t)]=\vec{0}$, $E[\xi(t)\xi(s)^T]=\frac{1}{2}\delta(t-s)I$, and $E[v(0)\xi(t)^T]=\vec{0}$.
 This describes a system in which each bird changes averages the differences between its own opinions and those of its neighbors and changes its opinion to minimize these differences.
@@
 \begin{equation}
-\dot{v}=Lv+B(\vec{1}-v)+\xi =L_\text{s}v+B\vec{1}+n_i\xi \label{opinion_dynamics2}
+\dot{v}=Lv+B(\vec{1}-v)+D\xi =L_\text{s}v+B\vec{1}+D\xi \label{opinion_dynamics2}
 \end{equation}
@@
 To generate a matrix $L$, we distribute the individuals randomly in space (specifically, in a unit square).
-has a strategy $n_i$ that dictates how many neighbors it has, so that $n_i=\sum_{j\neq i}{\bf I}(A_{ij}\neq 0)$.
-Each individual then pays attention to its $n_i$ nearest neighbors.
+has a strategy $k_i$ that dictates how many neighbors it has, so that $k_i=\sum_{j\neq i}{\bf I}(A_{ij}\neq 0)$.
+Each individual then pays attention to its $k_i$ nearest neighbors.
 One individual is chosen to be a receiver; it and all those individuals within a radius $r$ of the receiver perceive, so that $B_{ii}=1$ for all those birds within this radius.
@@
 This is meant to simulate a situation in which such that information is present only intermittently and otherwise the birds are moving randomly through space and restructuring their network.
+If, instead of paying attention to the closest neighbors, a bird choose $k_i$ neighbors at random out of $N$ birds, $N_\text{I}$ of whom are informed, the number of informed neighbors follows a hypergeometric distribution:
+\begin{equation}
+P(j \text{ informed neighbors} )=\frac{{N_\text{I}\choose j}{N-N_\text{I}\choose k_i-j}  }{{N\choose k_i}}. \label{hypergeom}
+\end{equation} 
+This distribution is not an exact description of the probability of $j$ informed neighbors in our model  since it ignores spatial proximity, but it is a useful approximation for considering the advantages and disadvantages of each strategy $k_i$.
 
@@
 $1$ minus the probability of being the least informed over many signaling events.
-If the signal is information that resources are available, we assume that whichever individual is most aware of the signal will get access to the resource, i.e.
+If the signal is information that resources are available, we consider two levels of resource abundance.
+If there are scarce resources, we assume that whichever individual is most aware of the signal will get access to the resource, i.e.
 the individual $i$ such that $q_i$ is highest.
@@
 the probability of being the most informed over many signaling events.
-The behavior that the birds can change to improve their fitness is the number of other birds they pay attention to, $n_i$.
-In particular, each individual has a strategy $n_i$ that dictates how many neighbors it has, so that $n_i=\sum_{j\neq i}{\bf I}(A_{ij}\neq 0)$.
+If there are abundant resources, we assume that the first four individuals to the resources can share it, so that fitness is proportional to the probability of being one of the four most informed birds.
+Finally, we consider a combination of predation and resources: fitness is proportional to the $1$ minus the probability of being predated times the probability of being the most informed.
+The behavior that the birds can change to improve their fitness is the number of other birds they pay attention to, $k_i$.
+In particular, each individual has a strategy $k_i$ that dictates how many neighbors it has, so that $k_i=\sum_{j\neq i}{\bf I}(A_{ij}\neq 0)$.
 
@@
 In a homogeneous flock, where every bird is using the same strategy, as both the strategy and the radius of the signal change, both $\mathscr{H}_2$ and the correlation length are affected.
-Figure \ref{compare_props} shows that there is a relationship between two group properties: the $\mathscr{H}_2$ norm and the correlation length of the flock.
-%%WHY? intuition for why each changes as they do with strategy and radius?!?!?
+Figure \ref{compare_props} shows that both $\mathscr{H}_2$ and correlation length are maximized at intermediate strategies when the radius is low and at low strategies when the radius is high.
+We prove in the Appendix (Appendix Claim \ref{cov_equivalence}) that if we use a symmetric matrix to calculate $\rho$ and all birds have the same number of neighbors $n$, then the covariance matrices used to calculate $\rho$ and $\lambda$ are identical up to a constant, which suggests that the two properties cannot be optimized independently.
 
 \subsection{There are multiple ESS, which depend on the content of the signal. }
-Surprisingly, even without imposing costs on paying attention to more neighbors, paying attention to the whole flock is not usually the best strategy.
+Surprisingly, even without imposing costs on paying attention to more neighbors, paying attention to the whole flock is not always the best strategy.
 The ESS strategy of the number of neighbors to pay attention strongly depends on which selection pressure is being applied.
+
+Those birds that can perceive the environmental signal will always have better information than those that cannot.
+Selection, therefore, acts to improve a bird's ability to learn about the signal when it cannot perceive it.
+In Appendix Claim \ref{exp_approx} we show that an uninformed bird's opinion can be approximated by $\sum_{\text{informed }j}L_{ij}$, that is the total weight given to informed birds.
+Being the most well-informed uninformed bird, therefore, amounts to having the highest proportion of informed neighbors and being the least well-informed amounts to having the smallest proportion of informed neighbors.
+As described above, the hypergeometric distribution is a good approximation for the probability distribution of the number of informed neighbors a bird with strategy $k_i$ will have.
+As shown in Figure \ref{hypergeo}, under the hypergeometric distribution, having many neighbors makes it unlikely that few of those neighbors will be uninformed, but it also makes it unlikely that all of them will be uninformed (Figure \ref{hypergeo}).
+
+These observations explain the ESS number of neighbors we find.
+
 If a focal bird pays attention to many others, its opinion will be similar to theirs.
@@
 Figure \ref{ESS}a shows the ESS strategy when selection is due to knowledge about resources is always lower than the ESS strategy when selection is due to knowledge about predators (Figure \ref{ESS}a).
+%%a different explanation: Having many neighbors means that your opinion will be close to the average opinion of the group.
+Even if some of your many neighbors or informed, others will not be, and your opinion will be somewhere in between.
+If you have few neighbors, on the other hand, your opinion will often be at the extremes of the group: low, if you have few informed neighbors, and high, if you have many informed neighbors.
+To avoid being in the bottom of the group, therefore, you should avoid having few neighbors.
+To try to be in the top, on the other hand, having few neighbors is a good strategy.
+I think maybe this a more intuitive way to say it?
 
@@
 
-Figure \ref{ESS_v_H2} shows how the ESS strategy changes as a function of the $\mathscr{H}_2$ robustness.
+%Figure \ref{ESS_v_H2} shows how the ESS strategy changes as a function of the $\mathscr{H}_2$ robustness.
 
@@
 
-\subsection{Linear algebra background }
+\subsection{Linear algebra claims and theorems }
+
+\begin{claim} 
+\label{exp_approx}
+Suppose. $v(t)$ satisfies $\dot{v}=Av+f$ with $v(0)=v_0$, where $A$ has no $0$ eigenvalues. If $V$ is the matrix of eigenvectors of $A$ and $\Lambda$ is diagonal with the eigenvalues of $A$, then
+\begin{equation} \label{diffeq_solution}
+v(t)=V\exp(t\Lambda)V^{-1}v_0+V\Lambda^{-1}(\exp(t\Lambda)-I)V^{-1}f.
+\end{equation}
+We can approximate $v(t)\approx V\exp(t\Lambda)V^{-1}v_0+V(tI+\frac{t^2}{2}A)V^{-1}f$ if the eigenvalues are not too far from $0$.
+In the case that $A=L_\text{f}$ and $f=B\vec{1}$, this suggests that for $i$ that cannot perceive the signal, $v_i(t)\approx \frac{t^2}{2}\sum_{\text{informed }j }L_{ij}.$ This approximation is reasonable for $i$ such that $B_{ii}=0$, but worse for $i$ such that $B_{ii}\neq 0$.
+\end{claim}
+
+\begin{pf}
+First we show that (\ref{diffeq_solution}) solves the differential equation. Note that if $A=V\Lambda V^{-1}$.
+\begin{align*}
+\dot{v}&=V\Lambda \exp(t\Lambda)V^{-1}v_0+V\Lambda^{-1}(\Lambda\exp(t\Lambda))V^{-1}f
+\\&=V\Lambda V^{-1}V\exp(t\Lambda)V^{-1}v_0+V\exp(t\Lambda)V^{-1}f
+\\&=AV\exp(t\Lambda)V^{-1}v_0+V\exp(t\Lambda)V^{-1}f-VV^{-1}f+VV^{-1}f
+\\&=AV\exp(t\Lambda)V^{-1}v_0+V(\exp(t\Lambda)-I)V^{-1}f+f
+\\&=AV\exp(t\Lambda)V^{-1}v_0+V\Lambda V^{-1}V \Lambda^{-1}(\exp(t\Lambda)-I)V^{-1}f+f
+\\&=AV\exp(t\Lambda)V^{-1}v_0+A\Lambda^{-1}(\exp(t\Lambda)-I)V^{-1}f+f
+\\&=Av(t)+f.
+\end{align*}
+If $g(x)=\frac{1}{x}(\exp(ax)-1)$, then $g(x)\approx a+\frac{a^2}{2}x$ for $x$ close to $0$.
+Then, if the eigenvalues of $A$ are not too far from $0$, $\Lambda^{-1}(\exp(t\Lambda)-I)\approx  tI+\frac{t^2}{2}\Lambda$.
+If $A=L_\text{f}$, then all eigenvalues of $A$ are negative.
+If $N_\text{I}$ is the number of informed individuals, i.e.
+$i$ such that $B_{ii}\neq0$, then there are $N_\text{I}$ eigenvalues of $A$ close to $-2$ such that, if  $k$ is the index of one of these eigenvalues, $V_{ik}$ is large if and only if $i$ is informed.
+The rest of the $N-N_\text{I}$ eigenvalues are between $-2$ and $0$ and, if $k$ is the index of one of these eigenvalues, $V_{ik}$ is large if and only if $i$ is uninformed.
+Therefore, the eigenvalue approximation is worst for the informed individuals and best for the uninformed individuals.
+Using the eigenvalue approximation, $v(t)$ becomes
+\begin{align*}
+v(t)=V\exp(t\Lambda)V^{-1}v_0+tVV^{-1}B\vec{1}+\frac{t^2}{2}V\Lambda V^{-1}B\vec{1}=V\exp(t\Lambda)V^{-1}v_0+tB\vec{1}+\frac{t^2}{2}L_\text{f}B\vec{1}
+\end{align*}
+Ignoring the informed individuals and the initial conditions, this gives $v_i(t)\approx \sum_{\text{informed } j}L_{ij}$.
+\end{pf}
 
 \begin{theorem} \label{lyapeq_theorem}
-Given any positive symmetric matrix $Q>0$ there exists a unique positive symmetric matrix $\Sigma>0$ satisfyying $A\Sigma+\Sigma A^T=Q$ if and only if the system $\dot{x}=Ax$ is globally asymptotically stable.
+Given any positive symmetric matrix $Q>0$ there exists a unique positive symmetric ] matrix $\Sigma>0$ satisfyying $A\Sigma+\Sigma A^T=Q$ if and only if the system $\dot{x}=Ax$ is globally asymptotically stable.
 \end{theorem}
+\begin{pf}
+Proof available elsewhere.
+\end{pf}
 
@@
 \\ &= A\Sigma_t+\Sigma_tA^T+vE[y(t)^T]+E[y(t)]v^T+QQ^T
-\\ \Rightarrow \lim_{t\to\infty} \dot{\Sigma}&=A\Sigma+\Sigma A-vv^T(A^{-1})^T-A^{-1}vv^T+QQ^T
-\\ \Rightarrow A\Sigma+\Sigma A&=vv^T(A^{-1})^T+A^{-1}vv^T-QQ^T \text{ since $\dot{\Sigma}=0$ at equilibrium}.
+\\ \Rightarrow \lim_{t\to\infty} \dot{\Sigma}&=A\Sigma+\Sigma A^T-vv^T(A^{-1})^T-A^{-1}vv^T+QQ^T
+\\ \Rightarrow A\Sigma+\Sigma A^T&=vv^T(A^{-1})^T+A^{-1}vv^T-QQ^T \text{ since $\dot{\Sigma}=0$ at equilibrium}.
 \end{align*}
@@
 
-%%ORGANIZE INTRODUCTION TO THE ROTATION MATRICES. 
-We define a matrix $Q\in\R^{N-1}\times\R^N$ such that each row of $Q$ is orthogonal to $\vec{1}$ and the rows are pairwise orthogonal.
+\begin{fact}
+If $V=\sum_iv_i/N$ and $C=\left(E[(v_i-V)(v_j-V)]\right)_{ij}$, then $\sum_{ij}C_{ij}=0$.
+\end{fact}
+\begin{pf}
+\begin{align*}
+\sum_{ij}E[(v_i-V)(v_j-V)]&=\sum_{ij}\left(E[v_iv_j]-E[v_iV]-E[v_jV]+E[VV]\right)
+\\&=\sum_{ij}E[v_iv_j]-N\sum_iE[v_iV]-N\sum_jE[v_jV]+N^2E[VV]
+\\&=\sum_{ij}E[v_iv_j]-\sum_{ij}E[v_iv_j]-\sum_{ij}E[v_iv_j]+\sum_{ij}E[v_iv_j]
+\\&=0
+\end{align*}
+\end{pf}
+
+In the following sections of the Appendix, transforming an opinion vector $x$ will be useful.
+To do so, we define a matrix $Q\in\R^{N-1}\times\R^N$ such that each row of $Q$ is orthogonal to $\vec{1}$ and the rows are pairwise orthogonal.
 Young et al.
 show in \cite{Young:2010fk} show that this equivalent to finding $Q$ such that $Q\vec{1}=\vec{0}$, $QQ^T=I_{N-1}$, and $Q^TQ=I_N-\frac{1}{N}\vec{1}_N\vec{1}_N^T$.
-Since the rows of $R$ are orthonormal, $RR^T=I$.
-Further, since the rows of $R$ are orthonormal, $R$ is invertible, which implies that $R^T=R^{-1}$.
-Then we can write
+We next define a rotation matrix $R\in\R^N\times\R^N$ such that the last row of $R$ is $\frac{1}{\sqrt{N}}\vec{1}^T$ and the other rows are given by the rows of $Q$.
+Since the rows of $R$ are orthonormal, $RR^T=I$ and $R$ is invertible, which implies that $R^T=R^{-1}$.
+If $z=Q^TQx$, then $z_i=v_i-\frac{1}{N}\sum_iv_i$.
+Similarly, if $\bar{y}=Rx$, then $\bar{y}_N=\frac{1}{\sqrt{N}}\sum_ix_i$ and $y=(\bar{y}_1,\dots,\bar{y}_{N-1})^T=Qx$ gives the component of $v$ that is orthogonal to the consensus vector $\vec{1}$.
+We can also translate between $y$ and $z$ by noting that $z=Q^Ty$ and $y=QQ^Tz$ or $y$ is given by the first $N-1$ coordinates of $Rz$.
 
-Again, we will use a rotation matrix $R\in\R^N\times\R^N$ such that the last row of $R$ is $\frac{1}{\sqrt{N}}\vec{1}$ and the other rows are normal and orthogonal to $\vec{1}$.
-Above, we found a $Q\in\R^{N-1}\times\R^N$ such that each row of $Q$ was orthogonal to $\vec{1}$ and $R\in\R^N\times\R^N$ be the matrix with $\frac{1}{\sqrt{N}}\vec{1}^T$ as its last row and the first $N-1$ rows identical to the rows of $Q$.
-Then $R$ is a rotation matrix, so $RR^T=R^TR=I$ and $\det(R)=1$.
-This allows us to
-\[
- \boxed{\bar{y}=Rx$, $y=Qx=\bar{y}_{i=1}^{N-1}$ (or $y=RQ^TQx$), $z=Q^TQx$, $V=\frac{1}{N}\sum_ix_i}
- \]
-
- Depending on the model framework, we will find either $E[yy^T]$ or $E[zz^T]$. This are interchangeable since $E[yy^t]=E[Rzz^TR^T]=RE[zz^T]R^T$ and $E[zz^T]=E[R^Tyy^TR]=R^TE[yy^T]R$.
+\begin{equation} \label{transformations}
+ \boxed{V=\frac{1}{N}\sum_ix_i ,\ z=Q^TQx=x-V\vec{1}, \ \bar{y}=Rx,\ y=Qx=\bar{y}_{i=1}^{N-1}}
+ \end{equation}
 
@@
 \begin{equation}
-\dot{x_i}=\frac{\sum_{\text{neighbors $j$ of $i$ } }(x_j-x_i)}{K_i} +\sqrt{n_i}\xi_i=\sum_j\bar{A}_{ij}x_j-x_i+\sqrt{n_i}\xi_i=\sum_jL_{ij}x_j+\sqrt{n_i}\xi_i.
+\dot{x_i}=\frac{\sum_{\text{neighbors $j$ of $i$ } }(x_j-x_i)}{k_i} +\sqrt{k_i}\xi_i=\sum_j\bar{A}_{ij}x_j-x_i+\sqrt{k_i}\xi_i=\sum_jL_{ij}x_j+\sqrt{k_i}\xi_i.
 \end{equation}
@@
 If $b\neq 0$, $L_\text{s}$ and therefore $\tilde{L}_\text{s}$ have all negative eigenvalues and $\bar{y}=(0, 0, \dots 0, \sqrt{N})^T$ is a globally asymptotically stable equilibrium.
-Therefore, by Claim \ref{lyapeq_derivation} and Theorem \ref{lyapeq_theorem}, $\lim_{t\to\infty}E[\bar{y}(t)\bar{y}(t)^T]=\Sigma_{\bar{y}}$ where $\tilde{L}_\text{s}\Sigma_{\bar{y}}+\Sigma_{\bar{y}}\tilde{L}_\text{s}^T=-RD^2R^T$.
+Therefore, by Theorem \ref{lyapeq_theorem} and Claim \ref{lyapeq_derivation}, $\lim_{t\to\infty}E[\bar{y}(t)\bar{y}(t)^T]=\Sigma_{\bar{y}}$ where $\tilde{L}_\text{s}\Sigma_{\bar{y}}+\Sigma_{\bar{y}}\tilde{L}_\text{s}^T=-RD^2R^T$.
 We can  find $\Sigma_y=\lim_{t\to\infty}E[y(t)y(t)^T]$
@@
 Table \ref{correspondence} shows how to find the covariance of $y$ depending on the framework being used and whether or not a signal is present.
+
 \begin{claim}\label{cov_equivalence}If we use $P$ instead of $L$ with the noisy dynamics framework and all birds have the same number of neighbors $n$, then $C_y=\frac{2}{n}\Sigma_y$.
@@
 \end{pf}
+
 A similar proof shows the equivalence of the two covariance matrices in the presence of a signal, as long as $P_\text{s}$ is used in the noisy dynamical system and all birds have the same number of neighbors.
@@
 \\$\bar{A}$ & normalized adjacency matrix & $\bar{A}_{ij}=\left\{\begin{array}{l l l }
-\frac{1}{K_i} & \text{ if $i$ pays attention to $j$, where $K_i=$ \# of birds $i$ pays attention to}
+\frac{1}{k_i} & \text{ if $i$ pays attention to $j$, where $k_i=$ \# of birds $i$ pays attention to}
 \\ 0 & \text{ otherwise }
@@
 \begin{figure}
-\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/H2_v_corrlength.pdf}
-\caption{\label{compare_props} }
+\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/H2_and_corrlength.pdf}
+\caption{\label{compare_props}  Both $\mathscr{H}_2$ and correlation length are maximized at intermediate strategies when there are few informed birds and at low strategies when there are many informed birds.
+In (a), we show the average $\mathscr{H}_2$ robustness over $1000$ signaling events, where all birds use the same strategy and the signal is available to all birds within a circle of a given radius.
+In (b), we show the average correlation length over $1000$ signaling events, where all birds use the same strategy and the signal is available to all birds within a circle of a given radius.
+}
 \end{figure}
@@
 \includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/ESSfigure.pdf}
-\caption{\label{ESS} The ESS number of neighbors is always higher if selection is due to predation than due to resources.
-When selection is due to predation, there can be multiple ESS strategies, which are a non-monotonic function of the radius of the signal.
-In (a), we show the (possibly multiple) ESS strategies as a function of the radius of the signal.
-In (b) we show the relative fitness of an invader as a function of the invader and resident strategies when $r=0.1$.
-The dots show the multiple ESSs.
-For (c) and (d), we choose a focal resident bird and consider the situation where neither the focal bird nor the invader has the environmental signal.The total of the stacked bars is the total probability that the invader will be predated when neither the focal bird nor the invader have the signal.
-We further break down these cases into five cases: there is a single informed bird and neither focal bird, only the focal resident, only the invader, or both pay attention to it or there are multiple informed birds.
-The horizontal dashed line shows $1/N$, the probability of being predated if the strategies are neutral.
-In (c), the invader has one fewer neighbor than the resident.
-In (d), the invader has one more neighbor than the resident.
-Parameters:  in all panels $N=20$, $\beta=1$, $T=1$, in (b)-(d) $r=0.1$.
+\caption{\label{ESS} The ESS number of neighbors is always higher if selection is due to predation than due to resources or to both.
+When selection is due to predation, the (possibly multiple) ESS strategies increase with the radius of the signal.
+In (a), we show the relative fitness of an invader as a function of the invader and resident strategies when the radius $r=0.1$, where fitness depends on escaping predation.
+The circles indicate the ESS strategies, i.e.
+those resident strategies such that no invader has a higher fitness than the resident.
+In (b), we show the (possibly multiple) ESS strategies as a function of the radius of the signal.
+The circles show the ESS when selection is due to predation, the squares when selection is due to access to scarce resources, and the diamonds when selection is due to both.
+The diamonds and squares overlap.
+There are no ESS when selection is due to access to abundant resources.
+Parameters:  in both panels $N=20$, $b=1$, $T=1$.
+}
+\end{figure}
+
+%\begin{figure}
+%\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/robustness_vs_ESS.pdf}
+%\caption{\label{ESS_v_H2}
+%%The correlation length of a homogeneous flock depends on both the strategy being used and the radius of the signal, but it depends much more strongly on the radius than the strategy.
+The ESS number of neighbors is an increasing function of correlation length of a homogeneous group.
+In (a), we show the average correlation length of a homogeneous flock as a function of the strategy being used and the radius of the signal.
+In (b), each point corresponds to a value of the radius of signal: the horizontal axis shows the correlation length of a flock paying attention to $N/2$ neighbors for that radius and the vertical axis shows the ESS number of neighbors for that radius.
+Parameters: $N=20$, $\beta=1$, $T=1$.
+%}
+%\end{figure}
+
+\begin{figure}
+\includegraphics[width=4in]{/Users/eleanorbrush/Desktop/greedyoptneighbors_radius=1.pdf}
+\caption{\label{greedyopt} When the birds perform greedy optimization and change their number of neighbors to improve their own fitness, they settle on strategies close to the ESS strategies for that selection regimes.
+When selection is due to predation, they settle on high strategies; when selection is due to access to few resources or to both resources and predation, they settle on low strategies; when selection is due to access to abundant resources, they settle in intermediate strategies.
+Only under selection due to generous access to resources does this optimization lead to an increase in both $\mathscr{H}_2$ robustness and correlation length.
+The upper row shows results from selection based to predation, the second row shows results from selection based on access to scarce resources, the third row shows results from selection based on both resources and predation, and the last row shows results from selection based on generous to abundant resources.
+In each panel, the  horizontal axis represents the number of times the birds are allowed to choose optimal strategies.
+The first column shows one example of how the birds' strategies change over time.
+The second column shows, for various initial conditions, how the $\mathscr{H}_2$ robustness changes over time, with the black line showing the average of the $20$ gray lines.
+The third column shows, for various initial conditions, how the correlation length changes over time, with the black line shows the average of the $20$ gray lines.
+Parameters:  $N=20$, $b=1$, $r=.1$, $T=1$.
 }
@@
 \begin{figure}
-\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/robustness_vs_ESS.pdf}
-\caption{\label{ESS_v_H2}
-%The correlation length of a homogeneous flock depends on both the strategy being used and the radius of the signal, but it depends much more strongly on the radius than the strategy.
-The ESS number of neighbors is an increasing function of correlation length of a homogeneous group.
-In (a), we show the average correlation length of a homogeneous flock as a function of the strategy being used and the radius of the signal.
-In (b), each point corresponds to a value of the radius of signal: the horizontal axis shows the correlation length of a flock paying attention to $N/2$ neighbors for that radius and the vertical axis shows the ESS number of neighbors for that radius.
-Parameters: $N=20$, $\beta=1$, $T=1$.
+\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/hypergeometric.pdf}
+\caption{\label{hypergeo} If a fixed number of birds are informed and birds choose neighbors randomly without regard to spatial proximity, the probability of having any number of informed neighbors follows the hypergeometric distribution.
+A bird with more neighbors is less likely to have a low proportion of informed neighbors, but is also less likely to have a high proportion of informed neighbors.
+In (a), we show the hypergeometric distribution, giving the probability of having a number of informed neighbors, normalized by the total number of neighbors, for two strategies, $15$ in blue and $5$ in red.
+In (b), we consider a group made up of $19$ birds using strategy $15$ and $1$ using strategy $5$.
+For each bird, we draw a number of informed neighbors from the hypergeometric distribution  $1000$ times and rank the birds according to the proportion of their neighbors that are informed.
+We show the frequency with which the bird using $5$ neighbors and one of the birds using $15$ neighbors is ranked $1^\text{st}$ through $6^\text{th}$.
+The bird with $5$ neighbors is ranked $1^\text{st}$ much more often, but is also ranked $6^\text{th}$ more often.
+Parameters: $N=20$.
 }
@@
 
-\begin{figure}
-%\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/greedyoptneighbors.pdf}
-\caption{\label{greedyopt} When the birds choose the optimal number of neighbors, given the strategies the rest of the flock are using, they settle on higher strategies when selection is due to predation than when selection is due to resources.
-Under selection due to predation, the $\mathscr{H}_2$ robustness and correlation length of the flock tend to decrease, whereas under selection due to resources, $\mathscr{H}_2$ robustness is relatively constant and correlation length tends to increase.
-The upper row shows results from implementing selection based to predation and the lower row shows results from implementing selection based on resources.
-In each panel, the  horizontal axis represents the number of times the birds are allowed to choose optimal strategies.
-The first column shows one example of how the birds' strategies change over time.
-The second column shows, for many initial conditions, how the $\mathscr{H}_2$ robustness changes over time.
-The third column shows, for many initial conditions, how the correlation length changes over time.
-Parameters:  $N=20$, $\beta=1$, $r=.1$, $T=1$.
-}
-\end{figure}
-
-\nocite{*}
+%\nocite{*}
 \bibliographystyle{plain}

### 2015-04-14 ESS_3regimes.m

@@ -1,0 +1,156 @@
+numworkers=str2num(getenv('PROCS')); %#ok<ST2NM>
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
+
+numsigs_permove=str2num(getenv('numsigs')); %#ok<ST2NM>
+nummoves=str2num(getenv('nummoves'));%#ok<ST2NM>
+
+N=20;
+b=1;
+
+strats=0:1:(N-1);
+L=length(strats);
+
+Tvals=[1 .01 .1 10];
+Nt=length(Tvals);
+
+radvals=[0:.1:1.4]; %#ok<NBRAK>
+% radvals=.1;
+Nr=length(radvals);
+
+ESSseaten=cell(Nt,Nr);
+ESSsgettoeat=cell(Nt,Nr);
+ESSsboth=cell(Nt,Nr);
+
+storefitnesseaten=cell(Nt,Nr);
+storerhoeaten=cell(Nt,Nr);
+
+storefitnessgettoeat=cell(Nt,Nr);
+storerhogettoeat=cell(Nt,Nr);
+
+storefitnessboth=cell(Nt,Nr);
+storerhoboth=cell(Nt,Nr);
+
+t0=tic;
+
+for q=1:Nt
+    for p=1:Nr
+        T=Tvals(q);
+        radius=radvals(p);
+
+        fitnesseaten=zeros(L,L);
+        rhoeaten=zeros(L,L);
+
+        fitnessgettoeat=zeros(L,L);
+        rhogettoeat=zeros(L,L);
+        
+        fitnessboth=zeros(L,L);
+        rhoboth=zeros(L,L);
+
+        featen=zeros(L,L,N-1);
+        geaten=zeros(L,L,N-1);
+
+        fgettoeat=zeros(L,L,N-1);
+        ggettoeat=zeros(L,L,N-1);
+        
+        fboth=zeros(L,L,N-1);
+        gboth=zeros(L,L,N-1);
+        
+        parfor ind=1:(L*L*(N-1))
+%         parfor ind=1:5
+            [u,v,k]=ind2sub([L,L,N-1],ind);
+            % rows are residents, columns are invaders, transpose to plot
+            % normally
+            resident=strats(u);
+            invader=strats(v);
+
+            strategy=resident*ones(1,N);
[…]

### 2015-04-14 ESS_4regimes.m

@@ -1,0 +1,187 @@
+numworkers=str2num(getenv('PROCS')); %#ok<ST2NM>
+dellacluster=parcluster('local');
+dellacluster.JobStorageLocation=strcat('/scratch/network/brush/tmp/',getenv('SLURM_JOB_ID'));
+dellapool=parpool(dellacluster, numworkers) ;
+
+numsigs_permove=str2num(getenv('numsigs')); %#ok<ST2NM>
+nummoves=str2num(getenv('nummoves'));%#ok<ST2NM>
+
+N=20;
+b=1;
+
+strats=1:1:(N-1);
+L=length(strats);
+
+Tvals=[1 .01 10];
+Nt=length(Tvals);
+
+radvals=[0:.1:1.4]; %#ok<NBRAK>
+% radvals=.1;
+Nr=length(radvals);
+
+ESSseaten=cell(Nt,Nr);
+ESSsgettoeat=cell(Nt,Nr);
+ESSsboth=cell(Nt,Nr);
+ESSsgenerous=cell(Nt,Nr);
+
+storefitnesseaten=cell(Nt,Nr);
+storerhoeaten=cell(Nt,Nr);
+
+storefitnessgettoeat=cell(Nt,Nr);
+storerhogettoeat=cell(Nt,Nr);
+
+storefitnessboth=cell(Nt,Nr);
+storerhoboth=cell(Nt,Nr);
+
+storefitnessgenerous=cell(Nt,Nr);
+storerhogenerous=cell(Nt,Nr);
+
+t0=tic;
+
+for q=1:Nt
+    for p=1:Nr
+        T=Tvals(q);
+        radius=radvals(p);
+
+        fitnesseaten=zeros(L,L);
+        rhoeaten=zeros(L,L);
+
+        fitnessgettoeat=zeros(L,L);
+        rhogettoeat=zeros(L,L);
+        
+        fitnessboth=zeros(L,L);
+        rhoboth=zeros(L,L);
+        
+        fitnessgenerous=zeros(L,L);
+        rhogenerous=zeros(L,L);
+
+        featen=zeros(L,L,N-1);
+        geaten=zeros(L,L,N-1);
+
+        fgettoeat=zeros(L,L,N-1);
+        ggettoeat=zeros(L,L,N-1);
+        
+        fboth=zeros(L,L,N-1);
+        gboth=zeros(L,L,N-1);
+        
+        fgenerous=zeros(L,L,N-1);
[…]

### 2015-04-14 ESS_figures_v10.m

@@ -10,17 +10,21 @@ labfontsz=9;
 
 markersz=3;
 lw=2;
+meanlw=3;
+thinlw=1;
 axislw=.5;
 
-qualcols=cbrewer('qual','Set1',5);
-qualcols=qualcols([1 3 2 4 5],:);
+qualcols=cbrewer('qual','Set1',3);
+qualcols=qualcols([3 1 2],:);
+twocols=cbrewer('qual','Set1',2);b
 seqcols = cbrewer('seq','Reds',11);
 seqcols2 = cbrewer('seq','Reds',7);
 seqcols3 = cbrewer('seq','Reds',15);
+seqcols4 = cbrewer('seq','Reds',22);
+greys=cbrewer('seq','Greys',3);
 
 
-xoffset=.1;
-yoffset=.06;
+
 
 %% load ESS data
 
@@ -75,7 +79,33 @@ resstrats_high = 2:19;
 invaderstrats = resstrats_high-1;
 
 [costs_to_low,~] = costs_by_category(N,resstrats_high,invaderstrats,nummoves,radius,b,T);
+%% find probability of being ranked for low and high strategies
+divisions=(2.5:N/5:N);
+Nd=length(divisions);
+      
+low=4;
+high=16;
+strategy1=low*ones(N,1);
+strategy1(1)=high;
+scores=signalingevents_returnscores(strategy1,numsigs_permove,nummoves,radius,b,T);
+rank=zeros(N,size(scores,2));
+for i=1:size(scores,2)
+    [~,~,ranked]=unique(scores(:,i));
+    rank(:,i)=ranked;
+end
+[lowres_low_count,~]=hist(rank(2,:),divisions);
+[lowres_high_count,~]=hist(rank(1,:),divisions);
 
+strategy2=high*ones(N,1);
+strategy2(1)=low;
+scores=signalingevents_returnscores(strategy2,numsigs_permove,nummoves,radius,b,T);
+rank=zeros(N,size(scores,2));
+for i=1:size(scores,2)
+    [~,~,ranked]=unique(scores(:,i));
+    rank(:,i)=ranked;
+end
+[highres_low_count,~]=hist(rank(1,:),divisions);
+[highres_high_count,~]=hist(rank(2,:),divisions);
 
 %% figure: ESS strategies and prob eaten comparison together
 figure
@@ -86,6 +116,10 @@ w=6.83;
 h=.5*w;
 set(gcf,'Units','inches');
[…]

### 2015-04-14 binomial_experiment.m

@@ -1,0 +1,208 @@
+N=20;
+k2=15;
+k1=5;
+radius=.1;
+strategy=k2*ones(N-1,1);
+strategy=[k1;strategy];
+nummoves=1000;
+m=15;
+p=(m)/N;
+freqs=zeros(N,nummoves);
+ranks=zeros(N,nummoves);
+for i=1:nummoves
+%     draw=binornd(strategy,p);
+    draw=hygernd(N,m,strategy);
+    freq=draw./strategy;
+    freqs(:,i)=freq;
+    [~,~,rank]=unique(-freq);
+    ranks(:,i)=rank;
+end
+eaten=zeros(N,nummoves);
+gettoeat=zeros(N,nummoves);
+for i=1:N
+   eaten(i,:)=(ranks(i,:)==max(ranks));
+   gettoeat(i,:)=(ranks(i,:)==min(ranks));
+end
+eaten=sum(eaten,2);
+gettoeat=sum(gettoeat,2);
+% bins1=(1:(k1+1))/k1;
+% bins1=bins1-(1/k1)/2;
+% bins2=(1:(k2+1))/k2;
+% bins2=bins2-(1/k2)/2;
+bins1=(0:k1)/k1;
+bins2=(0:k2)/k2;
+%%
+figure
+set(gcf,'Color','w')
+w=6.83;
+h=1/3*w;
+set(gcf,'Units','inches');
+set(gcf,'Position',[11.5 3 w h]);
+
+subplot(1,2,1)
+[c1,~]=hist(freqs(1,:),bins1);
+c1=c1/sum(c1);
+[c2,~]=hist(col(freqs(2:end,:)),bins2);
+c2=c2/sum(c2);
+plot(bins2,c2,'-o','LineWidth',lw,'Color',twocols(2,:));
+hold on;
+plot(bins1,c1,'-o','LineWidth',lw,'Color',twocols(1,:))
+box off
+xlabel('Proportion of informed neighbors','FontName',fontname,'FontSize',textfontsz)
+ylabel('Probability','FontName',fontname,'FontSize',textfontsz)
+leg=legend({num2str(k2),num2str(k1)});
+legend boxoff
+v=get(leg,'Position');
+set(leg,'Position',[.2 v(2:4)]) 
+set(gca,'FontName',fontname,'FontSize',labfontsz)
+
+subplot(1,2,2)
+M=max(max(ranks));
+width=.4;
[…]

### 2015-04-14 expected_spin.m

@@ -12,6 +12,7 @@ if nargin==2
 end
 
 initialconditions=.5*ones(N,1);
+% initialconditions=.5*ones(N,1)+normrnd(0,.01,N,1);
 
 [~,vals]=eig(A);
 w=find(sigfig(diag(vals),13)==0,1);

### 2015-04-14 greedyopt_analysis_v2.m

@@ -1,0 +1,189 @@
+% load /Users/eleanorbrush/Desktop/greedyopt_rad=0.1_T=1_nummoves=1000_numpermove=1_timesteps=200.mat
+
+
+%%
+
+% evolution_eaten=zeros(its,N,timesteps)
+its=size(evolution_eaten,1);
+timesteps=size(evolution_eaten,3);
+hold on
+
+plot(1:its,var(evolution_eaten(:,:,end),[],2))
+plot(1:its,var(evolution_gettoeat(:,:,end),[],2),'red')
+legend({'Predation','Resources'})
+
+figure
+hist(col(evolution_eaten(:,:,end)))
+
+%%
+figure
+set(gcf,'Color','w')
+v=get(gcf,'Position');
+ratio=v(4)/v(3);
+w=6.83;
+h=4/3*w;
+set(gcf,'Units','inches');
+set(gcf,'Position',[.5 1 w h]);
+fontname='TimesNewRoman';
+textfontsz=12;
+markersz=3;
+
+xoffset=.08;
+yoffset=-.02;
+
+letters={'(a)','(b)','(c)','(d)','(e)','(f)','(g)','(h)','(i)','(j)','(k)','(l)'};
+
+timesteps=size(evolution_eaten,3);
+M=max([max(maxt_eaten),max(maxt_gettoeat),max(maxt_both),max(maxt_generous)]);
+M=min(timesteps,M+5);
+
+rhomax=max(max(max(rhos_eaten)),max(max(rhos_gettoeat)))+.03;
+corrlengthmax=max(max(max(corrlengths_eaten)),max(max(corrlengths_eaten)))+.08;
+
+subplot(4,3,1)
+i=2;
+plot(1:timesteps,reshape(evolution_eaten(i,:,:),N,[]),'LineWidth',thinlw)
+set(gca,'xlim',[0 M])
+box off
+set(gca,'xtick',0:25:M,'ytick',0:5:N,'tickdir','out')
+set(gca,'FontName',fontname,'FontSize',textfontsz)
+xlabel('Timesteps','FontName',fontname,'FontSize',textfontsz)
+ylabel('Number of neighbors','FontName',fontname,'FontSize',textfontsz)
+apos=get(gca,'Position');
[…]

### 2015-04-14 greedyopt_convergence_check.m

@@ -1,0 +1,88 @@
+evolution_eaten=zeros(its,N,timesteps);
+changeorder_eaten=zeros(its,N,timesteps);
+num=1;
+timesteps=10;
+
+% init=randi([1 N-1],1,N);
+init=19*ones(N,1);
+init(6)=11;init(11)=16;
+
+t=1;
+timesteps2=timesteps;
+first_stop=0;
+while t<=timesteps2
+
+if t==1
+    strategy=init;
+%     strategy=2*ones(1,N);
+%     strategy(1)=10; 
+    evolution_eaten(num,:,t)=strategy;
+    t=t+1;
+else
+    strategy=evolution_eaten(num,:,t-1);
+    strategy=reshape(strategy,N,1);
+
+%         i=randsample(1:N,1);
+%         for i=1:N
+    changeorder=randsample(1:N,N,'false');
+    changeorder_eaten(num,:,t)=changeorder;     
+    for i=changeorder
+        [probeaten, ~,~,~]=signalingevents_new_v2(strategy,numsigs_permove,nummoves,radius,b,T);
+        perf=1-probeaten;
+
+        choosestrat=zeros(1,3);
+        choosestrat(2)=perf(i);
+
+        if strategy(i)==N-1
+            choosestrat(3)=perf(i);
+        else 
+            upone=strategy;
+            upone(i)=strategy(i)+1;
+
+            [probeaten, ~,~,~]=signalingevents_new_v2(upone,numsigs_permove,nummoves,radius,b,T);
+            choosestrat(3)=1-probeaten(i);
+        end
+
+        if strategy(i)==0
+            choosestrat(1)=perf(i);
+        else
+            downone=strategy;
+            downone(i)=strategy(i)-1;
+
+            [probeaten, ~,~,~]=signalingevents_new_v2(downone,numsigs_permove,nummoves,radius,b,T);
+            choosestrat(1)=1-probeaten(i);   
+        end 
+
+        if choosestrat(1)>choosestrat(2)
+            strategy(i)=strategy(i)-1;
+        end
+        if choosestrat(3)>choosestrat(2)
+            strategy(i)=strategy(i)+1;
+        end
+
+    end
+
+    evolution_eaten(num,:,t)=strategy;
+
+    if sum(evolution_eaten(num,:,t)==evolution_eaten(num,:,t-1))==N
+        evolution_eaten(num,:,(t+1):end)=repmat(col(strategy),1,timesteps-t);
+        if first_stop==0
[…]

### 2015-04-14 greedyopt_convergence_explanation.m

@@ -172,3 +172,9 @@ else
 %     end
 end
 end
+
+%%
+i=1;
+strategy=evolution_eaten(i,:,end);
+strategy=sort(strategy);
+[probeaten,~]=signalingevents(strategy,numsigs_permove,nummoves,radius,b,T);

### 2015-04-14 greedyopt_v8.m

@@ -15,34 +15,62 @@ radvals=[.1 .2 .7];
 % Nr=length(radvals);
 T=1;
 
-timesteps=200;
+timesteps=100;
 its=20;
 
 evolution_eaten=zeros(its,N,timesteps);
-groupconsensus_eaten=zeros(its,timesteps);
-groupconsensusforced_eaten=zeros(its,timesteps);
+H2s_eaten=zeros(its,timesteps);
+H2sforced_eaten=zeros(its,timesteps);
 corrlengths_eaten=zeros(its,timesteps);
 corrlengthsforced_eaten=zeros(its,timesteps);
 disconnected_eaten=zeros(its,timesteps);
+rhos_eaten=zeros(its,timesteps);
+rhosforced_eaten=zeros(its,timesteps);
 changeorder_eaten=zeros(its,N,timesteps);
 
 evolution_gettoeat=zeros(its,N,timesteps);
-groupconsensus_gettoeat=zeros(its,timesteps);
-groupconsensusforced_gettoeat=zeros(its,timesteps);
+H2s_gettoeat=zeros(its,timesteps);
+H2sforced_gettoeat=zeros(its,timesteps);
 corrlengths_gettoeat=zeros(its,timesteps);
 corrlengthsforced_gettoeat=zeros(its,timesteps);
 disconnected_gettoeat=zeros(its,timesteps);
+rhos_gettoeat=zeros(its,timesteps);
+rhosforced_gettoeat=zeros(its,timesteps);
 changeorder_gettoeat=zeros(its,N,timesteps);
+
+evolution_both=zeros(its,N,timesteps);
+H2s_both=zeros(its,timesteps);
+H2sforced_both=zeros(its,timesteps);
+corrlengths_both=zeros(its,timesteps);
+corrlengthsforced_both=zeros(its,timesteps);
+disconnected_both=zeros(its,timesteps);
+rhos_both=zeros(its,timesteps);
+rhosforced_both=zeros(its,timesteps);
+changeorder_both=zeros(its,N,timesteps);
+
+evolution_generous=zeros(its,N,timesteps);
+H2s_generous=zeros(its,timesteps);
+H2sforced_generous=zeros(its,timesteps);
+corrlengths_generous=zeros(its,timesteps);
+corrlengthsforced_generous=zeros(its,timesteps);
+disconnected_generous=zeros(its,timesteps);
+rhos_generous=zeros(its,timesteps);
[…]

### 2015-04-14 groupprops.m

@@ -44,7 +44,9 @@ for i=1:nummoves
     problem=find(sigfig(diag(vals),13)==0,1);
     
 %     noise=eye(N);
-    noise=diag(strategy);
+    strategynoise=strategy;
+    strategynoise(strategy==0)=1;
+    noise=diag(strategynoise);
     
     S=.5*(Abar+transpose(Abar));
     P=S;
@@ -57,6 +59,7 @@ for i=1:nummoves
     Cov=Wtilde*inv(Lambdatilde)*transpose(Wtilde); %#ok<*MINV>
 
     todivide=repmat(diag(-Cov),1,N).*repmat(reshape(diag(-Cov),1,[]),N,1);
+    todivide(sigfig(Cov,13)==0)=1;
     todivide=power(todivide,.5);
     Corr=-Cov./todivide;
     dvec=[dvec; col(d)]; %#ok<AGROW>
@@ -81,20 +84,25 @@ for i=1:nummoves
         beta(allreceivers)=b;
         B=diag(beta);
         Lf=L-B;
-        Pf=P-B; 
+        Pf=P-B;
+        [~,vals]=eig(Pf);
+        problem2=find(sigfig(diag(vals),13)==0,1);
         Lftilde=Qfull*Lf*transpose(Qfull);
         Bbar=Qfull*B;
         [~,vals]=eig(Lf);
-        problem2=find(sigfig(diag(vals),13)==0,1);
+        problem3=find(sigfig(diag(vals),13)==0,1);
 
-        if ~isempty(problem2)
+        if ~isempty(problem3) || ~isempty(problem2)
             nooneislistening=nooneislistening+1;
             H2vec_forced(i,j)=Inf;
         else
             sigmaz_forced=lyap(Lftilde,Qfull*noise*transpose(Qfull));
             sigmay_forced=sigmaz_forced(1:end-1,1:end-1);
             H_forced=sqrt(trace(sigmay_forced));
+%             if imag(H_forced)==0
             H2vec_forced(i,j)=H_forced;
+%             else H2vec_forced(i,j)=Inf;
+%             end
         end
         
             toinvert=-Pf-1/(b*sum(allreceivers))*B*ones(N,1)*ones(1,N)*B;
@@ -102,6 +110,7 @@ for i=1:nummoves
[…]

### 2015-04-14 signalingevents.m

@@ -48,9 +48,9 @@ maxmat=repmat(maxvals,N,1);
 maxscorer=zeros(N,size(scores,2));
 
 for i=1:size(scores,2)
-%     look=find(cols==i);
-%     maxscorer(rows(look),i)=1/size(look,1)/numsigs_tot;
-    maxscorer(rows(cols==i),i)=1/numsigs_tot;
+    look=find(cols==i);
+    maxscorer(rows(look),i)=1/size(look,1)/numsigs_tot;
+%     maxscorer(rows(cols==i),i)=1/numsigs_tot;
 end
 probgettoeat=sum(maxscorer,2);
 % end

### 2015-04-14 signalingevents_hypergeo_parallel.m

@@ -1,0 +1,92 @@
+function [probeaten, probgettoeat, both, probgettoeat_generous]=signalingevents_hypergeo_parallel(strategy,numsigs_permove,nummoves,m)
+N=max(size(strategy));
+
+numsigs_tot=numsigs_permove*nummoves;
+
+
+scores=zeros(N,numsigs_permove,nummoves);
+
+parfor i=1:nummoves
+% for i=1:nummoves
+    
+    for j=1:numsigs_permove
+        draw=hygernd(N,m,strategy);
+        freq=draw./strategy;
+        
+        scores(:,j,i)=freq;
+    end
+end
+
+scores=reshape(scores,N,[]);
+
+[minvals,~]=min(scores);
+minmat=repmat(minvals,N,1);
+[rows,cols]=find(abs(scores-minmat)<0.00001);
+minscorer=zeros(N,size(scores,2));
+
+for i=1:size(scores,2)
+    look=find(cols==i);
+    minscorer(rows(look),i)=1/size(look,1)/numsigs_tot;
+%     minscorer(rows(cols==i),i)=1/numsigs_tot;
+end
+probeaten=sum(minscorer,2);
+
+[maxvals,~]=max(scores);
+maxmat=repmat(maxvals,N,1);
+[rows,cols]=find(abs(scores-maxmat)<0.00001);
+maxscorer=zeros(N,size(scores,2));
+
+for i=1:size(scores,2)
+    look=find(cols==i);
+    maxscorer(rows(look),i)=1/size(look,1)/numsigs_tot;
+%     maxscorer(rows(cols==i),i)=1/numsigs_tot;
+end
+probgettoeat=sum(maxscorer,2);
+% end
+
+scores_first=scores(:,1:nummoves/2);
+scores_next=scores(:,(nummoves/2+1):end);
+
+[minvals_first,~]=min(scores_first);
+minmat_first=repmat(minvals_first,N,1);
+[rows_first,cols_first]=find(abs(scores_first-minmat_first)<0.00001);
+minscorer_first=zeros(N,size(scores_first,2));
+
+[maxvals_next,~]=max(scores_next);
+maxmat_next=repmat(maxvals_next,N,1);
+[rows_next,cols_next]=find(abs(scores_next-maxmat_next)<0.00001);
+maxscorer_next=zeros(N,size(scores_next,2));
+
+for i=1:size(scores_first,2)
+    look=find(cols_first==i);
[…]

### 2015-04-14 signalingevents_new_v2.m

@@ -1,0 +1,111 @@
+function [probeaten, probgettoeat, both, probgettoeat_generous]=signalingevents_new_v2(strategy,numsigs_permove,nummoves,radius,b,T)
+N=max(size(strategy));
+
+numsigs_tot=numsigs_permove*nummoves;
+
+
+scores=zeros(N,numsigs_permove,nummoves);
+
+% parfor i=1:nummoves
+for i=1:nummoves
+    positions=unifrnd(0,1,N,2);
+    d=squareform(pdist(positions));
+
+    M=zeros(N);
+    for ind=1:N
+        [~, order]=sort(d(ind,:));
+        neighbors=order(2:strategy(ind)+1);
+        M(ind,neighbors)=1/strategy(ind);
+        M(ind,ind)=-sum(M(ind,neighbors));
+    end
+%     M(1:N+1:end)=-1; %sets diagonal equal to -1
+%     for ind=1:N
+%         if strategy(ind)==0
+%             M(ind,ind)=0;
+%         end
+%     end
+    receivers=randsample(N,numsigs_permove,'true');
+    for j=1:numsigs_permove
+        beta=zeros(N,1);
+        receiver=receivers(j);
+        allreceivers=d(receiver,:)<=radius;
+        beta(allreceivers)=b;
+        v=real(expected_spin(M,T,beta));
+       
+        scores(:,j,i)=v;
+    end
+end
+
+scores=reshape(scores,N,[]);
+
+[minvals,~]=min(scores);
+minmat=repmat(minvals,N,1);
+[rows,cols]=find(abs(scores-minmat)<0.00001);
+minscorer=zeros(N,size(scores,2));
+
+for i=1:size(scores,2)
+    look=find(cols==i);
+    minscorer(rows(look),i)=1/size(look,1)/numsigs_tot;
+%     minscorer(rows(cols==i),i)=1/numsigs_tot;
+end
+probeaten=sum(minscorer,2);
+
+[maxvals,~]=max(scores);
+maxmat=repmat(maxvals,N,1);
+[rows,cols]=find(abs(scores-maxmat)<0.00001);
+maxscorer=zeros(N,size(scores,2));
+
+for i=1:size(scores,2)
+    look=find(cols==i);
+    maxscorer(rows(look),i)=1/size(look,1)/numsigs_tot;
+%     maxscorer(rows(cols==i),i)=1/numsigs_tot;
+end
+probgettoeat=sum(maxscorer,2);
+% end
+
[…]

### 2015-04-14 signalingevents_new_v2_parallel_hypogeom.m

@@ -1,0 +1,118 @@
+function [probeaten, probgettoeat, both, probgettoeat_generous]=signalingevents_new_v2_parallel_hypogeom(strategy,numsigs_permove,nummoves,radius,b,T)
+N=max(size(strategy));
+
+numsigs_tot=numsigs_permove*nummoves;
+
+
+scores=zeros(N,numsigs_permove,nummoves);
+
+parfor i=1:nummoves
+% for i=1:nummoves
+    positions=unifrnd(0,1,N,2);
+    d=squareform(pdist(positions));
+
+    M=zeros(N);
+    for ind=1:N
+%         [~, order]=sort(d(ind,:));
+%         neighbors=order(2:strategy(ind)+1);
+        neighbors=randsample(setdiff(1:N,ind),strategy(ind),'false');
+        M(ind,neighbors)=1/strategy(ind);
+        M(ind,ind)=-sum(M(ind,neighbors));
+    end
+%     M(1:N+1:end)=-1; %sets diagonal equal to -1
+%     for ind=1:N
+%         if strategy(ind)==0
+%             M(ind,ind)=0;
+%         end
+%     end
+    receivers=randsample(N,numsigs_permove,'true');
+    for j=1:numsigs_permove
+        beta=zeros(N,1);
+        receiver=receivers(j);
+        numreceivers=sum(d(receiver,:)<=radius);
+        if numreceivers>=N-1
+            allreceivers=ones(N,1);
+        else allreceivers=zeros(N,1);
+            allreceivers([receiver randsample(setdiff(1:N,receiver),numreceivers,'false')])=1;
+        end
+%         allreceivers=d(receiver,:)<=radius;
+        beta(logical(allreceivers))=b;
+        v=real(expected_spin(M,T,beta));
+       
+        scores(:,j,i)=v;
+    end
+end
+
+scores=reshape(scores,N,[]);
+
+[minvals,~]=min(scores);
+minmat=repmat(minvals,N,1);
+[rows,cols]=find(abs(scores-minmat)<0.00001);
+minscorer=zeros(N,size(scores,2));
+
+for i=1:size(scores,2)
+    look=find(cols==i);
+    minscorer(rows(look),i)=1/size(look,1)/numsigs_tot;
+%     minscorer(rows(cols==i),i)=1/numsigs_tot;
+end
+probeaten=sum(minscorer,2);
+
[…]

### 2015-04-14 signalingevents_parallel_new.m

@@ -1,5 +1,5 @@
-% function [probeaten, probgettoeat]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T)
-% N=max(size(strategy));
+function [probeaten, probgettoeat, both]=signalingevents_parallel_new(strategy,numsigs_permove,nummoves,radius,b,T)
+N=max(size(strategy));
 
 numsigs_tot=numsigs_permove*nummoves;
 
@@ -62,3 +62,29 @@ for i=1:size(scores,2)
 end
 probgettoeat=sum(maxscorer,2);
 % end
+
+scores_first=scores(:,1:nummoves/2);
+scores_next=scores(:,(nummoves/2+1):end);
+
+[minvals_first,~]=min(scores_first);
+minmat_first=repmat(minvals_first,N,1);
+[rows_first,cols_first]=find(abs(scores_first-minmat_first)<0.00001);
+minscorer_first=zeros(N,size(scores_first,2));
+
+[maxvals_next,~]=max(scores_next);
+maxmat_next=repmat(maxvals_next,N,1);
+[rows_next,cols_next]=find(abs(scores_next-maxmat_next)<0.00001);
+maxscorer_next=zeros(N,size(scores_next,2));
+
+for i=1:size(scores_first,2)
+    look=find(cols_first==i);
+    minscorer_first(rows_first(look),i)=1/size(look,1)/(numsigs_tot/2);
+end
+for i=1:size(scores_next,2)
+    look=find(cols_next==i);
+    maxscorer_next(rows_next(look),i)=1/size(look,1)/(numsigs_tot/2);
+end
+
+probeaten_first=sum(minscorer_first,2);
+probgettoeat_next=sum(maxscorer_next,2);
+both=(1-probeaten_first).*probgettoeat_next;

### 2015-04-14 signalingevents_returnscores.m

@@ -1,13 +1,13 @@
-% function [probeaten, probgettoeat]=signalingevents_parallel(strategy,numsigs_permove,nummoves,radius,b,T)
-% N=max(size(strategy));
+function scores=signalingevents_returnscores(strategy,numsigs_permove,nummoves,radius,b,T)
+N=max(size(strategy));
 
 numsigs_tot=numsigs_permove*nummoves;
 
 
 scores=zeros(N,numsigs_permove,nummoves);
 
-parfor i=1:nummoves
-% for i=1:nummoves
+% parfor i=1:nummoves
+for i=1:nummoves
     positions=unifrnd(0,1,N,2);
     d=squareform(pdist(positions));
 
@@ -37,28 +37,3 @@ parfor i=1:nummoves
 end
 
 scores=reshape(scores,N,[]);
-
-[minvals,~]=min(scores);
-minmat=repmat(minvals,N,1);
-[rows,cols]=find(abs(scores-minmat)<0.00001);
-minscorer=zeros(N,size(scores,2));
-
-for i=1:size(scores,2)
-    look=find(cols==i);
-    minscorer(rows(look),i)=1/size(look,1)/numsigs_tot;
-%     minscorer(rows(cols==i),i)=1/numsigs_tot;
-end
-probeaten=sum(minscorer,2);
-
-[maxvals,~]=max(scores);
-maxmat=repmat(maxvals,N,1);
-[rows,cols]=find(abs(scores-maxmat)<0.00001);
-maxscorer=zeros(N,size(scores,2));
-
-for i=1:size(scores,2)
-    look=find(cols==i);
-    maxscorer(rows(look),i)=1/size(look,1)/numsigs_tot;
-%     maxscorer(rows(cols==i),i)=1/numsigs_tot;
-end
-probgettoeat=sum(maxscorer,2);
-% end

### 2015-04-14 talk_figures.m

@@ -1,0 +1,128 @@
+figure
+set(gcf,'Color','w')
+w=6.83;
+h=.5*w;
+set(gcf,'Units','inches');
+set(gcf,'Position',[11.5 3 w h]);
+
+N=20;
+strategy=3*ones(N,1);
+positions=unifrnd(0,1,N,2);
+d=squareform(pdist(positions));
+
+M=zeros(N);
+for ind=1:N
+    [~, order]=sort(d(ind,:));
+    neighbors=order(2:strategy(ind)+1);
+    M(ind,neighbors)=1/strategy(ind);
+    M(ind,ind)=-sum(M(ind,neighbors));
+end
+tograph=sparse(M);
+gplot(tograph,positions,'-o')
+f=findobj(gcf,'Marker','o');
+set(f,'MarkerSize',8,'Color','w','MarkerFaceColor','r','MarkerEdgeColor','r','LineWidth',3)
+
+hold on
+delta=0.8;
+delta2=0.2;
+arrowlength=10;arrow_intangle=15;arrow_angle=30;lw=2;
+for i=1:N
+    for j=setdiff(1:N,i)
+        if M(i,j)~=0
+            v=positions(j,:)-positions(i,:);
+            start=positions(j,:)-delta*v;
+            arrowend=positions(j,:)-delta2*v;
+            arrow(start,arrowend,5,15,30,'EdgeColor','w','MarkerEdgeColor','w','LineWidth',lw)
+        end
+    end
+end
+box off
+gplot(tograph,positions,'-o')
+f=findobj(gcf,'Marker','o');
+set(f,'MarkerSize',8,'Color','w','MarkerFaceColor','r','MarkerEdgeColor','r','LineWidth',3)
+
+axis off
+
+set(gcf,'PaperSize',[w h]);
+set(gcf,'PaperPosition',[0 0 w h]);
+
+filename='/Users/eleanorbrush/Desktop/network6.pdf';
+print(filename,'-dpdf','-r300')
+
+%%
+N=20;radius=.1;b=1;T=1;nummoves=1000;numsigs_permove=1;
+strategy=6*ones(N,1);
+
+
+Q=zeros(N-1,N);
+Q(1,1:2)=[1/sqrt(2) -1/sqrt(2)];
+for j=2:(N-1)
+   v=rand(1,N);
+   v=v-sum(v)/N;
+   for k=1:(j-1)
[…]

### 2015-04-19 EBrush_starling_information.tex
figuring hypergeometric results and weird aliasing in hypergeometric_really

--- 2015-04-14 paper_skeleton.tex
@@
 
+\begin{center}
+{\bf \LARGE{Evolution of Information Gathering Strategies}}
+\vspace{10pt}
+%\\ Eleanor Brush, Naomi Leonard, Simon Levin
+\end{center}
+
+%\tableofcontents
 %
-%\begin{center}
-%{\bf \LARGE{Evolution of Information Gathering Strategies}}
-%\vspace{10pt}
-%\\ Eleanor Brush
-%\\ October 31, 2014
-%\end{center}
-
-\tableofcontents
-
-\section{Abstract}
-Birds in flocks, like animals in many types of social groups, can use their peers to learn about the environment and update their opinions about where to move in that environment.
+%\section{Abstract}
+%Birds in flocks, like animals in many types of social groups, can use their peers to learn about the environment and update their opinions about where to move in that environment.
 Previous work on starlings has measured the number of other birds an individuals pays attention to and shown that this number of ``neighbors" leads to a social network that is conducive to the whole flock reaching consensus.
@@
 This distribution is not an exact description of the probability of $j$ informed neighbors in our model  since it ignores spatial proximity, but it is a useful approximation for considering the advantages and disadvantages of each strategy $k_i$.
+{\red NOTE: I'M RUNNING AN ANALYSIS WHERE FITNESS IS BASED SOLELY ON THIS DISTRIBUTION, NOT ON THE FULL SPATIAL STRUCTURE OF THE MODEL.
+I SHOULD HAVE RESULTS SOON.
+I EXPECT THEM TO BE SIMILAR TO THE FULL MODEL, BUT I WANTED TO MAKE SURE AND TO USE IT AS A WAY OF EXPLAINING THE FULL MODEL RESULTS.}
 
@@
 
-\subsection{There are multiple ESS, which depend on the content of the signal. }
-Surprisingly, even without imposing costs on paying attention to more neighbors, paying attention to the whole flock is not always the best strategy.
-The ESS strategy of the number of neighbors to pay attention strongly depends on which selection pressure is being applied.
+\subsection{The ESS number of neighbors depend on the content of the signal. }
 
@@
 As shown in Figure \ref{hypergeo}, under the hypergeometric distribution, having many neighbors makes it unlikely that few of those neighbors will be uninformed, but it also makes it unlikely that all of them will be uninformed (Figure \ref{hypergeo}).
+Consequently, a bird with more neighbors will be less likely to be the least informed, but will also be less likely to be the best informed.
 
-These observations explain the ESS number of neighbors we find.
+These observations explain the ESS number of neighbors we find, which depends on which selection pressure is being applied.
+Figure \ref{ESS} shows that the ESS number of neighbors when selection is due to access to scarce resources is always $1$.
+The ESS number of neighbors when selection is due to both access to resources and predation is also always $1$.
+On the other hand, the ESS number of neighbors when selection is due to predation is always high, although not necessarily as high as possible.
+There is no ESS when selection is due to access to abundant resources.
 
-If a focal bird pays attention to many others, its opinion will be similar to theirs.
-This will prevent it from being the most informed and hence the first to locate resources.
-Thus, the uninvadable strategy is to pay attention to few neighbors when selection is due to knowledge about resources.
-On the other hand, having many neighbors will prevent the focal bird from being the least informed and thus the most likely to be eaten.
-Thus, the uninvadable strategy is to pay attention to many neighbors when selection is due to knowledge about predators.
-Figure \ref{ESS}a shows the ESS strategy when selection is due to knowledge about resources is always lower than the ESS strategy when selection is due to knowledge about predators (Figure \ref{ESS}a).
-%%a different explanation: Having many neighbors means that your opinion will be close to the average opinion of the group.
-Even if some of your many neighbors or informed, others will not be, and your opinion will be somewhere in between.
-If you have few neighbors, on the other hand, your opinion will often be at the extremes of the group: low, if you have few informed neighbors, and high, if you have many informed neighbors.
-To avoid being in the bottom of the group, therefore, you should avoid having few neighbors.
-To try to be in the top, on the other hand, having few neighbors is a good strategy.
-I think maybe this a more intuitive way to say it?
+\subsection{The predation ESS increases as the signal becomes more public.}
+The other feature of the ESS number of neighbors when selection is due to predation is that it increases as the signal becomes more public (Figure \ref{ESS}).
+When trying to avoid begin the least well-informed, having too few neighbors is always disadvantageous because a bird with too few neighbors runs the risk of having too few or no informed neighbors.
+When the signal is not widely available, having too many neighbors is also disadvantageous because, even if a bird paying attention to many neighbors pays attention to all the informed birds, the total weight given to those informed birds will be lower than a bird with fewer total neighbors.
+As more can perceive the signal, the costs of paying attention to uninformed birds disappear and the ESS is to pay attention to all other birds in the flock.
 
-The ESS strategies when selection is due to knowledge about predators show two interesting features.
-First, there are many parameters for which there are multiple ESS strategies.
-It might seem that if a strategy outperforms all others when it is the resident strategy, it should also outperform other strategies as an invader.
-However, when there are multiple ESS strategies, if we take any two of those ESS strategies, each is uninvadable by the other.
-This indicates how important the social environment set by the resident strategy is.
-The pairwise invisibility plot in Figure \ref{ESS}b shows that low strategies are always costly and that high strategies are only costly when the resident strategy is high.
-
-If a bird has the environmental signal, it will not be predated, regardless of the number of neighbors it has.
-The differences in performance only appear when comparing two birds, neither of whom have the signal.
-To understand the costs of having too few or too many neighbors, we compare the performance of two focal birds, a resident and an invader, neither of whom has the signal.
-There are two main cases: there can be a single informed bird or multiple informed birds.
-We further break down the first case depending on which of these two focal birds pays attention to the informed bird: neither, the resident, the invader, or both.
-For each of these cases, we find the probability that the invader is predated.
-An invader with one fewer neighbor than the resident always has probability of being predated greater than chance ($1/N$) because the resident often has an informed neighbor while the invader how no or fewer informed neighbors and thus a less informed opinion (Figure \ref{ESS}c).
-The total probability of being predated stays more or less constant regardless of the resident strategy, with a slight decrease for high resident strategies (Figure \ref{ESS}b and c).
-An invader with one more neighbor than the resident is only more likely than chance to be predated when the resident strategy is high enough  (Figure \ref{ESS}d).
-This invader performs most poorly when both it and the focal resident pay attention to the informed bird(s).
-In this case, an invader with one more neighbor has more naive neighbors and thus a less informed opinion.
-This is increasingly problematic as the resident strategy increases and both the focal birds are likely to pay attention to the informed bird(s).
-
-\subsection{The predation ESS is to pay attention to more neighbors when the signal is more public.}
-These opposing forces caused by the costs of not paying attention to informed neighbors  and the costs of paying attention to more uninformed neighbors  also explains the second feature of the ESS strategies: they are non-monotonic with respect to the the radius of the signal, i.e.
-how public the signal is (Figure \ref{ESS}a).
-Figure \ref{ESS}a shows that there are essentially three categories of radii: When the radius is low, the signal is private, with only a few birds having true information.
-When the radius is intermediate, the signal is public, with many birds having true information.
-When the radius is high, the signal is universal, with essentially all birds in the flock having true information.
-When the signal is private, paying attention to too few birds is costly because with more neighbors it is more likely to have a truly informed neighbor, but paying attention to too many uninformed birds can also be costly.
-This leads to a high but not maximal ESS strategy.
-When the signal is public, the costs of paying attention to uninformed birds are reduced and the ESS strategy increases to become as high as possible.
-When the signal is universal, not paying attention to informed birds does not impose a cost.
-Additionally, in a group in which the residents pay attention to all of their neighbors, any mutant strategy can get just as much information by paying attention to fewer neighbors so the maximal strategy is no longer an ESS (Figure \ref{ESS}a).
-
-%Figure \ref{ESS_v_H2} shows how the ESS strategy changes as a function of the $\mathscr{H}_2$ robustness.
-
-\subsection{Learned strategies }
-The equilibrium strategies when the individuals are allowed to learn are similar to the ESS strategies: learned strategies when selection is due to predation are always higher than learned strategies when selection is due to resources (Figure \ref{greedyopt}).
-%Surprisingly, even when the group starts with heterogeneous strategies, they often reach a homogeneous equilibrium.
-Under selection due to predation, the $\mathscr{H}_2$ robustness and correlation length of the flock tend to decrease, whereas under selection due to resources, $\mathscr{H}_2$ robustness is relatively constant and correlation length tends to increase (Figure \ref{greedyopt}).
+\subsection{The birds construct an optimal network when they learn strategies to optimize access to abundant resources. }
+Learned strategies when selection is due to predation are high and learned strategies when selection is due to access to scarce resources or both resources and predation are low (Figure \ref{greedyopt}).
+Whereas there is no ESS when selection is due to access to abundant resources, when we let a heterogenous flock learn strategies, we find that most birds in the flock decrease their strategies until they have three or four neighbors under this selection regime.
+As the birds choose new strategies, they change the average properties of the networks they construct.
+When they choose strategies according to predation, access to scarce resources, or both, both $\rho$ and $\lambda$ decrease or, at best, remain constant.
+However, when they choose strategies to get access to abundant resources, both $\rho$ and $\lambda$ increase (Figure \ref{greedyopt}).
+%because this is the convergence stable strategy
 
 \section{Discussion}
-explore vs. exploit 
-risk aversion vs. risk tolerance
-
-expect single ESS strategies but social information depends on the social environment set by the resident strategy
-expect ESS strategies to decrease as the signal becomes more public, but we find the opposite
+%explore vs. exploit 
+%risk aversion vs. risk tolerance
+%
+%Surprisingly, even without imposing costs on paying attention to more neighbors, paying attention to the whole flock is not always the best strategy.
+Surprisingly, when the signal becomes more public, more social information is better.
+%
+%expect single ESS strategies but social information depends on the social environment set by the resident strategy
+%expect ESS strategies to decrease as the signal becomes more public, but we find the opposite
 
@@
 
-\begin{table}
-\caption{\label{correspondence} $P^\dagger$ is the inverse of $P$ with the $0$ eigenvector / value removed. }
-\ra{1.3}
-\begin{tabular}{@{}l|llll@{}}
-& Without signal & With signal 
-\\\hline From noisy dynamics &  $\bar{L}\Sigma_y+\Sigma_y\bar{L}^T=-QD^2Q^T$ & $\tilde{L}_\text{s}\Sigma_{\bar{y}}+\Sigma_{\bar{y}}\tilde{L}_\text{s}^T=-RD^2R^T$, $\Sigma_y=(\Sigma_{\bar{y}})_{i,j=1}^{N-1}$
-\\\hline From Boltzmann equilibrium & $C_y=-QP^{\dagger}Q^T$ & $C_y=(\tilde{P}_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot })^\dagger$
-\\ & & or $C_y=Q(-P_\text{s}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB)^{\dagger}Q^T$
-\end{tabular}
-\end{table}
-
-\begin{table}
-\caption{ \label{matrices} Table of matrices and variables used in the text.}
-\ra{1.3}
-\begin{tabular}{@{}lllll@{}}
-$A$ & adjacency matrix & $A_{ij}=\left\{\begin{array}{l l l}
-1 & \text{ if $i$ pays attention to $j$ }
-\\0 & \text{ otherwise }
-\end{array}\right.$
-\\$\bar{A}$ & normalized adjacency matrix & $\bar{A}_{ij}=\left\{\begin{array}{l l l }
-\frac{1}{k_i} & \text{ if $i$ pays attention to $j$, where $k_i=$ \# of birds $i$ pays attention to}
-\\ 0 & \text{ otherwise }
-\end{array}\right. $
-\\$B$ & diagonal signal matrix & $B_{ij}=\left\{\begin{array}{l l l}
-b & \text{ if $i=j$ and $i$ perceives the environmental signal }
-\\0 & \text{ otherwise }
-\end{array}\right. $
-\\ $\lambda$ & correlation length
-\\ $\Lambda$ & diagonal matrix of eigenvalues of $P$
-\\$L$ & opinion dynamics w/o environmental signal &$L=\bar{A}-I$
-\\$L_\text{s}$ & opinion dynamics with environmental signal & $L_\text{s}=L-B$
-\\$\bar{L}$ & dynamics of deviations from consensus w/o signal & $\bar{L}=QLQ^T$
-\\$\tilde{L}_\text{s}$ & dynamics of deviations from consensus w/ signal & $\tilde{L}=RL_\text{s}R^T$
-\\$S$ & symmetrized interaction matrix & $S=\frac{1}{2}(\bar{A}+\bar{A}^T)$
-\\$P$ & energy interactions w/o environmental signal & $P_{ij}=\left\{\begin{array}{lll}
-S_{ij} & \text{ for } j\neq i
-\\-\sum_kS_{ik} & \text{ for } j=i
-\end{array}\right.
-$
-\\$P_\text{s}$ & energy interactions with environmental signal &  $P_\text{s}=P-B$
-\\$\tilde{P}_\text{s}$
-\\$Q$ & rotates opinion vector away from consensus & $Q\in \R^{N-1}\times\R^N \text{ such that } Q\vec{1}=\vec{0}, \ QQ^T=I_{N-1}, \text{ and } Q^TQ=I_N-\frac{1}{N}\vec{1}_N\vec{1}_N^T $
-\\$R$ & full rotation matrix whose last row is $\frac{1}{\sqrt{N}}\vec{1}^T$ & $\left(\begin{array}{cc} Q \\ \frac{1}{\sqrt{N}}\vec{1}^T \end{array}\right)$
-\\$\rho$ & $\mathscr{H}_2$ robustness & $\lim_{t\to\infty}\frac{1}{E[||y||]}$ where $y=Qx$
-\\$W$ & matrix of eigenvectors of $P$
-\end{tabular}
-\end{table}
+\newpage
+%\nocite{*}
+\bibliographystyle{plain}
+\bibliography{info_evo}
 
@@
 \section{Figures }
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/H2_and_corrlength.pdf}
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/ESSfigure.pdf}
@@
 
-%\begin{figure}
-%\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/robustness_vs_ESS.pdf}
-%\caption{\label{ESS_v_H2}
-%%The correlation length of a homogeneous flock depends on both the strategy being used and the radius of the signal, but it depends much more strongly on the radius than the strategy.
-The ESS number of neighbors is an increasing function of correlation length of a homogeneous group.
-In (a), we show the average correlation length of a homogeneous flock as a function of the strategy being used and the radius of the signal.
-In (b), each point corresponds to a value of the radius of signal: the horizontal axis shows the correlation length of a flock paying attention to $N/2$ neighbors for that radius and the vertical axis shows the ESS number of neighbors for that radius.
-Parameters: $N=20$, $\beta=1$, $T=1$.
-%}
-%\end{figure}
-
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[width=4in]{/Users/eleanorbrush/Desktop/greedyoptneighbors_radius=1.pdf}
-\caption{\label{greedyopt} When the birds perform greedy optimization and change their number of neighbors to improve their own fitness, they settle on strategies close to the ESS strategies for that selection regimes.
+\caption{\label{greedyopt} When the birds change their number of neighbors to improve their own fitness, they settle on strategies close to the ESS strategies for that selection regimes.
 When selection is due to predation, they settle on high strategies; when selection is due to access to few resources or to both resources and predation, they settle on low strategies; when selection is due to access to abundant resources, they settle in intermediate strategies.
@@
 The third column shows, for various initial conditions, how the correlation length changes over time, with the black line shows the average of the $20$ gray lines.
-Parameters:  $N=20$, $b=1$, $r=.1$, $T=1$.
+{\red NOTE: I'M RUNNING THIS FOR MORE TIMESTEPS TO SEE IF THE BOTTOM TWO ROWS CONVERGE MORE RELIABLY.} Parameters:  $N=20$, $b=1$, $r=.1$, $T=1$.
 }
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/hypergeometric.pdf}
@@
 
-%\nocite{*}
-\bibliographystyle{plain}
-\bibliography{info_evo}
+\newpage
+\section{Tables}
+\begin{table}[ht]
+\caption{ \label{matrices} Table of matrices and variables used in the text.}
+\ra{1.3}
+\begin{tabular}{@{}lllll@{}}
+$A$ & adjacency matrix & $A_{ij}=\left\{\begin{array}{l l l}
+1 & \text{ if $i$ pays attention to $j$ }
+\\0 & \text{ otherwise }
+\end{array}\right.$
+\\$\bar{A}$ & normalized adjacency matrix & $\bar{A}_{ij}=\left\{\begin{array}{l l l }
+\frac{1}{k_i} & \text{ if $i$ pays attention to $j$}
+\\ 0 & \text{ otherwise }
+\end{array}\right. $
+\\$B$ & diagonal signal matrix & $B_{ij}=\left\{\begin{array}{l l l}
+b & \text{ if $i=j$ and $i$ perceives the environmental signal }
+\\0 & \text{ otherwise }
+\end{array}\right. $
+\\ $\lambda$ & correlation length
+\\ $\Lambda$ & diagonal matrix of eigenvalues of $P$
+\\$L$ & opinion dynamics w/o  signal &$L=\bar{A}-I$
+\\$L_\text{s}$ & opinion dynamics with  signal & $L_\text{s}=L-B$
+\\$\bar{L}$ & rotated opinion dynamics w/o  signal & $\bar{L}=QLQ^T$
+\\$\tilde{L}_\text{s}$ & rotated opinion dynamics w/ signal & $\tilde{L}=RL_\text{s}R^T$
+\\$S$ & symmetrized interaction matrix & $S=\frac{1}{2}(\bar{A}+\bar{A}^T)$
+\\$P$ & energy interactions w/o  signal & $P_{ij}=\left\{\begin{array}{lll}
+S_{ij} & \text{ for } j\neq i
+\\-\sum_kS_{ik} & \text{ for } j=i
+\end{array}\right.
+$
+\\$P_\text{s}$ & energy interactions with  signal &  $P_\text{s}=P-B$
+\\$\tilde{P}_\text{s}$
+\\$Q$ & $N-1\times N$ rotation matrix & $Q\in \R^{N-1}\times\R^N \text{such that } Q\vec{1}=\vec{0}, QQ^T=I_{N-1}, Q^TQ=I_N-\frac{1}{N}\vec{1}_N\vec{1}_N^T $
+\\$R$ & full rotation matrix  & $\left(\begin{array}{cc} Q \\ \frac{1}{\sqrt{N}}\vec{1}^T \end{array}\right)$
+\\$\rho$ & $\mathscr{H}_2$ robustness & $\lim_{t\to\infty}\frac{1}{E[||y||]}$ where $y=Qx$
+\\$W$ & matrix of eigenvectors of $P$
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\caption{\label{correspondence} $P^\dagger$ is the inverse of $P$ with the $0$ eigenvector / value removed. }
+\ra{1.3}
+\begin{tabular}{@{}l|llll@{}}
+& Without signal & With signal 
+\\\hline From noisy dynamics &  $\bar{L}\Sigma_y+\Sigma_y\bar{L}^T=-QD^2Q^T$ & $\tilde{L}_\text{s}\Sigma_{\bar{y}}+\Sigma_{\bar{y}}\tilde{L}_\text{s}^T=-RD^2R^T$, $\Sigma_y=(\Sigma_{\bar{y}})_{i,j=1}^{N-1}$
+\\\hline From Boltzmann equilibrium & $C_y=-QP^{\dagger}Q^T$ & $C_y=(\tilde{P}_\text{s}-\frac{1}{\tilde{P}_{\text{s}NN}}\tilde{P}_{\text{s}\cdot N}\tilde{P}_{\text{s}N\cdot })^\dagger$
+\\ & & or $C_y=Q(-P_\text{s}-\frac{1}{bN_\text{I}}B\vec{1}\vec{1}^TB)^{\dagger}Q^T$
+\end{tabular}
+\end{table}
 

### 2015-04-19 binomial_experiment_v2.m

@@ -1,12 +1,12 @@
 N=20;
 k2=15;
-k1=5;
-radius=.1;
+k1=17;
+% radius=.1;
 strategy=k2*ones(N-1,1);
 strategy=[k1;strategy];
 nummoves=1000;
-m=15;
-p=(m)/N;
+m=3;
+% p=(m)/N;
 freqs=zeros(N,nummoves);
 ranks=zeros(N,nummoves);
 for i=1:nummoves
@@ -77,7 +77,7 @@ set(gcf,'PaperSize',[w h]);
 set(gcf,'PaperPosition',[0 0 w h]);
 
 filename='/Users/eleanorbrush/Desktop/hypergeometric.pdf';
-print(filename,'-dpdf','-r300')
+% print(filename,'-dpdf','-r300')
 
 %%
 N=50;

### 2015-05-26 paper_skeleton.tex
about to rewrite paper_skeleton a bit

@@
 
-%
-%\begin{center}
-%{\bf \LARGE{Evolution of Information Gathering Strategies}}
-%\vspace{10pt}
-%\\ Eleanor Brush
-%\\ October 31, 2014
-%\end{center}
+\begin{center}
+{\bf \LARGE{Evolution of Information Gathering Strategies}}
+\vspace{10pt}
+%\\ Eleanor Brush, Naomi Leonard, Simon Levin
+\end{center}
 
 \tableofcontents
-
-\section{Abstract}
-Birds in flocks, like animals in many types of social groups, can use their peers to learn about the environment and update their opinions about where to move in that environment.
+%
+%\section{Abstract}
+%Birds in flocks, like animals in many types of social groups, can use their peers to learn about the environment and update their opinions about where to move in that environment.
 Previous work on starlings has measured the number of other birds an individuals pays attention to and shown that this number of ``neighbors" leads to a social network that is conducive to the whole flock reaching consensus.
@@
 As shown in Figure \ref{hypergeo}, under the hypergeometric distribution, having many neighbors makes it unlikely that few of those neighbors will be uninformed, but it also makes it unlikely that all of them will be uninformed (Figure \ref{hypergeo}).
+Consequently, a bird with more neighbors will be less likely to be the least informed, but will also be less likely to be the best informed.
 
-These observations explain the ESS number of neighbors we find.
+These observations explain the ESS number of neighbors we find, which depends on which selection pressure is being applied.
+Figure \ref{ESS} shows that the ESS number of neighbors when selection is due to access to scarce resources is always $1$.
+The ESS number of neighbors when selection is due to both access to resources and predation is also always $1$.
+On the other hand, the ESS number of neighbors when selection is due to predation is always high, although not necessarily as high as possible.
+There is no ESS when selection is due to access to abundant resources.
 
-If a focal bird pays attention to many others, its opinion will be similar to theirs.
-This will prevent it from being the most informed and hence the first to locate resources.
-Thus, the uninvadable strategy is to pay attention to few neighbors when selection is due to knowledge about resources.
-On the other hand, having many neighbors will prevent the focal bird from being the least informed and thus the most likely to be eaten.
-Thus, the uninvadable strategy is to pay attention to many neighbors when selection is due to knowledge about predators.
-Figure \ref{ESS}a shows the ESS strategy when selection is due to knowledge about resources is always lower than the ESS strategy when selection is due to knowledge about predators (Figure \ref{ESS}a).
-%%a different explanation: Having many neighbors means that your opinion will be close to the average opinion of the group.
-Even if some of your many neighbors or informed, others will not be, and your opinion will be somewhere in between.
-If you have few neighbors, on the other hand, your opinion will often be at the extremes of the group: low, if you have few informed neighbors, and high, if you have many informed neighbors.
-To avoid being in the bottom of the group, therefore, you should avoid having few neighbors.
-To try to be in the top, on the other hand, having few neighbors is a good strategy.
-I think maybe this a more intuitive way to say it?
+\subsection{The predation ESS increases as the signal becomes more public.}
+The other feature of the ESS number of neighbors when selection is due to predation is that it increases as the signal becomes more public (Figure \ref{ESS}).
+When trying to avoid begin the least well-informed, having too few neighbors is always disadvantageous because a bird with too few neighbors runs the risk of having too few or no informed neighbors.
+When the signal is not widely available, having too many neighbors is also disadvantageous because, even if a bird paying attention to many neighbors pays attention to all the informed birds, the total weight given to those informed birds will be lower than a bird with fewer total neighbors.
+As more can perceive the signal, the costs of paying attention to uninformed birds disappear and the ESS is to pay attention to all other birds in the flock.
 
-The ESS strategies when selection is due to knowledge about predators show two interesting features.
-First, there are many parameters for which there are multiple ESS strategies.
-It might seem that if a strategy outperforms all others when it is the resident strategy, it should also outperform other strategies as an invader.
-However, when there are multiple ESS strategies, if we take any two of those ESS strategies, each is uninvadable by the other.
-This indicates how important the social environment set by the resident strategy is.
-The pairwise invisibility plot in Figure \ref{ESS}b shows that low strategies are always costly and that high strategies are only costly when the resident strategy is high.
-
-If a bird has the environmental signal, it will not be predated, regardless of the number of neighbors it has.
-The differences in performance only appear when comparing two birds, neither of whom have the signal.
-To understand the costs of having too few or too many neighbors, we compare the performance of two focal birds, a resident and an invader, neither of whom has the signal.
-There are two main cases: there can be a single informed bird or multiple informed birds.
-We further break down the first case depending on which of these two focal birds pays attention to the informed bird: neither, the resident, the invader, or both.
-For each of these cases, we find the probability that the invader is predated.
-An invader with one fewer neighbor than the resident always has probability of being predated greater than chance ($1/N$) because the resident often has an informed neighbor while the invader how no or fewer informed neighbors and thus a less informed opinion (Figure \ref{ESS}c).
-The total probability of being predated stays more or less constant regardless of the resident strategy, with a slight decrease for high resident strategies (Figure \ref{ESS}b and c).
-An invader with one more neighbor than the resident is only more likely than chance to be predated when the resident strategy is high enough  (Figure \ref{ESS}d).
-This invader performs most poorly when both it and the focal resident pay attention to the informed bird(s).
-In this case, an invader with one more neighbor has more naive neighbors and thus a less informed opinion.
-This is increasingly problematic as the resident strategy increases and both the focal birds are likely to pay attention to the informed bird(s).
-
-\subsection{The predation ESS is to pay attention to more neighbors when the signal is more public.}
-These opposing forces caused by the costs of not paying attention to informed neighbors  and the costs of paying attention to more uninformed neighbors  also explains the second feature of the ESS strategies: they are non-monotonic with respect to the the radius of the signal, i.e.
-how public the signal is (Figure \ref{ESS}a).
-Figure \ref{ESS}a shows that there are essentially three categories of radii: When the radius is low, the signal is private, with only a few birds having true information.
-When the radius is intermediate, the signal is public, with many birds having true information.
-When the radius is high, the signal is universal, with essentially all birds in the flock having true information.
-When the signal is private, paying attention to too few birds is costly because with more neighbors it is more likely to have a truly informed neighbor, but paying attention to too many uninformed birds can also be costly.
-This leads to a high but not maximal ESS strategy.
-When the signal is public, the costs of paying attention to uninformed birds are reduced and the ESS strategy increases to become as high as possible.
-When the signal is universal, not paying attention to informed birds does not impose a cost.
-Additionally, in a group in which the residents pay attention to all of their neighbors, any mutant strategy can get just as much information by paying attention to fewer neighbors so the maximal strategy is no longer an ESS (Figure \ref{ESS}a).
-
-%Figure \ref{ESS_v_H2} shows how the ESS strategy changes as a function of the $\mathscr{H}_2$ robustness.
-
-\subsection{Learned strategies }
-The equilibrium strategies when the individuals are allowed to learn are similar to the ESS strategies: learned strategies when selection is due to predation are always higher than learned strategies when selection is due to resources (Figure \ref{greedyopt}).
-%Surprisingly, even when the group starts with heterogeneous strategies, they often reach a homogeneous equilibrium.
-Under selection due to predation, the $\mathscr{H}_2$ robustness and correlation length of the flock tend to decrease, whereas under selection due to resources, $\mathscr{H}_2$ robustness is relatively constant and correlation length tends to increase (Figure \ref{greedyopt}).
+\subsection{The birds construct an optimal network when they learn strategies to optimize access to abundant resources. }
+Learned strategies when selection is due to predation are high and learned strategies when selection is due to access to scarce resources or both resources and predation are low (Figure \ref{greedyopt}).
+Whereas there is no ESS when selection is due to access to abundant resources, when we let a heterogenous flock learn strategies, we find that most birds in the flock decrease their strategies until they have three or four neighbors under this selection regime.
+As the birds choose new strategies, they change the average properties of the networks they construct.
+When they choose strategies according to predation, access to scarce resources, or both, both $\rho$ and $\lambda$ decrease or, at best, remain constant.
+However, when they choose strategies to get access to abundant resources, both $\rho$ and $\lambda$ increase (Figure \ref{greedyopt}).
+%because this is the convergence stable strategy
 
 \section{Discussion}
-explore vs. exploit 
-risk aversion vs. risk tolerance
+%explore vs. exploit 
+%risk aversion vs. risk tolerance
+%
+%Surprisingly, even without imposing costs on paying attention to more neighbors, paying attention to the whole flock is not always the best strategy.
+Surprisingly, when the signal becomes more public, more social information is better.
+%
+%expect single ESS strategies but social information depends on the social environment set by the resident strategy
+%expect ESS strategies to decrease as the signal becomes more public, but we find the opposite
 
-expect single ESS strategies but social information depends on the social environment set by the resident strategy
-expect ESS strategies to decrease as the signal becomes more public, but we find the opposite
+%it is by no means guaranteed that selection acting on individuals' learning abilities will lead them to use strategies that are optimal from the group's perspective, but it is possible.
 

### 2015-05-26 ESS_4regimes_hypogeom_really.m

@@ -6,13 +6,13 @@ dellapool=parpool(dellacluster, numworkers) ;
 numsigs_permove=str2num(getenv('numsigs')); %#ok<ST2NM>
 nummoves=str2num(getenv('nummoves'));%#ok<ST2NM>
 
-N=20;
+N=40;
 b=1;
 
 strats=1:1:(N-1);
 L=length(strats);
 
-Tvals=[1 .01 10];
+Tvals=1;
 Nt=length(Tvals);
 
 mvals=[1:2:N]; %#ok<NBRAK>
@@ -22,7 +22,7 @@ Nm=length(mvals);
 ESSseaten=cell(Nt,Nm);
 ESSsgettoeat=cell(Nt,Nm);
 ESSsboth=cell(Nt,Nm);
-ESSsgenerous=cell(Nt,Nm);
+% ESSsgenerous=cell(Nt,Nm);
 
 storefitnesseaten=cell(Nt,Nm);
 storerhoeaten=cell(Nt,Nm);
@@ -33,14 +33,14 @@ storerhogettoeat=cell(Nt,Nm);
 storefitnessboth=cell(Nt,Nm);
 storerhoboth=cell(Nt,Nm);
 
-storefitnessgenerous=cell(Nt,Nm);
-storerhogenerous=cell(Nt,Nm);
+% storefitnessgenerous=cell(Nt,Nm);
+% storerhogenerous=cell(Nt,Nm);
 
 t0=tic;
 
 for q=1:Nt
     for p=1:Nm
-        T=Tvals(q);
+%         T=Tvals(q);
         m=mvals(p);
 
         fitnesseaten=zeros(L,L);
@@ -52,8 +52,8 @@ for q=1:Nt
         fitnessboth=zeros(L,L);
         rhoboth=zeros(L,L);
         
-        fitnessgenerous=zeros(L,L);
-        rhogenerous=zeros(L,L);
+%         fitnessgenerous=zeros(L,L);
+%         rhogenerous=zeros(L,L);
 
         featen=zeros(L,L,N-1);
         geaten=zeros(L,L,N-1);
@@ -64,8 +64,8 @@ for q=1:Nt
         fboth=zeros(L,L,N-1);
         gboth=zeros(L,L,N-1);
         
-        fgenerous=zeros(L,L,N-1);
-        ggenerous=zeros(L,L,N-1);
+%         fgenerous=zeros(L,L,N-1);
+%         ggenerous=zeros(L,L,N-1);
         
         parfor ind=1:(L*L*(N-1))
[…]

### 2015-05-26 figures.m

@@ -1,0 +1,451 @@
+%% setup
+N = 20;
+nummoves = 1000;
+numsigs_permove = 1;
+b = 1;
+
+fontname='Times';
+textfontsz=12;
+labfontsz=9;
+
+markersz=3;
+lw=2;
+meanlw=3;
+thinlw=1;
+axislw=.5;
+
+qualcols=cbrewer('qual','Set1',3);
+qualcols=qualcols([3 1 2],:);
+twocols=cbrewer('qual','Set1',2);
+seqcols = cbrewer('seq','Reds',11);
+seqcols2 = cbrewer('seq','Reds',7);
+seqcols3 = cbrewer('seq','Reds',15);
+seqcols4 = cbrewer('seq','Reds',22);
+greys=cbrewer('seq','Greys',3);
+Ncols=cbrewer('qual','Set1',N);
+%% figure: H2 / correlation length heat maps
+figure
+set(gcf,'Color','w')
+w=6.83;
+h=1/4*w;
+set(gcf,'Units','inches');
+set(gcf,'Position',[11.5 3 w h]);
+
+xoffset=.08;
+yoffset=.06;
+
+stratvals=1:19;
+
+subplot(1,2,1)
+imagesc(radvals,stratvals,rhos_forced)
+set(gca,'ydir','normal')
+set(gca,'FontName',fontname,'FontSize',labfontsz)
+set(gca,'ytick',[1 10 19],'yticklabel',[1 10 19])
+caxis manual
+colormap(seqcols)
+cb=colorbar;
+set(cb,'FontName',fontname,'FontSize',labfontsz)
+ylab=ylabel(cb,'Robustness','FontName',fontname,'FontSize',labfontsz);
+set(ylab,'Rotation',270);
+v=get(ylab,'Position');
+set(ylab,'Position',[v(1)+.1 v(2) v(3)])
+
+xlabel('Radius','FontName',fontname,'FontSize',textfontsz)
+ylabel('Strategy','FontName',fontname,'FontSize',textfontsz)
+
+apos=get(gca,'Position');
+annotation('textbox',[apos(1)-xoffset apos(2)+apos(4)+yoffset .01 .04],'String','(a)','FitBoxToText','on','FontSize',textfontsz,'FontName',fontname,'EdgeColor','none','VerticalAlignment','middle','HorizontalAlignment','left')
+
+subplot(1,2,2)
+imagesc(radvals,stratvals,corrlengths_forced)
+set(gca,'ydir','normal')
[…]

### 2015-05-26 greedyopt_v8.m

@@ -87,7 +87,7 @@ for num=1:its
         changeorder=randsample(1:N,N,'false');
         changeorder_eaten(num,:,t)=changeorder;     
         for i=changeorder
-            [probeaten, ~,~,~]=signalingevents_new_v2_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+            [probeaten, ~,~,~]=signalingevents_4regimes_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
             perf=1-probeaten;
             
             choosestrat=zeros(1,3);
@@ -99,7 +99,7 @@ for num=1:its
                 upone=strategy;
                 upone(i)=strategy(i)+1;
 
-                [probeaten, ~,~,~]=signalingevents_new_v2_parallel(upone,numsigs_permove,nummoves,radius,b,T);
+                [probeaten, ~,~,~]=signalingevents_4regimes_parallel(upone,numsigs_permove,nummoves,radius,b,T);
                 choosestrat(3)=1-probeaten(i);
             end
             
@@ -109,7 +109,7 @@ for num=1:its
                 downone=strategy;
                 downone(i)=strategy(i)-1;
 
-                [probeaten, ~,~,~]=signalingevents_new_v2_parallel(downone,numsigs_permove,nummoves,radius,b,T);
+                [probeaten, ~,~,~]=signalingevents_4regimes_parallel(downone,numsigs_permove,nummoves,radius,b,T);
                 choosestrat(1)=1-probeaten(i);   
             end 
             
@@ -168,7 +168,7 @@ for num=1:its
         changeorder=randsample(1:N,N,'false');
         changeorder_gettoeat(num,:,t)=changeorder;     
         for i=changeorder
-            [~, probgettoeat,~,~]=signalingevents_new_v2_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
+            [~, probgettoeat,~,~]=signalingevents_4regimes_parallel(strategy,numsigs_permove,nummoves,radius,b,T);
             perf=probgettoeat;
             
             choosestrat=zeros(1,3);
@@ -180,7 +180,7 @@ for num=1:its
                 upone=strategy;
                 upone(i)=strategy(i)+1;
 
-                [~, probgettoeat,~,~]=signalingevents_new_v2_parallel(upone,numsigs_permove,nummoves,radius,b,T);
[…]

### 2015-05-26 signalingevents_4regimes_parallel.m

@@ -1,5 +1,5 @@
-% function [probeaten, probgettoeat, both, probgettoeat_generous]=signalingevents_hypergeo_parallel(strategy,numsigs_permove,nummoves,m)
-% N=max(size(strategy));
+function [probeaten, probgettoeat, both, probgettoeat_generous]=signalingevents_4regimes_parallel(strategy,numsigs_permove,nummoves,radius,b,T)
+N=max(size(strategy));
 
 numsigs_tot=numsigs_permove*nummoves;
 
@@ -8,12 +8,31 @@ scores=zeros(N,numsigs_permove,nummoves);
 
 % parfor i=1:nummoves
 for i=1:nummoves
-    
+    positions=unifrnd(0,1,N,2);
+    d=squareform(pdist(positions));
+
+    M=zeros(N);
+    for ind=1:N
+        [~, order]=sort(d(ind,:));
+        neighbors=order(2:strategy(ind)+1); 
+        M(ind,neighbors)=1/strategy(ind);
+        M(ind,ind)=-sum(M(ind,neighbors));
+    end
+%     M(1:N+1:end)=-1; %sets diagonal equal to -1
+%     for ind=1:N
+%         if strategy(ind)==0
+%             M(ind,ind)=0;
+%         end
+%     end
+    receivers=randsample(N,numsigs_permove,'true');
     for j=1:numsigs_permove
-        draw=hygernd(N,m,strategy);
-        freq=draw./strategy;
-        
-        scores(:,j,i)=freq;
+        beta=zeros(N,1);
+        receiver=receivers(j);
+        allreceivers=d(receiver,:)<=radius;
+        beta(allreceivers)=b;
+        v=real(expected_spin(M,T,beta));
+       
+        scores(:,j,i)=v;
     end
 end
 
@@ -75,7 +94,7 @@ maxscorer_generous=zeros(N,size(scores,2));
 for i=1:size(scores,2)
     [x,~,z]=unique(scores(:,i));
     [~,o]=sort(x);
-    cutoff_now=min(cutoff,length(x)-1);
+%     cutoff_now=min(cutoff,length(x)-1);
 %     maxscorer_generous(o(z)>=o(end-cutoff_now),i)=1/numsigs_tot;
     choose=zeros(N,1);
     count=0;

### 2015-05-26 signalingevents_new_v2.m

@@ -94,7 +94,7 @@ maxscorer_generous=zeros(N,size(scores,2));
 for i=1:size(scores,2)
     [x,~,z]=unique(scores(:,i));
     [~,o]=sort(x);
-    cutoff_now=min(cutoff,length(x)-1);
+%     cutoff_now=min(cutoff,length(x)-1);
 %     maxscorer_generous(o(z)>=o(end-cutoff_now),i)=1/numsigs_tot;
     choose=zeros(N,1);
     count=0;

### 2015-05-26 signalingevents_new_v2_parallel.m

@@ -14,7 +14,7 @@ parfor i=1:nummoves
     M=zeros(N);
     for ind=1:N
         [~, order]=sort(d(ind,:));
-        neighbors=order(2:strategy(ind)+1);
+        neighbors=order(2:strategy(ind)+1); %#ok<PFBNS>
         M(ind,neighbors)=1/strategy(ind);
         M(ind,ind)=-sum(M(ind,neighbors));
     end
@@ -94,7 +94,7 @@ maxscorer_generous=zeros(N,size(scores,2));
 for i=1:size(scores,2)
     [x,~,z]=unique(scores(:,i));
     [~,o]=sort(x);
-    cutoff_now=min(cutoff,length(x)-1);
+%     cutoff_now=min(cutoff,length(x)-1);
 %     maxscorer_generous(o(z)>=o(end-cutoff_now),i)=1/numsigs_tot;
     choose=zeros(N,1);
     count=0;

### 2015-05-27 paper_skeleton.tex
rewriting appendices

@@
 \usepackage{lscape}
+\usepackage[numbers,sort&compress]{natbib}
+\DeclareRobustCommand{\cite}[1]{\citep{#1}}
 \captionsetup{justification=RaggedRight, singlelinecheck=false}
@@
 
+  \setcounter{tocdepth}{1}
+
 \begin{document}
@@
 \tableofcontents
+
 %
@@
 
-\subsection{Modeling interactions and group-level properties }
-There are two ways to describe equilibrium properties of a flock sharing information over a network.
+\subsection{Opinion dynamics } 
+In our model, each bird has a scalar value, $v_i$, that represents its opinion about the environment.
+Each bird also has a strategy, $k_i$, that represents the number of its peers whose opinions it observes, which it uses to update its own opinion.
+Specifically, a bird observes the $k_i$ nearest birds, which form the neighborhood $\mathscr{N}_i$, and the focal bird changes its opinion to be close to the average of its neighbors':
+\begin{equation*}
+\dot{v_i}=\frac{1}{k_i}\sum_{j \in \mathscr{N}_i}(v_j-v_i).
+\end{equation*}
+The strategy of choosing $k_i$ nearest neighbors will be the focus of our analysis. It will be useful to write down the matrices describing the interaction between pairs of birds.  If 
+\begin{equation*}
+A_{ij}=\left\{\begin{array}{l l l}
+\frac{1}{k_i} & \text{ if $j$ is in $\mathscr{N}_i$ }
+\\0 & \text{ otherwise }
+\end{array}\right. \text{ and }  L_{ij}=\left\{\begin{array}{l l l}
+A_{ij} & \text{ if $j\neq i$ }
+\\-\sum_l A_{il} & \text{ if $j=i$ }
+\end{array}\right.,
+\end{equation*}
+then $L$ can be used to write a system of differential equations describing the vector of opinions $v=(v_1,\dots,v_N)$:
+\begin{equation*}
+\dot{v}=Lv.
+\end{equation*}
+As written, the system will reach consensus with each $v_i=\alpha$ for some $\alpha\in\R$, the exact value of which depends on the initial conditions $v(0)$.
+To incorporate environmental information, we assume that there is an environmental signal such that birds that can perceive the signal change their opinions to be close to $1$.
+Thus, having an opinion closer to $1$ means a bird is more informed and having an opinion farther from $1$ means a bird is less informed.
+The opinion dynamics of a bird that do not receive the signal are unchanged: $v_i=\sum_jL_{ij}v_j$.
+To describe the importance of the environmental signal, we introduce a parameter $b$, so that the opinion dynamics of a bird that \emph{do} receive the signal are given by
+\begin{equation*}
+\dot{v}_i=\sum_jL_{ij}v_j+b(1-v_i).
+\end{equation*}
+If $B$ is the diagonal matrix with $B_{ii}=b$ if $i$ can perceive the signal and $B_{ii}=0$ otherwise, and $L_\text{s}=L-B$, then
+\begin{equation}
+\dot{v}=Lv+B(\vec{1}-v)=L_\text{s}v+B\vec{1}. \label{opdynamics_signal}
+\end{equation}
+Note that $L\vec{1}=\vec{0}$, so $L$ has (at least) one eigenvalue equal to $0$.
+If at least one bird receives an environmental signal, then generally $L_\text{s}$ will have all negative non-zero eigenvalues.
+Then Claim \ref{exp_approx} in the Appendix shows the solution $v(t)$ to this set of ODEs.
+
+\subsection{Individual-level fitness }
+To find the optimal strategies from the individuals' perspectives, we need a measure of individual-level fitness.
+In our model, birds with more information about the environment have higher fitness.
+A signaling event is determined by the interaction matrix $A$ and by which birds receive the signal.
+For a single signaling event, the receivers and the birds to whom they are closest will generally be the most informed.
+To assess the fitness consequences of the number of nearest neighbors $k_i$, we need to determine how well birds with different strategies perform over many signaling events.
+
+We assume that signals appear relatively rarely compared to the timescale on which the birds in a flock move.
+Specifically, at each signaling event, we will assign all birds a new position by drawing a position randomly from a unit square.
+Each bird then pays attention to its $k_i$ nearest neighbors, which determines the adjacency matrix $A$ and therefore the matrix $L$.
+To determine which birds receive the signal, we pick a focal bird at random.
+Then, all birds within a radius $r$ of the focal bird receive the signal, which determines the matrix $B$.
+To fully specify the opinion dynamics, we assume that $v_i(0)=0.5$ when the signal appears for all birds.
+Eq.
+(\ref{opdynamics_signal}) then determines the opinion dynamics.
+We assume that the fitness consequences of the opinion dynamics occur a short time after the signal appears.
+Specifically, fitness is determined at time $t=1$, i.e.
+by $v_i(1)$.
+We repeat this process $1000$ times to model many independent signaling events.
+
+Birds that receive the signal will always be well informed.
+It is their performance when they do \emph{not} receive the signal that In Claim \ref{exp_approx} in the Appendix, we show that, for $i$ that do not receive the signal,
+\begin{align*}
+v_i(1)&\approx v_{i}(0)+\frac{1}{2}\frac{\text{$\#$ of $i$'s neighbors that receive the signal}}{k_i}
+\\&=\frac{1}{2}\left(1+\frac{\text{$\#$ of $i$'s neighbors that receive the signal}}{k_i}\right).
+\end{align*}
+%Therefore, the probability of having an opinion $x$ is 
+%\begin{equation*}
+%P\big(v_i(1) =x\big)=P\big(\text{$i$ has $(2x-1)k_i$  informed neighbors}\big).
+%\end{equation*}
+%for $x=\frac{1}{2},\frac{1}{2}\frac{k_i+1}{k_1},\dots,\frac{1}{2}\frac{k_i+k_i}{k_i}=1$.
+To improve its performance, a bird should increase the probability of having a high proportion of informed neighbors.
+If $N_\text{r}$ birds receive the signal in a flock of size $N$ and each bird chooses $k_i$ neighbors at random (instead of its closet neighbors), the number of informed neighbors bird $i$ has follows a hypergeometric distribution:
+\begin{align}
+P(\text{$i$ has $j$  informed neighbors} )&=\frac{{N_\text{I}\choose j}{N-N_\text{I}\choose k_i-j}  }{{N\choose k_i}} \label{hypergeom}
+\end{align} 
+This gives us an approximation of the distribution $P(v_i(1)=x)$ as a function of $k_i$, which will be useful in considering the advantages and disadvantages of each strategy.
+
+How exactly environmental information is translated into fitness depends on what the information is about.
+We consider signals that provide information about two types of content, predators and food, which lets us consider four selection regimes.
+If the signal is information that a predator is present, we assume that whichever individual is least informed about the signal will be predated.
+The probability of bird $i$ being eaten is therefore
+\begin{equation*}
+P(\text{$i$ is eaten})=\frac{\# \text{ of signaling events in which } v_i(1)=\min_i\{v_i(1)\}}{1000}.
+\end{equation*}
+The probability of $i$ surviving is $P(\text{$i$ survives})=1-P(\text{$i$ is eaten})$ and $i$'s fitness is proportional to this probability.
+If the signal is information that resources are available, we consider two levels of resource abundance.
+If there are scarce resources, we assume that whichever individual is most informed will get access to the resource.
+The probability that bird $i$ gets to eat is therefore
+\begin{equation*}
+P(\text{$i$ gets to eat})=\frac{\# \text{ of signaling events in which } v_i(1)=\max_i\{v_i(1)\}}{1000}
+\end{equation*}
+and $i$'s fitness is proportional to this probability.
+If there are abundant resources, we assume that the first four individuals to the resources can share it, so that the probability that bird $i$ gets the resource is the probability of that $v_i(1)$ is one of the four highest opinions over all signaling events and $i$'s fitness is proportional to this probability.
+Finally, we consider a combination of predation and access to scarce resources.
+To do well, a bird must both survive and get to eat.
+In this case, $i$'s fitness is proportional to $P(\text{$i$ survives})\times P(\text{$i$ gets to eat})$.
+
+\subsection{Optimization methods }
+We use two different methods for analyzing the evolutionary dynamics of the strategies $k_i$.
+First, we identify the evolutionarily stable strategy.
+The ESS is that strategy such that, if all members of the group use it, no other strategy can invade.
+Equivalently, if all members of the group use the ESS, no individual has an incentive to use another strategy.
+The probability of a new strategy invading the group depends on the relative fitness of the invader and resident strategies.
+The details on to find the ESS in a finite group are provided in the Appendix.
+Assuming the group all uses the same strategy allows us to exhaustively search over all strategies to find the ESS.
+However, it is a rather strict constraint.
+Our second method of optimization allows us to consider how the birds should optimize their strategies in the context of a heterogeneous population.
+Given a random set of initial strategies, we allow each bird to choose the strategy that would give it the highest fitness, given the rest of the strategies being used.
+We repeat this process until the birds reach an equilibrium set of strategies.
+We repeat this over many initial sets of strategies to find average properties of this greedy optimization process.
+
+\subsection{Group-level properties }
+In our model, the information each bird can gather from its neighbors about an environmental signal has direct consequences on its fitness.
+It may also be the case that being in a flock whose members have the same opinion is beneficial.
+There are two ways to measure how well a flock can reach consensus, correlation length and $\mathscr{H}_2$ robustness, both of which require that we introduce noise into the opinion dynamics of the flock.
+
+In the absence of an environmental signal, the flock will eventually reach a consensus state such that $\lim_{t\to\infty}v(t)=\alpha\vec{1}$ for some $\alpha\in\R$, where $\alpha$ depends on the initial conditions.
+In the presence of an environmental signal, $\lim_{t\to\infty}v(t)=\vec{1}$.
+If we introduce noise into the opinion dynamics, the steady state opinion vector will no longer be deterministic but will instead be described by a probability distribution.
+We will take the distribution of the steady state opinion vector, $\lim_{t\to\infty}v(t)$, to be known for the moment and will come back to it shortly.
+
+Since the consensus value $\alpha$ is not always well-defined and since we are interested in the group's ability to reach consensus, but not the actual value of consensus, we subtract the average opinion from the opinion vector.
+If $V=\sum_iv_i/N$, then we define $z=v-V\vec{1}$.
+At steady state, $\lim_{t\to\infty}E[z(t)]=\vec{0}$ as long as the flock is connected.
+If we know the distribution of opinion vectors at steady state, we can also find the distribution of $z$ at steady state, and therefore the covariance matrix $C_z=\lim_{t\to\infty}E[z(t)z(t)^T]$.
+The correlation matrix $\bar{C}$ is defined by $\bar{C}_{ij}=C_{zij}/\sqrt{C_{zii}C_{zjj}}$.
+By plotting the correlations $\{\bar{C}_{ij}\}$ against the distance between the birds, we can find the correlation length $\lambda$ at which the correlations cross $0$.
+(As shown in Fact \ref{sum_cov}, $\sum_{ij}C_{zij}=0$, so there must be a distance at which the covariances and hence the correlations drop below $0$.)
+
+The second group-level property, $\mathscr{H}_2$ robustness requires another manipulation of the opinion vector.
+We define a rotation matrix $R\in\R^N\times\R^N$ such that the last row of $R$ is equal to $\frac{1}{\sqrt{N}}\vec{1}^T$ and each of the first $N-1$ rows are normal and orthogonal to $\vec{1}$.
+We will denote the first $N-1$ rows of $R$ as $Q\in\R^{N-1}\times\R^N$.
+%Note that since $R^TR=RR^T=I_N$, $Q^TQ=I_N-\frac{1}{N}\vec{1}\vec{1}^T$ and $QQ^T=I_{N-1}$. 
+We now define $\bar{y}=Rz$.
+Then $\bar{y}_N=0$ and if $y=(\bar{y}_1\dots\bar{y}_{N-1})^T$ then $y=Qz=Q(v-V\vec{1})=Qv$, so that $y$ represents the component of $v$ that is orthogonal to the consensus vector $\vec{1}$ \cite{Young:2010fk,Young:2013kx}.
+The $\mathscr{H}_2$ robustness $\rho$ is defined as
+\begin{equation*}
+\rho=\frac{1}{\lim_{t\to\infty}\sqrt{E[y(t)^Ty(t)]}}.
+\end{equation*}
+$\rho$ measures how far away from consensus the opinion vector is on average at steady state. The close relationship between these two measures is discussed in more detail in the Appendix.
+%Since $y=Qz$, if the same distribution of the opinion vector is used to find both $C_z$ and $E[||y||]$, then $E[||y||]=E[y^Ty]=\Tr(E[y^Ty])=\Tr(E[yy^T])=Tr(QC_zQ^T)$, so the two group properties are closely related.
+
+What we have not specified so far is how to find the distribution of the opinion vector at steady state, which is needed to find both $\rho$ and $\lambda$.
+The same probability distribution could be used to find both.
+However, the two measures come from two different bodies of literature, $\lambda$ from statistical physics and $\rho$ from control theory, and the conventional probability distribution for each measure is different.
+
+The Boltzmann distribution is the standard way in which to write an equilibrium distribution in statistical physics.
+In physics, those states with the highest energy are least likely.
+In our system, those sets of opinions that are least in consensus will be less likely.
+The distribution is only well-defined if there are symmetric interactions.
+Since $L$ is only symmetric if $A$ is, we symmetrize $A$ by defining $S=\frac{1}{2}(A+A^T)$ and
+\begin{equation*}
+P_{ij}=\left\{\begin{array}{ll} S_{ij} & \text{ if $j\neq i$}
+\\ -\sum_lS_{il} & \text{ if $j=i$}
+\end{array}\right..
+\end{equation*}
+If $A$ is symmetric, then $S=A$ and $P=L$. To incorporate the environmental signal, we can define $P_\text{s}=P-B$.
+We can then use the following expressions to represent the ``energy'' of an opinion vector:
+%\begin{equation*}
+$H(v)=-v^TPv \text{ without the signal and } H_\text{s}(v)=-(v-\vec{1})^TP_\text{s}(v-\vec{1}) \text{ with the signal. }$%\label{Hamiltonian}
+%\end{equation*}
+This gives the Boltzmann distributions
+\begin{equation*}
+P(v)=\frac{1}{Z}\exp\left(-\frac{1}{2}H(v)\right) \text{ without the signal and } P_\text{s}(v)=\frac{1}{Z_\text{s}}\exp\left(-\frac{1}{2}H_\text{s}(v)\right) \text{ with the signal.
+}\label{Boltzmann}
+\end{equation*}
+where $Z=\int_\R^N\exp\left(-\frac{1}{2}H(v)\right)dv$ and $Z_\text{s}=\int_\R^N\exp\left(-\frac{1}{2}H_\text{s}(v)\right)dv$.
+Given these probability distributions, we can find $C_z$ with and without the signal, and therefore the correlation length $\lambda$ with and without the signal.
+Details on these calculations are provided in the Appendix.
+
+In control theory, it is conventional to write down the noisy dynamics leading to a steady state distribution rather than to assume the form of the steady state distribution.
+Specifically, we consider the noisy opinion dynamics of the form
+\begin{equation*}
+\dot{v}=Lv+K\xi \text{ without the signal and } \dot{v}=L_\text{s}v+B\vec{1}+D\xi \text{ with the signal,}
+\end{equation*}
+where $\xi$ such that $E[\xi(t)]=\vec{0}$ and $E[\xi(t)\xi(s)^T]=\frac{1}{2}\delta(t-s)I$ and $D$ is a diagonal matrix with $\sqrt{k_i}$ along the diagonal.
+This choice of $D$ assumes that a bird with more neighbors makes noisier estimates of the many opinions it is trying to estimate.
+\citet{Young:2010fk} showed that if $\Sigma_y=\lim_{t\to\infty}E[y(t)y(t)^T]$ than, in the absence of the signal $\Sigma_y$ satisfies the Lyapunov equation
+\begin{equation*}
+\bar{L}\sigma_y+\sigma_y\bar{L}=-QD^2Q^T,
+\end{equation*}
+where $\bar{L}=QLQ^T$. A proof of this claim and details on how to calculate $\Sigma_y$ with a signal present are provided in the appendix. Once we find $\Sigma_y$, then 
+\begin{equation*}
+\rho=\frac{1}{\lim_{t\to\infty}\sqrt{E[y(t)^Ty(t)]}}=
+%\frac{1}{\lim_{t\to\infty}\sqrt{E[\Tr(y(t)^Ty(t))]}}=
+\frac{1}{\lim_{t\to\infty}\sqrt{E[\Tr(y(t)y(t)^T)]}}=
+\frac{1}{\lim_{t\to\infty}\sqrt{E[\Tr(\Sigma_y)]}}.
+\end{equation*}
+
+%%%%%%%%%%
+%The first measurement of the flock's ability to reach consensus is what has been called the $\mathscr{H}_2$ robustness of the network, which measures how far away from consensus the flock is at equilibrium.
+Specifically, we consider the noisy opinion dynamics of the form
+%\begin{equation*}
+%\dot{v}_i=\sum_jL_{ij}v_j+\sqrt{k_i}\xi_i,
+%\end{equation*}
+%where $\xi_i$ is a random variable such that the vector $\xi(t)$ satisfies $E[\xi(t)]=\vec{0}$ and $E[\xi(t)\xi(s)^T]=\frac{1}{2}\delta(t-s)I$ where $I$ is the $N\times N$ identity matrix.
+In the absence of noise, the system would reach consensus with each $v_i=\alpha$ for some $\alpha\in\R$.
+To consider deviations from this consensus state, it is useful to define a matrix $Q\in\R^{N-1}\times\R^N$ such that each row of $Q$ is orthogonal to $\vec{1}$ and the rows are pairwise orthogonal.
+Young et al.
+show in \cite{Young:2010fk,Young:2013kx} that this is equivalent to a $Q$ such that $QQ^T=I_{N-1}$ and $Q^TQ=I_N-\frac{1}{N}\vec{1}\vec{1}^T$.
+%
+%There are two ways to describe equilibrium properties of a flock sharing information over a network.
 The first is to write down a set of differential equations with noise to describe how each bird's opinion changes over time and then to analyze the distribution of these opinions after the dynamics have reached equilibrium \cite{Young:2010fk,Young:2013kx}.
 The second is to write down an explicitly probability distribution of the opinions at equilibrium and to analyze this distribution.
-WIth either approach, we consider a vector of opinions $v$ where $v_i$ a a scalar that represents the $i^\text{th}$'s bird opinion about an unspecified behavior, for example where or how fast to fly.
+With either approach, we consider a vector of opinions $v$ where $v_i$ a a scalar that represents the $i^\text{th}$ bird's opinion about an unspecified behavior, for example where or how fast to fly.
 Both approaches rely on a network of connections between the birds, which we will describe first.
 Let $A$ be the adjacency matrix such that
-
-\begin{equation}
-A_{ij}=\left\{\begin{array}{l l l}
-1 & \text{ if $i$ pays attention to $j$ }
-\\0 & \text{ otherwise }
-\end{array}\right.
-\end{equation}
-If $k_i$ is the number of neighbors node $i$ has, then $k_i=\sum_kA_{ik}$.
+%
+%\begin{equation}
+%A_{ij}=\left\{\begin{array}{l l l}
+%1 & \text{ if $i$ pays attention to $j$ }
+%\\0 & \text{ otherwise }
+%\end{array}\right.
+%\end{equation}
+%If $k_i$ is the number of neighbors node $i$ has, then $k_i=\sum_kA_{ik}$.
 We assume that having more neighbors means paying less attention to each so we normalize $A$ to find $\bar{A}$: $\bar{A}_{ij}=\frac{A_{ij}}{k_i}$.
-
-The Laplacian of this matrix is given by $L=\bar{A}-I$, which is symmetric only if $A$ is.
+%
+%
+%The Laplacian of this matrix is given by $L=\bar{A}-I$, which is symmetric only if $A$ is.
 We can symmetrize the interactions by defining $S=\frac{1}{2}(A+A^T)$ and $P$ such that $P_{ij}=S_{ij}$ for $j\neq i$ and $P_{ii}=-\sum_kS_{ik}$.
@@
 It is more common to use $L$ in a dynamics system approach, so that
-\begin{equation}
-\dot{v}_i=\frac{1}{k_i}\sum_i\left(v_j-v_i\right)+\sqrt{k_i}\xi_i
-\end{equation}
-\begin{equation}
-\dot{v}=Lv+D\xi \label{opinion_dynamics}
-\end{equation} 
-where $D$ is a diagonal matrix with entires $\{\sqrt{k_i}\}$ and $E[\xi(t)]=\vec{0}$, $E[\xi(t)\xi(s)^T]=\frac{1}{2}\delta(t-s)I$, and $E[v(0)\xi(t)^T]=\vec{0}$.
+%\begin{equation}
+%\dot{v}_i=\frac{1}{k_i}\sum_i\left(v_j-v_i\right)+\sqrt{k_i}\xi_i
+%\end{equation}
+%\begin{equation}
+%\dot{v}=Lv+D\xi \label{opinion_dynamics}
+%\end{equation} 
+%where $D$ is a diagonal matrix with entires $\{\sqrt{k_i}\}$ and $E[\xi(t)]=\vec{0}$, $E[\xi(t)\xi(s)^T]=\frac{1}{2}\delta(t-s)I$, and $E[v(0)\xi(t)^T]=\vec{0}$.
 This describes a system in which each bird changes averages the differences between its own opinions and those of its neighbors and changes its opinion to minimize these differences.
 Our choice of $D$ assumes that birds with more neighbors make noisier estimates of how to change their opinions.
-
-The Boltzmann distribution in statistical physics is a standard way in which to write an equilibrium distribution.
+%
+%The Boltzmann distribution in statistical physics is a standard way in which to write an equilibrium distribution.
 In physics, those states with the highest energy are least likely.
@@
 Specifically, we will use the following expression to represent the ``energy'' of an opinion vector
-\begin{equation}
-H(v)=-v^TPv \label{Hamiltonian}
-\end{equation}
-and the Boltzmann distribution
-\begin{equation}
-P(v)=\frac{1}{Z}\exp\left(-\frac{1}{2}H(v)\right) \label{Boltzmann}
-\end{equation}
-where $Z=\int_\R^N\exp\left(-\frac{1}{2}H(v)\right)dv$.
-
-With the dynamical approach, in the absence of noise, the system would reach consensus with each $v_i=\alpha$ for some $\alpha\in\R$.
+%\begin{equation}
+%H(v)=-v^TPv \label{Hamiltonian}
+%\end{equation}
+%and the Boltzmann distribution
+%\begin{equation}
+%P(v)=\frac{1}{Z}\exp\left(-\frac{1}{2}H(v)\right) \label{Boltzmann}
+%\end{equation}
+%where $Z=\int_\R^N\exp\left(-\frac{1}{2}H(v)\right)dv$.
+%
+%With the dynamical approach, in the absence of noise, the system would reach consensus with each $v_i=\alpha$ for some $\alpha\in\R$.
 Similarly, with the statistical physics approach, consensus states with each $v_i=\alpha$ are most likely.
@@
 There is therefore a close correspondence between the two approaches.
-
-These covariance matrices allow us to define two group properties.
+%
+%These covariance matrices allow us to define two group properties.
 The first is $\mathscr{H}_2$ robustness, as used in \cite{Young:2010fk,Young:2013kx}.
@@
 Specifically,
-\begin{equation}
-\rho=\frac{1}{\lim_{t\to\infty}E[||y||]}=\frac{1}{\lim_{t\to\infty}\Tr(\Sigma_y)}. \label{H2def}
-\end{equation}
-The second property is the correlation length of the flock.
+%\begin{equation}
+%\rho=\frac{1}{\lim_{t\to\infty}E[||y||]}=\frac{1}{\lim_{t\to\infty}\Tr(\Sigma_y)}. \label{H2def}
+%\end{equation}
+%The second property is the correlation length of the flock.
 This is the distance over which birds' opinions tend to be positively correlated with each other.
 If we know the distances between the birds and the covariance matrix $C_z$, we can plot the correlations $\{E[z_iz_j]\}$ between each pair of birds' deviations from consensus against the distance between the birds and then identify the correlation length $\lambda$ by finding the distance at which these correlations become $0$.
-
-So far, we have considered birds whose opinions only depend on the opinions of the rest of the flock.
+%
+%So far, we have considered birds whose opinions only depend on the opinions of the rest of the flock.
 We can additionally incorporate environmental information by assuming that there is an environmental signal such that more informed birds have opinions closer to $1$.
-If we let $B$ be a matrix with $B_ii=1$ if $i$ can perceive the signal and $0$ otherwise, then the dynamical system becomes
-\begin{equation}
-\dot{v}=Lv+B(\vec{1}-v)+D\xi =L_\text{s}v+B\vec{1}+D\xi \label{opinion_dynamics2}
-\end{equation}
-where $L_\text{s}=L-B$. Similarly, the statistical physics model becomes
-\begin{equation}
-P(v)=\frac{1}{Z}\exp\left(-\frac{1}{2}(v-\vec{1})^T(-P+B)(v-\vec{1})\right)=\frac{1}{Z}\exp\left(\frac{1}{2}(v-\vec{1})^T(-P_\text{s})(v-\vec{1})\right)
-\end{equation}
-where $P_\text{s}=P-B$.
+If we let $B$ be a matrix with $B_{ii}=1$ if $i$ can perceive the signal and $0$ otherwise, then the dynamical system becomes
+%\begin{equation}
+%\dot{v}=Lv+B(\vec{1}-v)+D\xi =L_\text{s}v+B\vec{1}+D\xi \label{opinion_dynamics2}
+%\end{equation}
+%where $L_\text{s}=L-B$. Similarly, the statistical physics model becomes
+%\begin{equation}
+%P(v)=\frac{1}{Z}\exp\left(-\frac{1}{2}(v-\vec{1})^T(-P+B)(v-\vec{1})\right)=\frac{1}{Z}\exp\left(\frac{1}{2}(v-\vec{1})^T(-P_\text{s})(v-\vec{1})\right)
+%\end{equation}
+%where $P_\text{s}=P-B$.
 Again, we can find the covariance matrices $\Sigma_z$, $\Sigma_y$, $C_z$, and $C_y$.
@@
 We define $\rho$ and $\lambda$ as above using the matrices $\Sigma_y$ and $C_z$ in the presence of a signal.
-
-\subsection{Individual-level fitness }
-To find the optimal strategies from the individuals' perspectives, we need a measure of individual-level fitness.
-We use the dynamical system to describe how well each bird learns about the environmental signal.
-For a given matrix $L$, the expected opinion $q_i$ of each individual satisfies the equations
-\begin{equation}
-\dot{q}=L_\text{s}q+B\vec{1}. \label{expected_opinion}
-\end{equation}
-To generate a matrix $L$, we distribute the individuals randomly in space (specifically, in a unit square).
-has a strategy $k_i$ that dictates how many neighbors it has, so that $k_i=\sum_{j\neq i}{\bf I}(A_{ij}\neq 0)$.
-Each individual then pays attention to its $k_i$ nearest neighbors.
-One individual is chosen to be a receiver; it and all those individuals within a radius $r$ of the receiver perceive, so that $B_{ii}=1$ for all those birds within this radius.
-We do this $1000$ times and find $q(1)$ for each matrix, which we refer to as a signaling event.
-This is meant to simulate a situation in which such that information is present only intermittently and otherwise the birds are moving randomly through space and restructuring their network.
-If, instead of paying attention to the closest neighbors, a bird choose $k_i$ neighbors at random out of $N$ birds, $N_\text{I}$ of whom are informed, the number of informed neighbors follows a hypergeometric distribution:
-\begin{equation}
-P(j \text{ informed neighbors} )=\frac{{N_\text{I}\choose j}{N-N_\text{I}\choose k_i-j}  }{{N\choose k_i}}. \label{hypergeom}
-\end{equation} 
-This distribution is not an exact description of the probability of $j$ informed neighbors in our model  since it ignores spatial proximity, but it is a useful approximation for considering the advantages and disadvantages of each strategy $k_i$.
-{\red NOTE: I'M RUNNING AN ANALYSIS WHERE FITNESS IS BASED SOLELY ON THIS DISTRIBUTION, NOT ON THE FULL SPATIAL STRUCTURE OF THE MODEL.
-I SHOULD HAVE RESULTS SOON.
-I EXPECT THEM TO BE SIMILAR TO THE FULL MODEL, BUT I WANTED TO MAKE SURE AND TO USE IT AS A WAY OF EXPLAINING THE FULL MODEL RESULTS.}
-
-We consider signals that provide information about two types of content: predators and food.
-If the signal is information that a predator is present, we assume that whichever individual is least aware of the signal will be predated, i.e.
-the individual $i$ such that $q_i$ is lowest.
-Then, fitness is proportional to the probability of surviving, i.e.
-$1$ minus the probability of being the least informed over many signaling events.
-If the signal is information that resources are available, we consider two levels of resource abundance.
-If there are scarce resources, we assume that whichever individual is most aware of the signal will get access to the resource, i.e.
-the individual $i$ such that $q_i$ is highest.
-In this case, fitness is proportional to the probability of getting the resource, i.e.
-the probability of being the most informed over many signaling events.
-If there are abundant resources, we assume that the first four individuals to the resources can share it, so that fitness is proportional to the probability of being one of the four most informed birds.
-Finally, we consider a combination of predation and resources: fitness is proportional to the $1$ minus the probability of being predated times the probability of being the most informed.
-The behavior that the birds can change to improve their fitness is the number of other birds they pay attention to, $k_i$.
-In particular, each individual has a strategy $k_i$ that dictates how many neighbors it has, so that $k_i=\sum_{j\neq i}{\bf I}(A_{ij}\neq 0)$.
-
-\subsection{Optimization methods }
-To understand what strategies we might expect to find, we are interested in identifying the optimal strategies.
-We does this in two ways.
-First, using the framework of adaptive dynamics in finite populations, we identify the evolutionarily stable strategy.
-This framework assumes that there is a homogeneous population into which a mutant individual tries to invade.
-If strategies change on a learning rather than an evolutionary timescale, or if there is just a lot of variation in the population, we might expect individuals to try to optimize their strategies in the context of a heterogeneous population.
-This is our second method of optimization.
-Given a random set of initial strategies over the group, we allow each bird to choose the strategy that would be best given the rest of the strategies being used and repeat this process until the birds reach an equilibrium set of strategies.
-We repeat this over many initial sets of strategies to find average properties of this optimization process.
+%
+%
+%
 
@@
 
-\subsection{The ESS number of neighbors depend on the content of the signal. }
+\subsection{The ESS number of neighbors depends on the content of the signal. }
 
@@
 Selection, therefore, acts to improve a bird's ability to learn about the signal when it cannot perceive it.
-In Appendix Claim \ref{exp_approx} we show that an uninformed bird's opinion can be approximated by $\sum_{\text{informed }j}L_{ij}$, that is the total weight given to informed birds.
+In Appendix Claim \ref{exp_approx} we show that an uninformed bird's opinion can be approximated by $v_i(1)\approx v(0)+\frac{1}{2}\sum_{\text{informed }j}L_{ij}$.
 Being the most well-informed uninformed bird, therefore, amounts to having the highest proportion of informed neighbors and being the least well-informed amounts to having the smallest proportion of informed neighbors.
@@
 
+%caveats: scalar opinion, agnostic to what that opinion refers to in reality, static network as flock responds to environmental signal
+
+%expect our results to hold for metric neighborhoods as well as topological networks
 \newpage
@@
 We can approximate $v(t)\approx V\exp(t\Lambda)V^{-1}v_0+V(tI+\frac{t^2}{2}A)V^{-1}f$ if the eigenvalues are not too far from $0$.
-In the case that $A=L_\text{f}$ and $f=B\vec{1}$, this suggests that for $i$ that cannot perceive the signal, $v_i(t)\approx \frac{t^2}{2}\sum_{\text{informed }j }L_{ij}.$ This approximation is reasonable for $i$ such that $B_{ii}=0$, but worse for $i$ such that $B_{ii}\neq 0$.
+In the case that $A=L_\text{f}$ and $f=B\vec{1}$, this suggests that for $i$ that cannot perceive the signal,  for small $t$, $v_i(t)\approx v_i(0)+\frac{t^2}{2}\sum_{\text{informed }j }L_{ij}.$ This approximation is reasonable for $i$ such that $B_{ii}=0$, but worse for $i$ such that $B_{ii}\neq 0$.
 \end{claim}
@@
 The rest of the $N-N_\text{I}$ eigenvalues are between $-2$ and $0$ and, if $k$ is the index of one of these eigenvalues, $V_{ik}$ is large if and only if $i$ is uninformed.
-Therefore, the eigenvalue approximation is worst for the informed individuals and best for the uninformed individuals.
-Using the eigenvalue approximation, $v(t)$ becomes
+%NEEDS SUPPORT
+Therefore, the eigenvalue approximation is worst for the informed individuals and best for the uninformed individuals. Using the eigenvalue approximation, $v(t)$ becomes 
 \begin{align*}
@@
 \end{align*}
-Ignoring the informed individuals and the initial conditions, this gives $v_i(t)\approx \sum_{\text{informed } j}L_{ij}$.
+Assuming $t$ is small, for uninformed individuals, this gives $v_i(t)\approx v_i(0)+ \frac{t^2}{2}\sum_{\text{informed } j}L_{ij}$.
 \end{pf}
@@
 \begin{theorem} \label{lyapeq_theorem}
-Given any positive symmetric matrix $Q>0$ there exists a unique positive symmetric ] matrix $\Sigma>0$ satisfyying $A\Sigma+\Sigma A^T=Q$ if and only if the system $\dot{x}=Ax$ is globally asymptotically stable.
+Given any positive symmetric matrix $Q>0$ there exists a unique positive symmetric matrix $\Sigma>0$ satisfyying $A\Sigma+\Sigma A^T=Q$ if and only if the system $\dot{x}=Ax$ is globally asymptotically stable.
 \end{theorem}
@@
 
-\begin{fact}
+\begin{fact} \label{sum_cov}
 If $V=\sum_iv_i/N$ and $C=\left(E[(v_i-V)(v_j-V)]\right)_{ij}$, then $\sum_{ij}C_{ij}=0$.
@@
 \begin{equation}
-r=\lim_{t\to\infty}\frac{1}{E[||y||]}=\lim_{t\to\infty}\frac{1}{\Tr E[yy^T]}. \label{H2_robustness}
+\rho=\lim_{t\to\infty}\frac{1}{E[||y||]}=\lim_{t\to\infty}\frac{1}{\Tr E[yy^T]}. \label{H2_robustness}
 \end{equation}
@@
 %\nocite{*}
-\bibliographystyle{plain}
+%\bibliographystyle{plain}
+\bibliographystyle{plainnat}
 \bibliography{info_evo}
